### 2013-12-19 Writing/MS.tex
first commit

\documentclass[]{article}
\usepackage[margin=.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{brush_style}
\title{Species persistence under climate and fishing}
\author{Emma Fuller, Eleanor Brush, Malin Pinsky}
\date{}

\begin{document}
\maketitle

\section{Abstract}

When the climate changes, the area in which organisms can survive and reproduce will move through space. 
This change does not occur in isolation but rather appears on a background of other disturbances. We use an 
integrodifference model explicitly accounting for dispersal, reproduction, a shifting environment, and 
harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
persistence. We found the harvesting rate and the rate of environmental shift that are needed to allow the 
population to persist, studied how these critical parameters depend on the growth rate and dispersal behavior 
of the population, and measured the interactions between the stressors. In particular, we found positive 
synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. 
Finally, we introduced two conservation techniques into simulations of the population model -- threshold 
population levels below which harvesting is not allowed and marine protected areas (MPAs) -- and found that 
the robustness of the population to the two stressors improved.

\section{Introduction}

There are many stressors that can disturb an ecosystem. Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}. However, disturbances rarely occur in isolation.  It is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}. A perturbation that has little effect when affecting a system individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}. Synergistic interactions between stressors are especially worrying if an animal population can persist in the face of a single stressor but not in the presence of multiple stressors (e.g.  \citet{Pelletieretal2006}). However, it remains hard to predict when disturbances are likely to interact in ways not predicted by the effects of individual perturbations.

Climate change is a prime example of a stressor occurring in a background of multiple other, possibly interacting, disturbances. These include pollution, habitat fragmentation, invasive species, and over-exploitation \citep{Wilcoveetal1998, Salaetal2000, MEA2005}. It is expected that  species will move their home ranges in response to climate change. Many marine fish are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}. These species are faced not only with temperature-driven range range shifts but also with anthropogenic stressors including ocean acidification and commercial fishing \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}. Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact in the future.

Many methods have been proposed to help mitigate the effects of harvesting on fish populations. Marine protected areas (MPAs), which provide stepping stones to help species keep up with a changing environment, have been suggested as a key form of climate insurance \citep{Thomasetal2012, Hannahetal2007}. MPAs are frequently recommended for both conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}. Threshold harvesting levels below which no fish may be harvested have also been proposed as a means of both optimizing yield and conserving resources \citep{Landeetal1995, Landeetal1997}.

A wide range of models have been used to study the effect of climate
change and harvesting on fish populations.  A common approach to predicting future population distributions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs). These statistical models have been
used to find correlations between the presence or absence of fish
species and biophysical properties of the environment such as mean or maximum
temperatures, rainfall, or salinity, to explain and predict how species
ranges will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}. Similarly,
projection models of future climate conditions have been used to predict
where conditions will be favorable for fish populations in the future
\citep{Cheungetal2008, Cheungetal2009}.  Despite these models' widespread
adoption, they have  been criticized for being oversimplified
as they lack species interactions, dispersal and reproductive processes
\citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.

Some recent work on range shifts has addressed these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}, which have been highlighted as important factors in fish populations' vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.  These models only address one disturbance, climate-driven range shifts, and do not consider other anthropogenic stressors.  The models of Zhou and Kot
\citep{ZhouKot2011} and Berestycki et al. \citep{Berestyckietal2009} are of
intermediate complexity (sensu
\citep{Gaylordetal2005}).  Work that consider both climate-driven
range shifts and fishing is often case-specific and detailed, including
multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011,
Barangeetal2011, Howardetal2013}. These cumulative
impacts are important for management and conservation planning
\citep{Allisonetal2009}.  However, these models are
so complex that understanding the relative importance of particular
drivers, disturbances, and interactions is difficult (but see
\citep{Nyeetal2013} for an approach using ecosystem-level models to discern
relative importance of disturbances).

We analyzed an integrodifference model to study the effect of temperature-driven range shifts and
harvesting, two important interacting disturbances, on the persistence of a marine population. In particular, we extended the model of
Zhou and Kot \citep{ZhouKot2011}, which describes a fish population
reproducing and dispersing while their range shifts, by
additionally including harvesting pressure and introducing two different
conservation measures. We found that the disturbances interact
synergistically. We also found that MPAs can help a species persist with
higher harvesting pressure, but they do not change the maximum speed of
habitat shift with which a species can keep up.

\section{Methods}

We studied the dynamics of a fish population constrained to a single one-dimensional habitat patch by their inability to reproduce outside of the patch.  This viable habitat patch (here after `patch') is shifting at a fixed velocity and fish at each point in the patch can be harvested.  Given the reproductive rate and average dispersal distance, we determined the climate velocity and harvesting rate that would drive the population extinct.  We implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to 
determine how these management strategies affect population persistence.

\subsection{The Model }

In our model, the adults from the current year produce offspring according to a recruitment function and these 
offspring disperse across the one-dimensional world according to a dispersal kernel to become the next 
generation's adults. The adults are subjected to harvesting before they produce offspring so that only a 
proportion of the fish survive to reproduce. We incorporated these processes-- recruitment, harvesting, and 
dispersal-- into an integrodifference model to describe how the population changes over time. If $n_t(x)$ 
is the density of fish at position $x$ at time $t$, then the density of fish at the next generation is given by

\begin{equation}
n_{t+1}(x)=\int^{\frac{L}{2}+ct}_{-\frac{L}{2}+ct}k(x-y)f((1-h)n_t(y))dy \label{integrodifference}
\end{equation}

\noindent where $h$ is the proportion of adults harvested, $f(n)$ is the recruitment function giving the number of 
offspring produced by a population of size $n$, $k(x-y)$ is the dispersal kernel giving the probability of the 
larvae traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it 
shifts across space. We provide a list of variables and functions in Table 1. The functional form of the 
recruitment function only affects persistence through how quickly recruitment increases when the population 
increases from $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$. We therefore do not specify a 
recruitment function here. Similarly, the population's ability to persist only depends on how quickly harvesting 
increases when a small number of fish are introduced where they were previously absent, $h'(0)$. We 
therefore only considered a proportional harvesting function.

If a population of fish is going to persist in the face of a shifting environment, it must move along with the patch 
of environment in which it can reproduce. At equilibrium, the population will be described by a traveling wave, 
where the density of fish at a given point in space will change but the density of fish at a location relative to the 
shifting  patch will not. We therefore seek to describe how the population is distributed over the viable 
patch as it shifts through the world. We write this as $n^*(\bar{x})$, the density of fish at each point $\bar{x}$ in 
the shifting frame $\left[-\frac{L}{2}, \frac{L}{2}\right]$ As in \citet{ZhouKot2011}, the traveling wave $n^*$ must 
satisfy

\begin{equation}
n^*(\bar{x})=\int^{\frac{L}{2}}_{-\frac{L}{2}}k(\bar{x}+c-\bar{y})f((1-h))n^*(\bar{y}))d
\bar{y} \label{traveling_pulse}
\end{equation}

One solution to Equation \ref{traveling_pulse} is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it. If a population becomes very small (or if we introduce a small 
population), one of two things can happen. First, the population may crash and the trivial traveling pulse 
without any fish may appear again. Second, those small numbers may increase and form a stable population. 
In this sense, a small population can be thought of as a perturbation to the trivial traveling pulse. If the trivial 
pulse is stable, the system will return to the trivial pulse even after a perturbation in the form of the introduction 
of a small population. If a population is to persist, even when it is small it must be able to avoid extinction and 
grow. For this to be the case, the trivial pulse must be unstable to small perturbations.

For the separable Gaussian kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy

\begin{equation}
n^*(x)=\sum^\infty_{i=1}
a_i(x)\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f((1-H(n^*(y)))n^*(y))dy=\sum^\infty_{i=1}m_ia_i(x), \label{separable_integrodifference}
\end{equation}

\noindent where the $m_i$ satisfy the recursive equations

\begin{equation}
m_i=\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f\left((1-h)\sum^\infty_{i=1}m_ia_i(x)\right)
dy. \label{recursive_m}
\end{equation}

\noindent Equation \ref{recursive_m} allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.

In general, we would like to know the rate of environmental shift and the harvesting rate such that if the 
environment moves more quickly or we harvest more severely then the population will not be able to persist. 
We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$. 
We can find these rates by finding the parameters that make the trivial pulse unstable. Evaluating stability is in 
general difficult to do analytically. It becomes easier if the dispersal kernel is separable into its dependence on the 
source of larvae and its dependence on the destination of the larvae,
i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$. We used 
simulations to analyze dispersal kernels that are not amenable to this method.

In our analyses, we used a Gaussian kernel,

\[k(x-y)=\frac{1}{\sqrt{2\sigma^2\pi}}e^{\frac{-(x-y)^2}{2\sigma^2}}\]

\noindent To make stability easier to analyze, we approximated this kernel, as described in the Appendix. In our 
simulations, we used an analytically intractable Laplace kernel,

\[k(x-y)=\frac{1}{2}be^{-b\vert x-y\vert}\]

\noindent For each kernel, we can find the expected distance an offspring will travel, $\langle d \rangle$, which we will use to compare the two kernels.  For each kernel, the population's ability to persist depends on properties of the population itself: how far larvae 
tend to disperse ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment: the 
length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
$). If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than 
the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.

\subsection{Simulations }

We used simulations to extend the basic integrodifference model in two ways that make it analytically 
intractable. First, we examined the sensitivity of the model to choice of dispersal kernel by using a Laplace 
dispersal kernel. Second, we examined harvesting rules more complex than harvesting a constant proportion of the population. Whereas the 
analytical results do not depend on the functional form of recruitment $f$, to perform simulations we must 
specify a recruitment function. We chose to use a Beverton-Holt model,

\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]

\noindent where $R_0$ is the intrinsic growth rate and $K$ is the carrying capacity.  In the first generation, we seeded the world with 50 individuals at a single point, as in 
\citep{ZhouKot2011}. We first allowed the population to reach equilibrium without harvesting or climate shift.  We then 
added harvesting pressure, allowed the population to again reach equilibrium, and finally added climate 
change by moving the viable patch.

We added harvesting pressure by harvesting a constant proportion of the population, in order to confirm our analytical results. 
We also extended our analytical work by evaluating the effect of a threshold harvest rule and marine protected 
areas (MPAs). With a threshold rule, we evaluated the population at each point in space to determine how 
much harvesting should occur. If the population abundance was below the designated threshold, no 
harvesting occurred. If the population exceeded the threshold, then a proportion of the `surplus' individuals 
were harvested. To implement an MPA management strategy, we used two commonly advocated approaches: 
MPAs designed to improve fisheries yields and MPAs designed for primarily for conservation. These networks 
of MPAs were introduced into our simulations by designating segments of space in which harvesting was 
forbidden (i.e.~harvesting rates were equal to $0$). The conservation MPAs had a length of twice the average 
dispersal distance and had a distance of 4 times the average dispersal distance between them. Fisheries MPAs had a length of 1/3 of the 
average dispersal distance and had a distance of 2/3 of the average dispersal distance between them.

In our simulations, we defined the critical rate of environmental shift and the critical harvest rate as model runs 
in which the total population at equilibrium was less than 0.001. Equilibrium is calculated as the mean of 300 time steps once the difference in biomass between time step $t$ and $t+1$ no greater than $0.0360515$. Equilibrium biomass and catch are defined as the total amount of fish present and harvested, respectively, at equilibrium.

\subsection{Calculating Synergy }

We wanted to understand whether the two stressors interact 
additively, synergistically, or antagonistically \citep{Crainetal2008} for both our analytical model and simulations. In order to quantify the effect of the 
stressors, we found the total biomass of the population when it reached an equilibrium traveling pulse and 
compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.

We measured the effects of the stressors by comparing the equilibrium biomass of the stressed population to 
the equilibrium biomass of the unstressed population. We use $B_0$ to denote the equilibrium biomass 
without either stressor, $B_\text{h}$ the equilibrium biomass with harvesting but a constant environment, $B_\text{c}$ the 
equilibrium biomass with a shifting environment but no harvesting, and $B_\text{hc}$ the equilibrium biomass with 
both stressors. For each stressor or combination of stressors, we found the drop in  biomass caused 
by stressor $s$,

\[E_\text{s}=B_0-B_\text{s}.\]

\noindent If the stressors do not interact, the drop caused by both stressors would be the sum of the drops caused by 
either individually. The synergy is therefore defined as

\[S = E_\text{hc}-\left(E_\text{h}-E_\text{c}\right).\]

\noindent If the stressors aggravate each other, the effect of both stressors is worse than would be expected from 
considering either stressor individually, and synergy is positive. If the stressors alleviate each other, the effect 
of both stressors is better than would be expected from considering either stressor individually, and synergy is 
negative. If the effect of both stressors is exactly as expected from considering either stressor individually, 
there is no interaction and no synergy.

\section{Results}

\subsection{Interactions Between Stressors }

The equilibrium biomass of the population decreases as either the harvesting pressure 
increases or the environmental shifts more quickly (Figure \ref{eqbiomass}). If the equilibrium biomass is $0$, this 
indicates that the harvesting pressure has exceeded the critical harvesting rate $h^*$ and the environmental 
is shifting more quickly than the critical rate of environmental shift $c^*$. As the harvesting rate $h$ increases, 
the critical rate of environmental shift $c^*$ decreases: the environment must move more slowly to 
accommodate the population growing more slowly (Figure \ref{critical}). Conversely, as the rate of 
environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{critical}). This 
means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
the environment starts shifting. The simulations replicate the analytical results with the critical speed $c^*$ 
declining as the critical harvest rate $h^*$ increases and vice versa.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.5\textwidth]{plots/eqbiomass.pdf}
\caption{The equilibrium biomass of the population decreases as either the harvesting rate or rate of environmental shift increases.  If either harvesting rate or rate of environmental shift are too high the population is driven to extinction. The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the equilibrium biomass.   These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
\label{eqbiomass}
\end{center}
\end{figure}

It is always the case that increasing the intrinsic growth rate, $R_0$, of the population increases the critical 
speed $c^*$ and the critical harvesting rate $h^*$, since a population that grows more quickly can recover 
more quickly from losses caused by these disturbances. However, whether or not dispersing farther is better 
depends on how quickly the environment is shifting (Figure \ref{critical}). When the environment is shifting 
slowly, dispersing farther is detrimental since many larvae will disperse too far away from the viable patch. 
When the environment is shifting quickly, on the other hand, dispersing farther can help the population persist 
because some larvae will disperse into the space that will become viable shortly in the future.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.5\textwidth]{plots/critical_rates.pdf}
\caption{As the rate of environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases.  In other words, less harvesting pressure is required to drive the population extinct.  Increasing the growth rate always increases the critical harvesting rate.  On the other hand, increasing average dispersal distance increases the critical harvesting rate depends on how quickly the environment is shifting: it is more advantageous to disperse farther when the environment is shifting more quickly.  The rate of environmental shift is plotted on the x-axis and the critical harvesting rate is plotted on the y-axis.  Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.  Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.  These results are from a Gaussian dispersal kernel with $L=1$.}
\label{critical}
\end{center}
\end{figure}

We found positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
{Synergy}).  In other words, a doubly stressed population loses more biomass than would be predicted from either stressor individually.  The stressors interact most strongly when they are both high, shortly before they drive the population extinct.  Our simulations with a Laplace kernel produce similar 
results. In the Appendix, we show analogous results for a sinusoidal dispersal kernel.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.75\textwidth]{plots/synergy.pdf}
\caption{There is positive synergy between the two stressors: a doubly stressed population loses more biomass than would be predicted from either stressor individually.  The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.  These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
\label{Synergy}
\end{center}
\end{figure}

\subsection{Management Strategies }

We found that thresholds of any amount alleviate the effects of harvesting, and the ability of the population to 
persist is recovered (Figure \ref{thresholds}). Thus when thresholds are in place, the harvesting rate no longer 
determines the critical rate of environmental shift $c^*$. We also examined the effect of marine protected 
areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and 
climate change parameters where the fish population could survive. With MPAs in place, the population had a 
slightly higher abundance along the edges of the patch where the population is limited by harvesting, which 
translated into a slightly increased critical harvest rate (Figure \ref{MPAs}). Additionally MPAs increased 
overall catch at the highest harvest rates under which the population could survive (Figure \ref{MPAs}).

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.45\textwidth]{plots/eqbiomass_sim.pdf}
\includegraphics[width=.45\textwidth]{plots/eqbiomass_thresh.pdf}
\caption{With threshold harvesting rules, such that all biomass above a threshold level was harvested, the population always reached an equilibrium biomass greater than $0$ so that threshold harvesting rules increased biomass at equilibrium and improved the population's ability to persist. Additionally, threshold harvesting rules remove the interaction between harvesting and environmental shift: the critical rate of environmental shift no longer depends on harvesting. The figure on the left corresponds to no thresholds and the figure on the right corresponds to threshold harvesting rule.  The x-axis shows the rate of environmental shift, the y-axis shows either the proportion of the population harvested or the threshold above which harvesting was allowed, and the color indicates the equilibrium biomass.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$. }
\label{thresholds}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.45\textwidth]{plots/eqbiomass_sim.pdf}
\includegraphics[width=.45\textwidth]{plots/eqbiomass_mpa.pdf}
%\includegraphics[width=.45\textwidth]{plots/eqbiomass_mpa_v2.pdf}
\caption{MPAs improved the population's ability to persist.  The figure on the left corresponds to no MPAs and the figure on the right corresponds to MPAs designed to optimize fishing yields.  MPAs designed to optimize conservation give similar results.  The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the equilibrium biomass.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.}
\label{MPAs}
\end{center}
\end{figure}

\section{Discussion}

We extended the model of Zhou and Kot \citep{ZhouKot2011} to study a fish population subject to both the climate-driven range shifts already included and harvesting. Previous work has highlighted the importance of 
reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]
{Hastingsetal2005} and the integrodifference model explicitly includes these processes. We found that 
the two disturbances interact synergistically with each other in some cases. However, we also found that 
placing thresholds on harvesting intensity can alleviate interactions between the two stressors and that spatial 
management increased the maximum harvesting rate at which the population could survive.

We found that the synergy between the two stressors is positive and that the synergy is greatest in the region 
of parameter space where the equilibrium biomass is smallest. Additionally, we found that the higher the 
growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a 
population can adjust to harvest and climate change. We find similar results from the analytically derived 
biomass and the simulation derived biomass. This indicates that this result is robust to changes in the 
dispersal kernel. Similarly, in experiments, \cite{Moraetal2007} found no interaction between harvesting and 
habitat fragmentation but did find an interaction between harvesting and warming temperature \citep
{Moraetal2007}. We chose to measure the effect of each stressor by the drop in biomass caused by the 
stressor, as in \cite{Crainetal2008}. In general, measuring synergy additively is more conservative than 
measuring synergy multiplicatively: stressors that have positive multiplicative synergy may have zero additive 
synergy \citep{Crainetal2008, Foltetal1999}. Since we found small levels of positive additive synergy between 
the two stressors, other measures of synergy might show even higher levels of interaction. This synergy and 
the fact that it most severely affects those populations whose persistence is most tenuous is worrisome from a 
conservation perspective. Harvesting levels or rate of environment shift that are sustainable individually 
together can drive a population to extinction. It is therefore important to take the effects of both stressors into 
consideration when designing conservation and management strategies.

The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the 
population's biomass at equilibrium and improved its ability to persist. These management strategies also 
affect how the two stressors interact with each other.  Protected areas have 
been advanced as a way to help organisms keep pace with range shifts as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, 
Hannahetal2007,Botsfordetal2001, Gaylordetal2005, 
HastingsBotsford2003,Thomasetal2012}.  While MPAs only change harvesting practices 
and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting 
and range shifts will help to better implement and place protected areas.

The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.  However, it  ignores many of the complexities present in marine fisheries. We do 
not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist 
over time. However, we found that qualitatively similar results about the interaction between climate and 
harvesting would hold for a model with a recruitment function with Allee effect. We also did not include age 
structure in our model. The effects of both harvesting and climate change may be different across different age 
classes; including this level of complexity is left for future work. Additionally, we did not include any 
mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.

In addition to these species-specific extensions, this modeling framework can easily be adjusted to consider 
species interactions, especially predator-prey pairs. Predatory species affect marine populations and 
considering a set of species that included both predators and prey would be equivalent to imposing additional 
stressors on the prey species whose interactions we would like to measure \citep{Lingetal2009, 
Gurevitchetal2000}. Considering multiple species would not only allow us to understand the effects of multiple 
stressors on prey populations but would also afford us the opportunity to study how community composition 
would change as a result of climate change, as has been observed in some marine systems \citep
{Urbanetal2012}.

Finally, our results suggest that particular combinations of harvesting and rate of environmental shift will affect 
some species more than others. Indeed, \citet{Perryetal2005} found that fish that shifted in response to 
warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, 
smaller sizes at maturity). Using a simple mechanistic model like the one we presented can incorporate these 
species interactions and multiple disturbances to understand whether specific life histories are likely to be 
selected over others as harvesting and/or range shifts increase.

\bibliography{fish}
\section{Appendix}

As in Zhou et al. \citep{ZhouKot2011}, let $k(x-y)$ be a dispersal kernel and let $f(y)$ be a recruitment function.  The integrodifference model describing the population over time is given by 
\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y))dy. \label{integro} \end{equation}

To find a traveling pulse, we are only interested in the population density as a function of the location within the patch rather than absolute position, $\overline{x}\equiv x-ct$.

\begin{equation}n^*(\overline{x})\equiv n^*(x-ct)=n_t(x).   \label{trav} \end{equation}

Then (\ref{integro}) giveas s us an expression for $n^*$:

\begin{align*}
n^*(\overline{x}-c)&=\int_{-L/2}^{L/2}k(\overline{x}-\overline{y})f(n^*(\overline{y}))d\overline y 
\\ \Rightarrow n^*(\overline{x})&=\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})f(n^*(\overline{y}))d\overline y \tag{*} \label{pulse}
\end{align*}
If $f(0)=0$, $n^*(\overline{x})\equiv 0$ for all $\overline{x}\in[-L/2,L/2]$ is a trivial solution to this problem, i.e. if there are no fish anywhere there won't be at any time in the future. The population can be said to be persistent if the trivial traveling pulse is unstable since even when there are very small population levels, the population won't crash to $0$.  To evaluate stability (i.e. persistence), we will introduce a small perturbation to the traveling pulse $n^*(\overline{x})$,

\begin{align*}
n_t(x)&=n^*(\overline{x})+\xi_t(x)
\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)f'(n^*(\overline{y}))\xi_t(y)dy \text{ by linearizing around the traveling pulse and using (\ref{pulse})}
\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)f'(0)\xi_t(y)dy \text{ if we're interested in the stability of the trivial traveling pulse}
\end{align*}
If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\R$ and $u:[-L/2,L/2]\to\R$, then 
\begin{align*}
\lambda u(x-ct-c)&=f'(0)\int_{-L/2+ct}^{L/2+ct}k(x-y)u(y-ct)dy
\\ \lambda u(\overline{x})&=f'(0)\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})u(\overline{y})dy
\end{align*}

Define the integral operator
$$ \psi_f(g)(x)=\int_{-L/2}^{L/2}f'(0)k(x+c-y)g(y)dy. $$
so that the perturbation to the traveling pulse will satisfy 
\begin{equation} \psi_f(u)(x)=\lambda u(x) \label{eigen} \end{equation}
Then the trivial traveling pulse is unstable when the dominant eigenvalue of $\psi_f$ is greater than $1$.

Let $f$ denote the recruitment function, let $h$ denote a harvesting function and let $m(y)=f(y-h(y))$, i.e. $m$ denotes the number of offspring after the adults have been harvested.  Note that $m'(0)=f'(0)(1-h'(0))$, assuming $h(0)=0$ (which must be the case).

Suppose $u$ is an eigenfunction of $\psi_f$ with eigenvalue $\lambda$.  Then 
\begin{align*}
\psi_m(u)(x)&=\int_{-L/2}^{L/2}m'(0)k(x+c-y)u(y)dy
\\&=(1-h'(0))\int_{-L/2}^{L/2}f'(0)k(x+c-y)u(y)dy
\\&=(1-h'(0))\psi_f(u)(x)
\\&=(1-h'(0))\lambda u(x)
\end{align*}
so that $u$ is also an eigenfuction of $\psi_m$, now with eigenvalue $(1-h'(0))\lambda$.

\subsection{Separable dispersal kernels }
Jentzsch's theorem shows that there is an eigenfunction $u$, provided that the kernel $k$ satisfy some properties.  Finding the eigenfunction is in general a hard problem to solve.  It becomes easier if the kernel $k$ is separable, i.e. there are functions $a_n,b_n$ such that $k(x-y)=\sum_{n=1}^\infty a_n(x)b_n(y)$.  In that case, (\ref{eigen}) becomes
\begin{align*}
\lambda u(x)&=f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right)
\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right)
\\ \Rightarrow \lambda d_k&=f'(0)\sum_{n=1}^\infty A_{nk}d_n  \tag{**} \label{problem}
\end{align*}
where
\begin{equation*}
A_{nk}=\int_{-L/2}^{L/2}a_n(x)b_k(x-c)dx \text{ and } d_k=\int_{-L/2}^{L/2}b_k(x-c)u(x)dx
\end{equation*}

\subsection{Gaussian dispersal kernel }
The Gaussian dispersal kernel is given by
$$k(|x-y|)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.$$
As in \citep{Latore:1998fk}, this separable kernel can be written as
$$k(|x-y|)=\sum_{n=0}^\infty a_n(x)b_n(y)$$
where
$$a_n(x)=b_n(x)=\frac{1}{\sqrt{2n!\sqrt{D\pi}}}e^{-x^2/4D}\left(\frac{x}{\sqrt{2D}}\right)^n.$$

As a first approximation to $k$ we ignore all but the $0^{th}$ terms for $a_n$ and $b_n$ so that Equation \ref{problem} becomes
\begin{align*}
\lambda d_0(c)&=f'(0)A_{00}(c)d_0(c)
\\ \Rightarrow \lambda&=R_0(1-h)A_{00}(c)
\\\text{ where } A_{00}(c)&=2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]
\end{align*}
where $\text{erf}$ is the error function.  The critical rate of environmental shift $c^*$ and the critical harvesting rate $h^*$ are those values of $c$ and $h$, respectively, that make $\lambda=1$.

\subsection{Sinusoidal dispersal kernel }
A sinusoidal dispersal kernel is given by 
$$k(x-y)=\left\{\begin{array}{ccccc}
\frac{w}{2}\cos(w(x-y)) & , & |x-y|\leq\frac{\pi}{2w}
\\ 0 & , & |x-y|>\frac{\pi}{2w}
\end{array}\right.
$$
where $L$ is the length of the patch and we assume $\frac{\pi}{2w}>L,c<\frac{\pi}{2w}-L$.

In this case, $k(x-y)=\frac{w}{2}\cos(wx)\cos(w(y-c))+\frac{w}{2}\sin(wx)\sin(w(y-c))$ so that $A_{ij}$ and $d_i$ can be found for $i,j=1,2$ and (\ref{problem}) reduces to 
$$\lambda^2-\left(\frac{R_0(1-h)wL}{2}\cos(wc)\right)\lambda+\frac{R_0^2(1-h)^2}{16}\left(w^2L^2-\sin^2(wL)\right)=0.$$

If we solve for $\lambda$,we find
\begin{equation} \lambda=R_0(1-h)\left[\frac{wL\cos(wc)}{4}+\frac{1}{4}\sqrt{\sin^2(wL)-w^2L^2\sin^2(wc)}\right]. \label{cosine} \end{equation}

Zhou et al. \citep{ZhouKot2011} solve for the critical speed, $c^*$, at the population will be driven extinct:
$$c^*=c^*(R_0)=\frac{1}{w}\cos^{-1}\left[\frac{16+R_0^2(1-h)^2(w^2L^2-\sin^2(wL))}{8R_0(1-h)wL}\right].$$
Similarly, we can solve for the critical harvesting rate, $h^*$, at which the population will be driven extinct:
$$
h^*=1-\frac{1}{R_0}\cdot\frac{4wL}{w^2L^2-\sin^2(wL)}\left[\cos(wc)-\sqrt{\cos^2(wc)-1+\frac{\sin^2(wL)}{w^2L^2}}\right] 
$$

\subsection{Approximate Critical Harvesting Proportions}
~\\We will use the following Taylor series to make approximations of the critical harvesting proportions under the two dispersal kernels:
\begin{align*}
\cos(x)&=1-\frac{x^2}{2}
\\ \cos^2(x)&=1-x^2
\\ \sin^2(x)&=x^2-\frac{x^4}{3}
\\ erf(x)&=\frac{2}{\sqrt{\pi}}(x-\frac{x^3}{3})
\\ \exp(x)&=1+x+\frac{x^2}{2}
\end{align*}
For the sinusoidal kernel we found 
\begin{equation}
h^*=1-\frac{1}{R_0}\cdot\frac{4wL}{w^2L^2-\sin^2(wL)}\left[\cos(wc)-\sqrt{\cos^2(wc)-1+\frac{\sin^2(wL)}{w^2L^2}}\right] 
\end{equation} 
Using the Taylor series and the fact that $w=\frac{\sqrt{\frac{\pi^2}{4}-2}}{\sigma}$ where $\sigma^2$ is the variance of the sinusoidal kernel,
\begin{align*}
h^*&\sim 1-\frac{1}{R_0}\cdot\frac{12wL}{w^4L^4}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
%\\&=1-\frac{1}{R_0}\cdot\frac{12}{w^3L^3}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
%\\&=1-\frac{1}{R_0}\cdot\frac{12\sigma^3}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[1-\frac{(\frac{\pi^2}{4}-2)c^2}{2\sigma^2}-\sqrt{1-\frac{(\frac{\pi^2}{4}-2)c^2}{\sigma^2}-\frac{(\frac{\pi^2}{4}-2)L^2}{3\sigma^2}}\right]
%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[2\sigma^2-(\frac{\pi^2}{4}-2)c^2-\frac{2}{\sqrt{3}}\sigma\sqrt{3\sigma^2-(\frac{\pi^2}{4}-2)3c^2-(\frac{\pi^2}{4}-2)L^2}\right]
%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^34(\frac{\pi^2}{4}-2)^{3/2}}\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
%\\&=1-\frac{12}{R_0L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
\\&=1-\frac{1}{R_0}\cdot\frac{4\sqrt{3}}{L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sqrt{3}\sigma^2-(\pi^2-8)\sqrt{3}c^2-4\sigma\sqrt{12\sigma^2-(\pi^2-8)(3c^2+L^2)}\right]
\end{align*}
For the Gaussian kernel we found 
\begin{equation}
h^*=1-\frac{2\sqrt{2}\exp\left(\frac{c^{2}}{8D}\right)}{R_0\left[erf\left(\frac{L-c}{2\sqrt{2D}}\right)-erf\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]}
\end{equation} 
Using the Taylor series and the fact that $D=\frac{\sigma^2}{2}$ where $\sigma^2$ is the variance of the exponential kernel,
\begin{align*}
h^*&\sim 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\sqrt{\pi}\left[\frac{L-c}{2\sqrt{2D}}-\frac{(L-c)^3}{3(2\sqrt{2D})^3}-\frac{-L-c}{2\sqrt{2D}}+\frac{(-L-c)^3}{3(2\sqrt{2D})^3)}\right]}
%\\&= 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\left[\frac{2L}{2\sqrt{2D}}+\frac{-L^3+3L^2c-3Lc^2+c^3-L^3-3L^2c-3Lc^2-c^3}{3(2\sqrt{2D})^3}\right]}
%\\&= 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\left[\frac{L}{\sqrt{2D}}-2\frac{L^3+3Lc^2}{3(2\sqrt{2D})^3}\right]}
%\\&= 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0L\left[\frac{48D-2(L^2+3c^2)}{3(2\sqrt{2D})^3}\right]}
%\\&= 1-\frac{3*32D\sqrt{D\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0L\left(48D-2(L^2+3c^2)\right)}
%\\&= 1-\frac{3*32D\sqrt{D\pi}(\frac{128D^2+16c^2D+c^4}{128D^2})}{R_0L\left(48D-2(L^2+3c^2)\right)}
%\\&= 1-\frac{3\sqrt{\pi}(128D^2+16c^2D+c^4)}{4R_0L\sqrt{D}\left(48D-2(L^2+3c^2)\right)}
%\\&= 1-\frac{3\sqrt{2\pi}(32\sigma^4+8c^2\sigma^2+c^4)}{4R_0L\sigma\left(24\sigma^2-2(L^2+3c^2)\right)}
\\&= 1-\frac{1}{R_0}\cdot\frac{3\sqrt{2\pi}}{8L}\frac{(32\sigma^4+8c^2\sigma^2+c^4)}{\sigma\left(12\sigma^2-(L^2+3c^2)\right)}
\end{align*}
In the case of both kernels, the critical harvesting proportion can be approximated by a function that looks like 
\begin{equation}
h^*\sim1- \frac{1}{R_0}\cdot C(L)f(\sigma^2,c^2,L^2+3c^2)
\end{equation}
where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.

\end{document}

### 2013-12-19 Simluations/Aspatial_fast/Basic_Simulation.R

@@ -1,0 +1,113 @@
+rm(list=ls())
+library(plyr)
+library(zoo)
+
+# load parameters, functions
+setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
+source("Parameters.R")
+source("Functions.R")
+
+# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
+mpa.yes = null.yes
+mpa.no = null.no
+
+# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
+
+move_window = 100
+
+# initializing the population with no pressure (no harvesting, no climate)
+init<-array(0,c(w,maxt)) 
+init[which(patch==0.55),1]=50
+MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
+for(t in 2:maxt){
+	output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+	init[,t]= output[[1]]
+	MPA.current = output[[3]]
+	}
+
+# standard for equilibrium is the difference in the final step of the initialization 
+init.diff <- diff(colSums(init))[149]
+
+
+# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
+# using while loops. Only saving two time steps 
+
+
+for(q in 1:length(speeds)){
+  
+  for(j in 1:length(harvests)){
+    init.h<-array(0,c(w,1))
+    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+    init.h[,1]=next.pop[[1]]
+    MPA.current = next.pop[[3]]
+    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
+    roll.mean = 1	# set to start
+    
+    T = 1
+    
+    while(h.diff > init.diff | roll.mean > init.diff){
+      T = T + 1
[…]

### 2013-12-19 Simluations/Aspatial_fast/Data/plots/generate_plots.R

@@ -1,0 +1,157 @@
+library(ggplot2)
+library(ggthemes)
+###############     
+## Load data ##
+###############
+
+setwd('~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast/Data')
+
+noThresh <- read.csv('nothresh/noMPAnotThresh_add_2013-12-09.csv')
+consMPA <- read.csv('nothresh/consMPAnotThresh_add_2013-12-09.csv')
+fishMPA <- read.csv('nothresh/fishMPAnotThresh_add_2013-12-09.csv')
+thresh_noMPA <- read.csv('thresh/noMPA_Thresh_add_2013-12-10.csv')
+
+#########################################################
+## Fig 1. Equil.pop as a function of harvest and speed ##
+#########################################################
+
+ggplot(noThresh, aes(x = speed, y = Equil.pop, group = harvest)) + geom_line(aes(color = harvest)) + geom_point(aes(color = harvest)) + theme_few() + scale_colour_gradient(high = "red", low = "green") + ylab("Equilibrium biomass") + xlab("Speed") 
+
+ggsave(file="plots/x_Speed.pdf")
+
+
+ggplot(noThresh, aes(x = harvest, y = Equil.pop, group = speed)) + geom_line(aes(color = speed)) + geom_point(aes(color = speed)) + theme_few() + scale_colour_gradient(high = "red", low = "green") +xlab("Percentage harvested") + ylab("Equilibrium biomass")
+
+ggsave(file="plots/x_harvest.pdf")
+
+############################################
+## Figure 3: Synergy in no thresh, no mpa ##
+# S = E_(h+c) - (E_h - E_c)               ##
+############################################
+
+h_only <- data.frame(E_h = noThresh$Equil.pop[which(noThresh$speed==0)], harvest =  noThresh$harvest[which(noThresh$speed==0)])
+c_only <- data.frame(E_c = noThresh$Equil.pop[which(noThresh$harvest==0)], speed = noThresh$speed[which(noThresh$harvest==0)])
+
+# pristine (no disturbance)
+pristine <- noThresh$Equil.pop[which(noThresh$speed==0 & noThresh$harvest==0)]
+
+# setting up data to loop through
+modnothresh <- noThresh
+modnothresh <- modnothresh[-which(modnothresh$speed==0 | modnothresh$harvest ==0),]
+
[…]

### 2013-12-19 Simluations/Aspatial_fast/Functions.R

@@ -1,0 +1,86 @@
+# m function doesn't quite work yet. right now there's no way to disperse in front of the patch and have the patch move on top of you, so harder to get population persistence. also why you get lower harvests, can only harvest in the patch. 
+
+# maybe solution is to make dispersal kernal as big as the patch plus the biggest possible displacement (displacement from the fastest speed)
+
+# thought I figured it out with doubling the space ahead of the patch for it to shift into so I can keep track of fish dispersed ahead of patch. But can't maintain a population with no harvesting and no speed. So something is wrong.. 
+
+# i might have messed up the dispersal kernel.. should see what it looked like on other versions of the code. 
+
+# solution, the world units were an index, and instead need to be at the same scale as the step size.. So each the patch needs to be equal to 1 on the world units, and 101 steps need to make up the world.. still don't completely understand, but can now get a population to grow. 
+
+# figured it out by summing rows across dispersal matrix. Found that the dispersal step was knocking down the population hugely each time step such that it could never grow. 
+
+#Laplace dispersal kernel
+k<-function(x,y,b) return(1/2*b*exp(-b*abs(x-y)))    
+
+# dispersal matrix
+d<-array(,c(w,w)) #dispersal probabilities to point i from every point j
+for(i in 1:w) d[i,]=k(world[i],world,b)
+
+#Beverton-Holt recruitment
+f<-function(n,R0,K) return(R0*n/(1+(R0-1)/K*n))
+
+#Standard error removing NAs
+stderr <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
+
+m <- function(n, s, Fthresh = NA, Fharv = NA, mpa.yes = NA, mpa.no = NA, MPA.current=NA){
+	
+	# calculate how far the patch will move through the population (if speed !=0)
+	displaced = ifelse(s>0,s/step_size,1)
+	
+	# assign population that will still be inside the patch to moved patch
+	next_n = n[displaced:length(n)]
+	
+	# fill in newly existing patch with 0s
+	next_n = c(next_n,rep(0,length.out=(displaced-1)))
+	
+	# Move MPAS
[…]

### 2013-12-19 Simluations/Aspatial_fast/MPAcons_SIM.R

@@ -1,0 +1,113 @@
+rm(list=ls())
+library(plyr)
+library(zoo)
+
+# load parameters, functions
+setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
+source("Parameters.R")
+source("Functions.R")
+
+# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
+mpa.yes = cons.yes
+mpa.no = cons.no
+
+# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
+
+move_window = 100
+
+# initializing the population with no pressure (no harvesting, no climate)
+init<-array(0,c(w,maxt)) 
+init[which(patch==0.55),1]=50
+MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
+for(t in 2:maxt){
+	output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+	init[,t]= output[[1]]
+	MPA.current = output[[3]]
+	}
+
+# standard for equilibrium is the difference in the final step of the initialization 
+init.diff <- diff(colSums(init))[149]
+
+
+# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
+# using while loops. Only saving two time steps 
+
+
+for(q in 1:length(speeds)){
+  
+  for(j in 1:length(harvests)){
+    init.h<-array(0,c(w,1))
+    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+    init.h[,1]=next.pop[[1]]
+    MPA.current = next.pop[[3]]
+    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
+    roll.mean = 1	# set to start
+    
+    T = 1
+    
+    while(h.diff > init.diff | roll.mean > init.diff){
+      T = T + 1
[…]

### 2013-12-19 Simluations/Aspatial_fast/MPAfish_SIM.R

@@ -1,0 +1,114 @@
+rm(list=ls())
+library(plyr)
+library(zoo)
+
+# load parameters, functions
+setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
+source("Parameters.R")
+source("Functions.R")
+
+# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
+mpa.yes = fish.yes
+mpa.no = fish.no
+
+# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
+
+move_window = 100
+
+# initializing the population with no pressure (no harvesting, no climate)
+init<-array(0,c(w,maxt)) 
+init[which(patch==0.55),1]=50
+MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
+for(t in 2:maxt){
+	output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+	init[,t]= output[[1]]
+	MPA.current = output[[3]]
+	}
+
+# standard for equilibrium is the difference in the final step of the initialization 
+init.diff <- diff(colSums(init))[149]
+
+
+# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
+# using while loops. Only saving two time steps 
+
+
+for(q in 1:length(speeds)){
+  
+  for(j in 1:length(harvests)){
+    init.h<-array(0,c(w,1))
+    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+    init.h[,1]=next.pop[[1]]
+    MPA.current = next.pop[[3]]
+    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
+    roll.mean = 1	# set to start
+    
+    T = 1
+    
+    while(h.diff > init.diff | roll.mean > init.diff){
+      T = T + 1
[…]

### 2013-12-19 Simluations/Aspatial_fast/Parameters.R

@@ -1,0 +1,38 @@
+######################################
+## Parameters & Building Structures ##
+######################################
+
+step_size=0.01 #distance between points in space
+b=.5 #parameter for Laplace dispersal kernel
+R0=5 #growth parameter for recruitment
+K=100 #carrying capacity parameter for juvenile density dependence
+maxt = 150 # number of generations to run for initializing of harvest
+# speeds = seq(0,1,.1)
+speeds = seq(0,0.5,0.02)
+harvests = seq(0,0.15,0.02)
+
+# harvests = seq(0,1,.1)
+#thresholds = seq(0,1,.1)
+f_ind = 1 #per capita reproductive rate
+
+patch = seq(0,1,by=step_size)
+world = seq(-.51,1.5, by = step_size)
+w = length(world)
+
+# haven't dealt with MPAs yet !
+
+cons.yes = rep(1,2*b/step_size)
+cons.no = rep(0,4*b/step_size)
+fish.yes = rep(1,floor((1/3*b)/step_size))  # had to round because not complete step size. Rounded down. 
+fish.no = rep(0,floor((2/3*b)/step_size))
+
+null.yes = rep(0,length(world))
+null.no = rep(0, length(world))
+
+# build dataframes
+
+summaries <- data.frame(Equil.pop = rep(NA,length=length(speeds)*length(harvests)), Equil.pop.sd = rep(NA, length=length(speeds)*length(harvests)), Equil.pop.se = rep(NA, length=length(speeds)*length(harvests)), Equil.harvest = rep(NA,length=length(speeds)*length(harvests)),Equil.harvest.sd = rep(NA,length=length(speeds)*length(harvests)), Equil.harvest.se = rep(NA,length=length(speeds)*length(harvests)), speed=rep(NA,length=length(speeds)*length(harvests)), harvest = rep(NA,length=length(speeds)*length(harvests)), thresh=rep(NA,length=length(speeds)*length(harvests)), steps.harv = rep(NA,length=length(speeds)*length(harvests)),steps.move = rep(NA,length=length(speeds)*length(harvests)))
+
+
+# index for row number
+rownumber <- matrix(seq(1:(length(harvests)*length(speeds))),ncol=length(speeds))

### 2013-12-19 Simluations/Aspatial_fast/Run_sims.R

@@ -1,0 +1,11 @@
+rm(list=setdiff(ls(), "notify"))
+setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
+
+source('noMPA_SIM.R',verbose=TRUE)
+notify('noMPA_SIM finished!')
+
+source('MPAfish_SIM.R', verbose=TRUE)
+notify('MPAfish_SIM finished!')
+
+source('MPAcons_SIM.R', verbose=TRUE)
+notify('MPAcons_SIM finished!')

### 2013-12-19 Simluations/Aspatial_fast/noMPA_SIM.R

@@ -1,0 +1,114 @@
+#rm(list=ls())
+library(plyr)
+library(zoo)
+
+# load parameters, functions
+setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
+source("Parameters.R")
+source("Functions.R")
+
+# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
+mpa.yes = null.yes
+mpa.no = null.no
+
+# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
+
+move_window = 100
+
+# initializing the population with no pressure (no harvesting, no climate)
+init<-array(0,c(w,maxt)) 
+init[which(patch==0.55),1]=50
+MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
+for(t in 2:maxt){
+	output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+	init[,t]= output[[1]]
+	MPA.current = output[[3]]
+	}
+
+# standard for equilibrium is the difference in the final step of the initialization 
+init.diff <- diff(colSums(init))[149]
+
+
+# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
+# using while loops. Only saving two time steps 
+
+
+for(q in 1:length(speeds)){
+  
+  for(j in 1:length(harvests)){
+    init.h<-array(0,c(w,1))
+    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+    init.h[,1]=next.pop[[1]]
+    MPA.current = next.pop[[3]]
+    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
+    roll.mean = 1	# set to start
+    
+    T = 1
+    
+    while(h.diff > init.diff | roll.mean > init.diff){
+      T = T + 1
[…]

### 2013-12-19 Simluations/Aspatial_fast/thresh_Parameters.R

@@ -1,0 +1,37 @@
+######################################
+## Parameters & Building Structures ##
+######################################
+
+step_size=0.01 #distance between points in space
+b=.5 #parameter for Laplace dispersal kernel
+R0=5 #growth parameter for recruitment
+K=100 #carrying capacity parameter for juvenile density dependence
+maxt = 150 # number of generations to run for initializing of harvest
+# speeds = seq(0,1,.1)
+speeds = seq(0,0.5,0.02)
+harvests = 1
+thresholds = seq(0,1,by=0.1)
+
+f_ind = 1 #per capita reproductive rate
+
+patch = seq(0,1,by=step_size)
+world = seq(-.51,1.5, by = step_size)
+w = length(world)
+
+# haven't dealt with MPAs yet !
+
+cons.yes = rep(1,2*b/step_size)
+cons.no = rep(0,4*b/step_size)
+fish.yes = rep(1,floor((1/3*b)/step_size))  # had to round because not complete step size. Rounded down. 
+fish.no = rep(0,floor((2/3*b)/step_size))
+
+null.yes = rep(0,length(world))
+null.no = rep(0, length(world))
+
+# build dataframes
+
+summaries <- data.frame(Equil.pop = rep(NA,length=length(speeds)*length(thresholds)), Equil.pop.sd = rep(NA, length=length(speeds)*length(thresholds)), Equil.pop.se = rep(NA, length=length(speeds)*length(thresholds)), Equil.harvest = rep(NA,length=length(speeds)*length(thresholds)),Equil.harvest.sd = rep(NA,length=length(speeds)*length(thresholds)), Equil.harvest.se = rep(NA,length=length(speeds)*length(thresholds)), speed=rep(NA,length=length(speeds)*length(thresholds)), harvest = rep(NA,length=length(speeds)*length(thresholds)), thresh=rep(NA,length=length(speeds)*length(thresholds)), steps.harv = rep(NA,length=length(speeds)*length(thresholds)),steps.move = rep(NA,length=length(speeds)*length(thresholds)))
+
+
+# index for row number
+rownumber <- matrix(seq(1:(length(thresholds)*length(speeds))),ncol=length(speeds))

### 2013-12-19 Simluations/Aspatial_fast/thresh_noMPA_SIM.R

@@ -1,0 +1,114 @@
+#rm(list=ls())
+library(plyr)
+library(zoo)
+
+# load parameters, functions
+setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
+source("thresh_Parameters.R")
+source("Functions.R")
+
+# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
+mpa.yes = null.yes
+mpa.no = null.no
+
+# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
+
+move_window = 100
+
+# initializing the population with no pressure (no harvesting, no climate)
+init<-array(0,c(w,maxt)) 
+init[which(patch==0.55),1]=50
+MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
+for(t in 2:maxt){
+  output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+  init[,t]= output[[1]]
+  MPA.current = output[[3]]
+}
+
+# standard for equilibrium is the difference in the final step of the initialization 
+init.diff <- diff(colSums(init))[149]
+
+
+# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
+# using while loops. Only saving two time steps 
+
+
+for(q in 1:length(speeds)){
+  
+  for(j in 1:length(thresholds)){
+    init.h<-array(0,c(w,1))
+    next.pop <- m(n=init[,maxt], s = 0, Fthresh = thresholds[j], Fharv = 1, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+    init.h[,1]=next.pop[[1]]
+    MPA.current = next.pop[[3]]
+    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
+    roll.mean = 1	# set to start
+    
+    T = 1
+    
+    while(h.diff > init.diff | roll.mean > init.diff){
+      T = T + 1
[…]

### 2014-01-10 fish_figures.R
code for making figures

@@ -1,0 +1,446 @@
+library(fields)
+library(lattice)
+library(colorspace)
+setwd("/Users/eleanorbrush/Desktop")
+source("/Users/eleanorbrush/Desktop/biomass_synergy.R")
+source("/Users/eleanorbrush/Desktop/critical_parameters.R")
+
+
+### parameters from Zhou
+r=3
+w=pi/4
+L=1
+
+####
+#the world
+step_size=0.01
+L=1
+world=matrix(seq(-L/2,L/2,by=step_size),nrow=1)
+emmaR0=5
+emmaK=100
+
+#####
+#### plotting parameters
+cex.lab=3
+cex.axis=2
+line=5
+lwd=5
+yaxs='i'
+xaxs='i'
+cex.lab=1.5
+cex.axis=1
+line=1
+oma=c(0,0,0,2)
+
+##########
+##########
+##### wireframe of the equilibrium biomass as a function of c and h
+r=emmaR0
+K=emmaK
+sig2=1
+tol=.00001
+diff=.01
+cvals=seq(0,1,by=diff)
+hvals=seq(0,1,by=diff)
+
+
+# ebm=eqbiomass(cvals,hvals,r,K,sig2,tol=tol)
+load("/Users/eleanorbrush/Desktop/ebm.rdata")
+
+#which params to work with
+H=sum(apply(ebm,2,max)>tol)+2
+hvals=hvals[1:H]
+C=length(cvals)
+toplot_ebm=ebm[1:C,1:H]
+
+#breaks and colors
+lower=.001
+cuts=c(0,lower,seq(.5,max(ebm)+.5,by=.5))
+# s=5
+# maxgray=length(cuts)-2+s-1
+# vecgray=(s:maxgray)/maxgray
+# mycols=c('black',gray(vecgray))
+# colorbar.labels=as.character(round(cuts[-1],digits=1))
+# mycolorkey=list(col=mycols,at=cuts,height=1,lty=1)
+# myalpha=.7
+mycols=(sequential_hcl(length(cuts)-2,c.=c(0,0),l=c(30,100)))
+mycols=c('black',mycols)
+
+#axis labels
+numxlabs=11
+numylabs=3
+xdiff=.1
+ydiff=.2
[…]

### 2014-01-10 Writing/MS.tex
small changes to methods

@@
 \begin{equation}
-m_i=\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f\left((1-h)\sum^\infty_{i=1}m_ia_i(x)\right)
+m_i=\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f\left((1-h)\sum^\infty_{j=1}m_ja_j(x)\right)
 dy. \label{recursive_m}
@@
 the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.
+For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
+$$R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.$$
+A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
+\begin{equation}
+h^*\sim1- \frac{1}{R_0}\cdot C(L)f(\sigma^2,c^2,L^2+3c^2)
+\end{equation}
+where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
 

### 2014-01-12 fish_figures.R
figures code

@@ -114,7 +114,7 @@ for(i in 1:L){
 toplot_ebm=ebm
 
 #breaks and colors
-lower=min(ebm)
+lower=min(ebm)+1
 cuts=c(0,seq(lower,max(ebm)+.5,length.out=100))
 # s=5
 # maxgray=length(cuts)-2+s-1
@@ -174,7 +174,7 @@ for(i in 1:L){
 
 toplot_ebm=ebm
 #breaks and colors
-lower=min(ebm)
+lower=min(ebm)+1
 cuts=c(0,seq(lower,max(ebm)+.5,length.out=100))
 # s=5
 # maxgray=length(cuts)-2+s-1
@@ -297,7 +297,7 @@ for(i in 1:L){
 toplot_ebm=ebm
 
 #breaks and colors
-lower=min(ebm)
+lower=min(ebm)+1
 cuts=c(0,seq(lower,max(ebm)+.5,length.out=100))
 # s=5
 # maxgray=length(cuts)-2+s-1
@@ -346,7 +346,7 @@ cvals=matrix(seq(0,1,by=0.01),nrow=1)
 
 nr=3
 ns=2
-rvals=c(rep(3,ns),rep(7,ns),rep(10,ns))
+rvals=c(rep(3,ns),rep(5,ns),rep(10,ns))
 rcols=c(rep('black',ns),rep('red',ns),rep('blue',ns))
 xvals=c(.1,.25)
 sigvals=rep((xvals*sqrt(2*pi))^2,nr)

### 2014-01-12 Writing/MS.tex
figures captions and arrangements; some of malins comments about abstract and model description

@@
 
-When the climate changes, the area in which organisms can survive and reproduce will move through space. 
-This change does not occur in isolation but rather appears on a background of other disturbances. We use an 
+When the climate changes, the area in which organisms can survive and reproduce moves through space. 
+This change does not occur in isolation but rather appears on a background of other disturbances. We used an 
 integrodifference model explicitly accounting for dispersal, reproduction, a shifting environment, and 
 harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
-persistence. We found the harvesting rate and the rate of environmental shift that are needed to allow the 
+persistence. We found threshold rates of harvesting and of environmental shift that are needed to allow the 
 population to persist, studied how these critical parameters depend on the growth rate and dispersal behavior 
-of the population, and measured the interactions between the stressors. In particular, we found positive 
+of the population, and measured the interactions between the stressors. In particular, we found low but positive 
 synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. 
 Finally, we introduced two conservation techniques into simulations of the population model -- threshold 
-population levels below which harvesting is not allowed and marine protected areas (MPAs) -- and found that 
-the robustness of the population to the two stressors improved.
+harvest rules and marine protected areas (MPAs) -- and found that 
+these approaches could mitigate the negative interaction of the two stressors.
 
@@
 
-We studied the dynamics of a fish population constrained to a single one-dimensional habitat patch by their inability to reproduce outside of the patch.
-This viable habitat patch (here after `patch') is shifting at a fixed velocity and fish at each point in the patch can be harvested.
-Given the reproductive rate and average dispersal distance, we determined the climate velocity and harvesting rate that would drive the population extinct.
-We implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to
+We studied the dynamics of a fish population constrained to a single, one-dimensional habitat patch by their inability to reproduce outside of the patch.
+This viable habitat patch (here after `patch') is shifting at a fixed velocity and fish at each point in space can be harvested.
+Given the reproductive rate and average dispersal distance, first we determined the climate velocity and harvesting rate that would drive the population extinct.
+We then implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to
 determine how these management strategies affect population persistence.
@@
 
-In our model, the adults from the current year produce offspring according to a recruitment function and these 
+In our model, the adults from the current year produce offspring according to a recruitment function, and these 
 offspring disperse across the one-dimensional world according to a dispersal kernel to become the next 
@@
 \noindent where $h$ is the proportion of adults harvested, $f(n)$ is the recruitment function giving the number of 
-offspring produced by a population of size $n$, $k(x-y)$ is the dispersal kernel giving the probability of the 
-larvae traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it 
-shifts across space. We provide a list of variables and functions in Table 1. The functional form of the 
+offspring produced by a population of size $n$ (accounting for density dependence), $k(x-y)$ is the dispersal kernel giving the probability of a 
+larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it 
+shifts across space. We provide a list of variables and functions in Table 1.  As we explain below, the functional form of the 
 recruitment function only affects persistence through how quickly recruitment increases when the population 
-increases from $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$. We therefore do not specify a 
+size is near (but above) $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$. We therefore do not specify a 
 recruitment function here. Similarly, the population's ability to persist only depends on how quickly harvesting 
 increases when a small number of fish are introduced where they were previously absent, $h'(0)$. We 
-therefore only considered a proportional harvesting function.
+therefore only considered a proportional harvesting function initially.
 
 If a population of fish is going to persist in the face of a shifting environment, it must move along with the patch 
-of environment in which it can reproduce. At equilibrium, the population will be described by a traveling wave, 
+ in which it can reproduce. At equilibrium, the population will be described by a traveling wave, 
 where the density of fish at a given point in space will change but the density of fish at a location relative to the 
@@
 
-For the separable Gaussian kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
+We would like to know the rate of environmental shift and the harvesting rate such that as long as the 
+environment moves more slowly or we harvest less severely than those parameters, then the population will be able to persist. 
+We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$. 
+We can find these rates by finding the parameters that make the trivial pulse unstable. Evaluating stability in this kind of model is in 
+general difficult to do analytically. It becomes easier if the dispersal kernel is separable into its dependence on the 
+source of larvae and its dependence on the destination of the larvae,
+i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.  In our analyses, we used one such separable kernel, the Gaussian kernel given by
+
+\[k(x-y)=\frac{1}{\sqrt{2\sigma^2\pi}}e^{\frac{-(x-y)^2}{2\sigma^2}}\]
+
+\noindent To make stability easier to analyze, we approximated this kernel, as described in the Appendix.
+Analytical results for a separable sinusoidal kernel are also described in the Appendix.
+We used
+simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
+
+For each kernel, the population's ability to persist depends on properties of the population itself: the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment: the
+length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
+$). If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than 
+the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.
+For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
+$$R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.$$
+A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
+\begin{equation}
+h^*\sim1- \frac{1}{R_0}\cdot C(L)f(\sigma^2,c^2,L^2+3c^2)
+\end{equation}
+where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
+
+\citet{ZhouKot2011} only consider whether a shifting environment will drive a population extinct or not.
+To quantify the effects of both a shifting environment and harvesting pressure, we find the total biomass in the equilibrium traveling wave.
+As in \citet{Latore:1998fk}, for a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
 
@@
 equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.
+Whereas population persistence does not depend on the functional form of recruitment, $f$, equilibrium biomass does depend on what recruitment function we use.
+We chose to use a Beverton-Holt recruitment function,
 
-In general, we would like to know the rate of environmental shift and the harvesting rate such that if the 
-environment moves more quickly or we harvest more severely then the population will not be able to persist. 
-We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$. 
-We can find these rates by finding the parameters that make the trivial pulse unstable. Evaluating stability is in 
-general difficult to do analytically. It becomes easier if the dispersal kernel is separable into its dependence on the 
-source of larvae and its dependence on the destination of the larvae,
-i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$. We used 
-simulations to analyze dispersal kernels that are not amenable to this method.
-
-In our analyses, we used a Gaussian kernel,
-
-\[k(x-y)=\frac{1}{\sqrt{2\sigma^2\pi}}e^{\frac{-(x-y)^2}{2\sigma^2}}\]
-
-\noindent To make stability easier to analyze, we approximated this kernel, as described in the Appendix. In our 
-simulations, we used an analytically intractable Laplace kernel,
-
-\[k(x-y)=\frac{1}{2}be^{-b\vert x-y\vert}\]
-
-\noindent For each kernel, we can find the expected distance an offspring will travel, $\langle d \rangle$, which we will use to compare the two kernels.
-For each kernel, the population's ability to persist depends on properties of the population itself: how far larvae
-tend to disperse ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment: the 
-length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
-$). If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than 
-the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.
-For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
-$$R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.$$
-A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
-\begin{equation}
-h^*\sim1- \frac{1}{R_0}\cdot C(L)f(\sigma^2,c^2,L^2+3c^2)
-\end{equation}
-where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
+\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
 
@@
 We used simulations to extend the basic integrodifference model in two ways that make it analytically 
-intractable. First, we examined the sensitivity of the model to choice of dispersal kernel by using a Laplace 
-dispersal kernel. Second, we examined harvesting rules more complex than harvesting a constant proportion of the population. Whereas the 
-analytical results do not depend on the functional form of recruitment $f$, to perform simulations we must 
-specify a recruitment function. We chose to use a Beverton-Holt model,
+intractable. First, we examined the sensitivity of the model to choice of dispersal kernel by using the Laplace 
+dispersal kernel,
+
+\[ k(x-y)=\frac{1}{2}be^{-b|x-y|}\]
+
+\noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
+Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
+specify a recruitment function.  We chose to use a Beverton-Holt function,
 
@@
 The equilibrium biomass of the population decreases as either the harvesting pressure 
-increases or the environmental shifts more quickly (Figure \ref{eqbiomass}). If the equilibrium biomass is $0$, this 
+increases or the environmental shifts more quickly (Figure \ref{baseline}). If the equilibrium biomass is $0$, this 
 indicates that the harvesting pressure has exceeded the critical harvesting rate $h^*$ and the environmental 
@@
 the critical rate of environmental shift $c^*$ decreases: the environment must move more slowly to 
-accommodate the population growing more slowly (Figure \ref{critical}). Conversely, as the rate of 
-environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{critical}). This 
+accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
+environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}). This 
 means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
@@
 \begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.5\textwidth]{plots/eqbiomass.pdf}
-\caption{The equilibrium biomass of the population decreases as either the harvesting rate or rate of environmental shift increases.
-If either harvesting rate or rate of environmental shift are too high the population is driven to extinction.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the equilibrium biomass.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
-\label{eqbiomass}
-\end{center}
+\begin{subfigure}{3in}
+\subcaption{\label{biomass}}
+\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
+\end{subfigure}
+\begin{subfigure}{3in}
+\subcaption{\label{rates}}
+\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
+\end{subfigure}
+\caption{(a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+(b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
+Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
+These results are from an approximated Gaussian dispersal kernel with $L=1$.}
+\label{baseline}
 \end{figure}
@@
 more quickly from losses caused by these disturbances. However, whether or not dispersing farther is better 
-depends on how quickly the environment is shifting (Figure \ref{critical}). When the environment is shifting 
+depends on how quickly the environment is shifting (Figure \ref{baseline}). When the environment is shifting 
 slowly, dispersing farther is detrimental since many larvae will disperse too far away from the viable patch. 
@@
 because some larvae will disperse into the space that will become viable shortly in the future.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.5\textwidth]{plots/critical_rates.pdf}
-\caption{As the rate of environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases.
-In other words, less harvesting pressure is required to drive the population extinct.
-Increasing the growth rate always increases the critical harvesting rate.
-On the other hand, increasing average dispersal distance increases the critical harvesting rate depends on how quickly the environment is shifting: it is more advantageous to disperse farther when the environment is shifting more quickly.
-The rate of environmental shift is plotted on the x-axis and the critical harvesting rate is plotted on the y-axis.
-Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
-Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-These results are from a Gaussian dispersal kernel with $L=1$.}
-\label{critical}
-\end{center}
-\end{figure}
+This affects the critical harvesting rate: at a low rate of environmental shift, populations that disperse less can be harvested more severely than those that disperse further, whereas at a high rate of environmental shift, populations that disperse further can be harvested more severely.
 
@@
 \begin{center}
-\includegraphics[width=.75\textwidth]{plots/synergy.pdf}
-\caption{There is positive synergy between the two stressors: a doubly stressed population loses more biomass than would be predicted from either stressor individually.
+\includegraphics[width=3in]{plots/synergy.pdf}
+\caption{Positive synergy between the two stressors.
 The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
+These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
 \label{Synergy}
@@
 \begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.45\textwidth]{plots/eqbiomass_sim.pdf}
-\includegraphics[width=.45\textwidth]{plots/eqbiomass_thresh.pdf}
-\caption{With threshold harvesting rules, such that all biomass above a threshold level was harvested, the population always reached an equilibrium biomass greater than $0$ so that threshold harvesting rules increased biomass at equilibrium and improved the population's ability to persist.
-Additionally, threshold harvesting rules remove the interaction between harvesting and environmental shift: the critical rate of environmental shift no longer depends on harvesting.
-The figure on the left corresponds to no thresholds and the figure on the right corresponds to threshold harvesting rule.
-The x-axis shows the rate of environmental shift, the y-axis shows either the proportion of the population harvested or the threshold above which harvesting was allowed, and the color indicates the equilibrium biomass.
+
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
+\end{subfigure}
+\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+(a) No management.
+(b) Threshold harvesting levels.
+(c) MPAs.
 These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
@@
 \label{thresholds}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=.45\textwidth]{plots/eqbiomass_sim.pdf}
-\includegraphics[width=.45\textwidth]{plots/eqbiomass_mpa.pdf}
-%\includegraphics[width=.45\textwidth]{plots/eqbiomass_mpa_v2.pdf}
-\caption{MPAs improved the population's ability to persist.
-The figure on the left corresponds to no MPAs and the figure on the right corresponds to MPAs designed to optimize fishing yields.
-MPAs designed to optimize conservation give similar results.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the equilibrium biomass.
-These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.}
-\label{MPAs}
-\end{center}
 \end{figure}

### 2014-01-17 Writing/MS.tex
working on the discussion

@@
 
-In our model, the adults from the current year produce offspring according to a recruitment function, and these 
-offspring disperse across the one-dimensional world according to a dispersal kernel to become the next 
-generation's adults. The adults are subjected to harvesting before they produce offspring so that only a 
-proportion of the fish survive to reproduce. We incorporated these processes-- recruitment, harvesting, and 
-dispersal-- into an integrodifference model to describe how the population changes over time. If $n_t(x)$ 
+In the model of \cite{ZhouKot2011}, the adults from the current year produce offspring according to a recruitment function, and these offspring disperse across the one-dimensional world according to a dispersal kernel to become the next generation's adults.
+We extend this model by additionally subjecting the adults to harvesting before they produce offspring so that only a proportion of the fish survive to reproduce.
+These processes-- recruitment, harvesting, and dispersal-- are incorporated into an integrodifference model to describe how the population changes over time.
+If $n_t(x)$
 is the density of fish at position $x$ at time $t$, then the density of fish at the next generation is given by
@@
 We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$. 
-We can find these rates by finding the parameters that make the trivial pulse unstable. Evaluating stability in this kind of model is in 
+We found these rates by finding the parameters that make the trivial pulse unstable. Evaluating stability in this kind of model is in 
 general difficult to do analytically. It becomes easier if the dispersal kernel is separable into its dependence on the 
@@
 \citet{ZhouKot2011} only consider whether a shifting environment will drive a population extinct or not.
-To quantify the effects of both a shifting environment and harvesting pressure, we find the total biomass in the equilibrium traveling wave.
+To quantify the effects of both a shifting environment and harvesting pressure, we found the total biomass in the equilibrium traveling wave.
 As in \citet{Latore:1998fk}, for a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
@@
 The stressors interact most strongly when they are both high, shortly before they drive the population extinct.
-Our simulations with a Laplace kernel produce similar
-results. In the Appendix, we show analogous results for a sinusoidal dispersal kernel.
+We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results.
 
@@
 We found that thresholds of any amount alleviate the effects of harvesting, and the ability of the population to 
-persist is recovered (Figure \ref{thresholds}). Thus when thresholds are in place, the harvesting rate no longer 
+persist is recovered (Figure \ref{management}). Thus when thresholds are in place, the harvesting rate no longer 
 determines the critical rate of environmental shift $c^*$. We also examined the effect of marine protected 
@@
 slightly higher abundance along the edges of the patch where the population is limited by harvesting, which 
-translated into a slightly increased critical harvest rate (Figure \ref{MPAs}). Additionally MPAs increased 
-overall catch at the highest harvest rates under which the population could survive (Figure \ref{MPAs}).
+translated into a slightly increased critical harvest rate (Figure \ref{management}). Additionally MPAs increased 
+overall catch at the highest harvest rates under which the population could survive (Figure \ref{management}).
 
@@
 }
-\label{thresholds}
+\label{management}
 \end{figure}
@@
 
-We extended the model of Zhou and Kot \citep{ZhouKot2011} to study a fish population subject to both the climate-driven range shifts already included and harvesting.
-Previous work has highlighted the importance of
-reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]
-{Hastingsetal2005} and the integrodifference model explicitly includes these processes. We found that 
-the two disturbances interact synergistically with each other in some cases. However, we also found that 
-placing thresholds on harvesting intensity can alleviate interactions between the two stressors and that spatial 
-management increased the maximum harvesting rate at which the population could survive.
+We extended a spatial to study a fish population subject to both the climate-driven range shifts and harvesting.
+Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.
+In particular, we derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
+For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
+More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
+That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
+The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.
 
-We found that the synergy between the two stressors is positive and that the synergy is greatest in the region 
-of parameter space where the equilibrium biomass is smallest. Additionally, we found that the higher the 
-growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a 
-population can adjust to harvest and climate change. We find similar results from the analytically derived 
-biomass and the simulation derived biomass. This indicates that this result is robust to changes in the 
-dispersal kernel. Similarly, in experiments, \cite{Moraetal2007} found no interaction between harvesting and 
-habitat fragmentation but did find an interaction between harvesting and warming temperature \citep
-{Moraetal2007}. We chose to measure the effect of each stressor by the drop in biomass caused by the 
-stressor, as in \cite{Crainetal2008}. In general, measuring synergy additively is more conservative than 
-measuring synergy multiplicatively: stressors that have positive multiplicative synergy may have zero additive 
-synergy \citep{Crainetal2008, Foltetal1999}. Since we found small levels of positive additive synergy between 
-the two stressors, other measures of synergy might show even higher levels of interaction. This synergy and 
-the fact that it most severely affects those populations whose persistence is most tenuous is worrisome from a 
-conservation perspective. Harvesting levels or rate of environment shift that are sustainable individually 
-together can drive a population to extinction. It is therefore important to take the effects of both stressors into 
-consideration when designing conservation and management strategies.
+%% discussion of critical rates and curved persistence boundary
 
-The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the 
-population's biomass at equilibrium and improved its ability to persist. These management strategies also 
-affect how the two stressors interact with each other.  Protected areas have 
-been advanced as a way to help organisms keep pace with range shifts as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, 
-Hannahetal2007,Botsfordetal2001, Gaylordetal2005, 
-HastingsBotsford2003,Thomasetal2012}.  While MPAs only change harvesting practices 
-and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting 
-and range shifts will help to better implement and place protected areas.
+To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.
+We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.
+We found similar results from the analytically derived biomass and the simulation derived biomass.
+This indicates that this result is robust to changes in the dispersal kernel.
+We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.
+We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
+In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.
+Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
 
-The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.  However, it  ignores many of the complexities present in marine fisheries. We do 
-not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist 
-over time. However, we found that qualitatively similar results about the interaction between climate and 
-harvesting would hold for a model with a recruitment function with Allee effect. We also did not include age 
-structure in our model. The effects of both harvesting and climate change may be different across different age 
-classes; including this level of complexity is left for future work. Additionally, we did not include any 
-mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
+While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
+synergistic effects were quite small.
+However, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
+Additionally, the fact that we found additive synergy means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
+Synergy between harvesting and climate changes, e.g.
+warming temperature or a more variable climate, has been identified in experiments \citep{Moraetal2007}, in the demography and size of specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk}.
+Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
 
-In addition to these species-specific extensions, this modeling framework can easily be adjusted to consider 
-species interactions, especially predator-prey pairs. Predatory species affect marine populations and 
-considering a set of species that included both predators and prey would be equivalent to imposing additional 
-stressors on the prey species whose interactions we would like to measure \citep{Lingetal2009, 
-Gurevitchetal2000}. Considering multiple species would not only allow us to understand the effects of multiple 
-stressors on prey populations but would also afford us the opportunity to study how community composition 
-would change as a result of climate change, as has been observed in some marine systems \citep
-{Urbanetal2012}.
+The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.
+Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
+Our results, that spatial management increased the maximum harvesting rate at which the population could survive, corroborate these previous findings.
+The management strategies also affect how the two stressors interact with each other.
+Specifically, harvesting thresholds can alleviate interactions between the two stressors.
+While the management strategies only change harvesting practices and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement and place protected areas.
 
-Finally, our results suggest that particular combinations of harvesting and rate of environmental shift will affect 
-some species more than others. Indeed, \citet{Perryetal2005} found that fish that shifted in response to 
-warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, 
-smaller sizes at maturity). Using a simple mechanistic model like the one we presented can incorporate these 
-species interactions and multiple disturbances to understand whether specific life histories are likely to be 
-selected over others as harvesting and/or range shifts increase.
+The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
+However, it  ignores many of the complexities present in marine fisheries.
+We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
+However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect.
+We also did not include age structure in our model.
+The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work.
+Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
+In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
+By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
+
+Finally, our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
+Indeed, \citet{Perryetal2005} found that fish that shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity).
+Using a simple mechanistic model like the one we presented can incorporate these species interactions and multiple disturbances to understand whether specific life histories are likely to be selected over others as harvesting and/or range shifts increase.
 

### 2014-01-23 Simluations/Aspatial_fast/MPAcons_SIM.R
Changed MPA simulation parameters from 2/4 to 4/8, waiting for simulation to finish. Relates to issue #6

@@ -3,7 +3,7 @@ library(plyr)
 library(zoo)
 
 # load parameters, functions
-setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
+setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
 source("Parameters.R")
 source("Functions.R")
 

### 2014-01-23 Simluations/Aspatial_fast/Parameters.R

@@ -21,8 +21,8 @@ w = length(world)
 
 # haven't dealt with MPAs yet !
 
-cons.yes = rep(1,2*b/step_size)
-cons.no = rep(0,4*b/step_size)
+cons.yes = rep(1,4*b/step_size)
+cons.no = rep(0,8*b/step_size)
 fish.yes = rep(1,floor((1/3*b)/step_size))  # had to round because not complete step size. Rounded down. 
 fish.no = rep(0,floor((2/3*b)/step_size))
 

### 2014-01-23 Writing/MS.tex
Working on intro revisions, finished page 1 of Malin edits for #1

@@
 
-There are many stressors that can disturb an ecosystem.
-Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
+There are many stressors that can disturb an ecosystem, ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
 However, disturbances rarely occur in isolation.
 It is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
-A perturbation that has little effect when affecting a system individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}.
+A perturbation that has little effect when affecting a system may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}.
 Synergistic interactions between stressors are especially worrying if an animal population can persist in the face of a single stressor but not in the presence of multiple stressors (e.g.
 \citet{Pelletieretal2006}).
-However, it remains hard to predict when disturbances are likely to interact in ways not predicted by the effects of individual perturbations.
+However, it remains difficult to predict when disturbances are likely to interact in ways not anticipated by the effects of individual perturbations.
 
-Climate change is a prime example of a stressor occurring in a background of multiple other, possibly interacting, disturbances.
-These include pollution, habitat fragmentation, invasive species, and over-exploitation \citep{Wilcoveetal1998, Salaetal2000, MEA2005}.
-It is expected that  species will move their home ranges in response to climate change.
-Many marine fish are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}.
+Climate change is a prime example of a stressor occurring in a background of multiple other, possibly interacting, disturbances such as pollution, habitat fragmentation, invasive species, and over-exploitation \citep{Wilcoveetal1998, Salaetal2000, MEA2005}.
+Climate change is expected to drive range shifts;  many marine fish, for example, are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}.
 These species are faced not only with temperature-driven range range shifts but also with anthropogenic stressors including ocean acidification and commercial fishing \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}.
-Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact in the future.
+Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact.
 
-Many methods have been proposed to help mitigate the effects of harvesting on fish populations.
+Many management tactics have been proposed to help mitigate the effects of harvesting on fish populations.
 Marine protected areas (MPAs), which provide stepping stones to help species keep up with a changing environment, have been suggested as a key form of climate insurance \citep{Thomasetal2012, Hannahetal2007}.
@@
 A common approach to predicting future population distributions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
-These statistical models have been
+In terrestrial systems these statistical models have been
 used to find correlations between the presence or absence of fish
@@
 where conditions will be favorable for fish populations in the future
-\citep{Cheungetal2008, Cheungetal2009}.  Despite these models' widespread
+\citep{Cheungetal2008, Cheungetal2009, Hareetal2010, Lenoiretal2011}.  Despite these models' widespread
 adoption, they have  been criticized for being oversimplified

### 2014-01-23 Writing/MS.tex
Finished intro edits, this closes #1

--- 2014-01-17 Writing/MS.tex
@@
 Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
-However, disturbances rarely occur in isolation.
-It is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
-A perturbation that has little effect when affecting a system individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}.
-Synergistic interactions between stressors are especially worrying if an animal population can persist in the face of a single stressor but not in the presence of multiple stressors (e.g.
+However, disturbances rarely occur in isolation and it is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
+A perturbation that has little effect when it occurs individually may still amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}.
+Synergistic interactions between stressors can be especially problematic if an animal population can persist in the face of a single stressor but the combination of multiple stressors suffices to drive the population extinct (i.e.
 \citet{Pelletieretal2006}).
-However, it remains hard to predict when disturbances are likely to interact in ways not predicted by the effects of individual perturbations.
+However it remains difficult to predict when disturbances are likely to interact in ways not anticipated when perturbations are measured individually.
 
-Climate change is a prime example of a stressor occurring in a background of multiple other, possibly interacting, disturbances.
-These include pollution, habitat fragmentation, invasive species, and over-exploitation \citep{Wilcoveetal1998, Salaetal2000, MEA2005}.
-It is expected that  species will move their home ranges in response to climate change.
-Many marine fish are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}.
-These species are faced not only with temperature-driven range range shifts but also with anthropogenic stressors including ocean acidification and commercial fishing \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}.
+Climate change is a prime example of multiple, possibly interacting, disturbances.
+Species that are likely to undergo (or already undergoing) shifts in range and phenology are subject to many other disturbances including pollution, habitat fragmentation, over-exploitation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005}.
+A common approach to predicting future population distrubtions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
+These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to explain and predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
+Despite these models' widespread adoption, SDMs have frequently been criticized as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
+Recent work on range shifts has addressed some of these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}.
+However these models only address one disturbance: that of climate-driven range shifts.
+Not only does climate change disturb ecological systems via range shifts (for example), but these range shifts may interact with anthropogenetic stressors already present.
+Marine fish, for example, are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and projected to move more  \citep{Kelletal2005, Mackenzieetal2007}.
+These organisms not only are faced with ocean acidification and temperature driven range shifts \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}, but are often harvested commercially.
+
 Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact in the future.
+Predictions for how climate change will drive shifts in marine systems have focused on how physiological limitations \citep{PortnerKnust2007} and global fish distributions \citep{Cheungetal2008, Cheungetal2009, Hareetal2010, Lenoiretal2011} may change.
+Work considering the joint impacts of climate and fishing often consider climate fluctuations (large anomalies around the mean) rather than directional changes in climate \citep{WaltersParma1996, KingMcFarlane2006}.
+Work that does consider the effects of climate-driven range shifts on fishing typically are case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}.
+These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, however these models are so complex that understanding the relative importance of particular drivers, disturbances, and interactions is difficult (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances).
+Here we develop models of intermediate complexity \citep[sensu][]{Gaylordetal2005} that can help clarify the importance of two interacting disturbances: temperature-driven range shifts and harvest.
+Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in micosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
+This empirical work underscores the importance of understanding how range shifts and harvesting interact and, to our knowledge, there has been limited work examining interactions between harvesting and range shifts.
 
-Many methods have been proposed to help mitigate the effects of harvesting on fish populations.
-Marine protected areas (MPAs), which provide stepping stones to help species keep up with a changing environment, have been suggested as a key form of climate insurance \citep{Thomasetal2012, Hannahetal2007}.
-MPAs are frequently recommended for both conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}.
-Threshold harvesting levels below which no fish may be harvested have also been proposed as a means of both optimizing yield and conserving resources \citep{Landeetal1995, Landeetal1997}.
+Here we construct an analytical model of a fish population subject to both harvesting and climate change-induced range-shift.
+We explicitly include two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal.
+Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
+We focus on determining how climate-driven range shifts and fishing interact to affect population persistence.
+We find that these disturbances interact additively, not multiplicatively which is often the default assumption when considering mortality.
 
-A wide range of models have been used to study the effect of climate
-change and harvesting on fish populations.
-A common approach to predicting future population distributions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
-These statistical models have been
-used to find correlations between the presence or absence of fish
-species and biophysical properties of the environment such as mean or maximum
-temperatures, rainfall, or salinity, to explain and predict how species
-ranges will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}. Similarly,
-projection models of future climate conditions have been used to predict
-where conditions will be favorable for fish populations in the future
-\citep{Cheungetal2008, Cheungetal2009}.  Despite these models' widespread
-adoption, they have  been criticized for being oversimplified
-as they lack species interactions, dispersal and reproductive processes
-\citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
-
-Some recent work on range shifts has addressed these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}, which have been highlighted as important factors in fish populations' vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-These models only address one disturbance, climate-driven range shifts, and do not consider other anthropogenic stressors.
-The models of Zhou and Kot
-\citep{ZhouKot2011} and Berestycki et al. \citep{Berestyckietal2009} are of
-intermediate complexity (sensu
-\citep{Gaylordetal2005}).  Work that consider both climate-driven
-range shifts and fishing is often case-specific and detailed, including
-multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011,
-Barangeetal2011, Howardetal2013}. These cumulative
-impacts are important for management and conservation planning
-\citep{Allisonetal2009}.  However, these models are
-so complex that understanding the relative importance of particular
-drivers, disturbances, and interactions is difficult (but see
-\citep{Nyeetal2013} for an approach using ecosystem-level models to discern
-relative importance of disturbances).
-
-We analyzed an integrodifference model to study the effect of temperature-driven range shifts and
-harvesting, two important interacting disturbances, on the persistence of a marine population. In particular, we extended the model of
-Zhou and Kot \citep{ZhouKot2011}, which describes a fish population
-reproducing and dispersing while their range shifts, by
-additionally including harvesting pressure and introducing two different
-conservation measures. We found that the disturbances interact
-synergistically. We also found that MPAs can help a species persist with
-higher harvesting pressure, but they do not change the maximum speed of
-habitat shift with which a species can keep up.
+We also examine the effect of marine protected areas (MPAs) on species persistence.
+Protected areas have been suggested as a key form of climate insurance and stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
+MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat is shifting rapidly.
+We find that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
 

### 2014-01-23 Writing/MS.tex
formatted latex for ecological applications, this closes #8
formatting for Eco Applications, reopen #8, figure links are now broken

@@
-\documentclass[]{article}
-\usepackage[margin=.75in]{geometry}
+\documentclass[12pt,english]{article}
+\usepackage{geometry}
+\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
 \usepackage[T1]{fontenc}
@@
 \bibliographystyle{plainnat}
-\usepackage{todonotes}
 \usepackage{graphicx}
@@
 \usepackage{brush_style}
+\usepackage{setspace}
+\usepackage{lineno}
+\linenumbers
 \title{Species persistence under climate and fishing}
@@
 \maketitle
+
+\begin{spacing}{1.9}
+\begin{flushleft}
 
@@
 these approaches could mitigate the negative interaction of the two stressors.  
+\hspace{10cm}
+
+\noindent {\bf Keywords:} Climate change, fishing, integrodifference model, synergy, multiple disturbances
 
@@
 
-\begin{figure}[htbp]
-\begin{subfigure}{3in}
-\subcaption{\label{biomass}}
-\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
-\end{subfigure}
-\begin{subfigure}{3in}
-\subcaption{\label{rates}}
-\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
-\end{subfigure}
-\caption{(a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-(b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
-Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
-Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-These results are from an approximated Gaussian dispersal kernel with $L=1$.}
-\label{baseline}
-\end{figure}
-
 It is always the case that increasing the intrinsic growth rate, $R_0$, of the population increases the critical 
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=3in]{plots/synergy.pdf}
-\caption{Positive synergy between the two stressors.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
-\label{Synergy}
-\end{center}
-\end{figure}
-
 \subsection{Management Strategies }
@@
 overall catch at the highest harvest rates under which the population could survive (Figure \ref{management}).
-
-\begin{figure}[htbp]
-
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
-\end{subfigure}
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
-\end{subfigure}
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
-\end{subfigure}
-\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
-(a) No management.
-(b) Threshold harvesting levels.
-(c) MPAs.
-These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
-}
-\label{management}
-\end{figure}
 
@@
 
+\section{ Acknowledgements}
+We'd like to thank our reviewers for providing constructive comments.
+
 \bibliography{fish}
+
+\section{Figures}
+
+\begin{figure}[htbp]
+\begin{subfigure}{3in}
+\subcaption{\label{biomass}}
+\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
+\end{subfigure}
+\begin{subfigure}{3in}
+\subcaption{\label{rates}}
+\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
+\end{subfigure}
+%\caption{(a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+(b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
+Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
+These results are from an approximated Gaussian dispersal kernel with $L=1$.}
+\label{baseline}
+\end{figure}
+
+\pagebreak
+
+Figure \ref{baseline}: (a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+(b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
+Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
+These results are from an approximated Gaussian dispersal kernel with $L=1$.
+
+\pagebreak
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=3in]{plots/synergy.pdf}
+%\caption{Positive synergy between the two stressors.
+The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
+These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
+\label{Synergy}
+\end{center}
+\end{figure}
+
+\pagebreak
+
+Figure \ref{Synergy}: Positive synergy between the two stressors.
+The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
+These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+
+\pagebreak
+
+\begin{figure}[htbp]
+
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
+\end{subfigure}
+%\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+(a) No management.
+(b) Threshold harvesting levels.
+(c) MPAs.
+These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
+}
+\label{management}
+\end{figure}
+
+\pagebreak
+
+Figure \ref{management}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+(a) No management.
+(b) Threshold harvesting levels.
+(c) MPAs.
+These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
+
 \section{Appendix}
@@
 
+\end{flushleft}
+\end{spacing}
+
 \end{document}

### 2014-01-23 fish_figures.R
Working on issue #7, possible difference due to different color scale?

@@ -95,6 +95,9 @@ dev.off()
 ##### wireframe of the equilibrium biomass as a function of c and h : SIMULATIONS
 
 sim=read.csv("~/Desktop/noMPAnotThresh_add_2013-12-09.csv")
+# for emma's local copy
+	sim=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv")
+
 
 cvals=unique(sim$speed)
 hvals=unique(sim$harvest)
@@ -156,6 +159,8 @@ dev.off()
 ##########
 ##### wireframe of the equilibrium biomass as a function of c and h : MPAs
 mpas=read.csv("~/Desktop/fishMPAnotThresh_add_2013-12-09.csv")
+# emma's local copy
+	mpas = read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv")
 
 cvals=unique(mpas$speed)
 hvals=unique(mpas$harvest)
@@ -216,9 +221,12 @@ dev.off()
 
 ##### wireframe of the equilibrium biomass as a function of c and h : MPAs v2
 mpas2=read.csv("~/Desktop/consMPAnotThresh_add_2013-12-09.csv")
+# emma's local copy
+	mpas2=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2013-12-09.csv")
+	
 
-cvals=unique(mpas$speed)
-hvals=unique(mpas$harvest)
+cvals=unique(mpas2$speed)
+hvals=unique(mpas2$harvest)
 
 H=length(hvals)
 C=length(cvals)
@@ -277,6 +285,8 @@ dev.off()
 ########## 
 ##### wireframe of the equilibrium biomass as a function of c and h : THRESHOLD
 thresh=read.csv("~/Desktop/noMPA_Thresh_add_2013-12-10.csv")
+# emma's local copy
+	thresh=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/thresh/noMPA_Thresh_add_2013-12-10.csv")
 
 cvals=unique(thresh$speed)
 threshvals=unique(thresh$thresh)

### 2014-01-23 Simluations/Aspatial_fast/thresh_noMPA_SIM.R
Figured out problem in thresholds, thresholds were only going from 0-1 instead of the maximum observered population along world when finished initializing, now re-running threshold simulations, this helps along #7

@@ -1,9 +1,9 @@
-#rm(list=ls())
+rm(list=ls())
 library(plyr)
 library(zoo)
 
 # load parameters, functions
-setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
+setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
 source("thresh_Parameters.R")
 source("Functions.R")
 
@@ -24,6 +24,14 @@ for(t in 2:maxt){
   init[,t]= output[[1]]
   MPA.current = output[[3]]
 }
+
+# thresholds need to be relative to maximum population along 1d world
+
+biggestP <- max(init[,maxt])
+
+thresholds <- thresholds * biggestP
+# add a buffer to last to make sure captures biggest population
+thresholds <- tail(thresholds,1) + 0.001
 
 # standard for equilibrium is the difference in the final step of the initialization 
 init.diff <- diff(colSums(init))[149]

### 2014-01-23 fish_figures.R

@@ -295,16 +295,16 @@ H=length(threshvals)
 C=length(cvals)
 L=length(thresh$Equil.pop)
 
-ebm=array(,c(C,H))
+ebmThresh=array(,c(C,H))
 
 for(i in 1:L){
 	tind=which(threshvals==thresh$thresh[i])
 	cind=which(cvals==thresh$speed[i])
-	ebm[cind,tind]=thresh$Equil.pop[i]
+	ebmThresh[cind,tind]=thresh$Equil.pop[i]
 }
 # threshvals=c(threshvals,0)
 # toplot_ebm=cbind(toplot_ebm,rep(0,C))
-toplot_ebm=ebm
+toplot_ebmThresh=ebmThresh
 
 #breaks and colors
 lower=min(ebm)+1
@@ -335,14 +335,14 @@ pdf(file='eqbiomass_thresh.pdf',width=6.83)
 
 # wireframe(toplot_ebm[C:1,H:1], par.settings = list(axis.line = list(col = "transparent")),xlab=list("Rate of environmental shift",rot=10),ylab=list("Threshold",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,screen=list(z=30,x=-80))
 par(oma=oma)
-image.plot(toplot_ebm[1:C,H:1],bigplot=c(.13,.85,.15,.95),smallplot=c(.91,.95,.15,.95))
-image(cvals,(threshvals),toplot_ebm[1:C,H:1],breaks=cuts,col=mycols,
+image.plot(toplot_ebmThresh[1:C,H:1],bigplot=c(.13,.85,.15,.95),smallplot=c(.91,.95,.15,.95))
+image(cvals,(threshvals),toplot_ebmThresh[1:C,H:1],breaks=cuts,col=mycols,
 	xlab="Rate of environmental shift",ylab="Threshold",
 	cex.lab=cex.lab,cex.axis=cex.axis,axes=F)
 axis(1)
 axis(2,at=seq(0,1,by=0.2),labels=rev(seq(0,1,by=0.2)))
 box()
-image.plot(cvals,(threshvals),toplot_ebm,legend.shrink=1,legend.width=.1,zlim=range(cuts),
+image.plot(cvals,(threshvals),toplot_ebmThresh,legend.shrink=1,legend.width=.1,zlim=range(cuts),
[…]

### 2014-01-23 fish_figures.R
undoing changes made to  because I figured out the real problem with the threshold simulations, see #7

@@ -295,16 +295,16 @@ H=length(threshvals)
 C=length(cvals)
 L=length(thresh$Equil.pop)
 
-ebmThresh=array(,c(C,H))
+ebm=array(,c(C,H))
 
 for(i in 1:L){
 	tind=which(threshvals==thresh$thresh[i])
 	cind=which(cvals==thresh$speed[i])
-	ebmThresh[cind,tind]=thresh$Equil.pop[i]
+	ebm[cind,tind]=thresh$Equil.pop[i]
 }
 # threshvals=c(threshvals,0)
 # toplot_ebm=cbind(toplot_ebm,rep(0,C))
-toplot_ebmThresh=ebmThresh
+toplot_ebm=ebm
 
 #breaks and colors
 lower=min(ebm)+1
@@ -335,14 +335,14 @@ pdf(file='eqbiomass_thresh.pdf',width=6.83)
 
 # wireframe(toplot_ebm[C:1,H:1], par.settings = list(axis.line = list(col = "transparent")),xlab=list("Rate of environmental shift",rot=10),ylab=list("Threshold",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,screen=list(z=30,x=-80))
 par(oma=oma)
-image.plot(toplot_ebmThresh[1:C,H:1],bigplot=c(.13,.85,.15,.95),smallplot=c(.91,.95,.15,.95))
-image(cvals,(threshvals),toplot_ebmThresh[1:C,H:1],breaks=cuts,col=mycols,
+image.plot(toplot_ebm[1:C,H:1],bigplot=c(.13,.85,.15,.95),smallplot=c(.91,.95,.15,.95))
+image(cvals,(threshvals),toplot_ebm[1:C,H:1],breaks=cuts,col=mycols,
 	xlab="Rate of environmental shift",ylab="Threshold",
 	cex.lab=cex.lab,cex.axis=cex.axis,axes=F)
 axis(1)
 axis(2,at=seq(0,1,by=0.2),labels=rev(seq(0,1,by=0.2)))
 box()
-image.plot(cvals,(threshvals),toplot_ebmThresh,legend.shrink=1,legend.width=.1,zlim=range(cuts),
+image.plot(cvals,(threshvals),toplot_ebm,legend.shrink=1,legend.width=.1,zlim=range(cuts),
[…]

### 2014-01-23 Simluations/Aspatial_fast/thresh_noMPA_SIM.R
finished running conservation MPA, new data present, helps #6

@@ -31,7 +31,7 @@ biggestP <- max(init[,maxt])
 
 thresholds <- thresholds * biggestP
 # add a buffer to last to make sure captures biggest population
-thresholds <- tail(thresholds,1) + 0.001
+tail(thresholds,1) <- tail(thresholds,1) + 0.001
 
 # standard for equilibrium is the difference in the final step of the initialization 
 init.diff <- diff(colSums(init))[149]
@@ -110,7 +110,7 @@ for(q in 1:length(speeds)){
     summaries[rownumber[j,q],] <- c(
       pop.mean, pop.sd, pop.se, 
       harv.mean, harv.sd, harv.se, 
-      speeds[q], 1,ifelse(exists(thresholds[j]),thresholds[j],NA),
+      speeds[q], 1,ifelse(exists("thresholds"),thresholds[j],NA),
       ncol(init.h), ncol(move))
     print(paste("harvest is ",round(j/length(thresholds),1)*100,"% done and speed is ",round(q/length(speeds),1)*100, "% done",sep=""))
     
@@ -119,4 +119,4 @@ for(q in 1:length(speeds)){
 }
 
 write.csv(summaries,file = paste("Data/noMPA_Thresh_add_",Sys.Date(),".csv",sep=""))
-notify('Simulation is finished!')
+#notify('Simulation is finished!')

### 2014-01-23 fish_figures.R
Changing  to point to new consMPA data, just need to run it now, #9

@@ -220,9 +220,9 @@ dev.off()
 
 
 ##### wireframe of the equilibrium biomass as a function of c and h : MPAs v2
-mpas2=read.csv("~/Desktop/consMPAnotThresh_add_2013-12-09.csv")
+mpas2=read.csv("~/Desktop/consMPAnotThresh_add_2014-01-23.csv")
 # emma's local copy
-	mpas2=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2013-12-09.csv")
+	mpas2=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv")
 	
 
 cvals=unique(mpas2$speed)

### 2014-01-23 Writing/MS.tex
Fixed broken figure links, close #8
Fixes #10

@@
 \noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
-Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
-specify a recruitment function.  We chose to use a Beverton-Holt function,
+Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, we must specify a recruitment function to perform simulations.
+We chose to use a Beverton-Holt function,
 
@@
 \noindent where $R_0$ is the intrinsic growth rate and $K$ is the carrying capacity.  In the first generation, we seeded the world with 50 individuals at a single point, as in 
-\citep{ZhouKot2011}. We first allowed the population to reach equilibrium without harvesting or climate shift.  We then 
+\citep{ZhouKot2011}. We first ran through 150 generations in order for the population to reach equilibrium without harvesting or climate shift.  We then 
 added harvesting pressure, allowed the population to again reach equilibrium, and finally added climate 
@@
 We added harvesting pressure by harvesting a constant proportion of the population, in order to confirm our analytical results.
-We also extended our analytical work by evaluating the effect of a threshold harvest rule and marine protected 
+We extended our work by evaluating the effect of a threshold harvest rule and marine protected
 areas (MPAs). With a threshold rule, we evaluated the population at each point in space to determine how 
@@
 were harvested. To implement an MPA management strategy, we used two commonly advocated approaches: 
-MPAs designed to improve fisheries yields and MPAs designed for primarily for conservation. These networks 
+MPAs designed to improve fisheries yields and MPAs designed for primarily for conservation [REFS]. These networks 
 of MPAs were introduced into our simulations by designating segments of space in which harvesting was 
-forbidden (i.e.~harvesting rates were equal to $0$). The conservation MPAs had a length of twice the average 
-dispersal distance and had a distance of 4 times the average dispersal distance between them. Fisheries MPAs had a length of 1/3 of the 
+forbidden (i.e.~harvesting rates were equal to $0$). The conservation MPAs had a length of 4 times the average 
+dispersal distance and had a distance of 8 times the average dispersal distance between them. Fisheries MPAs had a length of 1/3 of the 
 average dispersal distance and had a distance of 2/3 of the average dispersal distance between them.
+These distances were chosen for the following reasons: the goal of conservation-oriented MPAs is to protect source populations, ideally that are self sustaining under a static climate.
+Fisheries MPAs are focused on maximizing spillover of fish populations into fishable waters, thus many MPAs, each of which are smaller than the average dispersal distance, replicates this high spillover effect.
 
-In our simulations, we defined the critical rate of environmental shift and the critical harvest rate as model runs 
-in which the total population at equilibrium was less than 0.001.
-Equilibrium is calculated as the mean of 300 time steps once the difference in biomass between time step $t$ and $t+1$ no greater than $0.0360515$.
+Equilibrium is calculated as the mean of 300 time steps once the difference in biomass between time step $t$ and $t+1$ was no greater than $0.1$.
 Equilibrium biomass and catch are defined as the total amount of fish present and harvested, respectively, at equilibrium.
@@
 \end{subfigure}
-%\caption{(a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+\caption{
+%(a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
 These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
@@
 Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-These results are from an approximated Gaussian dispersal kernel with $L=1$.}
+These results are from an approximated Gaussian dispersal kernel with $L=1$.
+}
 \label{baseline}
@@
 \includegraphics[width=3in]{plots/synergy.pdf}
-%\caption{Positive synergy between the two stressors.
+\caption{
+%Positive synergy between the two stressors.
 The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
+These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+}
 \label{Synergy}
@@
 \end{subfigure}
-%\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+\caption{
+%The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
 (a) No management.
@@
 These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
+
+\pagebreak
 

### 2014-01-25 fish_figures.R
Fixed #7, #9

@@ -203,7 +203,7 @@ myaxes=list(arrows=FALSE,
 	z=list(at=seq(0,max(ebm),by=250),labels=as.character(seq(0,max(ebm),by=250))))
 
 quartz()
-pdf(file='eqbiomass_mpa.pdf',width=6.83)
+pdf(file='eqbiomass_fishmpa.pdf',width=6.83)
 
 # wireframe(toplot_ebm[C:1,H:1],xlab=list("Rate of environmental shift",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
 par(oma=oma)
@@ -266,7 +266,7 @@ myaxes=list(arrows=FALSE,
 	z=list(at=seq(0,max(ebm),by=250),labels=as.character(seq(0,max(ebm),by=250))))
 
 quartz()
-pdf(file='eqbiomass_mpa_v2.pdf',width=6.83)
+pdf(file='eqbiomass_consmpa.pdf',width=6.83)
 
 # wireframe(toplot_ebm[C:1,H:1],xlab=list("Rate of environmental shift",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
 par(oma=oma)
@@ -284,9 +284,9 @@ dev.off()
 ##########
 ########## 
 ##### wireframe of the equilibrium biomass as a function of c and h : THRESHOLD
-thresh=read.csv("~/Desktop/noMPA_Thresh_add_2013-12-10.csv")
+thresh=read.csv("~/Desktop/noMPA_Thresh_add_2014-01-24.csv")
 # emma's local copy
-	thresh=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/thresh/noMPA_Thresh_add_2013-12-10.csv")
[…]

### 2014-01-25 Writing/plots/figureCode/MPA_diffs.Rmd
comparing MPA scenarios to each other

# Comparing MPAs to each other, to no management

```{r required, echo=FALSE,message=FALSE}
library(fields)
library(lattice)
library(colorspace)
require(RCurl)
```

```{r loadData, echo=FALSE, message=FALSE}
#sim=read.csv("~/Desktop/noMPAnotThresh_add_2013-12-09.csv")
# for emma's local copy
	#sim=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv")

# for synced online github data
  x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL25vTVBBbm90VGhyZXNoX2FkZF8yMDEzLTEyLTA5LmNzdiIsImV4cGlyZXMiOjEzOTEyNjg5OTB9--e041d4ef68764b1fc49140c128bda685c6f740f3")
  sim = read.csv(textConnection(x))

  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2Zpc2hNUEFub3RUaHJlc2hfYWRkXzIwMTMtMTItMDkuY3N2IiwiZXhwaXJlcyI6MTM5MTI2OTAzM30%3D--f35f2593a7e901c900188fadca920fdd68a57e35")
  fishmpa = read.csv(textConnection(fish))

  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2NvbnNNUEFub3RUaHJlc2hfYWRkXzIwMTQtMDEtMjMuY3N2IiwiZXhwaXJlcyI6MTM5MTI2OTA3Mn0%3D--ffb8451d33daaeab48f8b3be3357cacbda0aadbb")
  consmpa = read.csv(textConnection(cons))
```

```{r plotFunctions, echo=FALSE,message=FALSE}
plotfig <- function(df){
  cvals=unique(df$speed)
  hvals=unique(df$harvest)

  H=length(hvals)
  C=length(cvals)
  L=length(df$Equil.pop)

  ebm=array(,c(C,H))

  for(i in 1:L){
    hind=which(hvals==df$harvest[i])
    cind=which(cvals==df$speed[i])
    ebm[cind,hind]=df$Equil.pop[i]
  }

  all <- list(ebm = ebm, cvals = cvals, hvals = hvals, C = C, H = H)
  return(all)
}

plotParams <- function(lis){
  ebm <- lis$ebm
  cvals <- lis$cvals
  hvals <- lis$hvals
  C <- lis$C
  H <- lis$H
  # plotting parameters
  cex.lab=3
  cex.axis=2
  line=5
  lwd=5
  yaxs='i'
  xaxs='i'
  cex.lab=1.5
  cex.axis=1
  line=1
  oma=c(0,0,0,2)
  #breaks and colors
    lower=min(ebm)+1
    cuts=c(0,seq(lower,1538.649+.5,length.out=100))
    mycols=(sequential_hcl(length(cuts)-2,c.=c(0,0),l=c(30,100)))
    mycols=c('black',mycols)

  #axis labels
    numxlabs=11
    numylabs=3
    xdiff=.1
    ydiff=.2
    myaxes=list(arrows=FALSE,
      x=list(at=rev(seq(1,C,length.out=(cvals[C]-cvals[1])/xdiff+1)),labels=as.character(round(seq(cvals[1],cvals[C],by=xdiff),2))),
      y=list(at=rev(seq(1,H,length.out=(hvals[H]-hvals[1])/ydiff)),labels=as.character(round(seq(hvals[1],hvals[H],by=ydiff),2))),
      z=list(at=seq(0,1538.649,by=250),labels=as.character(seq(0,1538.649,by=250))))

  par(oma=oma)
  image.plot(cvals,hvals,ebm,breaks=cuts,col=mycols,
    xlab="Rate of environmental shift",ylab="Harvesting rate",
  	cex.lab=cex.lab,cex.axis=cex.axis, 
  	yaxs=yaxs,xaxs=xaxs,axes=TRUE,
  	legend.shrink=1,legend.width=.1,zlim=range(cuts),
  	axis.args=list(at=myaxes$z$at,labels=myaxes$z$labels,cex.axis=cex.axis),
  	legend.args=list(text="Equilibrium biomass",cex=cex.lab,side=2,line=0.5,las=0),
  	bigplot=c(.13,.85,.15,.95),smallplot=c(.91,.95,.15,.95),horizontal=FALSE)
  box()
}
```

Comparing no management to several small MPAs

```{r diff_fishnoManagement, echo=FALSE,message=FALSE}
noMang <- plotfig(sim)
fish <- plotfig(fishmpa)
combined <- noMang
combined$ebm = fish$ebm - noMang$ebm
plotParams(combined)
```

Comparing no management to few large MPAs

```{r diff_consnoManagement, message=FALSE,echo=FALSE}
noMang <- plotfig(sim)
cons <- plotfig(consmpa)
combined <- noMang
combined$ebm = cons$ebm - noMang$ebm
plotParams(combined)
```

Comparing the MPAs to each other

```{r eachOther, message=FALSE, echo=FALSE, warning=FALSE}
combined <- cons
combined$ebm = cons$ebm - fish$ebm
plotParams(combined)
```

### 2014-01-25 fish_figures.R

@@ -349,6 +349,61 @@ image.plot(cvals,(threshvals),toplot_ebm,legend.shrink=1,legend.width=.1,zlim=ra
 
 dev.off()
 
+##########
+########## 
+##### Difference between MPAs
+
+fishmpa = mpas
+consmpa = mpas2
+
+cvals=unique(fishmpa$speed)
+hvals=unique(fishmpa$harvest)
+
+H=length(hvals)
+C=length(cvals)
+L=length(fishmpa$Equil.pop)
+
+ebmfish=array(,c(C,H))
+
+for(i in 1:L){
+	hind=which(hvals==fishmpa$harvest[i])
+	cind=which(cvals==fishmpa$speed[i])
+	ebmfish[cind,hind]=fishmpa$Equil.pop[i]
+}
+
+toplot_ebmfish=ebmfish
+#breaks and colors
+lower=min(ebmfish)
+cuts=c(0,seq(lower,max(ebmfish)+.5,length.out=100))
+mycols=(sequential_hcl(length(cuts)-2,c.=c(0,0),l=c(30,100)))
+mycols=c('black',mycols)
+
+#axis labels
+numxlabs=11
+numylabs=3
+xdiff=.1
+ydiff=.2
+myaxes=list(arrows=FALSE,
+	x=list(at=rev(seq(1,C,length.out=(cvals[C]-cvals[1])/xdiff+1)),labels=as.character(round(seq(cvals[1],cvals[C],by=xdiff),2))),
+	y=list(at=rev(seq(1,H,length.out=(hvals[H]-hvals[1])/ydiff)),labels=as.character(round(seq(hvals[1],hvals[H],by=ydiff),2))),
+	z=list(at=seq(0,max(ebm),by=250),labels=as.character(seq(0,max(ebm),by=250))))
+
+quartz()
+pdf(file='eqbiomass_diffmpa.pdf',width=6.83)
+
+# wireframe(toplot_ebm[C:1,H:1],xlab=list("Rate of environmental shift",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
+par(oma=oma)
+image.plot(cvals,hvals,toplot_ebm2,breaks=cuts,col=mycols,
[…]

### 2014-01-28 Writing/MS.tex
Working on catch story, updated Malin's comments

@@
 
-\begin{spacing}{1.9}
-\begin{flushleft}
+%\begin{spacing}{1.9}
+%\begin{flushleft}
 
@@
 We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$.
-We found these rates by finding the parameters that make the trivial pulse unstable. Evaluating stability in this kind of model is in 
-general difficult to do analytically. It becomes easier if the dispersal kernel is separable into its dependence on the 
+We found these rates by finding the parameters that make the trivial pulse unstable.
+Evaluating stability in this kind of model is in  general difficult to do analytically.
+It becomes easier if the dispersal kernel is separable into its dependence on the
 source of larvae and its dependence on the destination of the larvae,
@@
 
-\noindent where $R_0$ is the intrinsic growth rate and $K$ is the carrying capacity.  In the first generation, we seeded the world with 50 individuals at a single point, as in 
-\citep{ZhouKot2011}. We first ran through 150 generations in order for the population to reach equilibrium without harvesting or climate shift.  We then 
-added harvesting pressure, allowed the population to again reach equilibrium, and finally added climate 
-change by moving the viable patch.
+\noindent where $R_0$ is the intrinsic growth rate and $K$ is the carrying capacity.
+In the first generation, we seeded the world with 50 individuals at a single point, as in \citep{ZhouKot2011}.
+We first ran through 150 generations in order for the population to reach equilibrium without harvesting or climate shift.
+We then added harvesting pressure, allowed the population to again reach equilibrium, and finally added climate change by moving the viable patch.
 
@@
 harvesting occurred. If the population exceeded the threshold, then a proportion of the `surplus' individuals 
-were harvested. To implement an MPA management strategy, we used two commonly advocated approaches: 
-MPAs designed to improve fisheries yields and MPAs designed for primarily for conservation [REFS]. These networks 
-of MPAs were introduced into our simulations by designating segments of space in which harvesting was 
-forbidden (i.e.~harvesting rates were equal to $0$). The conservation MPAs had a length of 4 times the average 
-dispersal distance and had a distance of 8 times the average dispersal distance between them. Fisheries MPAs had a length of 1/3 of the 
-average dispersal distance and had a distance of 2/3 of the average dispersal distance between them.
-These distances were chosen for the following reasons: the goal of conservation-oriented MPAs is to protect source populations, ideally that are self sustaining under a static climate.
-Fisheries MPAs are focused on maximizing spillover of fish populations into fishable waters, thus many MPAs, each of which are smaller than the average dispersal distance, replicates this high spillover effect.
+were harvested
+
+MPAs are a form of management designed to check the impact of fishing on targeted populations and are typically designed to meet either conservation of fishery management goals \citep{Agardy1994, HollandBrazee1996, Gainesetal2010a}.
+To implement an MPA management strategy in our model, we examine the effect of both of these commonly advocated approaches.
+While both conservation and fisheries oriented MPA schemes align in their goal of maintaining a sustainable fished population, they differ in desired level of adult spillover.
+Fisheries-oriented MPAs are often designed such that they maximize adult spillover into fishable areas by creating many small reserves closely spaced \citep{HastingsBotsford2003}.
+The converse of this is the goal of conservation-oriented MPAs which seek to reduce adult spillover by minizing the ratio between the reserve edge length relative to area protected \citep{Gainesetal2010a}.
+
+Networks of MPAs were introduced into our simulations by designating segments of space in which harvesting was forbidden (i.e.
+harvesting rates were equal to $0$).
+Conservation-oriented MPAs, are frequently large and rarely part of a larger network of reserves \citep{HastingsBotsford2003}.
+For solitary reserves to be successful at protecting target species, they must encompass self-sustaining fish populations \citep{HastingsBotsford2006, Gainesetal2010a}.
+As such modeling studies estimate that isolated reserves must be at least as large as the average dispersal distance for the targeted fish species \citep{Lockwoodetal2002, HastingsBotsford2003, Botsfordetal2001, Gainesetal2010}.
+To implement conservation MPAs we created reserves with a length of 4 times the average dispersal distance and had a distance of 8 times the average dispersal distance between them to ensure that populations would be self sustiaining and not dependent on other dispersal for other reserves \citep{Lockwoodetal2002}.
+
+Previous work has shown that if MPAs are to benefit fisheries, the reserves should be broken into a network, closely spaced to maximize adult spillover into fishable areas and export of larvae from reserve to reserve \citep{ HastingsBotsford2003, Gaylordetal2005, Gainesetal2010a}.
+To mimic this management scheme, MPAs had a length of $\tfrac{1}{3}$ of the average dispersal distance and had a distance of $\tfrac{2}{3}$ of the average dispersal distance between them.
 
@@
 
-We found that thresholds of any amount alleviate the effects of harvesting, and the ability of the population to 
-persist is recovered (Figure \ref{management}). Thus when thresholds are in place, the harvesting rate no longer 
-determines the critical rate of environmental shift $c^*$. We also examined the effect of marine protected 
-areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and 
-climate change parameters where the fish population could survive. With MPAs in place, the population had a 
-slightly higher abundance along the edges of the patch where the population is limited by harvesting, which 
-translated into a slightly increased critical harvest rate (Figure \ref{management}). Additionally MPAs increased 
-overall catch at the highest harvest rates under which the population could survive (Figure \ref{management}).
+We found that when thresholds are in place, the harvesting rate no longer determines the critical rate of environmental shift $c^*$ (Figure \ref{management}).
+We also examined the effect of marine protected  areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and  climate change parameters where the fish population could survive.
+With MPAs in place, the population had a  slightly higher abundance along the edges of the patch where the population is limited by harvesting, which  translated into a slightly increased critical harvest rate (Figure \ref{management}).
+Additionally MPAs increased overall catch at the highest harvest rates under which the population could survive (Figure \ref{management}).
 
@@
 
-\end{flushleft}
-\end{spacing}
+%\end{flushleft}
+%\end{spacing}
 

### 2014-01-28 Writing/plots/figureCode/compareCatch.Rmd
Working on catch story, updated Malin's comments

# Comparing catch

```{r required, echo=FALSE,message=FALSE}
library(fields)
library(lattice)
library(colorspace)
require(RCurl)
require(ggplot2)
require(ggthemes)
```

```{r loadData, echo=FALSE, message=FALSE}
#sim=read.csv("~/Desktop/noMPAnotThresh_add_2013-12-09.csv")
# for emma's local copy
  #sim=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv")

# for synced online github data
  x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL25vTVBBbm90VGhyZXNoX2FkZF8yMDEzLTEyLTA5LmNzdiIsImV4cGlyZXMiOjEzOTEyNjg5OTB9--e041d4ef68764b1fc49140c128bda685c6f740f3")
  sim = read.csv(textConnection(x))

  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2Zpc2hNUEFub3RUaHJlc2hfYWRkXzIwMTMtMTItMDkuY3N2IiwiZXhwaXJlcyI6MTM5MTI2OTAzM30%3D--f35f2593a7e901c900188fadca920fdd68a57e35")
  fishmpa = read.csv(textConnection(fish))

  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2NvbnNNUEFub3RUaHJlc2hfYWRkXzIwMTQtMDEtMjMuY3N2IiwiZXhwaXJlcyI6MTM5MTI2OTA3Mn0%3D--ffb8451d33daaeab48f8b3be3357cacbda0aadbb")
  consmpa = read.csv(textConnection(cons))

# combine all data into one dataframe

sim$management = rep("noManagement",nrow(sim))
consmpa$management = rep("consMPA", nrow(consmpa))
fishmpa$management = rep("fishMPA", nrow(fishmpa))

allData <- rbind(sim,consmpa,fishmpa)
```

Map equilibrium harvest as function of speed of climate velocity and harvest rate.

```{r Compare_Management,echo = FALSE, fig.width=10, fig.height=10}
contourplot(Equil.harvest ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,3), drape=TRUE)

```

Looks like conservation MPAs do the best at the highest amount of harvest rates. Wonder if there's a difference in variance among these different MPAs. Subsetting the data to just look at harvest rate = 0.04 I find that the conservation MPA seems to have the most variance around average catch.

```{r variance, echo = FALSE, fig.width=10, fig.height=5}
subData <- subset(allData,harvest == 0.04 )
pd <- position_dodge(.1)
ggplot(subData, aes(x = speed, y = Equil.harvest))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.harvest - Equil.harvest.se, ymax = Equil.harvest + Equil.harvest.se), alpha = 0.35) + theme_bw()
```

Trying the same plot but a higher rate of catch: $h = 0.10$ I get,

```{r variance_high, echo = FALSE, fig.width=10, fig.height=5}
subData <- subset(allData,harvest == 0.10 )
pd <- position_dodge(.1)
ggplot(subData, aes(x = speed, y = Equil.harvest))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.harvest - Equil.harvest.se, ymax = Equil.harvest + Equil.harvest.se), alpha = 0.35) + theme_bw()
```

### 2014-01-29 Writing/plots/figureCode/compareCatch.Rmd
Looking at how catch varies, #13

@@
 
-Trying the same plot but a higher rate of catch: $h = 0.10$ I get,
+Trying the same plot but a higher rate of catch: $$h = 0.10$$ I get,
 
@@
 ```
+
+Looking at holding speed constant and examining catch as a function of harvest
+
+```{r harvestVar, echo = FALSE, fig.width=10, fig.height=5}
+subData <- subset(allData,speed==0.1)
+ggplot(subData, aes(x = harvest, y = Equil.harvest)) + facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.harvest - Equil.harvest.se, ymax = Equil.harvest + Equil.harvest.se), alpha = 0.35) + theme_bw()
+
+subData <- subset(allData,speed==0.2)
+ggplot(subData, aes(x = harvest, y = Equil.harvest)) + facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.harvest - Equil.harvest.se, ymax = Equil.harvest + Equil.harvest.se), alpha = 0.35) + theme_bw()
+
+subData <- subset(allData,speed==0.3)
+ggplot(subData, aes(x = harvest, y = Equil.harvest)) + facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.harvest - Equil.harvest.se, ymax = Equil.harvest + Equil.harvest.se), alpha = 0.35) + theme_bw()
+```

### 2014-01-29 Writing/MS.tex
more of malin's comments

--- 2014-01-23 Writing/MS.tex
@@
-\documentclass[12pt,english]{article}
-\usepackage{geometry}
-\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
+\documentclass[]{article}
+\usepackage[margin=.75in]{geometry}
 \usepackage[T1]{fontenc}
@@
 \bibliographystyle{plainnat}
+\usepackage{todonotes}
 \usepackage{graphicx}
@@
 \usepackage{brush_style}
-\usepackage{setspace}
-\usepackage{lineno}
-\linenumbers
 \title{Species persistence under climate and fishing}
@@
 
-\begin{spacing}{1.9}
-\begin{flushleft}
-
 \section{Abstract}
@@
 harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
-persistence. We found threshold rates of harvesting and of environmental shift that are needed to allow the 
-population to persist, studied how these critical parameters depend on the growth rate and dispersal behavior 
-of the population, and measured the interactions between the stressors. In particular, we found low but positive 
+persistence.
+We found threshold rates of harvesting and of environmental shift such that more extreme harvesting or shift will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior
+of the population.  We then measured the interaction between the stressors and we found low but positive 
 synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. 
@@
 these approaches could mitigate the negative interaction of the two stressors.
-\hspace{10cm}
-
-\noindent {\bf Keywords:} Climate change, fishing, integrodifference model, synergy, multiple disturbances
 
@@
 Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
-However, disturbances rarely occur in isolation and it is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
-A perturbation that has little effect when it occurs individually may still amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}.
-Synergistic interactions between stressors can be especially problematic if an animal population can persist in the face of a single stressor but the combination of multiple stressors suffices to drive the population extinct (i.e.
+However, disturbances rarely occur in isolation.
+It is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
+A perturbation that has little effect when affecting a system individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}.
+Synergistic interactions between stressors are especially worrying if an animal population can persist in the face of a single stressor but not in the presence of multiple stressors (e.g.
 \citet{Pelletieretal2006}).
-However it remains difficult to predict when disturbances are likely to interact in ways not anticipated when perturbations are measured individually.
+However, it remains hard to predict when disturbances are likely to interact in ways not predicted by the effects of individual perturbations.
 
-Climate change is a prime example of multiple, possibly interacting, disturbances.
-Species that are likely to undergo (or already undergoing) shifts in range and phenology are subject to many other disturbances including pollution, habitat fragmentation, over-exploitation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005}.
-A common approach to predicting future population distrubtions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
-These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to explain and predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
-Despite these models' widespread adoption, SDMs have frequently been criticized as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
-Recent work on range shifts has addressed some of these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}.
-However these models only address one disturbance: that of climate-driven range shifts.
-Not only does climate change disturb ecological systems via range shifts (for example), but these range shifts may interact with anthropogenetic stressors already present.
-Marine fish, for example, are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and projected to move more  \citep{Kelletal2005, Mackenzieetal2007}.
-These organisms not only are faced with ocean acidification and temperature driven range shifts \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}, but are often harvested commercially.
+Climate change is a prime example of a stressor occurring in a background of multiple other, possibly interacting, disturbances.
+These include pollution, habitat fragmentation, invasive species, and over-exploitation \citep{Wilcoveetal1998, Salaetal2000, MEA2005}.
+It is expected that  species will move their home ranges in response to climate change.
+Many marine fish are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}.
+These species are faced not only with temperature-driven range range shifts but also with anthropogenic stressors including ocean acidification and commercial fishing \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}.
+Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact in the future.
 
-Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact in the future.
-Predictions for how climate change will drive shifts in marine systems have focused on how physiological limitations \citep{PortnerKnust2007} and global fish distributions \citep{Cheungetal2008, Cheungetal2009, Hareetal2010, Lenoiretal2011} may change.
-Work considering the joint impacts of climate and fishing often consider climate fluctuations (large anomalies around the mean) rather than directional changes in climate \citep{WaltersParma1996, KingMcFarlane2006}.
-Work that does consider the effects of climate-driven range shifts on fishing typically are case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}.
-These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, however these models are so complex that understanding the relative importance of particular drivers, disturbances, and interactions is difficult (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances).
-Here we develop models of intermediate complexity \citep[sensu][]{Gaylordetal2005} that can help clarify the importance of two interacting disturbances: temperature-driven range shifts and harvest.
-Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in micosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
-This empirical work underscores the importance of understanding how range shifts and harvesting interact and, to our knowledge, there has been limited work examining interactions between harvesting and range shifts.
+Many methods have been proposed to help mitigate the effects of harvesting on fish populations.
+Marine protected areas (MPAs), which provide stepping stones to help species keep up with a changing environment, have been suggested as a key form of climate insurance \citep{Thomasetal2012, Hannahetal2007}.
+MPAs are frequently recommended for both conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}.
+Threshold harvesting levels below which no fish may be harvested have also been proposed as a means of both optimizing yield and conserving resources \citep{Landeetal1995, Landeetal1997}.
 
-Here we construct an analytical model of a fish population subject to both harvesting and climate change-induced range-shift.
-We explicitly include two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal.
-Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We focus on determining how climate-driven range shifts and fishing interact to affect population persistence.
-We find that these disturbances interact additively, not multiplicatively which is often the default assumption when considering mortality.
+A wide range of models have been used to study the effect of climate
+change and harvesting on fish populations.
+A common approach to predicting future population distributions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
+These statistical models have been
+used to find correlations between the presence or absence of fish
+species and biophysical properties of the environment such as mean or maximum
+temperatures, rainfall, or salinity, to explain and predict how species
+ranges will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}. Similarly,
+projection models of future climate conditions have been used to predict
+where conditions will be favorable for fish populations in the future
+\citep{Cheungetal2008, Cheungetal2009}.  Despite these models' widespread
+adoption, they have  been criticized for being oversimplified
+as they lack species interactions, dispersal and reproductive processes
+\citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
 
-We also examine the effect of marine protected areas (MPAs) on species persistence.
-Protected areas have been suggested as a key form of climate insurance and stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
-MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat is shifting rapidly.
-We find that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
+Some recent work on range shifts has addressed these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}, which have been highlighted as important factors in fish populations' vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
+These models only address one disturbance, climate-driven range shifts, and do not consider other anthropogenic stressors.
+The models of Zhou and Kot
+\citep{ZhouKot2011} and Berestycki et al. \citep{Berestyckietal2009} are of
+intermediate complexity (sensu
+\citep{Gaylordetal2005}).  Work that consider both climate-driven
+range shifts and fishing is often case-specific and detailed, including
+multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011,
+Barangeetal2011, Howardetal2013}. These cumulative
+impacts are important for management and conservation planning
+\citep{Allisonetal2009}.  However, these models are
+so complex that understanding the relative importance of particular
+drivers, disturbances, and interactions is difficult (but see
+\citep{Nyeetal2013} for an approach using ecosystem-level models to discern
+relative importance of disturbances).
+
+We analyzed an integrodifference model to study the effect of temperature-driven range shifts and
+harvesting, two important interacting disturbances, on the persistence of a marine population. In particular, we extended the model of
+Zhou and Kot \citep{ZhouKot2011}, which describes a fish population
+reproducing and dispersing while their range shifts, by
+additionally including harvesting pressure and introducing two different
+conservation measures. We found that the disturbances interact
+synergistically. We also found that MPAs can help a species persist with
+higher harvesting pressure, but they do not change the maximum speed of
+habitat shift with which a species can keep up.
 
@@
 
-\noindent To make stability easier to analyze, we approximated this kernel, as described in the Appendix.
+\noindent To find analytical results about the Gaussian kernel, we approximated the kernel, as described in the Appendix.
 Analytical results for a separable sinusoidal kernel are also described in the Appendix.
@@
 
-For each kernel, the population's ability to persist depends on properties of the population itself: the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment: the
+For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment-- the
 length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
@@
 \begin{equation}
-h^*\sim1- \frac{1}{R_0}\cdot C(L)f(\sigma^2,c^2,L^2+3c^2)
+h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\sigma^2,c^2,L^2+3c^2)
 \end{equation}
@@
 \citet{ZhouKot2011} only consider whether a shifting environment will drive a population extinct or not.
-To quantify the effects of both a shifting environment and harvesting pressure, we found the total biomass in the equilibrium traveling wave.
-As in \citet{Latore:1998fk}, for a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
+To quantify the effects of both a shifting environment and harvesting pressure, we additionally found the total biomass in the equilibrium traveling wave.
+For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy \citep{Latore:1998fk}
 
@@
 Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
-specify a recruitment function.  We chose to use a Beverton-Holt function,
-
-\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
+specify a recruitment function.  Again, we chose to use a Beverton-Holt function.
 
@@
 
-\[S = E_\text{hc}-\left(E_\text{h}-E_\text{c}\right).\]
+\[S = E_\text{hc}-\left(E_\text{h}+E_\text{c}\right).\]
 
@@
 
-The equilibrium biomass of the population decreases as either the harvesting pressure 
-increases or the environmental shifts more quickly (Figure \ref{baseline}). If the equilibrium biomass is $0$, this 
-indicates that the harvesting pressure has exceeded the critical harvesting rate $h^*$ and the environmental 
-is shifting more quickly than the critical rate of environmental shift $c^*$. As the harvesting rate $h$ increases, 
+As the harvesting rate $h$ increases, 
 the critical rate of environmental shift $c^*$ decreases: the environment must move more slowly to 
@@
 means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
-the environment starts shifting. The simulations replicate the analytical results with the critical speed $c^*$ 
+the environment starts shifting.    When the harvesting pressure has exceeded the critical harvesting rate $h^*$ or the environmental 
+is shifting more quickly than the critical rate of environmental shift $c^*$, the biomass of the population at equilibrium will be equal to $0$.
+Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure
+increases or the environmental shifts more quickly (Figure \ref{baseline}). The simulations replicate the analytical results with the critical speed $c^*$ 
 declining as the critical harvest rate $h^*$ increases and vice versa.
+
+\begin{figure}[htbp]
+\begin{subfigure}{3in}
+\subcaption{\label{rates}}
+\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
+\end{subfigure}
+\begin{subfigure}{3in}
+\subcaption{\label{biomass}}
+\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
+\end{subfigure}
+
+\caption{(a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
+Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
+(b) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+These results are from an approximated Gaussian dispersal kernel with $L=1$.}
+\label{baseline}
+\end{figure}
 
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=3in]{plots/synergy.pdf}
+\caption{Positive synergy between the two stressors.
+The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
+These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
+\label{Synergy}
+\end{center}
+\end{figure}
+
 \subsection{Management Strategies }
@@
 
+\begin{figure}[htbp]
+
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
+\end{subfigure}
+\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+(a) No management.
+(b) Threshold harvesting levels.
+(c) MPAs.
+These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
+}
+\label{management}
+\end{figure}
+
 \section{Discussion}
 
-We extended a spatial to study a fish population subject to both the climate-driven range shifts and harvesting.
+We extended a spatial model to study a fish population subject to both the climate-driven range shifts and harvesting.
 Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.
@@
 The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.
-
-%% discussion of critical rates and curved persistence boundary
 
@@
 Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
-Our results, that spatial management increased the maximum harvesting rate at which the population could survive, corroborate these previous findings.
+Our results, that spatial management increased the maximum harvesting rate and maximum climate velocity at which the population could survive, corroborate these previous findings.
 The management strategies also affect how the two stressors interact with each other.
@@
 
-\section{ Acknowledgements}
-We'd like to thank our reviewers for providing constructive comments.
-
 \bibliography{fish}
-
-\section{Figures}
-
-\begin{figure}[htbp]
-\begin{subfigure}{3in}
-\subcaption{\label{biomass}}
-\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
-\end{subfigure}
-\begin{subfigure}{3in}
-\subcaption{\label{rates}}
-\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
-\end{subfigure}
-%\caption{(a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-(b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
-Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
-Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-These results are from an approximated Gaussian dispersal kernel with $L=1$.}
-\label{baseline}
-\end{figure}
-
-\pagebreak
-
-Figure \ref{baseline}: (a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-(b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
-Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
-Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-These results are from an approximated Gaussian dispersal kernel with $L=1$.
-
-\pagebreak
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=3in]{plots/synergy.pdf}
-%\caption{Positive synergy between the two stressors.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
-\label{Synergy}
-\end{center}
-\end{figure}
-
-\pagebreak
-
-Figure \ref{Synergy}: Positive synergy between the two stressors.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-
-\pagebreak
-
-\begin{figure}[htbp]
-
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
-\end{subfigure}
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
-\end{subfigure}
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
-\end{subfigure}
-%\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
-(a) No management.
-(b) Threshold harvesting levels.
-(c) MPAs.
-These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
-}
-\label{management}
-\end{figure}
-
-\pagebreak
-
-Figure \ref{management}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
-(a) No management.
-(b) Threshold harvesting levels.
-(c) MPAs.
-These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
-
 \section{Appendix}
@@
 
-\end{flushleft}
-\end{spacing}
-
 \end{document}

### 2014-01-29 Writing/MS.tex
cleaned up conflicts

--- 2014-01-28 Writing/MS.tex
@@
 
-For each kernel, the population's ability to persist depends on properties of the population itself: the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment: the
-length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
+For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment-- the length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
 $). If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than 
@@
 \noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
-Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, we must specify a recruitment function to perform simulations.
-We chose to use a Beverton-Holt function,
-
-\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
+Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
+specify a recruitment function.  Again, we chose to use a Beverton-Holt function.
 
@@
 means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
-the environment starts shifting. The simulations replicate the analytical results with the critical speed $c^*$ 
+the environment starts shifting.    When the harvesting pressure has exceeded the critical harvesting rate $h^*$ or the environmental 
+is shifting more quickly than the critical rate of environmental shift $c^*$, the biomass of the population at equilibrium will be equal to $0$.
+Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure
+increases or the environmental shifts more quickly (Figure \ref{baseline}). The simulations replicate the analytical results with the critical speed $c^*$ 
 declining as the critical harvest rate $h^*$ increases and vice versa.
@@
 \begin{subfigure}{3in}
+\subcaption{\label{rates}}
+\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
+\end{subfigure}
+\begin{subfigure}{3in}
 \subcaption{\label{biomass}}
@@
 \end{subfigure}
-\begin{subfigure}{3in}
-\subcaption{\label{rates}}
-\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
-\end{subfigure}
-\caption{
-%(a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-(b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+
+\caption{(a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
 Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
 Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-These results are from an approximated Gaussian dispersal kernel with $L=1$.
-}
+(b) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+These results are from an approximated Gaussian dispersal kernel with $L=1$.}
+
 \label{baseline}
@@
 \caption{
-%Positive synergy between the two stressors.
+Positive synergy between the two stressors.
 The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
@@
 \caption{
-%The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
 (a) No management.
@@
 
-Figure \ref{management}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
-(a) No management.
-(b) Threshold harvesting levels.
-(c) MPAs.
-These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
+\section{Discussion}
+We extended a spatial model to study a fish population subject to both the climate-driven range shifts and harvesting.
+Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.
+In particular, we derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
+For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
+More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
+That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
+The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.
 
-\pagebreak
+To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.
+We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.
+We found similar results from the analytically derived biomass and the simulation derived biomass.
+This indicates that this result is robust to changes in the dispersal kernel.
+We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.
+We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
+In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.
+Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
+
+While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
+synergistic effects were quite small.
+However, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
+Additionally, the fact that we found additive synergy means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
+Synergy between harvesting and climate changes, e.g.
+warming temperature or a more variable climate, has been identified in experiments \citep{Moraetal2007}, in the demography and size of specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk}.
+Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
+
+The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.
+Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
+Our results, that spatial management increased the maximum harvesting rate and maximum climate velocity at which the population could survive, corroborate these previous findings.
+The management strategies also affect how the two stressors interact with each other.
+Specifically, harvesting thresholds can alleviate interactions between the two stressors.
+While the management strategies only change harvesting practices and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement and place protected areas.
+
+The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
+However, it  ignores many of the complexities present in marine fisheries.
+We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
+However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect.
+We also did not include age structure in our model.
+The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work.
+Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
+In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
+By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
+
+Finally, our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
+Indeed, \citet{Perryetal2005} found that fish that shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity).
+Using a simple mechanistic model like the one we presented can incorporate these species interactions and multiple disturbances to understand whether specific life histories are likely to be selected over others as harvesting and/or range shifts increase.
+
+\bibliography{fish}
 

### 2014-01-30 Writing/MS.tex
small typos and grammatical edits

@@
 Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
-However, disturbances rarely occur in isolation and it is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
+However, disturbances rarely occur in isolation, so it is important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
 A perturbation that has little effect when it occurs individually may still amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}.
@@
 Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
-specify a recruitment function.  Again, we chose to use a Beverton-Holt function.
-
-\noindent where $R_0$ is the intrinsic growth rate and $K$ is the carrying capacity.
+specify a recruitment function.
+Again, we chose to use a Beverton-Holt function.
 In the first generation, we seeded the world with 50 individuals at a single point, as in \citep{ZhouKot2011}.

### 2014-01-30 Writing/MS.tex
rearranged order of problems / results so that synergy comes before critical rates

@@
 Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
-However, disturbances rarely occur in isolation, so it is important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
-A perturbation that has little effect when it occurs individually may still amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}.
-Synergistic interactions between stressors can be especially problematic if an animal population can persist in the face of a single stressor but the combination of multiple stressors suffices to drive the population extinct (i.e.
+If disturbances interact synergistically, a perturbation that has little effect when it occurs individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008,Nyeetal2013,Gurevitchetal2000}.
+In the most extreme (and worrying) cases, synergistic interactions will lead multiple stressors to drive a population extinct when it could persist in the face of any single stressor (i.e.
 \citet{Pelletieretal2006}).
-However it remains difficult to predict when disturbances are likely to interact in ways not anticipated when perturbations are measured individually.
+If disturbances interact antagonistically, on the other hand, the effects of multiple stressors may be less than that predicted by any stressor individually.
+Since disturbances rarely occur in isolation, it is important to measure the synergy between  disturbances in order to understand how a system will be affected by their presence and when multiple disturbances will drive a population extinct  \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
 
-Climate change is a prime example of multiple, possibly interacting, disturbances.
-Species that are likely to undergo (or already undergoing) shifts in range and phenology are subject to many other disturbances including pollution, habitat fragmentation, over-exploitation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005}.
-A common approach to predicting future population distrubtions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
+Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008}.
+They therefore present an important case study of how disturbances interact in their effects on biological populations  and understanding these interactions will be crucial to managing populations subjected to these disturbances.
+In response to climate change, marine fish are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and projected to move more in the future \citep{Kelletal2005, Mackenzieetal2007}.
+Species that are likely to undergo or already undergoing shifts in range are also subject harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
+Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in microcosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
+This empirical work underscores the importance of understanding how range shifts and harvesting interact.
+
+A common approach to predicting future population distributions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
 These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to explain and predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
@@
 However these models only address one disturbance: that of climate-driven range shifts.
-Not only does climate change disturb ecological systems via range shifts (for example), but these range shifts may interact with anthropogenetic stressors already present.
-Marine fish, for example, are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and projected to move more  \citep{Kelletal2005, Mackenzieetal2007}.
-These organisms not only are faced with ocean acidification and temperature driven range shifts \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}, but are often harvested commercially.
 
-Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact in the future.
-Predictions for how climate change will drive shifts in marine systems have focused on how physiological limitations \citep{PortnerKnust2007} and global fish distributions \citep{Cheungetal2008, Cheungetal2009, Hareetal2010, Lenoiretal2011} may change.
 Work considering the joint impacts of climate and fishing often consider climate fluctuations (large anomalies around the mean) rather than directional changes in climate \citep{WaltersParma1996, KingMcFarlane2006}.
-Work that does consider the effects of climate-driven range shifts on fishing typically are case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}.
+When the effects of climate-driven range shifts on fishing are considered, the models are typically case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}.
 These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, however these models are so complex that understanding the relative importance of particular drivers, disturbances, and interactions is difficult (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances).
 Here we develop models of intermediate complexity \citep[sensu][]{Gaylordetal2005} that can help clarify the importance of two interacting disturbances: temperature-driven range shifts and harvest.
-Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in micosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
-This empirical work underscores the importance of understanding how range shifts and harvesting interact and, to our knowledge, there has been limited work examining interactions between harvesting and range shifts.
 
@@
 Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We focus on determining how climate-driven range shifts and fishing interact to affect population persistence.
-We find that these disturbances interact additively, not multiplicatively which is often the default assumption when considering mortality.
+We find that climate-driven range shifts and fishing interact synergistically, when their effects are measured by absolute drop in biomass.
+We also find how the threshold level of harvesting that suffices to drive the population extinct depends on quickly the range is shifting.
 
@@
 larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it 
-shifts across space. We provide a list of variables and functions in Table 1.  As we explain below, the functional form of the 
-recruitment function only affects persistence through how quickly recruitment increases when the population 
-size is near (but above) $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$. We therefore do not specify a 
-recruitment function here. Similarly, the population's ability to persist only depends on how quickly harvesting 
-increases when a small number of fish are introduced where they were previously absent, $h'(0)$. We 
-therefore only considered a proportional harvesting function initially.
+shifts across space.
+Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$ so we use such a kernel in our analytical expressions.
+The critical parameter of the recruitment function is how quickly recruitment increases when the population 
+size is near (but above) $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$.
+Since the population's ability to persist only depends on how quickly harvesting increases when a small number of fish are introduced where they were previously absent, $h'(0)$.
+We
+therefore only considered a proportional harvesting function initially.  We provide a list of variables and functions in Table 1.
 
-If a population of fish is going to persist in the face of a shifting environment, it must move along with the patch 
- in which it can reproduce. At equilibrium, the population will be described by a traveling wave, 
+At equilibrium, the population will be described by a traveling wave, 
 where the density of fish at a given point in space will change but the density of fish at a location relative to the 
-shifting  patch will not. We therefore seek to describe how the population is distributed over the viable 
-patch as it shifts through the world. We write this as $n^*(\bar{x})$, the density of fish at each point $\bar{x}$ in 
+shifting  patch will not. We sought to describe how the population is distributed over the viable 
+patch as it shifts through the world in order to study the size of the population at equilibrium and whether or not the population could persist.
+We wrote this as $n^*(\bar{x})$, the density of fish at each point $\bar{x}$ in
 the shifting frame $\left[-\frac{L}{2}, \frac{L}{2}\right]$ As in \citet{ZhouKot2011}, the traveling wave $n^*$ must 
@@
 \end{equation}
-
-One solution to Equation \ref{traveling_pulse} is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
-{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it. If a population becomes very small (or if we introduce a small 
-population), one of two things can happen. First, the population may crash and the trivial traveling pulse 
-without any fish may appear again. Second, those small numbers may increase and form a stable population. 
-In this sense, a small population can be thought of as a perturbation to the trivial traveling pulse. If the trivial 
-pulse is stable, the system will return to the trivial pulse even after a perturbation in the form of the introduction 
-of a small population. If a population is to persist, even when it is small it must be able to avoid extinction and 
-grow. For this to be the case, the trivial pulse must be unstable to small perturbations.
-
-We would like to know the rate of environmental shift and the harvesting rate such that as long as the 
-environment moves more slowly or we harvest less severely than those parameters, then the population will be able to persist.
-We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$.
-We found these rates by finding the parameters that make the trivial pulse unstable.
-Evaluating stability in this kind of model is in  general difficult to do analytically.
-It becomes easier if the dispersal kernel is separable into its dependence on the
-source of larvae and its dependence on the destination of the larvae,
-i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.  In our analyses, we used one such separable kernel, the Gaussian kernel given by
-
-\[k(x-y)=\frac{1}{\sqrt{2\sigma^2\pi}}e^{\frac{-(x-y)^2}{2\sigma^2}}\]
-
-\noindent To find analytical results about the Gaussian kernel, we approximated the kernel, as described in the Appendix.
-Analytical results for a separable sinusoidal kernel are also described in the Appendix.
-We used
-simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
-
-For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment-- the length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
-$). If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than 
-the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.
-For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
-$$R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.$$
-A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
-\begin{equation}
-h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\sigma^2,c^2,L^2+3c^2)
-\end{equation}
-where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
 
@@
 \noindent Equation \ref{recursive_m} allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
-equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.
-Whereas population persistence does not depend on the functional form of recruitment, $f$, equilibrium biomass does depend on what recruitment function we use.
-We chose to use a Beverton-Holt recruitment function,
+equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.  We chose to use a Beverton-Holt recruitment function,
 
 \[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
-
-\subsection{Simulations }
-
-We used simulations to extend the basic integrodifference model in two ways that make it analytically 
-intractable. First, we examined the sensitivity of the model to choice of dispersal kernel by using the Laplace 
-dispersal kernel,
-
-\[ k(x-y)=\frac{1}{2}be^{-b|x-y|}\]
-
-\noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
-Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
-specify a recruitment function.
-Again, we chose to use a Beverton-Holt function.
-In the first generation, we seeded the world with 50 individuals at a single point, as in \citep{ZhouKot2011}.
-We first ran through 150 generations in order for the population to reach equilibrium without harvesting or climate shift.
-We then added harvesting pressure, allowed the population to again reach equilibrium, and finally added climate change by moving the viable patch.
-
-We added harvesting pressure by harvesting a constant proportion of the population, in order to confirm our analytical results.
-We extended our work by evaluating the effect of a threshold harvest rule and marine protected
-areas (MPAs). With a threshold rule, we evaluated the population at each point in space to determine how 
-much harvesting should occur. If the population abundance was below the designated threshold, no 
-harvesting occurred. If the population exceeded the threshold, then a proportion of the `surplus' individuals 
-were harvested
-
-MPAs are a form of management designed to check the impact of fishing on targeted populations and are typically designed to meet either conservation of fishery management goals \citep{Agardy1994, HollandBrazee1996, Gainesetal2010a}.
-To implement an MPA management strategy in our model, we examine the effect of both of these commonly advocated approaches.
-While both conservation and fisheries oriented MPA schemes align in their goal of maintaining a sustainable fished population, they differ in desired level of adult spillover.
-Fisheries-oriented MPAs are often designed such that they maximize adult spillover into fishable areas by creating many small reserves closely spaced \citep{HastingsBotsford2003}.
-The converse of this is the goal of conservation-oriented MPAs which seek to reduce adult spillover by minizing the ratio between the reserve edge length relative to area protected \citep{Gainesetal2010a}.
-
-Networks of MPAs were introduced into our simulations by designating segments of space in which harvesting was forbidden (i.e.
-harvesting rates were equal to $0$).
-Conservation-oriented MPAs, are frequently large and rarely part of a larger network of reserves \citep{HastingsBotsford2003}.
-For solitary reserves to be successful at protecting target species, they must encompass self-sustaining fish populations \citep{HastingsBotsford2006, Gainesetal2010a}.
-As such modeling studies estimate that isolated reserves must be at least as large as the average dispersal distance for the targeted fish species \citep{Lockwoodetal2002, HastingsBotsford2003, Botsfordetal2001, Gainesetal2010}.
-To implement conservation MPAs we created reserves with a length of 4 times the average dispersal distance and had a distance of 8 times the average dispersal distance between them to ensure that populations would be self sustiaining and not dependent on other dispersal for other reserves \citep{Lockwoodetal2002}.
-
-Previous work has shown that if MPAs are to benefit fisheries, the reserves should be broken into a network, closely spaced to maximize adult spillover into fishable areas and export of larvae from reserve to reserve \citep{ HastingsBotsford2003, Gaylordetal2005, Gainesetal2010a}.
-To mimic this management scheme, MPAs had a length of $\tfrac{1}{3}$ of the average dispersal distance and had a distance of $\tfrac{2}{3}$ of the average dispersal distance between them.
-
-Equilibrium is calculated as the mean of 300 time steps once the difference in biomass between time step $t$ and $t+1$ was no greater than $0.1$.
-Equilibrium biomass and catch are defined as the total amount of fish present and harvested, respectively, at equilibrium.
 
@@
 We wanted to understand whether the two stressors interact 
-additively, synergistically, or antagonistically \citep{Crainetal2008} for both our analytical model and simulations. In order to quantify the effect of the 
+additively, synergistically, or antagonistically \citep{Crainetal2008}. In order to quantify the effect of the 
 stressors, we found the total biomass of the population when it reached an equilibrium traveling pulse and 
@@
 
+\subsection{Persistence }
+One possible equilibrium traveling wave that solves Equation \ref{traveling_pulse} is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
+{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.  If a population becomes very small (or if we introduce a small 
+population), one of two things can happen. First, the population may crash and the trivial traveling pulse 
+without any fish may appear again. Second, those small numbers may increase and form a stable population. 
+In this sense, a small population can be thought of as a perturbation to the trivial traveling pulse. If the trivial 
+pulse is stable, the system will return to the trivial pulse even after a perturbation in the form of the introduction 
+of a small population. If a population is to persist, even when it is small it must be able to avoid extinction and 
+grow. For this to be the case, the trivial pulse must be unstable to small perturbations.
+
+We would like to know the rate of environmental shift and the harvesting rate such that as long as the 
+environment moves more slowly or we harvest less severely than those parameters, then the population will be able to persist.
+We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$.
+We found these rates by finding the parameters that make the trivial pulse unstable.   In our analyses, we used a  separable kernel, the Gaussian kernel given by
+
+\[k(x-y)=\frac{1}{\sqrt{2\sigma^2\pi}}e^{\frac{-(x-y)^2}{2\sigma^2}}\]
+
+\noindent To find analytical results about the Gaussian kernel, we approximated the kernel, as described in the Appendix.
+Analytical results for a separable sinusoidal kernel are also described in the Appendix.
+We used
+simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
+
+For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment-- the length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
+$).
+Whereas equilibrium biomass depends on the function form of recruitment, population persistence only depends on the intrinsic growth rate $R_0$.
+If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than
+the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.
+For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
+$$R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.$$
+A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
+\begin{equation}
+h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\sigma^2,c^2,L^2+3c^2)
+\end{equation}
+where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
+
+\subsection{Simulations }
+
+We used simulations to extend the basic integrodifference model in two ways that make it analytically 
+intractable. First, we examined the sensitivity of the model to choice of dispersal kernel by using the Laplace 
+dispersal kernel,
+
+\[ k(x-y)=\frac{1}{2}be^{-b|x-y|}\]
+
+\noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
+Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
+specify a recruitment function.
+Again, we chose to use a Beverton-Holt function.
+In the first generation, we seeded the world with 50 individuals at a single point, as in \citep{ZhouKot2011}.
+We first ran through 150 generations in order for the population to reach equilibrium without harvesting or climate shift.
+We then added harvesting pressure, allowed the population to again reach equilibrium, and finally added climate change by moving the viable patch.
+Equilibrium is calculated as the mean of 300 time steps once the difference in biomass between time step $t$ and $t+1$ was no greater than $0.1$.
+Equilibrium biomass and catch are defined as the total amount of fish present and harvested, respectively, at equilibrium.
+
+We added harvesting pressure by harvesting a constant proportion of the population, in order to confirm our analytical results.
+We extended our work by evaluating the effect of a threshold harvest rule and marine protected
+areas (MPAs). With a threshold rule, we evaluated the population at each point in space to determine how 
+much harvesting should occur. If the population abundance was below the designated threshold, no 
+harvesting occurred. If the population exceeded the threshold, then a proportion of the `surplus' individuals 
+were harvested.
+
+MPAs are a form of management designed to check the impact of fishing on targeted populations and are typically designed to meet either conservation of fishery management goals \citep{Agardy1994, HollandBrazee1996, Gainesetal2010a}.
+To implement an MPA management strategy in our model, we examine the effect of both of these commonly advocated approaches.
+While both conservation and fisheries oriented MPA schemes align in their goal of maintaining a sustainable fished population, they differ in desired level of adult spillover.
+Fisheries-oriented MPAs are often designed such that they maximize adult spillover into fishable areas by creating many small reserves closely spaced \citep{HastingsBotsford2003}.
+The converse of this is the goal of conservation-oriented MPAs which seek to reduce adult spillover by minizing the ratio between the reserve edge length relative to area protected \citep{Gainesetal2010a}.
+
+Networks of MPAs were introduced into our simulations by designating segments of space in which harvesting was forbidden (i.e.
+harvesting rates were equal to $0$).
+Conservation-oriented MPAs, are frequently large and rarely part of a larger network of reserves \citep{HastingsBotsford2003}.
+For solitary reserves to be successful at protecting target species, they must encompass self-sustaining fish populations \citep{HastingsBotsford2006, Gainesetal2010a}.
+As such modeling studies estimate that isolated reserves must be at least as large as the average dispersal distance for the targeted fish species \citep{Lockwoodetal2002, HastingsBotsford2003, Botsfordetal2001, Gainesetal2010}.
+To implement conservation MPAs we created reserves with a length of 4 times the average dispersal distance and had a distance of 8 times the average dispersal distance between them to ensure that populations would be self sustiaining and not dependent on other dispersal for other reserves \citep{Lockwoodetal2002}.
+
+Previous work has shown that if MPAs are to benefit fisheries, the reserves should be broken into a network, closely spaced to maximize adult spillover into fishable areas and export of larvae from reserve to reserve \citep{ HastingsBotsford2003, Gaylordetal2005, Gainesetal2010a}.
+To mimic this management scheme, MPAs had a length of $\tfrac{1}{3}$ of the average dispersal distance and had a distance of $\tfrac{2}{3}$ of the average dispersal distance between them.
+
 \section{Results}
@@
 \subsection{Interactions Between Stressors }
+
+We found positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
+{Synergy}).
+In other words, a doubly stressed population loses more biomass than would be predicted from either stressor individually.
+The stressors interact most strongly when they are both high, shortly before they drive the population extinct.
+We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results.
 
@@
 
-We found positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
-{Synergy}).
-In other words, a doubly stressed population loses more biomass than would be predicted from either stressor individually.
-The stressors interact most strongly when they are both high, shortly before they drive the population extinct.
-We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results.
-
 \subsection{Management Strategies }
@@
 Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.
-In particular, we derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
-For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
-More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
-That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
-The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.
-
-%% discussion of critical rates and curved persistence boundary
-
 To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.
@@
 Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
+
+We also derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
+For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
+More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
+That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
+The curved line separating parameters that will allow the population to persist from those that won't is another indication of an interaction between the stressors.
+
+%% discussion of critical rates and curved persistence boundary
 

### 2014-01-31 Writing/MS.tex
more small edits throughout

@@
 There are many stressors that can disturb an ecosystem.
-Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
+Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}, however less work has been done to measure the effects of multiple stressors and the interactions between them.
 If disturbances interact synergistically, a perturbation that has little effect when it occurs individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008,Nyeetal2013,Gurevitchetal2000}.
-In the most extreme (and worrying) cases, synergistic interactions will lead multiple stressors to drive a population extinct when it could persist in the face of any single stressor (i.e.
+In the most extreme (and worrying) cases, synergistic interactions between multiple stressors will drive a population extinct even though it could persist in the face of any single stressor (i.e.
 \citet{Pelletieretal2006}).
 If disturbances interact antagonistically, on the other hand, the effects of multiple stressors may be less than that predicted by any stressor individually.
-Since disturbances rarely occur in isolation, it is important to measure the synergy between  disturbances in order to understand how a system will be affected by their presence and when multiple disturbances will drive a population extinct  \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
+Since disturbances rarely occur in isolation, it is important to measure the synergy between  disturbances in order to understand how a system will be affected by their presence and to understand when multiple disturbances will drive a population extinct  \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
 
 Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008}.
-They therefore present an important case study of how disturbances interact in their effects on biological populations  and understanding these interactions will be crucial to managing populations subjected to these disturbances.
-In response to climate change, marine fish are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and projected to move more in the future \citep{Kelletal2005, Mackenzieetal2007}.
-Species that are likely to undergo or already undergoing shifts in range are also subject harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
+They therefore present an important case study of how disturbances interact in their effects on biological populations.
+Further, understanding these interactions will be crucial to managing populations subjected to both of these disturbances.
+Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and they are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}.
+Species that are likely to undergo or already undergoing shifts in range are also subject to harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
 Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in microcosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
@@
 This viable habitat patch (here after `patch') is shifting at a fixed velocity and fish at each point in space can be harvested.
-Given the reproductive rate and average dispersal distance, first we determined the climate velocity and harvesting rate that would drive the population extinct.
-We then implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to
+We first measured the drop in biomass caused by range shifts, harvesting, and both stressors together in order to determine whether they interact synergistically.
+We then determined the climate velocity and harvesting rate that would drive the population extinct.
+We finally implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to
 determine how these management strategies affect population persistence.
@@
 larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it 
-shifts across space.
-Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$ so we use such a kernel in our analytical expressions.
-The critical parameter of the recruitment function is how quickly recruitment increases when the population 
+shifts across space.  We chose to use a Beverton-Holt recruitment function,
+
+\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
+
+\noindent Regardless of the exact functional form of the recruitment function, a critical parameter of the recruitment function in determining population persistence is how quickly recruitment increases when the population
 size is near (but above) $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$.
-Since the population's ability to persist only depends on how quickly harvesting increases when a small number of fish are introduced where they were previously absent, $h'(0)$.
-We
-therefore only considered a proportional harvesting function initially.  We provide a list of variables and functions in Table 1.
+Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$ and we use such a kernel in our analytical expressions.
+We provide a list of variables and functions in Table 1.
 
@@
 shifting  patch will not. We sought to describe how the population is distributed over the viable 
-patch as it shifts through the world in order to study the size of the population at equilibrium and whether or not the population could persist.
-We wrote this as $n^*(\bar{x})$, the density of fish at each point $\bar{x}$ in
-the shifting frame $\left[-\frac{L}{2}, \frac{L}{2}\right]$ As in \citet{ZhouKot2011}, the traveling wave $n^*$ must 
+patch as it shifts through the world in order to study the size of the population at equilibrium and whether or not the population could persist.  The traveling wave $n^*$ must 
 satisfy
@@
 
-\citet{ZhouKot2011} only consider whether a shifting environment will drive a population extinct or not.
+\noindent where $\bar{x}\in\left[-\frac{L}{2}, \frac{L}{2}\right]$ describes the position within the patch \citep{ZhouKot2011}.
+\citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct or not.
 To quantify the effects of both a shifting environment and harvesting pressure, we additionally found the total biomass in the equilibrium traveling wave.
-For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy \citep{Latore:1998fk}
+For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
 
@@
 
-\noindent Equation \ref{recursive_m} allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
-equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.  We chose to use a Beverton-Holt recruitment function,
-
-\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
+\noindent \citep{Latore:1998fk}. Equation \ref{recursive_m} allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
+equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.
 
@@
 
-We measured the effects of the stressors by comparing the equilibrium biomass of the stressed population to 
-the equilibrium biomass of the unstressed population. We use $B_0$ to denote the equilibrium biomass 
+We used $B_0$ to denote the equilibrium biomass 
 without either stressor, $B_\text{h}$ the equilibrium biomass with harvesting but a constant environment, $B_\text{c}$ the 

### 2014-01-31 Writing/MS.tex
small edits

@@
 
-When the climate changes, the area in which organisms can survive and reproduce moves through space. 
+When the climate changes, the habitat in which organisms can survive and reproduce moves through space. 
 This change does not occur in isolation but rather appears on a background of other disturbances. We used an 
-integrodifference model explicitly accounting for dispersal, reproduction, a shifting environment, and 
-harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
-persistence.
-We found threshold rates of harvesting and of environmental shift such that more extreme harvesting or shift will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior
-of the population.  We then measured the interaction between the stressors and we found low but positive 
-synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. 
-Finally, we introduced two conservation techniques into simulations of the population model -- threshold 
-harvest rules and marine protected areas (MPAs) -- and found that 
-these approaches could mitigate the negative interaction of the two stressors.  
+integrodifference model explicitly accounting for dispersal, reproduction, to examine how two disturbances, range shift and harvesting, interact and govern population persistence.
+We found threshold rates of harvesting and of environmental shift such that more increasing either disturbance will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior of the population.
+We then measured the interaction between the stressors and we found low but positive synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range.
+Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that these approaches could mitigate, under some circumstances, the negative interaction of the two stressors.
 \hspace{10cm}
@@
 Further, understanding these interactions will be crucial to managing populations subjected to both of these disturbances.
-Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and they are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}.
+Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and they are projected to continue moving in the future \citep{Kelletal2005, Mackenzieetal2007}.
 Species that are likely to undergo or already undergoing shifts in range are also subject to harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
@@
 We find that climate-driven range shifts and fishing interact synergistically, when their effects are measured by absolute drop in biomass.
-We also find how the threshold level of harvesting that suffices to drive the population extinct depends on quickly the range is shifting.
+We also find how the threshold level of harvesting that suffices to drive the population extinct depends on how quickly the range is shifting.
 
@@
 We then determined the climate velocity and harvesting rate that would drive the population extinct.
-We finally implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to
-determine how these management strategies affect population persistence.
+We finally implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to determine how these management strategies affect population persistence.
 
@@
 \noindent where $h$ is the proportion of adults harvested, $f(n)$ is the recruitment function giving the number of 
-offspring produced by a population of size $n$ (accounting for density dependence), $k(x-y)$ is the dispersal kernel giving the probability of a 
-larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it 
-shifts across space.  We chose to use a Beverton-Holt recruitment function,
+offspring produced by a population of size $n$ (accounting for density dependence), $k(x-y)$ is the dispersal kernel giving the probability of a  larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it  shifts across space.
+We chose to use a Beverton-Holt recruitment function,
 
@@
 \noindent where $\bar{x}\in\left[-\frac{L}{2}, \frac{L}{2}\right]$ describes the position within the patch \citep{ZhouKot2011}.
-\citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct or not.
-To quantify the effects of both a shifting environment and harvesting pressure, we additionally found the total biomass in the equilibrium traveling wave.
+\citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct, however to quantify the effects of both a shifting environment and harvesting pressure we must also find the total biomass in the equilibrium traveling wave.
 For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
@@
 
-We wanted to understand whether the two stressors interact 
-additively, synergistically, or antagonistically \citep{Crainetal2008}. In order to quantify the effect of the 
-stressors, we found the total biomass of the population when it reached an equilibrium traveling pulse and 
+In order to quantify whether the two stressors are interacting additively, synergistically, or antagonistically,  we found the total biomass of the population when it reached an equilibrium traveling pulse and
 compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.

### 2014-01-31 Writing/MS.tex
typos

--- 2014-01-31 Writing/MS.tex
@@
 harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
-persistence.
-We found threshold rates of harvesting and of environmental shift such that more extreme harvesting or shift will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior
-of the population.  We then measured the interaction between the stressors and we found low but positive 
+persistence.  We measured the interaction between the stressors and we found low but positive 
 synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range.
+We then found threshold rates of harvesting and of environmental shift such that more extreme harvesting or shift will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior
+of the population.   
 Finally, we introduced two conservation techniques into simulations of the population model -- threshold 
@@
 There are many stressors that can disturb an ecosystem.
-Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}, however less work has been done to measure the effects of multiple stressors and the interactions between them.
+Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}, but less work has been done to measure the effects of multiple stressors and the interactions between them.
 If disturbances interact synergistically, a perturbation that has little effect when it occurs individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008,Nyeetal2013,Gurevitchetal2000}.
@@
 Species that are likely to undergo or already undergoing shifts in range are also subject to harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
-Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in microcosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
+Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and microcosm experiments have demonstrated synergistic effects between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
 This empirical work underscores the importance of understanding how range shifts and harvesting interact.
@@
 
-Here we construct an analytical model of a fish population subject to both harvesting and climate change-induced range-shift.
-We explicitly include two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal.
+Here we constructed an analytical model of a fish population subject to both harvesting and climate change-induced range-shift.
+We explicitly included two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal.
 Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We find that climate-driven range shifts and fishing interact synergistically, when their effects are measured by absolute drop in biomass.
-We also find how the threshold level of harvesting that suffices to drive the population extinct depends on quickly the range is shifting.
+We found that climate-driven range shifts and fishing interact synergistically, when their effects are measured by absolute drop in biomass.
+We also found how the threshold level of harvesting that suffices to drive the population extinct depends on how quickly the range is shifting.
 
-We also examine the effect of marine protected areas (MPAs) on species persistence.
+We also examined the effect of threshold harvesting rules and marine protected areas (MPAs) on species persistence.
 Protected areas have been suggested as a key form of climate insurance and stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
 MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat is shifting rapidly.
-We find that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
+We found that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
 
@@
 We added harvesting pressure by harvesting a constant proportion of the population, in order to confirm our analytical results.
-We extended our work by evaluating the effect of a threshold harvest rule and marine protected
+We then evaluated the effect of a threshold harvest rule and marine protected
 areas (MPAs).  With a threshold rule, we evaluated the population at each point in space to determine how 

### 2014-01-31 Writing/MS.tex

--- 2014-01-31 Writing/MS.tex
@@
 
-Here we construct an analytical model of a fish population subject to both harvesting and climate change-induced range-shift.
-We explicitly include two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal.
+Here we constructed an analytical model of a fish population subject to both harvesting and climate change-induced range-shift.
+We explicitly included two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal.
 Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We find that climate-driven range shifts and fishing interact synergistically, when their effects are measured by absolute drop in biomass.
-We also find how the threshold level of harvesting that suffices to drive the population extinct depends on how quickly the range is shifting.
+We found that climate-driven range shifts and fishing interact synergistically, when their effects are measured by absolute drop in biomass.
+We also found how the threshold level of harvesting that suffices to drive the population extinct depends on how quickly the range is shifting.
 
-We also examine the effect of marine protected areas (MPAs) on species persistence.
+We also examined the effect of threshold harvesting rules and marine protected areas (MPAs) on species persistence.
 Protected areas have been suggested as a key form of climate insurance and stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
 MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat is shifting rapidly.
-We find that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
+We found that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
 

### 2014-01-31 Writing/MS.tex
Closes #15
transferred Malin's comments to issues

@@
 
-\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
+\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}.\]
 
@@
 increases or the environmental shifts more quickly (Figure \ref{baseline}). The simulations replicate the analytical results with the critical speed $c^*$ 
-declining as the critical harvest rate $h^*$ increases and vice versa.
+declining as the critical harvest rate $h^*$ increases and vice versa (Figure \ref{nomang}).
 
@@
 With MPAs in place, the population had a  slightly higher abundance along the edges of the patch where the population is limited by harvesting, which  translated into a slightly increased critical harvest rate (Figure \ref{management}).
-Additionally MPAs increased overall catch at the highest harvest rates under which the population could survive (Figure \ref{management}).
 
@@
 \includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
+\label{nomang}
 \end{subfigure}
@@
 \end{figure}
+
+\pagebreak
+
+The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+(a) No management.
+(b) Threshold harvesting levels.
+(c) MPAs.
+These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
 

### 2014-01-31 Writing/MS.tex
Edits out catch story, close #13

@@
 We then added harvesting pressure, allowed the population to again reach equilibrium, and finally added climate change by moving the viable patch.
-Equilibrium is calculated as the mean of 300 time steps once the difference in biomass between time step $t$ and $t+1$ was no greater than $0.1$.
-Equilibrium biomass and catch are defined as the total amount of fish present and harvested, respectively, at equilibrium.
+Equilibrium biomass is calculated as the mean biomass of 300 time steps once the difference in biomass between time step $t$ and $t+1$ was no greater than $0.1$.
 

### 2014-02-01 Writing/MS.tex
edited equation punctuation, close #17
small edits, close #18
Restructured discussion, needs wording edits, but organization is improved, closes #22

@@
 \begin{equation}
-n_{t+1}(x)=\int^{\frac{L}{2}+ct}_{-\frac{L}{2}+ct}k(x-y)f((1-h)n_t(y))dy \label{integrodifference}
+n_{t+1}(x)=\int^{\frac{L}{2}+ct}_{-\frac{L}{2}+ct}k(x-y)f((1-h)n_t(y))dy \label{integrodifference},
 \end{equation}
@@
 n^*(\bar{x})=\int^{\frac{L}{2}}_{-\frac{L}{2}}k(\bar{x}+c-\bar{y})f((1-h))n^*(\bar{y}))d
-\bar{y} \label{traveling_pulse}
+\bar{y}, \label{traveling_pulse}
 \end{equation}
@@
 
-\[k(x-y)=\frac{1}{\sqrt{2\sigma^2\pi}}e^{\frac{-(x-y)^2}{2\sigma^2}}\]
+\[k(x-y)=\frac{1}{\sqrt{2\sigma^2\pi}}e^{\frac{-(x-y)^2}{2\sigma^2}}.\]
 
@@
 \begin{equation}
-h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\sigma^2,c^2,L^2+3c^2)
+h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\sigma^2,c^2,L^2+3c^2),
 \end{equation}
@@
 
-\[ k(x-y)=\frac{1}{2}be^{-b|x-y|}\]
+\[ k(x-y)=\frac{1}{2}be^{-b|x-y|}.\]
 
@@
 The stressors interact most strongly when they are both high, shortly before they drive the population extinct.
-We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results.
+We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results which indicates that this result is robust to changes in the dispersal kernel.
+We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.
+We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
+In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.
+Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
 
@@
 
-We extended a spatial to study a fish population subject to both the climate-driven range shifts and harvesting.
-Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.
-To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.
-We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.
-We found similar results from the analytically derived biomass and the simulation derived biomass.
-This indicates that this result is robust to changes in the dispersal kernel.
-We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.
-We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
-In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.
-Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
-
-While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
-synergistic effects were quite small.
-However, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
-Additionally, the fact that we found additive synergy means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
+Knowing whether two disturbances are synergistic in their impacts for a given population is important for management.
+Climate change-driven range shifts and fishing form a potential, but largely unexamined, example of synergistic disturbances.
+We found a negative relationship between the critical harvesting rate and the rate of environmental shift.
+That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
+The curved line separating parameters that will allow the population to persist from those that won't is an indication of an interaction between the stressors (Figure \ref{biomass}).
+We find that synergy between these two disturbances is greatest in the region of parameter space where the equilibrium biomass is smallest (Figure \ref{Synergy}).
+While synergistic effects were quite small, the synergy is additive, which means that harvesting levels or climate velocities that are sustainable individually together can drive a population to extinction.
+Further, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome if population persistence is a management objective.
+Although our model is quite general, this suggests that it is marginal populations that are most at risk from the synergistic effects of these two disturbances.
+% the below needs some help, Eleanor can you flesh out with a better discussion of how our results compare to Mora, Planque and Kirby?
 Synergy between harvesting and climate changes, e.g.
 warming temperature or a more variable climate, has been identified in experiments \citep{Moraetal2007}, in the demography and size of specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk}.
-Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
 
-We also derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
-For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
-More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
-That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
-The curved line separating parameters that will allow the population to persist from those that won't is another indication of an interaction between the stressors.
+Our results also suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
+As shown in Figure \ref{baseline}, species with a shorter generation time and longer the average dispersal distance will better track a high rate of environmental shift relative to a species that has a long generation time and short dispersal distance.
+This is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity) \citep{Perryetal2005}.
 
-%% discussion of critical rates and curved persistence boundary
-
-The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.
-Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
-Our results, that spatial management increased the maximum harvesting rate at which the population could survive, corroborate these previous findings.
-The management strategies also affect how the two stressors interact with each other.
-Specifically, harvesting thresholds can alleviate interactions between the two stressors.
-While the management strategies only change harvesting practices and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement and place protected areas.
-
+Previous modeling work examining examining how harvest and range shift intersect has lacked either the generality to examine how these two disturbances are likely to jointly affect population persistence.
+Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, butthe degree of detail and case-specificity makes it difficult to draw general conclusions.
+We built a general model to examine how climate and harvesting interact to affect species persistence.
 The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
@@
 In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
-By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
+By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000} and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
 
-Finally, our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
-Indeed, \citet{Perryetal2005} found that fish that shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity).
-Using a simple mechanistic model like the one we presented can incorporate these species interactions and multiple disturbances to understand whether specific life histories are likely to be selected over others as harvesting and/or range shifts increase.
+We also examined frequently recommended management approaches: harvest thresholds and MPAs and whether these strategies would work to ensure species persistence.
+We found these two management strategies to both increase the population's biomass at equilibrium and improved its ability to persist.
+Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
+Our results, that spatial management increased the maximum harvesting rate at which the population could survive, corroborate these previous findings.
+Harvesting thresholds are already widely implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors.
+While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.
+
+Using a simple mechanistic model like the one we present here provides a useful framework for incorporating additional ecological complexities which can mediate species persistence under multiple disturbances.
+Exploring how species interactions, age structure, and additional disturbances (i.e.
+pollution, disease, physiological response to temperature) will improve our predictions and help us to understand whether specific life histories are likely to be selected over others as harvesting and/or range shifts increase.
 

### 2014-02-01 Writing/MS.tex
small edits

@@
 
-Previous modeling work examining examining how harvest and range shift intersect has lacked either the generality to examine how these two disturbances are likely to jointly affect population persistence.
-Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, butthe degree of detail and case-specificity makes it difficult to draw general conclusions.
+We also examined frequently recommended management approaches: harvest thresholds and MPAs and whether these strategies would work to ensure species persistence.
+We found these two management strategies to both increase the population's biomass at equilibrium and improved its ability to persist.
+Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
+Our results, that spatial management increased the maximum harvesting rate at which the population could survive, corroborate these previous findings.
+Harvesting thresholds are already widely implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors.
+While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.
+
+Previous modeling work examining examining how harvest and range shift intersect has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence.
+Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions.
 We built a general model to examine how climate and harvesting interact to affect species persistence.
@@
 
-We also examined frequently recommended management approaches: harvest thresholds and MPAs and whether these strategies would work to ensure species persistence.
-We found these two management strategies to both increase the population's biomass at equilibrium and improved its ability to persist.
-Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
-Our results, that spatial management increased the maximum harvesting rate at which the population could survive, corroborate these previous findings.
-Harvesting thresholds are already widely implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors.
-While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.
-
 Using a simple mechanistic model like the one we present here provides a useful framework for incorporating additional ecological complexities which can mediate species persistence under multiple disturbances.
 Exploring how species interactions, age structure, and additional disturbances (i.e.
-pollution, disease, physiological response to temperature) will improve our predictions and help us to understand whether specific life histories are likely to be selected over others as harvesting and/or range shifts increase.
+pollution, disease, physiological response to temperature) affect population viability will improve our predictions and help us to understand whether species will persist under predicted climate and harvesting regimes.
+Finally, this work can help make general predictions as to whether specific life histories are likely to be selected over others as harvesting and/or range shifts increase.
 

### 2014-02-01 Writing/MS.tex
Added explanations for Allee effect, close #20

@@
 Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions.
-We built a general model to examine how climate and harvesting interact to affect species persistence.
+Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.
+
 The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
 However, it  ignores many of the complexities present in marine fisheries.
-We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
-However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect.
+Allee effects, for example, have been suggested as an important factor in fish population dynamics.
+By leaving out Allee effects, populations which shrank to very low levels were still able for it to persist over time.
+However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with an Allee effect; i.e.
+rather than extinction occurring when $n(t)=0$, it would be when $n(t)=\text{the Allee threshold}$.
+Thus the absolute values of the critical climate velocity and harvesting rates would change, but their relationship to each other and general shape would remain the same.
+
 We also did not include age structure in our model.

### 2014-02-01 Writing/MS.tex
Adding explanation about age structure, close #21

@@
 We also did not include age structure in our model.
-The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work.
-Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
+The effects of both harvesting and climate change may be different across different age classes which may cause long-term changes in abundance or increased variance due to cohort resonance \citep{Botsfordetal2011, Rouyeretal2012}.
+Both of these effects can be expected to reduce population viability in the face of both fishing and range shifts; including this level of complexity is left for future work.
+Additionally, we did not allow any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
 In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
@@
 
-\section{ Acknowledgements}
+\section*{ Acknowledgements}
 We'd like to thank our reviewers for providing constructive comments.

### 2014-02-01 Writing/MS.tex
Added a few details to MPA section, fixes #19

@@
 We found these two management strategies to both increase the population's biomass at equilibrium and improved its ability to persist.
-Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
-Our results, that spatial management increased the maximum harvesting rate at which the population could survive, corroborate these previous findings.
-Harvesting thresholds are already widely implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors.
+Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
+Our results, that spatial management increased the maximum harvesting rate at which the population could survive, support the idea that MPAs could be used to reduce the impact of harvesting.
+However we did not find any evidence that MPAs increased the climate velocity under which the population could persist.
+The second management approach we investigated, harvesting thresholds, are already widely implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors.
 While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.

### 2014-02-01 Writing/MS.tex
remade table

@@
 Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$ and we use such a kernel in our analytical expressions.
-We provide a list of variables and functions in Table 1.
+We provide a list of variables and functions in Table \ref{variables}.
 
@@
 
-\[k(x-y)=\frac{1}{\sqrt{2\sigma^2\pi}}e^{\frac{-(x-y)^2}{2\sigma^2}}.\]
+\[k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.\]
 
@@
 \begin{equation}
-h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\sigma^2,c^2,L^2+3c^2),
+h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\langle d \rangle,c^2,L^2+3c^2),
 \end{equation}
@@
 
-The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+Figure \ref{management}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
 (a) No management.
@@
 \pagebreak
+
+\section{Tables}
+
+\begin{table}
+\begin{tabular}{@{}lllllll@{}}
+  Variable & Definition
+\\\cmidrule{1-1} \cmidrule{2-2}   
+$n_t(x)$ & density of fish at position $x$ at time $t$
+\\ $n^*(\overline{x})$ & density of fish at equilibrium at position $\overline{x}$ relative to the patch 
+\\ $k(x-y)$ & dispersal kernel, probability of larva traveling from position $y$ to position $x$
+\\ $\langle d \rangle $ & expected distance traveled by larva
+\\ $f(n)$ & recruitment function, number of offspring produced by a population of size $n$
+\\ $R_0$ & intrinsic growth rate, $R_0=f'(0)$
+\\ $h$ & proportion of adults harvested
+\\ $L$ & patch length
+\\ $c$ & rate of environmental shift
+\end{tabular}
+\caption{}
+\label{variables}
+\end{table}
 

### 2014-02-03 Writing/MS.tex
reordered presentation of results in discussion

@@
 Knowing whether two disturbances are synergistic in their impacts for a given population is important for management.
-Climate change-driven range shifts and fishing form a potential, but largely unexamined, example of synergistic disturbances.
-We found a negative relationship between the critical harvesting rate and the rate of environmental shift.
-That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
-The curved line separating parameters that will allow the population to persist from those that won't is an indication of an interaction between the stressors (Figure \ref{biomass}).
-We find that synergy between these two disturbances is greatest in the region of parameter space where the equilibrium biomass is smallest (Figure \ref{Synergy}).
-While synergistic effects were quite small, the synergy is additive, which means that harvesting levels or climate velocities that are sustainable individually together can drive a population to extinction.
-Further, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome if population persistence is a management objective.
-Although our model is quite general, this suggests that it is marginal populations that are most at risk from the synergistic effects of these two disturbances.
-% the below needs some help, Eleanor can you flesh out with a better discussion of how our results compare to Mora, Planque and Kirby?
+The co-occurrence of climate change-driven range shifts and fishing mean that these is the potential for synergistic interactions, which have been largely unexamined.
+Previous modeling work examining examining how harvest and range shift intersect has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence.
+Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions.
+Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.
+
+To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.
+We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.
+We found similar results from the analytically derived biomass and the simulation derived biomass.
+This indicates that this result is robust to changes in the dispersal kernel.
+We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.
+We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
+In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.
+Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
+
+While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
+synergistic effects were quite small.
+However, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
+Additionally, the fact that we found additive synergy means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
 Synergy between harvesting and climate changes, e.g.
 warming temperature or a more variable climate, has been identified in experiments \citep{Moraetal2007}, in the demography and size of specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk}.
+Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
+% the below needs some help, Eleanor can you flesh out with a better discussion of how our results compare to Mora, Planque and Kirby?
 
-Our results also suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
-As shown in Figure \ref{baseline}, species with a shorter generation time and longer the average dispersal distance will better track a high rate of environmental shift relative to a species that has a long generation time and short dispersal distance.
+We also derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
+For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
+More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
+That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
+The curved line separating parameters that will allow the population to persist from those that won't is another indication of an interaction between the stressors.
+
+Our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
+As shown in Figure \ref{baseline}, species with a shorter generation time and a longer average dispersal distance will better track a high rate of environmental shift relative to a species that has a long generation time and short dispersal distance.
 This is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity) \citep{Perryetal2005}.
@@
 
-Previous modeling work examining examining how harvest and range shift intersect has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence.
-Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions.
-Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.
-
 The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
 However, it  ignores many of the complexities present in marine fisheries.
-Allee effects, for example, have been suggested as an important factor in fish population dynamics.
-By leaving out Allee effects, populations which shrank to very low levels were still able for it to persist over time.
-However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with an Allee effect; i.e.
-rather than extinction occurring when $n(t)=0$, it would be when $n(t)=\text{the Allee threshold}$.
-Thus the absolute values of the critical climate velocity and harvesting rates would change, but their relationship to each other and general shape would remain the same.
-
+We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
+However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect.
 We also did not include age structure in our model.
-The effects of both harvesting and climate change may be different across different age classes which may cause long-term changes in abundance or increased variance due to cohort resonance \citep{Botsfordetal2011, Rouyeretal2012}.
-Both of these effects can be expected to reduce population viability in the face of both fishing and range shifts; including this level of complexity is left for future work.
-Additionally, we did not allow any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
+The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work.
+Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
 In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
-By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000} and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
+By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
 

### 2014-02-03 Writing/MS.tex
added Catherine and Will to acknowledgments
last of malins comments

@@
 We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$.
-We found these rates by finding the parameters that make the trivial pulse unstable.   In our analyses, we used a  separable kernel, the Gaussian kernel given by
+We found these rates by finding the parameters that make the trivial pulse unstable.   In our analyses, as in \citep{Latore:1998fk}, we used the separable Gaussian kernel given by
 
@@
 
-\[ k(x-y)=\frac{1}{2}be^{-b|x-y|}.\]
+\[ k(x-y)=\frac{1}{2}be^{-b|x-y|},\]
 
-\noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
+\noindent a commonly used model of larval dispersal \citep{Pinsky:2011fk}.
+Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
 Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
@@
 Knowing whether two disturbances are synergistic in their impacts for a given population is important for management.
-The co-occurrence of climate change-driven range shifts and fishing mean that these is the potential for synergistic interactions, which have been largely unexamined.
-Previous modeling work examining examining how harvest and range shift intersect has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence.
+The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.
+Previous work examining  how harvest and range shift interact has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence.
 Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions.
@@
 Additionally, the fact that we found additive synergy means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
-Synergy between harvesting and climate changes, e.g.
-warming temperature or a more variable climate, has been identified in experiments \citep{Moraetal2007}, in the demography and size of specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk}.
-Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
-% the below needs some help, Eleanor can you flesh out with a better discussion of how our results compare to Mora, Planque and Kirby?
+Synergy between harvesting and climate changes has been identified in experimental populations \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
+However, these results are not directly comparable to ours because they focus on different aspects of climate change, e.g.
+warming temperature \citep{Moraetal2007,Kirby:2009fk} or a more variable climate \citep{Planque:2010uq}.
+Additionally, in the experimental populations, synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
+Nevertheless, our results generally agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
 
@@
 \section*{ Acknowledgements}
-We'd like to thank our reviewers for providing constructive comments.
+We thank Catherine Offord and Will Scott for their contributions to an earlier version of this manuscript and our reviewers for providing constructive comments.
 
@@
 \begin{table}
+\caption{Table of variables used in the text}
 \begin{tabular}{@{}lllllll@{}}
@@
 \end{tabular}
-\caption{}
 \label{variables}
@@
 
-\section{Discussion}
-We extended a spatial model to study a fish population subject to both the climate-driven range shifts and harvesting.
+%\section{Discussion}
+%We extended a spatial model to study a fish population subject to both the climate-driven range shifts and harvesting.
 Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.
@@
 The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.
-
-To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.
+%
+%
+%To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.
 We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.
@@
 Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
-
-While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
+%
+%While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
 synergistic effects were quite small.
@@
 Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
-
-The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.
+%
+%The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.
 Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
@@
 While the management strategies only change harvesting practices and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement and place protected areas.
-
-The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
+%
+%The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
 However, it  ignores many of the complexities present in marine fisheries.
@@
 By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
-
-Finally, our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
+%
+%Finally, our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
 Indeed, \citet{Perryetal2005} found that fish that shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity).

### 2014-02-03 Writing/MS.tex
Tiny edits

--- 2014-02-03 Writing/MS.tex
@@
 
-%\begin{spacing}{1.9}
-%\begin{flushleft}
+\begin{spacing}{1.9}
+\begin{flushleft}
 
@@
 Whereas equilibrium biomass depends on the function form of recruitment, population persistence only depends on the intrinsic growth rate $R_0$.
-If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than
-the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.
+If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.
 For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
-$$R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.$$
+
+\[R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.\]
+
 A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
@@
 \section*{ Acknowledgements}
-We'd like to thank our reviewers for providing constructive comments.
+We thank Catherine Offord and Will Scott for their contributions to an earlier version of this manuscript and our reviewers for providing constructive comments.
+
+\end{flushleft}
+\end{spacing}
 
 \bibliography{fish}
-
+\pagebreak
 \section{Figures}
@@
 
-\caption{(a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+\caption{%(a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
 Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
@@
 These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-These results are from an approximated Gaussian dispersal kernel with $L=1$.}
+These results are from an approximated Gaussian dispersal kernel with $L=1$.
+}
 
@@
 
-Figure \ref{baseline}: (a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+\noindent {\bf Figure \ref{baseline}}: (a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
 These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
@@
 \caption{
-Positive synergy between the two stressors.
+%Positive synergy between the two stressors.
 The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
@@
 
-Figure \ref{Synergy}: Positive synergy between the two stressors.
+\noindent {\bf Figure \ref{Synergy}}: Positive synergy between the two stressors.
 The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
@@
 \caption{
-The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+%The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
 (a) No management.
@@
 
-Figure \ref{management}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+\noindent {\bf Figure \ref{management}}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
 (a) No management.
@@
 \section{Tables}
-
-\begin{table}
+\begin{table}[h]
+\caption{Table of variables used in the text}
 \begin{tabular}{@{}lllllll@{}}
@@
 
-\section{Discussion}
-We extended a spatial model to study a fish population subject to both the climate-driven range shifts and harvesting.
-Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.
-In particular, we derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
-For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
-More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
-That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
-The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.
-
-To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.
-We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.
-We found similar results from the analytically derived biomass and the simulation derived biomass.
-This indicates that this result is robust to changes in the dispersal kernel.
-We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.
-We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
-In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.
-Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
-
-While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
-synergistic effects were quite small.
-However, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
-Additionally, the fact that we found additive synergy means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
-Synergy between harvesting and climate changes, e.g.
-warming temperature or a more variable climate, has been identified in experiments \citep{Moraetal2007}, in the demography and size of specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk}.
-Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
-
-The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.
-Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
-Our results, that spatial management increased the maximum harvesting rate and maximum climate velocity at which the population could survive, corroborate these previous findings.
-The management strategies also affect how the two stressors interact with each other.
-Specifically, harvesting thresholds can alleviate interactions between the two stressors.
-While the management strategies only change harvesting practices and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement and place protected areas.
-
-The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
-However, it  ignores many of the complexities present in marine fisheries.
-We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
-However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect.
-We also did not include age structure in our model.
-The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work.
-Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
-In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
-By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
-
-Finally, our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
-Indeed, \citet{Perryetal2005} found that fish that shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity).
-Using a simple mechanistic model like the one we presented can incorporate these species interactions and multiple disturbances to understand whether specific life histories are likely to be selected over others as harvesting and/or range shifts increase.
-
-\bibliography{fish}
+\pagebreak
 
@@
 
-%\end{flushleft}
-%\end{spacing}
-
 \end{document}

### 2014-02-11 Writing/plots/figureCode/MPA_figs.Rmd
Updated consMPA simulations

--- 2014-01-25 Writing/plots/figureCode/MPA_diffs.Rmd
@@
-# Comparing MPAs to each other, to no management
+# Comparing MPAs to each other, to no management using figure plot code
 
@@
 # for synced online github data
-  x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL25vTVBBbm90VGhyZXNoX2FkZF8yMDEzLTEyLTA5LmNzdiIsImV4cGlyZXMiOjEzOTEyNjg5OTB9--e041d4ef68764b1fc49140c128bda685c6f740f3")
+x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL25vTVBBbm90VGhyZXNoX2FkZF8yMDEzLTEyLTA5LmNzdiIsImV4cGlyZXMiOjEzOTI2NDYxMDl9--46e45a3eb2faef7fc22b6124253c925bd2290de9")
 sim = read.csv(textConnection(x))
 
-  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2Zpc2hNUEFub3RUaHJlc2hfYWRkXzIwMTMtMTItMDkuY3N2IiwiZXhwaXJlcyI6MTM5MTI2OTAzM30%3D--f35f2593a7e901c900188fadca920fdd68a57e35")
+fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2Zpc2hNUEFub3RUaHJlc2hfYWRkXzIwMTMtMTItMDkuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE0NH0%3D--8eff4ee16d08efcf6b266c3e4e42adde7aeff969")
 fishmpa = read.csv(textConnection(fish))
 
-  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2NvbnNNUEFub3RUaHJlc2hfYWRkXzIwMTQtMDEtMjMuY3N2IiwiZXhwaXJlcyI6MTM5MTI2OTA3Mn0%3D--ffb8451d33daaeab48f8b3be3357cacbda0aadbb")
+cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2NvbnNNUEFub3RUaHJlc2hfYWRkXzIwMTQtMDEtMjMuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE2Nn0%3D--586d3cb31ac0f1fe3c49f48886e3796b5132bb77")
 consmpa = read.csv(textConnection(cons))
@@
 ```
+Results of all simulations
+
+```{r results}
+par(mfrow=c(1,3))
+noMang <- plotfig(sim); plotParams(noMang)
+fish <- plotfig(fishmpa); plotParams(fish)
+cons <- plotfig(consmpa); plotParams(cons)
+```
 
@@
 
-Comparing the MPAs to each other
+Comparing the MPAs to each other (cons - fish)
 

### 2014-02-12 Writing/MS.tex
Working on Malin's comments, small edits

@@
 
-Knowing whether two disturbances are synergistic in their impacts for a given population is important for management.
+Knowing whether two disturbances interact in their impacts for a given population is important for management.
 The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.
@@
 
-We also examined frequently recommended management approaches: harvest thresholds and MPAs and whether these strategies would work to ensure species persistence.
-We found these two management strategies to both increase the population's biomass at equilibrium and improved its ability to persist.
+We also examined whether frequently recommended management approaches ensure species persistence.
+We found increases in the population's biomass at equilibrium and an improved ability to persist.
 Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
@@
 The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work.
-Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
-In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
+Similarly, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
+Besides these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
 By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.

### 2014-02-12 Simluations/Aspatial_fast/Functions.R
Re-organizing, streamlining simulation code

@@ -85,3 +85,17 @@ m <- function(n, s, Fthresh = NA, Fharv = NA, mpa.yes = NA, mpa.no = NA, MPA.cur
  	lines(world,n2,lwd=2,col="blue")
  	return(list(n2,harv,MPA))
 }
+
+startOut <- function(w, maxt, mpa.yes,mpa.no,world){
+# initializing the population with no pressure (no harvesting, no climate)
+init<-array(0,c(w,maxt)) 
+init[which(patch==0.55),1]=50
+MPA.start = rep(c(mpa.yes,mpa.no),length.out=length(world))
+for(t in 2:maxt){
+	output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.start)
+	init[,t]= output[[1]]
+	MPA.current = output[[3]]
+	}
+
+return(list(init, MPA.start))
+}

### 2014-02-12 Simluations/Aspatial_fast/MPAcons_SIM.R

@@ -1,115 +1,0 @@
-rm(list=ls())
-library(plyr)
-library(zoo)
-
-# load parameters, functions
-setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
-source("Parameters.R")
-source("Functions.R")
-
-# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
-mpa.yes = cons.yes
-mpa.no = cons.no
-
-# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
-
-move_window = 100
-
-# initializing the population with no pressure (no harvesting, no climate)
-init<-array(0,c(w,maxt)) 
-init[which(patch==0.55),1]=50
-MPA.start = rep(c(mpa.yes,mpa.no),length.out=length(world))
-for(t in 2:maxt){
-	output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.start)
-	init[,t]= output[[1]]
-	MPA.current = output[[3]]
-	}
-
-# standard for equilibrium is the difference in the final step of the initialization 
-init.diff <- diff(colSums(init))[149]
-
-
-# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
-# using while loops. Only saving two time steps 
-
-
-for(q in 1:length(speeds)){
-  
-  for(j in 1:length(harvests)){
-    init.h<-array(0,c(w,1))
-    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.start)    # this resets the MPA across the world, otherwise would pick up where the last simulation left off. 
-    init.h[,1]=next.pop[[1]]
-    MPA.current = next.pop[[3]]
-    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
-    roll.mean = 1	# set to start
-    
-    T = 1
-    
[…]

### 2014-02-12 Simluations/Aspatial_fast/MPAfish_SIM.R

@@ -1,114 +1,0 @@
-rm(list=ls())
-library(plyr)
-library(zoo)
-
-# load parameters, functions
-setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
-source("Parameters.R")
-source("Functions.R")
-
-# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
-mpa.yes = fish.yes
-mpa.no = fish.no
-
-# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
-
-move_window = 100
-
-# initializing the population with no pressure (no harvesting, no climate)
-init<-array(0,c(w,maxt)) 
-init[which(patch==0.55),1]=50
-MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
-for(t in 2:maxt){
-	output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
-	init[,t]= output[[1]]
-	MPA.current = output[[3]]
-	}
-
-# standard for equilibrium is the difference in the final step of the initialization 
-init.diff <- diff(colSums(init))[149]
-
-
-# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
-# using while loops. Only saving two time steps 
-
-
-for(q in 1:length(speeds)){
-  
-  for(j in 1:length(harvests)){
-    init.h<-array(0,c(w,1))
-    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
-    init.h[,1]=next.pop[[1]]
-    MPA.current = next.pop[[3]]
-    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
-    roll.mean = 1	# set to start
-    
-    T = 1
-    
-    while(h.diff > init.diff | roll.mean > init.diff){
-      T = T + 1
[…]

### 2014-02-12 Simluations/Aspatial_fast/Make.R

@@ -1,0 +1,21 @@
+rm(list=ls())
+setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
+
+
+# load parameters, functions
+	source("Functions.R")
+	source("Parameters.R")
+	
+# run analysis
+	# choose threshold or no threshold
+	model = "noThresh"	# "noThresh"; "Thresh"
+	
+	# if no threshold, choose MPA: "null", "cons", "fish"
+	MPA = "fish"
+
+# analysis
+	if(model=="noThresh") {source("Parameters_nothresh.R","Sim_noThresh.R")} else {
+		if(model=="Thresh" & MPA == "null") {source("Parameters_thresh.R","Sim_thresh")} else {
+			warning("model needs to be 'noThresh' or 'Thresh', MPA needs to be 'null'")
+			}
+		}

### 2014-02-12 Simluations/Aspatial_fast/Parameters.R

@@ -7,19 +7,13 @@ b=.5 #parameter for Laplace dispersal kernel
 R0=5 #growth parameter for recruitment
 K=100 #carrying capacity parameter for juvenile density dependence
 maxt = 150 # number of generations to run for initializing of harvest
-# speeds = seq(0,1,.1)
 speeds = seq(0,0.5,0.02)
-harvests = seq(0,0.15,0.02)
 
-# harvests = seq(0,1,.1)
-#thresholds = seq(0,1,.1)
 f_ind = 1 #per capita reproductive rate
 
 patch = seq(0,1,by=step_size)
 world = seq(-.51,1.5, by = step_size)
 w = length(world)
-
-# haven't dealt with MPAs yet !
 
 cons.yes = rep(1,4*b/step_size)
 cons.no = rep(0,8*b/step_size)
@@ -29,10 +23,4 @@ fish.no = rep(0,floor((2/3*b)/step_size))
 null.yes = rep(0,length(world))
 null.no = rep(0, length(world))
 
-# build dataframes
-
-summaries <- data.frame(Equil.pop = rep(NA,length=length(speeds)*length(harvests)), Equil.pop.sd = rep(NA, length=length(speeds)*length(harvests)), Equil.pop.se = rep(NA, length=length(speeds)*length(harvests)), Equil.harvest = rep(NA,length=length(speeds)*length(harvests)),Equil.harvest.sd = rep(NA,length=length(speeds)*length(harvests)), Equil.harvest.se = rep(NA,length=length(speeds)*length(harvests)), speed=rep(NA,length=length(speeds)*length(harvests)), harvest = rep(NA,length=length(speeds)*length(harvests)), thresh=rep(NA,length=length(speeds)*length(harvests)), steps.harv = rep(NA,length=length(speeds)*length(harvests)),steps.move = rep(NA,length=length(speeds)*length(harvests)))
-
-
-# index for row number
-rownumber <- matrix(seq(1:(length(harvests)*length(speeds))),ncol=length(speeds))
+move_window = 100

### 2014-02-12 Simluations/Aspatial_fast/Parameters_nothresh.R

@@ -1,0 +1,20 @@
+# parameters for no-threshold simulations (just proportional harvesting)
+	harvests = seq(0,0.15,0.02)
+
+# build dataframes
+	summaries <- data.frame(
+		Equil.pop = rep(NA,length=length(speeds)*length(harvests)), 
+		Equil.pop.sd = rep(NA, length=length(speeds)*length(harvests)), 
+		Equil.pop.se = rep(NA, length=length(speeds)*length(harvests)), 
+		Equil.harvest = rep(NA,length=length(speeds)*length(harvests)),
+		Equil.harvest.sd = rep(NA,length=length(speeds)*length(harvests)), 
+		Equil.harvest.se = rep(NA,length=length(speeds)*length(harvests)), 
+		speed = rep(NA,length=length(speeds)*length(harvests)), 
+		harvest = rep(NA,length=length(speeds)*length(harvests)), 
+		thresh = rep(NA,length=length(speeds)*length(harvests)), 
+		steps.harv = rep(NA,length=length(speeds)*length(harvests)),
+		steps.move = rep(NA,length=length(speeds)*length(harvests)))
+
+
+# index for row number
+	rownumber <- matrix(seq(1:(length(harvests)*length(speeds))),ncol=length(speeds))

### 2014-02-12 Simluations/Aspatial_fast/Parameters_thresh.R

@@ -1,0 +1,21 @@
+# parameters for threshold simulations (no proportional harvesting)
+	harvests = 1
+	thresholds = seq(0,1,by=0.1)
+
+# build dataframes
+	summaries <- data.frame(
+		Equil.pop = rep(NA,length=length(speeds)*length(thresholds)), 
+		Equil.pop.sd = rep(NA, length=length(speeds)*length(thresholds)), 
+		Equil.pop.se = rep(NA, length=length(speeds)*length(thresholds)), 
+		Equil.harvest = rep(NA,length=length(speeds)*length(thresholds)),
+		Equil.harvest.sd = rep(NA,length=length(speeds)*length(thresholds)), 
+		Equil.harvest.se = rep(NA,length=length(speeds)*length(thresholds)), 
+		speed = rep(NA,length=length(speeds)*length(thresholds)), 
+		harvest = rep(NA,length=length(speeds)*length(thresholds)), 
+		thresh=rep(NA,length=length(speeds)*length(thresholds)), 
+		steps.harv = rep(NA,length=length(speeds)*length(thresholds)),
+		steps.move = rep(NA,length=length(speeds)*length(thresholds)))
+
+
+# index for row number
+	rownumber <- matrix(seq(1:(length(thresholds)*length(speeds))),ncol=length(speeds))

### 2014-02-12 Simluations/Aspatial_fast/Run_sims.R

@@ -1,11 +1,0 @@
-rm(list=setdiff(ls(), "notify"))
-setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
-
-source('noMPA_SIM.R',verbose=TRUE)
-notify('noMPA_SIM finished!')
-
-source('MPAfish_SIM.R', verbose=TRUE)
-notify('MPAfish_SIM finished!')
-
-source('MPAcons_SIM.R', verbose=TRUE)
-notify('MPAcons_SIM finished!')

### 2014-02-12 Simluations/Aspatial_fast/Sim_noThresh.R

@@ -1,43 +1,28 @@
-rm(list=ls())
-library(plyr)
-library(zoo)
-
-# load parameters, functions
-setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
-source("Parameters.R")
-source("Functions.R")
-
-# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
-mpa.yes = null.yes
-mpa.no = null.no
-
 # added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
 
-move_window = 100
-
-# initializing the population with no pressure (no harvesting, no climate)
-init<-array(0,c(w,maxt)) 
-init[which(patch==0.55),1]=50
-MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
-for(t in 2:maxt){
-	output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
-	init[,t]= output[[1]]
-	MPA.current = output[[3]]
+if(MPA=="cons") {mpa.yes=cons.yes; mpa.no=cons.no} else {
+	if(MPA=="fish") {mpa.yes=fish.yes; mpa.no=fish.no} else {
+		if(MPA=="null") {mpa.yes=null.yes; mpa.no=null.no} else{
+			if(exists("MPA")) warning(paste("MPA needs to be 'cons', 'fish', or 'null'.",sep=""))
+		}
 	}
+}
+
+# initialize the world, set MPA (cons.yes/no; fish.yes/no; null.yes/no)
+	output <- startOut(w,maxt,mpa.yes,mpa.no,world)
+	init <- output[[1]]
+	MPA.start <- output[[2]]
 
 # standard for equilibrium is the difference in the final step of the initialization 
-init.diff <- diff(colSums(init))[149]
+	init.diff <- diff(colSums(init))[(maxt-1)]
 
-
-# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
-# using while loops. Only saving two time steps 
-
[…]

### 2014-02-12 Simluations/Aspatial_fast/Sim_thresh.R

@@ -1,45 +1,23 @@
-rm(list=ls())
-library(plyr)
-library(zoo)
-
-# load parameters, functions
-setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
-source("thresh_Parameters.R")
-source("Functions.R")
-
-# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
-mpa.yes = null.yes
-mpa.no = null.no
 
 # added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
 
-move_window = 100
 
-# initializing the population with no pressure (no harvesting, no climate)
-init<-array(0,c(w,maxt)) 
-init[which(patch==0.55),1]=50
-MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
-for(t in 2:maxt){
-  output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
-  init[,t]= output[[1]]
-  MPA.current = output[[3]]
-}
+# initialize the world, set MPA (cons.yes/no; fish.yes/no; null.yes/no)
+	output <- startOut(w,maxt,mpa.yes,mpa.no,world)
+	init <- output[[1]]
+	MPA.start <- output[[2]]
 
 # thresholds need to be relative to maximum population along 1d world
-
-biggestP <- max(init[,maxt])
-
-thresholds <- thresholds * biggestP
+	biggestP <- max(init[,maxt])
+	thresholds <- thresholds * biggestP
 # add a buffer to last to make sure captures biggest population
-tail(thresholds,1) <- tail(thresholds,1) + 0.001
+	tail(thresholds,1) <- tail(thresholds,1) + 0.001
 
 # standard for equilibrium is the difference in the final step of the initialization 
-init.diff <- diff(colSums(init))[149]
-
+	init.diff <- diff(colSums(init))[149]
 
 # Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
 # using while loops. Only saving two time steps 
-
 
 for(q in 1:length(speeds)){
   
[…]

### 2014-02-12 Simluations/Aspatial_fast/noMPA_SIM.R

@@ -1,114 +1,0 @@
-#rm(list=ls())
-library(plyr)
-library(zoo)
-
-# load parameters, functions
-setwd("~/Desktop/MovingFish/Updated_simulations/ToRun/code/Aspatial_fast")
-source("Parameters.R")
-source("Functions.R")
-
-# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
-mpa.yes = null.yes
-mpa.no = null.no
-
-# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
-
-move_window = 100
-
-# initializing the population with no pressure (no harvesting, no climate)
-init<-array(0,c(w,maxt)) 
-init[which(patch==0.55),1]=50
-MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
-for(t in 2:maxt){
-	output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
-	init[,t]= output[[1]]
-	MPA.current = output[[3]]
-	}
-
-# standard for equilibrium is the difference in the final step of the initialization 
-init.diff <- diff(colSums(init))[149]
-
-
-# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
-# using while loops. Only saving two time steps 
-
-
-for(q in 1:length(speeds)){
-  
-  for(j in 1:length(harvests)){
-    init.h<-array(0,c(w,1))
-    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
-    init.h[,1]=next.pop[[1]]
-    MPA.current = next.pop[[3]]
-    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
-    roll.mean = 1	# set to start
-    
-    T = 1
-    
-    while(h.diff > init.diff | roll.mean > init.diff){
-      T = T + 1
[…]

### 2014-02-12 Simluations/Aspatial_fast/thresh_Parameters.R

@@ -1,37 +1,0 @@
-######################################
-## Parameters & Building Structures ##
-######################################
-
-step_size=0.01 #distance between points in space
-b=.5 #parameter for Laplace dispersal kernel
-R0=5 #growth parameter for recruitment
-K=100 #carrying capacity parameter for juvenile density dependence
-maxt = 150 # number of generations to run for initializing of harvest
-# speeds = seq(0,1,.1)
-speeds = seq(0,0.5,0.02)
-harvests = 1
-thresholds = seq(0,1,by=0.1)
-
-f_ind = 1 #per capita reproductive rate
-
-patch = seq(0,1,by=step_size)
-world = seq(-.51,1.5, by = step_size)
-w = length(world)
-
-# haven't dealt with MPAs yet !
-
-cons.yes = rep(1,2*b/step_size)
-cons.no = rep(0,4*b/step_size)
-fish.yes = rep(1,floor((1/3*b)/step_size))  # had to round because not complete step size. Rounded down. 
-fish.no = rep(0,floor((2/3*b)/step_size))
-
-null.yes = rep(0,length(world))
-null.no = rep(0, length(world))
-
-# build dataframes
-
-summaries <- data.frame(Equil.pop = rep(NA,length=length(speeds)*length(thresholds)), Equil.pop.sd = rep(NA, length=length(speeds)*length(thresholds)), Equil.pop.se = rep(NA, length=length(speeds)*length(thresholds)), Equil.harvest = rep(NA,length=length(speeds)*length(thresholds)),Equil.harvest.sd = rep(NA,length=length(speeds)*length(thresholds)), Equil.harvest.se = rep(NA,length=length(speeds)*length(thresholds)), speed=rep(NA,length=length(speeds)*length(thresholds)), harvest = rep(NA,length=length(speeds)*length(thresholds)), thresh=rep(NA,length=length(speeds)*length(thresholds)), steps.harv = rep(NA,length=length(speeds)*length(thresholds)),steps.move = rep(NA,length=length(speeds)*length(thresholds)))
-
-
-# index for row number
-rownumber <- matrix(seq(1:(length(thresholds)*length(speeds))),ncol=length(speeds))

### 2014-02-12 Simluations/Aspatial_fast/figureCode/MPA_diffs.Rmd
Moved figureCode to sit with simulation data

# Comparing catch

```{r required, echo=FALSE,message=FALSE}
library(fields)
library(lattice)
library(colorspace)
require(RCurl)
require(ggplot2)
require(ggthemes)
```

```{r loadData, echo=FALSE, message=FALSE}
#sim=read.csv("~/Desktop/noMPAnotThresh_add_2013-12-09.csv")
# for emma's local copy
  #sim=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv")

# for synced online github data
   x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL25vTVBBbm90VGhyZXNoX2FkZF8yMDEzLTEyLTA5LmNzdiIsImV4cGlyZXMiOjEzOTI2NDYxMDl9--46e45a3eb2faef7fc22b6124253c925bd2290de9")
  sim = read.csv(textConnection(x))

  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2Zpc2hNUEFub3RUaHJlc2hfYWRkXzIwMTMtMTItMDkuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE0NH0%3D--8eff4ee16d08efcf6b266c3e4e42adde7aeff969")
  fishmpa = read.csv(textConnection(fish))

  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2NvbnNNUEFub3RUaHJlc2hfYWRkXzIwMTQtMDEtMjMuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE2Nn0%3D--586d3cb31ac0f1fe3c49f48886e3796b5132bb77")
  consmpa = read.csv(textConnection(cons))

# combine all data into one dataframe
sim$management = rep("noManagement",nrow(sim))
consmpa$management = rep("consMPA", nrow(consmpa))
fishmpa$management = rep("fishMPA", nrow(fishmpa))

allData <- rbind(sim,consmpa,fishmpa)
```

Map equilibrium harvest as function of speed of climate velocity and harvest rate.

```{r Compare_Management,echo = FALSE, fig.width=10, fig.height=10}
contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,3), drape=TRUE)

```

Looks like conservation MPAs do the best, but there are some weird fluctuations at higher harvest rates, why is that? Can take slices to look more closely

```{r variance, echo = FALSE, fig.width=10, fig.height=5}
subData <- subset(allData,harvest == 0.04 )
pd <- position_dodge(.1)
ggplot(subData, aes(x = speed, y = Equil.pop))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()
```

Trying the same plot but a higher rate of catch: $h = 0.10$ I get,

```{r variance_high, echo = FALSE, fig.width=10, fig.height=5}
subData <- subset(allData,harvest == 0.10 )
pd <- position_dodge(.1)
ggplot(subData, aes(x = speed, y = Equil.pop))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()
```

Looking at holding speed constant and examining population as a function of harvest rate

```{r harvestVar, echo = FALSE, fig.width=10, fig.height=5}
subData <- subset(allData,speed==0.1)
ggplot(subData, aes(x = harvest, y = Equil.pop)) + facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()

subData <- subset(allData,speed==0.2)
ggplot(subData, aes(x = harvest, y = Equil.pop)) + facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()

subData <- subset(allData,speed==0.3)
ggplot(subData, aes(x = harvest, y = Equil.pop)) + facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()
```

Need to redo simulations and examine what's going on. Just looking at initialized conservation MPA simulation, see that the whole right side of the MPA has declined due to harvest on the outside.

```{r sim, echo=FALSE,fig.keep='none'}
setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
source("Parameters.R")
source("Functions.R")

# reducing range of speeds and harvests to look over
speeds = speeds[1:15]
harvests = harvests = harvests[6:8]

## copy and pasted from MPAcons_SIM.R
# set MPA (cons.yes/no; fish.yes/no; null.yes/no)
mpa.yes = cons.yes
mpa.no = cons.no

# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim

move_window = 100

# initializing the population with no pressure (no harvesting, no climate)
init<-array(0,c(w,maxt)) 
init[which(patch==0.55),1]=50
MPA.current = rep(c(mpa.yes,mpa.no),length.out=length(world))
for(t in 2:maxt){
  output = m(n=init[,t-1], s = 0, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
	init[,t]= output[[1]]
	MPA.current = output[[3]]
	}
```

```{r plots, echo=FALSE}
plot(world, init[,t],lwd=2,col="blue",bty="n",type='l',ylab="Biomass")
lines(world,MPA.current*9.15,col="grey",lty=2,lwd=2)
legend("bottomleft",legend=c("Fish Biomass","MPA boundary"),col=c("blue","grey"),lty=c(1,2),lwd=2,bty="n")
```

Almost certainly the issue is that for each run of the simulation, the MPA borders are not starting at the same location. So how the population does at the beginning will vary. That's the first thing that needs to be fixed.

It looks like the `MPA.current` variable is not being re-set and is just picking up where it left off for the previous run of the simulation. Have fixed this I think by adding a new variable called `MPA.start` that is the same and resets for each new simulation run (each loop of the for loop). Only changed this on MPAcons right now. Will see if that changes some of the oscillations.

```{r continuingSim,echo=FALSE}
# continuing on with the simulation
# standard for equilibrium is the difference in the final step of the initialization 
init.diff <- diff(colSums(init))[149]

# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
# using while loops. Only saving two time steps

```

### 2014-02-12 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd
Reorganizing things

# Effect of MPAs

```{r required, echo=FALSE,message=FALSE}
library(fields)
library(lattice)
library(colorspace)
require(RCurl)
require(ggplot2)
require(ggthemes)
```
Have noticed that there's a section of parameter space in which the `Equil.pop` varies in the `consMPA` simulations. Why is that?

```{r loadData, echo=FALSE, message=FALSE}
# for synced online github data
   x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxMy0xMi0wOS5jc3YiLCJleHBpcmVzIjoxMzkyODI5NjEzfQ%3D%3D--9fbc7c170a93836981fa63448bf378fbbdb13905")
  sim = read.csv(textConnection(x))

  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxMy0xMi0wOS5jc3YiLCJleHBpcmVzIjoxMzkyODI5NjQzfQ%3D%3D--a432e4025f1b9042246c8a3d6b49af3464a03546")
  fishmpa = read.csv(textConnection(fish))

  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-02-10.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMi0xMC5jc3YiLCJleHBpcmVzIjoxMzkyODI5NjYwfQ%3D%3D--00004e8d0186aa3a6f7f52bb264a406f147d1abf")
  consmpa = read.csv(textConnection(cons))

# combine all data into one dataframe
sim$management = rep("noManagement",nrow(sim))
consmpa$management = rep("consMPA", nrow(consmpa))
fishmpa$management = rep("fishMPA", nrow(fishmpa))

allData <- rbind(sim,consmpa,fishmpa)
```
Map equilibrium harvest as function of speed of climate velocity and harvest rate.

```{r Compare_Management,echo = FALSE, fig.width=10, fig.height=10}
contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,3), drape=TRUE)

```

Looks like conservation MPAs do the best, but there are some weird fluctuations at higher harvest rates, why is that? Can take slices to look more closely

```{r variance, echo = FALSE, fig.width=10, fig.height=5}
subData <- subset(allData,harvest == 0.04 )
pd <- position_dodge(.1)
ggplot(subData, aes(x = speed, y = Equil.pop))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()
```

Trying the same plot but a higher rate of catch: $h = 0.10$ I get,

```{r variance_high, echo = FALSE, fig.width=10, fig.height=5}
subData <- subset(allData,harvest == 0.10 )
pd <- position_dodge(.1)
ggplot(subData, aes(x = speed, y = Equil.pop))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()
```

The relevant section of parameter space is shown below

```{r sub, echo=FALSE,fig.width=15, fig.height=5}
subData <- subset(allData,harvest > 0.05 & harvest < 0.13 & management == "consMPA" & speed < 0.3)
contourplot(Equil.pop ~ speed * harvest, data = subData, region = TRUE, drape = TRUE)
```

```{r sim, echo=FALSE}
setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
source("Parameters.R")
source("Functions.R")
require(zoo)
require(plyr)

# adding in subset of parameters, so instead of running Parameters_nothresh.R, am manually entering
    harvests = seq(0.05,0.13,0.001)
    speeds = seq(0,0.21,.001)

# build dataframes
  summaries <- data.frame(
		Equil.pop = rep(NA,length=length(speeds)*length(harvests)), 
		Equil.pop.sd = rep(NA, length=length(speeds)*length(harvests)), 
		Equil.pop.se = rep(NA, length=length(speeds)*length(harvests)), 
		Equil.harvest = rep(NA,length=length(speeds)*length(harvests)),
		Equil.harvest.sd = rep(NA,length=length(speeds)*length(harvests)), 
		Equil.harvest.se = rep(NA,length=length(speeds)*length(harvests)), 
		speed = rep(NA,length=length(speeds)*length(harvests)), 
		harvest = rep(NA,length=length(speeds)*length(harvests)), 
		thresh = rep(NA,length=length(speeds)*length(harvests)), 
		steps.harv = rep(NA,length=length(speeds)*length(harvests)),
		steps.move = rep(NA,length=length(speeds)*length(harvests)))

# index for row number
	rownumber <- matrix(seq(1:(length(harvests)*length(speeds))),ncol=length(speeds))

MPA = "cons"

source("Sim_noThresh.R")

```

After re-running the model for

### 2014-02-13 Writing/MS.tex
some of malins comments

@@
 \linenumbers
-\title{Species persistence under climate and fishing}
+\title{Persistence of marine populations under climate and fishing}
 \author{Emma Fuller, Eleanor Brush, Malin Pinsky}
@@
 When the climate changes, the habitat in which organisms can survive and reproduce moves through space. 
-This change does not occur in isolation but rather appears on a background of other disturbances. We used an 
-integrodifference model explicitly accounting for dispersal, reproduction, to examine how two disturbances, range shift and harvesting, interact and govern population persistence.
-We found threshold rates of harvesting and of environmental shift such that more increasing either disturbance will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior of the population.
-We then measured the interaction between the stressors and we found low but positive synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range.
+This change does not occur in isolation but rather appears on a background of other disturbances. 
+In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we studied an integrodifference model that explicitly included the mechanisms of dispersal and reproduction.
+If the viable habitat moves too quickly or harvesting pressure is too great, the population is driven extinct.
+We found the rates of harvesting and environmental shift required to allow the population to persist and and studied how these critical parameters depend on the growth rate and dispersal behavior of the population.
+We then measured the interaction between the stressors.
+The stressors interact nearly additively: we found very low positive synergy at those levels of the stressors that almost drive the population extinct.
+Positive synergy suggests that harvesting may aggravate the population's sensitivity to a shifting range.
 Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that these approaches could mitigate, under some circumstances, the negative interaction of the two stressors.
@@
 These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, however these models are so complex that understanding the relative importance of particular drivers, disturbances, and interactions is difficult (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances).
-Here we develop models of intermediate complexity \citep[sensu][]{Gaylordetal2005} that can help clarify the importance of two interacting disturbances: temperature-driven range shifts and harvest.
+The degree of detail and case-specificity in these studies makes it difficult to draw general conclusions.
 
@@
 
-We found positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
+We found very low levels of positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
 {Synergy}).
-In other words, a doubly stressed population loses more biomass than would be predicted from either stressor individually.
+Where there is positive synergy, a doubly stressed population loses more biomass than would be predicted from either stressor individually.
 The stressors interact most strongly when they are both high, shortly before they drive the population extinct.
+However, the excess loss in biomass is extremely low, making it difficult to distinguish positive synergy from additive interactions.
 We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results which indicates that this result is robust to changes in the dispersal kernel.
-We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.
-We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
-In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.
-Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
 
@@
 The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.
-Previous work examining  how harvest and range shift interact has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence.
-Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions.
 Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.
@@
 
-While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
+The fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
+This means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
+However, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
 synergistic effects were quite small.
-However, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
-Additionally, the fact that we found additive synergy means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
 Synergy between harvesting and climate changes has been identified in experimental populations \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
-However, these results are not directly comparable to ours because they focus on different aspects of climate change, e.g.
+Additionally, in the experimental populations, synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
+While we did find (very) low levels of positive synergy, we did not find as much as might be predicted from these empirical studies.
+However, these previous results are not directly comparable to ours because they focus on different aspects of climate change, e.g.
 warming temperature \citep{Moraetal2007,Kirby:2009fk} or a more variable climate \citep{Planque:2010uq}.
-Additionally, in the experimental populations, synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
-Nevertheless, our results generally agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
+Additionally, while we can isolate the affects of climate shift and harvesting in our simple analytical model, there are other forces acting on real populations that may produce the observed synergistic effects.
+
+%e.g.?!?!
+
+%Nevertheless, our results generally agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
 
@@
 \section*{ Acknowledgements}
-We thank Catherine Offord and Will Scott for their contributions to an earlier version of this manuscript and our reviewers for providing constructive comments.
+We thank Catherine Offord and Will Scott for their contributions to an earlier version of this manuscript.
 

### 2014-02-19 Writing/MS.tex

@@
 Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We found that climate-driven range shifts and fishing interact synergistically, when their effects are measured by absolute drop in biomass.
-We also found how the threshold level of harvesting that suffices to drive the population extinct depends on how quickly the range is shifting.
+We found the level of harvesting and the rate of environmental shift that drive the population extinct and how this threshold level of depends on how quickly the range is shifting.
+We also found that climate-driven range shifts and fishing interact nearly additively, with very low positive synergy at more extreme levels of the stressors.
 
@@
 This viable habitat patch (here after `patch') is shifting at a fixed velocity and fish at each point in space can be harvested.
-We first measured the drop in biomass caused by range shifts, harvesting, and both stressors together in order to determine whether they interact synergistically.
-We then determined the climate velocity and harvesting rate that would drive the population extinct.
+We first determined the climate velocity and harvesting rate that would drive the population extinct.
+We then measured the drop in biomass caused by range shifts, harvesting, and both stressors together in order to determine whether they interact synergistically.
 We finally implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to determine how these management strategies affect population persistence.
@@
 \noindent where $\bar{x}\in\left[-\frac{L}{2}, \frac{L}{2}\right]$ describes the position within the patch \citep{ZhouKot2011}.
-\citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct, however to quantify the effects of both a shifting environment and harvesting pressure we must also find the total biomass in the equilibrium traveling wave.
-For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
-
-\begin{equation}
-n^*(x)=\sum^\infty_{i=1}
-a_i(x)\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f((1-H(n^*(y)))n^*(y))dy=\sum^\infty_{i=1}m_ia_i(x), \label{separable_integrodifference}
-\end{equation}
-
-\noindent where the $m_i$ satisfy the recursive equations
-
-\begin{equation}
-m_i=\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f\left((1-h)\sum^\infty_{j=1}m_ja_j(x)\right)
-dy. \label{recursive_m}
-\end{equation}
-
-\noindent \citep{Latore:1998fk}. Equation \ref{recursive_m} allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
-equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.
-
-\subsection{Calculating Synergy }
-
-In order to quantify whether the two stressors are interacting additively, synergistically, or antagonistically, we found the total biomass of the population when it reached an equilibrium traveling pulse and
-compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.
-
-We used $B_0$ to denote the equilibrium biomass 
-without either stressor, $B_\text{h}$ the equilibrium biomass with harvesting but a constant environment, $B_\text{c}$ the 
-equilibrium biomass with a shifting environment but no harvesting, and $B_\text{hc}$ the equilibrium biomass with 
-both stressors. For each stressor or combination of stressors, we found the drop in  biomass caused 
-by stressor $s$,
-
-\[E_\text{s}=B_0-B_\text{s}.\]
-
-\noindent If the stressors do not interact, the drop caused by both stressors would be the sum of the drops caused by 
-either individually. The synergy is therefore defined as
-
-\[S = E_\text{hc}-\left(E_\text{h}+E_\text{c}\right).\]
-
-\noindent If the stressors aggravate each other, the effect of both stressors is worse than would be expected from 
-considering either stressor individually, and synergy is positive. If the stressors alleviate each other, the effect 
-of both stressors is better than would be expected from considering either stressor individually, and synergy is 
-negative. If the effect of both stressors is exactly as expected from considering either stressor individually, 
-there is no interaction and no synergy.
 
@@
 where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
+
+\subsection{Calculating Synergy }
+
+\citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct.
+In order to quantify whether the two stressors are interacting additively, synergistically, or antagonistically, we found the total biomass of the population when it reached an equilibrium traveling pulse and
+compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.
+For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
+
+\begin{equation}
+n^*(x)=\sum^\infty_{i=1}
+a_i(x)\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f((1-H(n^*(y)))n^*(y))dy=\sum^\infty_{i=1}m_ia_i(x), \label{separable_integrodifference}
+\end{equation}
+
+\noindent where the $m_i$ satisfy the recursive equations
+
+\begin{equation}
+m_i=\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f\left((1-h)\sum^\infty_{j=1}m_ja_j(x)\right)
+dy. \label{recursive_m}
+\end{equation}
+
+\noindent \citep{Latore:1998fk}. Equation \ref{recursive_m} allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
+equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.
+
+We used $B_0$ to denote the equilibrium biomass 
+without either stressor, $B_\text{h}$ the equilibrium biomass with harvesting but a constant environment, $B_\text{c}$ the 
+equilibrium biomass with a shifting environment but no harvesting, and $B_\text{hc}$ the equilibrium biomass with 
+both stressors. For each stressor or combination of stressors, we found the drop in  biomass caused 
+by stressor $s$,
+
+\[E_\text{s}=B_0-B_\text{s}.\]
+
+\noindent If the stressors do not interact, the drop caused by both stressors would be the sum of the drops caused by 
+either individually. The synergy is therefore defined as
+
+\[S = E_\text{hc}-\left(E_\text{h}+E_\text{c}\right).\]
+
+\noindent If the stressors aggravate each other, the effect of both stressors is worse than would be expected from 
+considering either stressor individually, and synergy is positive. If the stressors alleviate each other, the effect 
+of both stressors is better than would be expected from considering either stressor individually, and synergy is 
+negative. If the effect of both stressors is exactly as expected from considering either stressor individually, 
+there is no interaction and no synergy.
 
@@
 \subsection{Interactions Between Stressors }
-
-We found very low levels of positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
-{Synergy}).
-Where there is positive synergy, a doubly stressed population loses more biomass than would be predicted from either stressor individually.
-The stressors interact most strongly when they are both high, shortly before they drive the population extinct.
-However, the excess loss in biomass is extremely low, making it difficult to distinguish positive synergy from additive interactions.
-We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results which indicates that this result is robust to changes in the dispersal kernel.
-
 As the harvesting rate $h$ increases, 
@@
 
+We found very low levels of positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
+{Synergy}).
+Where there is positive synergy, a doubly stressed population loses more biomass than would be predicted from either stressor individually.
+The stressors interact most strongly when they are both high, shortly before they drive the population extinct.
+However, the excess loss in biomass is extremely low, making it difficult to distinguish positive synergy from additive interactions.
+We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results which indicates that this result is robust to changes in the dispersal kernel.
+
 \subsection{Management Strategies }
@@
 Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.
+
+We derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
+For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
+More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
+That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
+The curved line separating parameters that will allow the population to persist from those that won't is an indication of an interaction between the stressors.
 
@@
 %Nevertheless, our results generally agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
-
-We also derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
-For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
-More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
-That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
-The curved line separating parameters that will allow the population to persist from those that won't is another indication of an interaction between the stressors.
 

### 2014-02-19 Writing/MS.tex
more of malins comments and cleaning up the appendix

@@
 
-\begin{equation}
+\begin{equation*}
 n_{t+1}(x)=\int^{\frac{L}{2}+ct}_{-\frac{L}{2}+ct}k(x-y)f((1-h)n_t(y))dy \label{integrodifference},
-\end{equation}
+\end{equation*}
 
@@
 \subsection{Persistence }
-One possible equilibrium traveling wave that solves Equation \ref{traveling_pulse} is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
+One possible equilibrium traveling wave that solves Equation (\ref{traveling_pulse}) is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
 {2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.  If a population becomes very small (or if we introduce a small 
@@
 A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
-\begin{equation}
+\begin{equation*}
 h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\langle d \rangle,c^2,L^2+3c^2),
-\end{equation}
+\end{equation*}
 where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
 
-\subsection{Calculating Synergy }
+\subsection{Calculating synergy }
 
@@
 n^*(x)=\sum^\infty_{i=1}
-a_i(x)\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f((1-H(n^*(y)))n^*(y))dy=\sum^\infty_{i=1}m_ia_i(x), \label{separable_integrodifference}
+a_i(x)\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f((1-h)n^*(y))dy=\sum^\infty_{i=1}m_ia_i(x), \label{separable_integrodifference}
 \end{equation}
@@
 
-\noindent \citep{Latore:1998fk}. Equation \ref{recursive_m} allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
-equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.
+\noindent \citep{Latore:1998fk}. Equation (\ref{recursive_m}) allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
+equilibrium traveling pulse by using these $m_i$ and integrating Equation (\ref{separable_integrodifference}).
 
@@
 
-\subsection{Interactions Between Stressors }
+\subsection{Interactions between stressors }
 As the harvesting rate $h$ increases, 
@@
 
-\subsection{Management Strategies }
+\subsection{Management strategies }
 
-We found that when thresholds are in place, the harvesting rate no longer determines the critical rate of environmental shift $c^*$ (Figure \ref{management}).
+Without any management strategies, we found that when the population is harvested more severely, slower rates of environmental shift will suffice to drive the population extinct.
+However, when thresholds are in place, a small population can always escape harvesting pressure and the critical rate of environmental shift $c^*$ no longer depends on the harvesting rate (Figure \ref{management}).
+In other words, as long as there is some threshold below which harvesting is not allowed, there is a constant critical rate of environmental shift that only depends on the growth rate, length of the viable patch, and average dispersal distance.
+
 We also examined the effect of marine protected  areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and  climate change parameters where the fish population could survive.
@@
 The curved line separating parameters that will allow the population to persist from those that won't is an indication of an interaction between the stressors.
+%say more about this line of evidence for an interaction
 
@@
 \\ $n^*(\overline{x})$ & density of fish at equilibrium at position $\overline{x}$ relative to the patch 
-\\ $k(x-y)$ & dispersal kernel, probability of larva traveling from position $y$ to position $x$
+\\ $k(x-y)$ & dispersal kernel, the probability of larva traveling from position $y$ to position $x$
 \\ $\langle d \rangle $ & expected distance traveled by larva
-\\ $f(n)$ & recruitment function, number of offspring produced by a population of size $n$
+\\ $f(n)$ & recruitment function, the number of offspring produced by a population of size $n$
 \\ $R_0$ & intrinsic growth rate, $R_0=f'(0)$
@@
 
+\appendix
 \section{Appendix}
+In Appendix \ref{stab}, we provide the details for assessing the persistence of a population with an integrodifference model and we discuss the effect of the harvesting function on population persistence.
+In Appendix \ref{sep}, we provide the details for assessing population persistence with separable dispersal kernels.
+In Appendix \ref{gausapp} and \ref{sinapp}, we derive expressions for the critical harvesting rate and rate of environmental shift for Gaussian and sinuosoidal dispersal kernels.
+In Appendix \ref{approxcrit}, we derive approximate expressions for these critical rates.
 
-As in Zhou et al. \citep{ZhouKot2011}, let $k(x-y)$ be a dispersal kernel and let $f(y)$ be a recruitment function.  The integrodifference model describing the population over time is given by 
-\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y))dy. \label{integro} \end{equation}
+\subsection{Determining stability \label{stab}}
+As in Zhou et al.
+\citep{ZhouKot2011}, let $k(x-y)$ be a dispersal kernel, let $f(n)$ be a recruitment function, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
+The integrodifference model describing the population over time is given by
+
+\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy. \label{integro} \end{equation}
 
@@
 
-\begin{equation}n^*(\overline{x})\equiv n^*(x-ct)=n_t(x).   \label{trav} \end{equation}
+\begin{equation*}
+n^*(\overline{x})\equiv n^*(x-ct)=n_t(x).   \label{trav} 
+\end{equation*}
 
-Then (\ref{integro}) giveas s us an expression for $n^*$:
+The integrodifference equation (\ref{integro}) gives us an expression for $n^*$:
 
+\begin{align}
+n^*(\overline{x}-c)&=\int_{-L/2}^{L/2}k(\overline{x}-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y})))d\overline y \notag
+\\ \Rightarrow n^*(\overline{x})&=\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y}))d\overline y  \label{pulse}
+\end{align}
+As long as $f(0)=0$, there is a trivial solution to this problem where $n^*(\overline{x})\equiv 0$ for all $\overline{x}\in[-L/2,L/2]$, i.e.
+there is a trivial traveling pulse with no fish in it.
+If the trivial traveling pulse is unstable, even very small populations will persist or grow and avoid crashing back to the trivial pulse.
+To evaluate stability a traveling pulse, we introduce a small perturbation to the traveling pulse $n^*(\overline{x})$ and see if this perturbation grows or shrinks over time:
+
+\begin{align}
+n_t(x)&=n^*(\overline{x})+\xi_t(x) \notag
+\\ \Rightarrow \xi_{t+1}(x)&=n_{t+1}(x)-n^*(\overline{x}) \notag
+\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(f(n_t(y)-g(n_t(y)))-f(n^*(\overline{y})-g(n^*(\overline{y}))\big)dy \text{ using (\ref{pulse})} \notag
+\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(n^*(\overline{y})))f'(n^*(\overline{y})-g(n^*(\overline{y}))\xi_t(y)dy \notag
+\\& \text{ by linearizing around the traveling pulse} \notag
+\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(0))f'(0)\xi_t(y)dy \text{ if $n^*(\overline{x})=0$} \label{perturb}
+\end{align}
+
+If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\R$ and $u:[-L/2,L/2]\to\R$, then the perturbation grows in time if and only if $\lambda >1$.
+Using Equation (\ref{perturb}), we can rewrite $\xi_t(x)$,
 \begin{align*}
-n^*(\overline{x}-c)&=\int_{-L/2}^{L/2}k(\overline{x}-\overline{y})f(n^*(\overline{y}))d\overline y 
-\\ \Rightarrow n^*(\overline{x})&=\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})f(n^*(\overline{y}))d\overline y \tag{*} \label{pulse}
-\end{align*}
-If $f(0)=0$, $n^*(\overline{x})\equiv 0$ for all $\overline{x}\in[-L/2,L/2]$ is a trivial solution to this problem, i.e.
-if there are no fish anywhere there won't be at any time in the future.
-The population can be said to be persistent if the trivial traveling pulse is unstable since even when there are very small population levels, the population won't crash to $0$.
-To evaluate stability (i.e.
-persistence), we will introduce a small perturbation to the traveling pulse $n^*(\overline{x})$,
-
-\begin{align*}
-n_t(x)&=n^*(\overline{x})+\xi_t(x)
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)f'(n^*(\overline{y}))\xi_t(y)dy \text{ by linearizing around the traveling pulse and using (\ref{pulse})}
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)f'(0)\xi_t(y)dy \text{ if we're interested in the stability of the trivial traveling pulse}
-\end{align*}
-If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\R$ and $u:[-L/2,L/2]\to\R$, then 
-\begin{align*}
-\lambda u(x-ct-c)&=f'(0)\int_{-L/2+ct}^{L/2+ct}k(x-y)u(y-ct)dy
-\\ \lambda u(\overline{x})&=f'(0)\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})u(\overline{y})dy
+\lambda u(x-ct-c)&=(1-g'(0))f'(0)\int_{-L/2+ct}^{L/2+ct}k(x-y)u(y-ct)dy
+\\ \\\Rightarrow \lambda u(\overline{x})&=(1-g'(0))f'(0)\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})u(\overline{y})dy
 \end{align*}
@@
 Define the integral operator
-$$ \psi_f(g)(x)=\int_{-L/2}^{L/2}f'(0)k(x+c-y)g(y)dy. $$
-so that the perturbation to the traveling pulse will satisfy 
+$$ \psi_f(u)(x)=(1-g'(0))f'(0)\int_{-L/2}^{L/2}k(x+c-y)u(y)dy. $$
+Then the perturbation to the traveling pulse will satisfy 
 \begin{equation} \psi_f(u)(x)=\lambda u(x) \label{eigen} \end{equation}
-Then the trivial traveling pulse is unstable when the dominant eigenvalue of $\psi_f$ is greater than $1$.
+$\lambda$ and $u$ are thus an eigenvalue and eigenvector of the functional operator $\psi_f$.  The trivial traveling pulse is unstable when the dominant eigenvalue of $\psi_f$ is greater than $1$.
 
-Let $f$ denote the recruitment function, let $h$ denote a harvesting function and let $m(y)=f(y-h(y))$, i.e.
-$m$ denotes the number of offspring after the adults have been harvested.
-Note that $m'(0)=f'(0)(1-h'(0))$, assuming $h(0)=0$ (which must be the case).
+The biomass in the equilibrium traveling wave depends on the specific functional form of the harvesting function $g(n)$.
+However, the persistence of the population only depends on $g'(0)$ so in this paper, we only considered a proportional harvesting function, i.e.
+the amount of fish harvested obeyed $g(n)=hn$.
+For this function, $g'(0)=h$.
 
-Suppose $u$ is an eigenfunction of $\psi_f$ with eigenvalue $\lambda$.  Then 
-\begin{align*}
-\psi_m(u)(x)&=\int_{-L/2}^{L/2}m'(0)k(x+c-y)u(y)dy
-\\&=(1-h'(0))\int_{-L/2}^{L/2}f'(0)k(x+c-y)u(y)dy
-\\&=(1-h'(0))\psi_f(u)(x)
-\\&=(1-h'(0))\lambda u(x)
-\end{align*}
-so that $u$ is also an eigenfuction of $\psi_m$, now with eigenvalue $(1-h'(0))\lambda$.
-
-\subsection{Separable dispersal kernels }
-Jentzsch's theorem shows that there is an eigenfunction $u$, provided that the kernel $k$ satisfy some properties.
-Finding the eigenfunction is in general a hard problem to solve.
+\subsection{Separable dispersal kernels \label{sep}}
+Jentzsch's theorem shows that there is an eigenfunction $u$, provided that the kernel $k$ satisfies some properties \citep{ZhouKot2011}.
+Finding the eigenfunctions and eigenvalues is in general a hard problem to solve.
 It becomes easier if the kernel $k$ is separable, i.e.
@@
 In that case, (\ref{eigen}) becomes
-\begin{align*}
-\lambda u(x)&=f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right)
-\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right)
-\\ \Rightarrow \lambda d_k&=f'(0)\sum_{n=1}^\infty A_{nk}d_n  \tag{**} \label{problem}
-\end{align*}
+\begin{align}
+\lambda u(x)&=f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right) \notag
+\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right) \notag
+\\ \Rightarrow \lambda d_k&=f'(0)\sum_{n=1}^\infty A_{nk}d_n  \label{problem}
+\end{align}
 where
@@
 \end{equation*}
+Finding the eigenvalues of (\ref{eigen}) then reduces to finding the eigenvalues of the matrix $(A_{nk})_{n,k=1}^\infty$.
 
-\subsection{Gaussian dispersal kernel }
+\subsection{Gaussian dispersal kernel \label{gausapp}}
 The Gaussian dispersal kernel is given by
-$$k(|x-y|)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.$$
+$$k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.$$
 As in \citep{Latore:1998fk}, this separable kernel can be written as
-$$k(|x-y|)=\sum_{n=0}^\infty a_n(x)b_n(y)$$
+$$k(x-y)=\sum_{n=0}^\infty a_n(x)b_n(y)$$
 where
@@
 
-As a first approximation to $k$ we ignore all but the $0^{th}$ terms for $a_n$ and $b_n$ so that Equation \ref{problem} becomes
+As a first approximation to $k$ we ignore all but the $0^{th}$ terms for $a_n$ and $b_n$ so that Equation (\ref{problem}) becomes
 \begin{align*}
-\lambda d_0(c)&=f'(0)A_{00}(c)d_0(c)
-\\ \Rightarrow \lambda&=R_0(1-h)A_{00}(c)
+\lambda d_0(c)&=(1-h)f'(0)A_{00}(c)d_0(c)
+\\ \Rightarrow \lambda&=(1-h)R_0A_{00}(c)
 \\\text{ where } A_{00}(c)&=2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]
@@
 
-\subsection{Sinusoidal dispersal kernel }
+\subsection{Sinusoidal dispersal kernel \label{sinapp}}
 A sinusoidal dispersal kernel is given by 
@@
 If we solve for $\lambda$,we find
-\begin{equation} \lambda=R_0(1-h)\left[\frac{wL\cos(wc)}{4}+\frac{1}{4}\sqrt{\sin^2(wL)-w^2L^2\sin^2(wc)}\right]. \label{cosine} \end{equation}
+\begin{equation*} \lambda=(1-h)R_0\left[\frac{wL\cos(wc)}{4}+\frac{1}{4}\sqrt{\sin^2(wL)-w^2L^2\sin^2(wc)}\right]. \label{cosine} \end{equation*}
 
-Zhou et al. \citep{ZhouKot2011} solve for the critical speed, $c^*$, at the population will be driven extinct:
+Zhou et al. \citep{ZhouKot2011} solve for the critical speed, $c^*$, at which the population will be driven extinct:
 $$c^*=c^*(R_0)=\frac{1}{w}\cos^{-1}\left[\frac{16+R_0^2(1-h)^2(w^2L^2-\sin^2(wL))}{8R_0(1-h)wL}\right].$$
@@
 
-\subsection{Approximate Critical Harvesting Proportions}
+\subsection{Approximate critical harvesting proportions \label{approxcrit}}
 ~\\We will use the following Taylor series to make approximations of the critical harvesting proportions under the two dispersal kernels:

### 2014-02-19 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd

@@
 require(ggthemes)
+require(plyr)
+require(reshape2)
 ```
@@
 
-```{r sim, echo=FALSE}
+The next step is to run a few of these simulations and save entire outputs (biomass over space/time). Will do $h = 0.10, s = 0.02-0.07$
+
+```{r sim, echo=FALSE,fig.keep='none',message=FALSE,results='hide'}
 setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
@@
 require(plyr)
+require(reshape2)
+require(lattice)
 
 # adding in subset of parameters, so instead of running Parameters_nothresh.R, am manually entering
-    harvests = seq(0.05,0.13,0.001)
-    speeds = seq(0,0.21,.001)
-
-# build dataframes
-  summaries <- data.frame(
-		Equil.pop = rep(NA,length=length(speeds)*length(harvests)), 
-		Equil.pop.sd = rep(NA, length=length(speeds)*length(harvests)), 
-		Equil.pop.se = rep(NA, length=length(speeds)*length(harvests)), 
-		Equil.harvest = rep(NA,length=length(speeds)*length(harvests)),
-		Equil.harvest.sd = rep(NA,length=length(speeds)*length(harvests)), 
-		Equil.harvest.se = rep(NA,length=length(speeds)*length(harvests)), 
-		speed = rep(NA,length=length(speeds)*length(harvests)), 
-		harvest = rep(NA,length=length(speeds)*length(harvests)), 
-		thresh = rep(NA,length=length(speeds)*length(harvests)), 
-		steps.harv = rep(NA,length=length(speeds)*length(harvests)),
-		steps.move = rep(NA,length=length(speeds)*length(harvests)))
-
-# index for row number
-	rownumber <- matrix(seq(1:(length(harvests)*length(speeds))),ncol=length(speeds))
-
+  harvests = 0.1
+  speeds = seq(0.02,0.1,.01)
   MPA = "cons"
 
-source("Sim_noThresh.R")
+# setting up correct MPA
+  if(MPA=="cons") {mpa.yes=cons.yes; mpa.no=cons.no} else {
+    if(MPA=="fish") {mpa.yes=fish.yes; mpa.no=fish.no} else {
+  		if(MPA=="null") {mpa.yes=null.yes; mpa.no=null.no} else{
+  			if(exists("MPA")) warning(paste("MPA needs to be 'cons', 'fish', or 'null'.",sep=""))
+  		}
+  	}
+  }
 
+# initialize the world, set MPA
+  output <- startOut(w,maxt,mpa.yes,mpa.no,world)
+	init <- output[[1]]
+	MPA.start <- output[[2]]
+
+# standard for equilibrium is the difference in the final step of the initialization 
+  init.diff <- diff(colSums(init))[(maxt-1)]
+
+# pre-allocating list to save simulations
+
+sims <- vector("list", length(speeds))
+times <- vector("list", length(speeds))
+
+for(q in 1:length(speeds)){
+
+  for(j in 1:length(harvests)){
+    init.h<-array(0,c(w,1))
+    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.start)
+    init.h[,1]=next.pop[[1]]
+    MPA.current = next.pop[[3]]
+    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
+    roll.mean = 1  # set to start
+
+    T = 1
+
+    while(h.diff > init.diff | roll.mean > init.diff){
+      T = T + 1
+      next.pop <- m(n=init.h[,T-1], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+      init.h <- cbind(init.h,next.pop[[1]])
+      MPA.current <- next.pop[[3]]
+      h.diff <- abs(sum(init.h[,T]-init.h[,T-1]))
+      if(ncol(init.h) > 100){
+      	rolls <- rollmean(colSums(init.h), k = move_window)
+      	diff.rolls <- diff(rolls)
+		    roll.mean <- diff.rolls[length(diff.rolls)]
+    }
+    }
+    # add in moving climate
+    move <- array(0,c(w,1))
+    next.pop <- m(n=init.h[,T], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+    move[,1] <- next.pop[[1]]
+    MPA.current <- next.pop[[3]]
+    h.diff <- abs(sum(init.h[,T]-move[,1]))
+    roll.mean = 1 # reset
+
+    T = 1
+
+    while(h.diff > init.diff & roll.mean > init.diff){
+      T = T + 1
+      next.pop <- m(n=move[,T-1], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+
+      move <- cbind(move, next.pop[[1]])
+      MPA.current <- next.pop[[3]]
+      h.diff <- abs(sum(move[,T]-move[,T-1]))
+      if(ncol(move) > 300){
+      	rolls <- rollmean(colSums(move), k = move_window)
+      	diff.rolls <- diff(rolls)
+		roll.mean <- diff.rolls[length(diff.rolls)]
+		print(T)
+    }
+    }
+
+      # simulate for 300 steps after equilibrium and take average
+      harv <- vector(mode = 'logical', length = 300)
+      for (i in (T+1):(T+300)){
+        next.pop <- m(n=move[,i-1], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+        move <- cbind(move, next.pop[[1]])
+        MPA.current = next.pop[[3]]
+        harv[(i-T)] <- sum(next.pop[[2]])
+
+      }
+    }
+
+  sims[[q]] = move[,(ncol(move)-299):ncol(move)]
+}
+
+# save the time series of population abundance 
+  series <- llply(sims,colSums)
+  df <- melt(series)
+  names(df) <- c("sums","speed")
+  df$speed=factor(df$speed)
+  df$time <- rep(seq(1,300),9)
+```
+After re-running the model, can see that it replicates the pattern of fluctuating total biomas
+```{r final-plot,echo=FALSE,message=FALSE,results='hide'}
+finals <- llply(series,mean)
+df_finals <- melt(finals)
+names(df_finals) <- c("totalPop","Speed")
+df_finals$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finals,type='l',lwd=2,xlab="Speed")
+```
+And when you look at the time-series population trends, it's as we expected.
+All the populations have the same pattern, that they oscillate up and down depending on whether or not they're in a reserve, but the slowest speeds go through the biggest oscillations which means that the average population is much lower.
+However I wonder if just extending the number of time steps over which I run the average would reduce this effect.
+```{r time-series,echo=FALSE,message=FALSE,results='hide'}
+  xyplot(sums ~ time | factor(speed), data = df, type='l',xlab="time",ylab="Population biomass",lwd=2)
+```
+After re-running the simulation (this time with 600 time steps for equilibrium)
+```{r second-sim,echo=FALSE,fig.keep='none',message=FALSE,results='hide'}
+# pre-allocating list to save simulations
+
+sims2 <- vector("list", length(speeds))
+times2 <- vector("list", length(speeds))
+
+for(q in 1:length(speeds)){
+
+  for(j in 1:length(harvests)){
+    init.h<-array(0,c(w,1))
+    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.start)
+    init.h[,1]=next.pop[[1]]
+    MPA.current = next.pop[[3]]
+    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
+    roll.mean = 1  # set to start
+
+    T = 1
+
+    while(h.diff > init.diff | roll.mean > init.diff){
+      T = T + 1
+      next.pop <- m(n=init.h[,T-1], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+      init.h <- cbind(init.h,next.pop[[1]])
+      MPA.current <- next.pop[[3]]
+      h.diff <- abs(sum(init.h[,T]-init.h[,T-1]))
+      if(ncol(init.h) > 100){
+        rolls <- rollmean(colSums(init.h), k = move_window)
+      	diff.rolls <- diff(rolls)
+		    roll.mean <- diff.rolls[length(diff.rolls)]
+    }
+    }
+    # add in moving climate
+    move <- array(0,c(w,1))
+    next.pop <- m(n=init.h[,T], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+    move[,1] <- next.pop[[1]]
+    MPA.current <- next.pop[[3]]
+    h.diff <- abs(sum(init.h[,T]-move[,1]))
+    roll.mean = 1 # reset
+
+    T = 1
+
+    while(h.diff > init.diff & roll.mean > init.diff){
+      T = T + 1
+      next.pop <- m(n=move[,T-1], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+
+      move <- cbind(move, next.pop[[1]])
+      MPA.current <- next.pop[[3]]
+      h.diff <- abs(sum(move[,T]-move[,T-1]))
+      if(ncol(move) > 300){
+      	rolls <- rollmean(colSums(move), k = move_window)
+      	diff.rolls <- diff(rolls)
+		roll.mean <- diff.rolls[length(diff.rolls)]
+		print(T)
+    }
+    }
+
+      # simulate for 600 steps after equilibrium and take average
+      harv <- vector(mode = 'logical', length = 600)
+      for (i in (T+1):(T+600)){
+        next.pop <- m(n=move[,i-1], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+        move <- cbind(move, next.pop[[1]])
+        MPA.current = next.pop[[3]]
+        harv[(i-T)] <- sum(next.pop[[2]])
+
+      }
+    }
+
+  sims2[[q]] = move[,(ncol(move)-599):ncol(move)]
+}
+
+# save the time series of population abundance 
+  series2 <- llply(sims2,colSums)
+  df2 <- melt(series2)
+  names(df2) <- c("sums","speed")
+  df2$time <- rep(seq(1,600),length(speeds))
 ```
 
-After re-running the model for
+```{r plots,echo=FALSE,message=FALSE,results='hide'}
+  xyplot(sums ~ time | factor(speed), data = df2, type='l',xlab="time",ylab="Population biomass",lwd=2)
+
+finals2 <- llply(series2,mean)
+df_finals2 <- melt(finals2)
+names(df_finals2) <- c("totalPop","Speed")
+df_finals2$Speed = speeds
+
+xyplot(totalPop ~ Speed, data=df_finals2,type='l',lwd=2,xlab="Speed")
+```
+
+This seems to matter a little bit
+
+```{r compare,echo=FALSE,message=FALSE,results='hide'}
+plot(speeds,df_finals$totalPop,pch=19,type="o",main="Mean")
+lines(speeds,df_finals2$totalPop,col="red",type="o",pch=19)
+```
+
+This is weird, shouldn't be like this, not sure why it is. Needs more work. Next step is to try 900 steps to see if the equilibrium varies
+
+```{r thirdSim, echo=FALSE,fig.keep='none'}
+
+sims3 <- vector("list", length(speeds))
+
+for(q in 1:length(speeds)){
+
+  for(j in 1:length(harvests)){
+    init.h<-array(0,c(w,1))
+    next.pop <- m(n=init[,maxt], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.start)
+    init.h[,1]=next.pop[[1]]
+    MPA.current = next.pop[[3]]
+    h.diff <- abs(sum(init[,maxt]-init.h[,1]))
+    roll.mean = 1  # set to start
+
+    T = 1
+
+    while(h.diff > init.diff | roll.mean > init.diff){
+      T = T + 1
+      next.pop <- m(n=init.h[,T-1], s = 0, Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+      init.h <- cbind(init.h,next.pop[[1]])
+      MPA.current <- next.pop[[3]]
+      h.diff <- abs(sum(init.h[,T]-init.h[,T-1]))
+      if(ncol(init.h) > 100){
+        rolls <- rollmean(colSums(init.h), k = move_window)
+        diff.rolls <- diff(rolls)
+		    roll.mean <- diff.rolls[length(diff.rolls)]
+    }
+    }
+    # add in moving climate
+    move <- array(0,c(w,1))
+    next.pop <- m(n=init.h[,T], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+    move[,1] <- next.pop[[1]]
+    MPA.current <- next.pop[[3]]
+    h.diff <- abs(sum(init.h[,T]-move[,1]))
+    roll.mean = 1 # reset
+
+    T = 1
+
+    while(h.diff > init.diff & roll.mean > init.diff){
+      T = T + 1
+      next.pop <- m(n=move[,T-1], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+
+      move <- cbind(move, next.pop[[1]])
+      MPA.current <- next.pop[[3]]
+      h.diff <- abs(sum(move[,T]-move[,T-1]))
+      if(ncol(move) > 300){
+      	rolls <- rollmean(colSums(move), k = move_window)
+      	diff.rolls <- diff(rolls)
+		roll.mean <- diff.rolls[length(diff.rolls)]
+		print(T)
+    }
+    }
+
+      # simulate for 600 steps after equilibrium and take average
+      harv <- vector(mode = 'logical', length = 900)
+      for (i in (T+1):(T+900)){
+        next.pop <- m(n=move[,i-1], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+        move <- cbind(move, next.pop[[1]])
+        MPA.current = next.pop[[3]]
+        harv[(i-T)] <- sum(next.pop[[2]])
+
+      }
+    }
+
+  sims3[[q]] = move[,(ncol(move)-899):ncol(move)]
+}
+
+# save the time series of population abundance 
+  series3 <- llply(sims3,colSums)
+  df3 <- melt(series3)
+  names(df3) <- c("sums","speed")
+  df3$time <- rep(seq(1,900),length(speeds))
+```
+
+```{r plots3,echo=FALSE,message=FALSE,results='hide'}
+  xyplot(sums ~ time | factor(speed), data = df3, type='l',xlab="time",ylab="Population biomass",lwd=2)
+
+finals3 <- llply(series3,mean)
+df_finals3 <- melt(finals3)
+names(df_finals3) <- c("totalPop","Speed")
+df_finals3$Speed = speeds
+
+xyplot(totalPop ~ Speed, data=df_finals3,type='l',lwd=2,xlab="Speed")
+```

### 2014-02-19 Writing/MS.tex
few more small changes

@@
 However, the excess loss in biomass is extremely low, making it difficult to distinguish positive synergy from additive interactions.
-We found similar analytical results for a sinusoidal dispersal kernel and our simulations with a Laplace kernel produce similar results which indicates that this result is robust to changes in the dispersal kernel.
+We found similar analytical results for a sinusoidal dispersal kernel, which indicates that this result is robust to changes in the dispersal kernel.
 
@@
 
-We derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.
 For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
@@
 pollution, disease, physiological response to temperature) affect population viability will improve our predictions and help us to understand whether species will persist under predicted climate and harvesting regimes.
-Finally, this work can help make general predictions as to whether specific life histories are likely to be selected over others as harvesting and/or range shifts increase.
+Finally, this work can help make general predictions as to whether specific life histories are likely to be selected over others as harvesting and range shifts increase.
 

### 2014-02-20 Writing/MS.tex
starting to do a close read through

@@
 
-When the climate changes, the habitat in which organisms can survive and reproduce moves through space. 
+When the climate changes, the areas with conditions in which organisms can survive and reproduce move. 
 This change does not occur in isolation but rather appears on a background of other disturbances. 
@@
 Positive synergy suggests that harvesting may aggravate the population's sensitivity to a shifting range.
-Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that these approaches could mitigate, under some circumstances, the negative interaction of the two stressors.
+Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could mitigate the negative interaction of the two stressors.
 \hspace{10cm}
@@
 If disturbances interact synergistically, a perturbation that has little effect when it occurs individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008,Nyeetal2013,Gurevitchetal2000}.
-In the most extreme (and worrying) cases, synergistic interactions between multiple stressors will drive a population extinct even though it could persist in the face of any single stressor (i.e.
+In the most extreme (and worrying) cases, synergistic interactions between multiple stressors will drive a population extinct even though it could persist in the face of any single stressor (e.g.
 \citet{Pelletieretal2006}).
-If disturbances interact antagonistically, on the other hand, the effects of multiple stressors may be less than that predicted by any stressor individually.
+If disturbances interact antagonistically, on the other hand, the effects of multiple stressors may be less than that predicted by the individual effects of the stressors.
 Since disturbances rarely occur in isolation, it is important to measure the synergy between  disturbances in order to understand how a system will be affected by their presence and to understand when multiple disturbances will drive a population extinct  \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
@@
 Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008}.
-They therefore present an important case study of how disturbances interact in their effects on biological populations.
+They therefore provide an important case study of how disturbances interact in their effects on biological populations.
 Further, understanding these interactions will be crucial to managing populations subjected to both of these disturbances.
@@
 Species that are likely to undergo or already undergoing shifts in range are also subject to harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
-Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in microcosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
+Synergistic interactions between overfishing and temperature-driven range shifts have been found in empirical case studies \citep{Lingetal2009} and synergistic interactions between warming temperatures, harvesting and connectivity have been identified in microcosm experiments\citep{Moraetal2007}.
 This empirical work underscores the importance of understanding how range shifts and harvesting interact.
 
-A common approach to predicting future population distributions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
+A common approach to predicting how populations will be distributed in future after climate-driven range shifts has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
 These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to explain and predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
@@
 Recent work on range shifts has addressed some of these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}.
-However these models only address one disturbance: that of climate-driven range shifts.
+However these models only address one disturbance, climate-driven range shifts.
 
-Work considering the joint impacts of climate and fishing often consider climate fluctuations (large anomalies around the mean) rather than directional changes in climate \citep{WaltersParma1996, KingMcFarlane2006}.
+Work on the joint impacts of climate and fishing often considers climate fluctuations (large anomalies around the mean) rather than directional changes in climate \citep{WaltersParma1996, KingMcFarlane2006}.
 When the effects of climate-driven range shifts on fishing are considered, the models are typically case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}.
@@
 
-Here we constructed an analytical model of a fish population subject to both harvesting and climate change-induced range-shift.
-We explicitly included two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal.
+Here we extended a previously studied model of a fish population subject to climate-driven range shift by also considering harvesting pressure.
+Reproduction and dispersal, two mechanistic processes central to species' responses to climate and fishing, are explicitly included.
 Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We found the level of harvesting and the rate of environmental shift that drive the population extinct and how this threshold level of depends on how quickly the range is shifting.
+We found the rate of harvesting and the rate of environmental shift that drive the population extinct and how the threshold harvesting level of depends on how quickly the range is shifting.
 We also found that climate-driven range shifts and fishing interact nearly additively, with very low positive synergy at more extreme levels of the stressors.
@@
 Using a simple mechanistic model like the one we present here provides a useful framework for incorporating additional ecological complexities which can mediate species persistence under multiple disturbances.
-Exploring how species interactions, age structure, and additional disturbances (i.e.
+Exploring how species interactions, age structure, and additional disturbances (e.g.
 pollution, disease, physiological response to temperature) affect population viability will improve our predictions and help us to understand whether species will persist under predicted climate and harvesting regimes.
@@
 The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
+This excess loss, on the order of $.001$, is small in comparison to the total biomass, which can be as large as $20$.
 These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.

### 2014-02-20 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd
small edits

--- 2014-02-12 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd
@@
 require(plyr)
+require(lattice)
+require(dplyr)
+require(reshape)
+steps = 300   # setting number of steps over which to take equilibrium 
+source("Figures/variable_steps.R")
+sims300 = sims
+df300 = df
+series300 = series
+notify("Simulation is finished!")
+```
+After re-running the model, can see that it replicates the pattern of fluctuating total biomas
+```{r final-plot,echo=FALSE,message=FALSE,results='hide'}
+finals300 <- llply(series300,mean)
+sd300 <- llply(series300,sd)
+sd300 <- melt(sd300)
+df_finals300 <- melt(finals300)
+df_finals300$sd = sd300$value
+names(df_finals300) <- c("totalPop","Speed","sd")
+df_finals300$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finals300,type='l',lwd=2,xlab="Speed", main = "300 steps")
+ggplot(df_finals300,aes(x = Speed, y = totalPop)) + geom_errorbar(aes(ymin=totalPop - sd, ymax = totalPop + sd)) + geom_line() + geom_point()
 
-# adding in subset of parameters, so instead of running Parameters_nothresh.R, am manually entering
-    harvests = seq(0.05,0.13,0.001)
-    speeds = seq(0,0.21,.001)
+```
+And when you look at the time-series population trends, it's as we expected.
+All the populations have the same pattern, that they oscillate up and down depending on whether or not they're in a reserve, but the slowest speeds go through the biggest oscillations which means that the average population is much lower.
+However I wonder if just extending the number of time steps over which I run the average would reduce this effect.
+```{r time-series,echo=FALSE,message=FALSE,results='hide'}
+  xyplot(sums ~ time | speed, data = df300, type='l',xlab="time",ylab="Population biomass",lwd=2, main = "300 steps")
+```
+After re-running the simulation (this time with 600 time steps for equilibrium)
+```{r second-sim,echo=FALSE,fig.keep='none',message=FALSE,results='hide'}
+# pre-allocating list to save simulations
 
-# build dataframes
-  summaries <- data.frame(
-		Equil.pop = rep(NA,length=length(speeds)*length(harvests)), 
-		Equil.pop.sd = rep(NA, length=length(speeds)*length(harvests)), 
-		Equil.pop.se = rep(NA, length=length(speeds)*length(harvests)), 
-		Equil.harvest = rep(NA,length=length(speeds)*length(harvests)),
-		Equil.harvest.sd = rep(NA,length=length(speeds)*length(harvests)), 
-		Equil.harvest.se = rep(NA,length=length(speeds)*length(harvests)), 
-		speed = rep(NA,length=length(speeds)*length(harvests)), 
-		harvest = rep(NA,length=length(speeds)*length(harvests)), 
-		thresh = rep(NA,length=length(speeds)*length(harvests)), 
-		steps.harv = rep(NA,length=length(speeds)*length(harvests)),
-		steps.move = rep(NA,length=length(speeds)*length(harvests)))
+steps = 600   # setting number of steps over which to take equilibrium 
+source("Figures/variable_steps.R")
+sims600 = sims
+df600 = df
+series600 = series
+notify("Simulation is finished!")
+```
 
-# index for row number
-	rownumber <- matrix(seq(1:(length(harvests)*length(speeds))),ncol=length(speeds))
+```{r plots,echo=FALSE,message=FALSE,results='hide'}
+  xyplot(sums ~ time | factor(speed), data = df600, type='l',xlab="time",ylab="Population biomass",lwd=2, main = "600 steps")
 
-MPA = "cons"
+finals600 <- llply(series600,mean)
+df_finals600 <- melt(finals600)
+names(df_finals600) <- c("totalPop","Speed")
+df_finals600$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finals600,type='l',lwd=2,xlab="Speed")
+```
 
-source("Sim_noThresh.R")
+This seems to matter a little bit
+
+```{r compare,echo=FALSE,message=FALSE,results='hide'}
+plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean")
+lines(speeds,df_finals600$totalPop,col="red",type="o",pch=19)
+```
+
+This is weird, shouldn't be like this, not sure why it is. Needs more work. Next step is to try 900 steps to see if the equilibrium varies
+
+```{r thirdSim, echo=FALSE,fig.keep='none'}
+steps = 900   # setting number of steps over which to take equilibrium 
+source("Figures/variable_steps.R")
+sims900 = sims
+df900 = df
+series900 = series
+
+xyplot(sums ~ time | factor(speed), data = df900, type='l',xlab="time",ylab="Population biomass",lwd=2, main="900 steps")
+
+finals900 <- llply(series900,mean)
+df_finals900 <- melt(finals900)
+names(df_finals900) <- c("totalPop","Speed")
+df_finals900$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finals900,type='l',lwd=2,xlab="Speed")
+
+plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean")
+lines(speeds,df_finals600$totalPop,col="red",type="o",pch=19)
+lines(speeds,df_finals900$totalPop,col="blue", type="o",pch=19)
 
@@
 
-After re-running the model for
+```{r fourthSim, echo=FALSE,fig.keep='none'}
+steps = 1200   # setting number of steps over which to take equilibrium 
+source("Figures/variable_steps.R")
+sims1200 = sims
+df1200 = df
+series1200 = series
+
+xyplot(sums ~ time | factor(speed), data = df1200, type='l',xlab="time",ylab="Population biomass",lwd=2, main="1200 steps")
+
+finals1200 <- llply(series1200,mean)
+df_finals1200 <- melt(finals1200)
+names(df_finals1200) <- c("totalPop","Speed")
+df_finals1200$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finals1200,type='l',lwd=2,xlab="Speed")
+
+plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean")
+lines(speeds,df_finals600$totalPop,col="red",type="o",pch=19)
+lines(speeds,df_finals900$totalPop,col="blue", type="o",pch=19)
+lines(speeds,df_finals1200$totalPop,col="green", type="o",pch=19)
+
+notify("simluation is done!")
+```
+
+```{r fifthSim, echo=FALSE,fig.keep='none'}
+steps = 1500   # setting number of steps over which to take equilibrium 
+source("Figures/variable_steps.R")
+sims1500 = sims
+df1500 = df
+series1500 = series
+
+xyplot(sums ~ time | factor(speed), data = df1500, type='l',xlab="time",ylab="Population biomass",lwd=2, main="1500 steps")
+
+finals1500 <- llply(series1500,mean)
+df_finals1500 <- melt(finals1500)
+names(df_finals1500) <- c("totalPop","Speed")
+df_finals1500$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finals1500,type='l',lwd=2,xlab="Speed")
+
+plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean")
+lines(speeds,df_finals600$totalPop,col="red",type="o",pch=19)
+lines(speeds,df_finals900$totalPop,col="blue", type="o",pch=19)
+lines(speeds,df_finals1200$totalPop,col="green", type="o",pch=19)
+lines(speeds,df_finals1500$totalPop,col="purple", type="o",pch=19)
+
+notify("simluation is done!")
+```
+
+```{r sixthSim, echo=FALSE,fig.keep='none'}
+steps = 1800   # setting number of steps over which to take equilibrium 
+source("Figures/variable_steps.R")
+sims1800 = sims
+df1800 = df
+series1800 = series
+
+xyplot(sums ~ time | factor(speed), data = df1800, type='l',xlab="time",ylab="Population biomass",lwd=2, main="1800 steps")
+
+finals1800 <- llply(series1800,mean)
+df_finals1800 <- melt(finals1800)
+names(df_finals1800) <- c("totalPop","Speed")
+df_finals1800$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finals1800,type='l',lwd=2,xlab="Speed")
+
+require(RColorBrewer)
+colz <- colorRampPalette(brewer.pal(6,"Dark2"))(6)
+
+plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean", col=colz[1], lwd=3, bty="n")
+lines(speeds,df_finals600$totalPop,col=colz[2],type="o",pch=19,lwd=3)
+lines(speeds,df_finals900$totalPop,col=colz[3], type="o",pch=19,lwd=3)
+lines(speeds,df_finals1200$totalPop,col=colz[4], type="o",pch=19,lwd=3)
+lines(speeds,df_finals1500$totalPop,col=colz[5], type="o",pch=19,lwd=3)
+lines(speeds,df_finals1800$totalPop,col=colz[6], type="o",pch=19,lwd=3)
+
+legend("topright",legend=seq(300,1800,300), lwd=3, pch=19,col=colz, title="Steps", bty="n")
+
+notify("simluation is done!")
+```
+
+```{r sixthSim, echo=FALSE,fig.keep='none'}
+steps = 2100   # setting number of steps over which to take equilibrium 
+source("Figures/variable_steps.R")
+sims2100 = sims
+df2100 = df
+series2100 = series
+
+xyplot(sums ~ time | factor(speed), data = df2100, type='l',xlab="time",ylab="Population biomass",lwd=2, main="2100 steps")
+
+finals2100 <- llply(series2100,mean)
+df_finals2100 <- melt(finals2100)
+names(df_finals2100) <- c("totalPop","Speed")
+df_finals2100$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finals2100,type='l',lwd=2,xlab="Speed")
+
+require(RColorBrewer)
+colz <- colorRampPalette(brewer.pal(6,"Dark2"))(7)
+
+plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean", col=colz[1], lwd=3, bty="n")
+lines(speeds,df_finals600$totalPop,col=colz[2],type="o",pch=19,lwd=3)
+lines(speeds,df_finals900$totalPop,col=colz[3], type="o",pch=19,lwd=3)
+lines(speeds,df_finals1200$totalPop,col=colz[4], type="o",pch=19,lwd=3)
+lines(speeds,df_finals1500$totalPop,col=colz[5], type="o",pch=19,lwd=3)
+lines(speeds,df_finals1800$totalPop,col=colz[6], type="o",pch=19,lwd=3)
+lines(speeds,df_finals2100$totalPop,col=colz[7], type="o",pch=19,lwd=3)
+
+legend("topright",legend=seq(300,2100,300), lwd=3, pch=19,col=colz, title="Steps", bty="n")
+
+notify("simluation is done!")
+```

### 2014-02-20 Writing/MS.tex
small edits

--- 2014-02-19 Writing/MS.tex
@@
 
-\begin{equation*}
+\begin{equation}
 n_{t+1}(x)=\int^{\frac{L}{2}+ct}_{-\frac{L}{2}+ct}k(x-y)f((1-h)n_t(y))dy \label{integrodifference},
-\end{equation*}
+\end{equation}
 
@@
 \subsection{Persistence }
-One possible equilibrium traveling wave that solves Equation (\ref{traveling_pulse}) is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
+One possible equilibrium traveling wave that solves Equation \ref{traveling_pulse} is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
 {2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.  If a population becomes very small (or if we introduce a small 
@@
 A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
-\begin{equation*}
+\begin{equation}
 h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\langle d \rangle,c^2,L^2+3c^2),
-\end{equation*}
+\end{equation}
 where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
 
-\subsection{Calculating synergy }
+\subsection{Calculating Synergy }
 
@@
 n^*(x)=\sum^\infty_{i=1}
-a_i(x)\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f((1-h)n^*(y))dy=\sum^\infty_{i=1}m_ia_i(x), \label{separable_integrodifference}
+a_i(x)\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f((1-H(n^*(y)))n^*(y))dy=\sum^\infty_{i=1}m_ia_i(x), \label{separable_integrodifference}
 \end{equation}
@@
 
-\noindent \citep{Latore:1998fk}. Equation (\ref{recursive_m}) allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
-equilibrium traveling pulse by using these $m_i$ and integrating Equation (\ref{separable_integrodifference}).
+\noindent \citep{Latore:1998fk}. Equation \ref{recursive_m} allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
+equilibrium traveling pulse by using these $m_i$ and integrating Equation \ref{separable_integrodifference}.
 
@@
 
-\subsection{Interactions between stressors }
-As the harvesting rate $h$ increases, 
-the critical rate of environmental shift $c^*$ decreases: the environment must move more slowly to 
+\subsection{Interactions Between Stressors }
+We find the critical climate velocity and harvest  rate to be inversely related: as the harvesting rate $h$ increases, 
+the critical climate velocity $c^*$ decreases as the environment must move more slowly to 
 accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
@@
 means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
-the environment starts shifting.    When the harvesting pressure has exceeded the critical harvesting rate $h^*$ or the environmental 
-is shifting more quickly than the critical rate of environmental shift $c^*$, the biomass of the population at equilibrium will be equal to $0$.
+the environment starts changing.
+When the climate velocity or harvesting pressure exceed their critical rates ($h^*, c^*$ respectively), the biomass of the population at equilibrium will be equal to $0$.
 Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure
-increases or the environmental shifts more quickly (Figure \ref{baseline}). The simulations replicate the analytical results with the critical speed $c^*$ 
-declining as the critical harvest rate $h^*$ increases and vice versa (Figure \ref{nomang}).
+increases or the environmental shifts more quickly (Figure \ref{baseline}).
+Our simulations confirm the analytical results with the critical speed $c^*$ declining as the critical harvest rate $h^*$ increases and vice versa (Figure \ref{nomang}).
 
@@
 
-\subsection{Management strategies }
+\subsection{Management Strategies }
 
-Without any management strategies, we found that when the population is harvested more severely, slower rates of environmental shift will suffice to drive the population extinct.
-However, when thresholds are in place, a small population can always escape harvesting pressure and the critical rate of environmental shift $c^*$ no longer depends on the harvesting rate (Figure \ref{management}).
-In other words, as long as there is some threshold below which harvesting is not allowed, there is a constant critical rate of environmental shift that only depends on the growth rate, length of the viable patch, and average dispersal distance.
-
+We found that when thresholds are in place, the harvesting rate no longer determines the critical rate of environmental shift $c^*$ (Figure \ref{management}).
 We also examined the effect of marine protected  areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and  climate change parameters where the fish population could survive.
@@
 The curved line separating parameters that will allow the population to persist from those that won't is an indication of an interaction between the stressors.
-%say more about this line of evidence for an interaction
 
@@
 \\ $n^*(\overline{x})$ & density of fish at equilibrium at position $\overline{x}$ relative to the patch 
-\\ $k(x-y)$ & dispersal kernel, the probability of larva traveling from position $y$ to position $x$
+\\ $k(x-y)$ & dispersal kernel, probability of larva traveling from position $y$ to position $x$
 \\ $\langle d \rangle $ & expected distance traveled by larva
-\\ $f(n)$ & recruitment function, the number of offspring produced by a population of size $n$
+\\ $f(n)$ & recruitment function, number of offspring produced by a population of size $n$
 \\ $R_0$ & intrinsic growth rate, $R_0=f'(0)$
@@
 
-\appendix
 \section{Appendix}
-In Appendix \ref{stab}, we provide the details for assessing the persistence of a population with an integrodifference model and we discuss the effect of the harvesting function on population persistence.
-In Appendix \ref{sep}, we provide the details for assessing population persistence with separable dispersal kernels.
-In Appendix \ref{gausapp} and \ref{sinapp}, we derive expressions for the critical harvesting rate and rate of environmental shift for Gaussian and sinuosoidal dispersal kernels.
-In Appendix \ref{approxcrit}, we derive approximate expressions for these critical rates.
 
-\subsection{Determining stability \label{stab}}
-As in Zhou et al.
-\citep{ZhouKot2011}, let $k(x-y)$ be a dispersal kernel, let $f(n)$ be a recruitment function, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
-The integrodifference model describing the population over time is given by
-
-\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy. \label{integro} \end{equation}
+As in Zhou et al. \citep{ZhouKot2011}, let $k(x-y)$ be a dispersal kernel and let $f(y)$ be a recruitment function.  The integrodifference model describing the population over time is given by 
+\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y))dy. \label{integro} \end{equation}
 
@@
 
-\begin{equation*}
-n^*(\overline{x})\equiv n^*(x-ct)=n_t(x).   \label{trav} 
-\end{equation*}
+\begin{equation}n^*(\overline{x})\equiv n^*(x-ct)=n_t(x).   \label{trav} \end{equation}
 
-The integrodifference equation (\ref{integro}) gives us an expression for $n^*$:
+Then (\ref{integro}) giveas s us an expression for $n^*$:
 
-\begin{align}
-n^*(\overline{x}-c)&=\int_{-L/2}^{L/2}k(\overline{x}-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y})))d\overline y \notag
-\\ \Rightarrow n^*(\overline{x})&=\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y}))d\overline y  \label{pulse}
-\end{align}
-As long as $f(0)=0$, there is a trivial solution to this problem where $n^*(\overline{x})\equiv 0$ for all $\overline{x}\in[-L/2,L/2]$, i.e.
-there is a trivial traveling pulse with no fish in it.
-If the trivial traveling pulse is unstable, even very small populations will persist or grow and avoid crashing back to the trivial pulse.
-To evaluate stability a traveling pulse, we introduce a small perturbation to the traveling pulse $n^*(\overline{x})$ and see if this perturbation grows or shrinks over time:
+\begin{align*}
+n^*(\overline{x}-c)&=\int_{-L/2}^{L/2}k(\overline{x}-\overline{y})f(n^*(\overline{y}))d\overline y 
+\\ \Rightarrow n^*(\overline{x})&=\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})f(n^*(\overline{y}))d\overline y \tag{*} \label{pulse}
+\end{align*}
+If $f(0)=0$, $n^*(\overline{x})\equiv 0$ for all $\overline{x}\in[-L/2,L/2]$ is a trivial solution to this problem, i.e.
+if there are no fish anywhere there won't be at any time in the future.
+The population can be said to be persistent if the trivial traveling pulse is unstable since even when there are very small population levels, the population won't crash to $0$.
+To evaluate stability (i.e.
+persistence), we will introduce a small perturbation to the traveling pulse $n^*(\overline{x})$,
 
-\begin{align}
-n_t(x)&=n^*(\overline{x})+\xi_t(x) \notag
-\\ \Rightarrow \xi_{t+1}(x)&=n_{t+1}(x)-n^*(\overline{x}) \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(f(n_t(y)-g(n_t(y)))-f(n^*(\overline{y})-g(n^*(\overline{y}))\big)dy \text{ using (\ref{pulse})} \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(n^*(\overline{y})))f'(n^*(\overline{y})-g(n^*(\overline{y}))\xi_t(y)dy \notag
-\\& \text{ by linearizing around the traveling pulse} \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(0))f'(0)\xi_t(y)dy \text{ if $n^*(\overline{x})=0$} \label{perturb}
-\end{align}
-
-If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\R$ and $u:[-L/2,L/2]\to\R$, then the perturbation grows in time if and only if $\lambda >1$.
-Using Equation (\ref{perturb}), we can rewrite $\xi_t(x)$,
 \begin{align*}
-\lambda u(x-ct-c)&=(1-g'(0))f'(0)\int_{-L/2+ct}^{L/2+ct}k(x-y)u(y-ct)dy
-\\ \\\Rightarrow \lambda u(\overline{x})&=(1-g'(0))f'(0)\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})u(\overline{y})dy
+n_t(x)&=n^*(\overline{x})+\xi_t(x)
+\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)f'(n^*(\overline{y}))\xi_t(y)dy \text{ by linearizing around the traveling pulse and using (\ref{pulse})}
+\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)f'(0)\xi_t(y)dy \text{ if we're interested in the stability of the trivial traveling pulse}
+\end{align*}
+If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\R$ and $u:[-L/2,L/2]\to\R$, then 
+\begin{align*}
+\lambda u(x-ct-c)&=f'(0)\int_{-L/2+ct}^{L/2+ct}k(x-y)u(y-ct)dy
+\\ \lambda u(\overline{x})&=f'(0)\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})u(\overline{y})dy
 \end{align*}
@@
 Define the integral operator
-$$ \psi_f(u)(x)=(1-g'(0))f'(0)\int_{-L/2}^{L/2}k(x+c-y)u(y)dy. $$
-Then the perturbation to the traveling pulse will satisfy 
+$$ \psi_f(g)(x)=\int_{-L/2}^{L/2}f'(0)k(x+c-y)g(y)dy. $$
+so that the perturbation to the traveling pulse will satisfy 
 \begin{equation} \psi_f(u)(x)=\lambda u(x) \label{eigen} \end{equation}
-$\lambda$ and $u$ are thus an eigenvalue and eigenvector of the functional operator $\psi_f$.  The trivial traveling pulse is unstable when the dominant eigenvalue of $\psi_f$ is greater than $1$.
+Then the trivial traveling pulse is unstable when the dominant eigenvalue of $\psi_f$ is greater than $1$.
 
-The biomass in the equilibrium traveling wave depends on the specific functional form of the harvesting function $g(n)$.
-However, the persistence of the population only depends on $g'(0)$ so in this paper, we only considered a proportional harvesting function, i.e.
-the amount of fish harvested obeyed $g(n)=hn$.
-For this function, $g'(0)=h$.
+Let $f$ denote the recruitment function, let $h$ denote a harvesting function and let $m(y)=f(y-h(y))$, i.e.
+$m$ denotes the number of offspring after the adults have been harvested.
+Note that $m'(0)=f'(0)(1-h'(0))$, assuming $h(0)=0$ (which must be the case).
 
-\subsection{Separable dispersal kernels \label{sep}}
-Jentzsch's theorem shows that there is an eigenfunction $u$, provided that the kernel $k$ satisfies some properties \citep{ZhouKot2011}.
-Finding the eigenfunctions and eigenvalues is in general a hard problem to solve.
+Suppose $u$ is an eigenfunction of $\psi_f$ with eigenvalue $\lambda$.  Then 
+\begin{align*}
+\psi_m(u)(x)&=\int_{-L/2}^{L/2}m'(0)k(x+c-y)u(y)dy
+\\&=(1-h'(0))\int_{-L/2}^{L/2}f'(0)k(x+c-y)u(y)dy
+\\&=(1-h'(0))\psi_f(u)(x)
+\\&=(1-h'(0))\lambda u(x)
+\end{align*}
+so that $u$ is also an eigenfuction of $\psi_m$, now with eigenvalue $(1-h'(0))\lambda$.
+
+\subsection{Separable dispersal kernels }
+Jentzsch's theorem shows that there is an eigenfunction $u$, provided that the kernel $k$ satisfy some properties.
+Finding the eigenfunction is in general a hard problem to solve.
 It becomes easier if the kernel $k$ is separable, i.e.
@@
 In that case, (\ref{eigen}) becomes
-\begin{align}
-\lambda u(x)&=f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right) \notag
-\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right) \notag
-\\ \Rightarrow \lambda d_k&=f'(0)\sum_{n=1}^\infty A_{nk}d_n  \label{problem}
-\end{align}
+\begin{align*}
+\lambda u(x)&=f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right)
+\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right)
+\\ \Rightarrow \lambda d_k&=f'(0)\sum_{n=1}^\infty A_{nk}d_n  \tag{**} \label{problem}
+\end{align*}
 where
@@
 \end{equation*}
-Finding the eigenvalues of (\ref{eigen}) then reduces to finding the eigenvalues of the matrix $(A_{nk})_{n,k=1}^\infty$.
 
-\subsection{Gaussian dispersal kernel \label{gausapp}}
+\subsection{Gaussian dispersal kernel }
 The Gaussian dispersal kernel is given by
-$$k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.$$
+$$k(|x-y|)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.$$
 As in \citep{Latore:1998fk}, this separable kernel can be written as
-$$k(x-y)=\sum_{n=0}^\infty a_n(x)b_n(y)$$
+$$k(|x-y|)=\sum_{n=0}^\infty a_n(x)b_n(y)$$
 where
@@
 
-As a first approximation to $k$ we ignore all but the $0^{th}$ terms for $a_n$ and $b_n$ so that Equation (\ref{problem}) becomes
+As a first approximation to $k$ we ignore all but the $0^{th}$ terms for $a_n$ and $b_n$ so that Equation \ref{problem} becomes
 \begin{align*}
-\lambda d_0(c)&=(1-h)f'(0)A_{00}(c)d_0(c)
-\\ \Rightarrow \lambda&=(1-h)R_0A_{00}(c)
+\lambda d_0(c)&=f'(0)A_{00}(c)d_0(c)
+\\ \Rightarrow \lambda&=R_0(1-h)A_{00}(c)
 \\\text{ where } A_{00}(c)&=2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]
@@
 
-\subsection{Sinusoidal dispersal kernel \label{sinapp}}
+\subsection{Sinusoidal dispersal kernel }
 A sinusoidal dispersal kernel is given by 
@@
 If we solve for $\lambda$,we find
-\begin{equation*} \lambda=(1-h)R_0\left[\frac{wL\cos(wc)}{4}+\frac{1}{4}\sqrt{\sin^2(wL)-w^2L^2\sin^2(wc)}\right]. \label{cosine} \end{equation*}
+\begin{equation} \lambda=R_0(1-h)\left[\frac{wL\cos(wc)}{4}+\frac{1}{4}\sqrt{\sin^2(wL)-w^2L^2\sin^2(wc)}\right]. \label{cosine} \end{equation}
 
-Zhou et al. \citep{ZhouKot2011} solve for the critical speed, $c^*$, at which the population will be driven extinct:
+Zhou et al. \citep{ZhouKot2011} solve for the critical speed, $c^*$, at the population will be driven extinct:
 $$c^*=c^*(R_0)=\frac{1}{w}\cos^{-1}\left[\frac{16+R_0^2(1-h)^2(w^2L^2-\sin^2(wL))}{8R_0(1-h)wL}\right].$$
@@
 
-\subsection{Approximate critical harvesting proportions \label{approxcrit}}
+\subsection{Approximate Critical Harvesting Proportions}
 ~\\We will use the following Taylor series to make approximations of the critical harvesting proportions under the two dispersal kernels:

### 2014-02-20 Writing/MS.tex

--- 2014-02-20 Writing/MS.tex
@@
 
-\subsection{Interactions between stressors }
-As the harvesting rate $h$ increases, 
-the critical rate of environmental shift $c^*$ decreases: the environment must move more slowly to 
-accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
+\subsection{Interactions Between Stressors }
+We find the critical climate velocity and harvest  rate to be inversely related: as the harvesting rate $h$ increases, 
+the critical climate velocity $c^*$ decreases as the environment must move more slowly to accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
 environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}). This 
 means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
-the environment starts shifting.    When the harvesting pressure has exceeded the critical harvesting rate $h^*$ or the environmental 
-is shifting more quickly than the critical rate of environmental shift $c^*$, the biomass of the population at equilibrium will be equal to $0$.
+the environment starts changing.
+When the climate velocity or harvesting pressure exceed their critical rates ($h^*, c^*$ respectively), the biomass of the population at equilibrium will be equal to $0$.
 Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure
-increases or the environmental shifts more quickly (Figure \ref{baseline}). The simulations replicate the analytical results with the critical speed $c^*$ 
-declining as the critical harvest rate $h^*$ increases and vice versa (Figure \ref{nomang}).
+increases or the environmental shifts more quickly (Figure \ref{baseline}).
+Our simulations confirm the analytical results with the critical speed $c^*$ declining as the critical harvest rate $h^*$ increases and vice versa (Figure \ref{nomang}).
 

### 2014-02-20 Simluations/Aspatial_fast/Make.R
Fixing MPA issue

@@ -3,19 +3,24 @@ setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspa
 
 
 # load parameters, functions
+  source("Parameters.R")
 	source("Functions.R")
-	source("Parameters.R")
 	
 # run analysis
 	# choose threshold or no threshold
 	model = "noThresh"	# "noThresh"; "Thresh"
 	
 	# if no threshold, choose MPA: "null", "cons", "fish"
-	MPA = "fish"
+	MPA = "cons"
 
 # analysis
-	if(model=="noThresh") {source("Parameters_nothresh.R","Sim_noThresh.R")} else {
-		if(model=="Thresh" & MPA == "null") {source("Parameters_thresh.R","Sim_thresh")} else {
+	if(model=="noThresh") {sapply(c("Parameters_nothresh.R","Sim_noThresh.R"),source,.GlobalEnv)} else {
+		if(model=="Thresh" & MPA == "null") {sapply(c("Parameters_thresh.R","Sim_thresh.R"),source,.GlobalEnv)} else {
 			warning("model needs to be 'noThresh' or 'Thresh', MPA needs to be 'null'")
 			}
 		}
+
+# 2014-02-10 
+# have run no-thresh cons
+
+

### 2014-02-20 Simluations/Aspatial_fast/Sim_noThresh.R

@@ -80,7 +80,8 @@ for(q in 1:length(speeds)){
       harv.mean <- mean(harv)
       harv.sd <- sd(harv)
       harv.se <- stderr(harv)
-      pop.mean <- mean(colSums(move)[(T+1):(T+300)])
+      #pop.mean <- mean(colSums(move)[(T+1):(T+300)])
+      pop.mean <- mean(max(colSums(move)[(T+1):(T+300)]),min(colSums(move)[(T+1):(T+300)])) # this deals with variability in taking mean of oscillation
       pop.sd <- sd(colSums(move)[(T+1):(T+300)])
       pop.se <- stderr(colSums(move)[(T+1):(T+300)])
        

### 2014-02-20 Writing/MS.tex
small edits, commiting just to make sure they dont get lost

--- 2014-02-20 Writing/MS.tex
@@
 
-When the climate changes, the areas with conditions in which organisms can survive and reproduce move. 
+When the climate changes, the habitat with suitable conditions in which organisms can survive and reproduce moves. 
 This change does not occur in isolation but rather appears on a background of other disturbances. 
@@
 Positive synergy suggests that harvesting may aggravate the population's sensitivity to a shifting range.
-Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could mitigate the negative interaction of the two stressors.
+Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could mitigate the interaction between the two stressors.
 \hspace{10cm}
@@
 We studied the dynamics of a fish population constrained to a single, one-dimensional habitat patch by their inability to reproduce outside of the patch.
-This viable habitat patch (here after `patch') is shifting at a fixed velocity and fish at each point in space can be harvested.
+This viable habitat patch (here after `patch') is shifting at a fixed velocity and the fish at each point in space can be harvested.
 We first determined the climate velocity and harvesting rate that would drive the population extinct.
@@
 
-In the model of \cite{ZhouKot2011}, the adults from the current year produce offspring according to a recruitment function, and these offspring disperse across the one-dimensional world according to a dispersal kernel to become the next generation's adults.
+In the model of \cite{ZhouKot2011}, the adults from the current year produce offspring according to a recruitment function and these offspring disperse across the one-dimensional world according to a dispersal kernel to become the next generation's adults.
 We extend this model by additionally subjecting the adults to harvesting before they produce offspring so that only a proportion of the fish survive to reproduce.
@@
 offspring produced by a population of size $n$ (accounting for density dependence), $k(x-y)$ is the dispersal kernel giving the probability of a  larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it  shifts across space.
+We provide a list of variables and functions in Table \ref{variables}.
 We chose to use a Beverton-Holt recruitment function,
@@
 
-\noindent Regardless of the exact functional form of the recruitment function, a critical parameter of the recruitment function in determining population persistence is how quickly recruitment increases when the population
+\noindent Regardless of the exact functional form of the recruitment function, the critical parameter in determining population persistence is how quickly recruitment increases when the population 
 size is near (but above) $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$.
 Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$ and we use such a kernel in our analytical expressions.
-We provide a list of variables and functions in Table \ref{variables}.
 
@@
 One possible equilibrium traveling wave that solves Equation (\ref{traveling_pulse}) is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
-{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.  If a population becomes very small (or if we introduce a small 
-population), one of two things can happen. First, the population may crash and the trivial traveling pulse 
-without any fish may appear again. Second, those small numbers may increase and form a stable population. 
-In this sense, a small population can be thought of as a perturbation to the trivial traveling pulse. If the trivial 
-pulse is stable, the system will return to the trivial pulse even after a perturbation in the form of the introduction 
-of a small population. If a population is to persist, even when it is small it must be able to avoid extinction and 
-grow. For this to be the case, the trivial pulse must be unstable to small perturbations.
+{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.  If a population is to persist, it must be able to avoid extinction and 
+grow even when it is small.   A small population can be thought of as a perturbation to the trivial traveling pulse. If the trivial 
+pulse is stable, the system will return to the trivial pulse even after the introduction 
+of a small population.
+If the trivial pulse is unstable, a small population may increase and form a stable population.
+Population persistence is therefore equivalent to the trivial traveling pulse being an unstable equilibrium.
 
-We would like to know the rate of environmental shift and the harvesting rate such that as long as the 
-environment moves more slowly or we harvest less severely than those parameters, then the population will be able to persist.
-We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$.
+We would like to know the rate of environmental shift and the harvesting rate such that  the population will be able to persist as long as the 
+environment moves more slowly or we harvest less severely than those parameters. We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$. 
 We found these rates by finding the parameters that make the trivial pulse unstable.   In our analyses, as in \citep{Latore:1998fk}, we used the separable Gaussian kernel given by
@@
 
-\noindent To find analytical results about the Gaussian kernel, we approximated the kernel, as described in the Appendix.
+\noindent To derive analytical expressions, we approximated the kernel, as described in the Appendix.
 Analytical results for a separable sinusoidal kernel are also described in the Appendix.
@@
 $).
-Whereas equilibrium biomass depends on the function form of recruitment, population persistence only depends on the intrinsic growth rate $R_0$.
+The population biomass at equilibrium depends on the function form of recruitment, but population persistence only depends on the intrinsic growth rate $R_0$.
 If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.

### 2014-02-20 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd

@@
 ```
+
+Idea, can I sub-sample the 2100 step run and take the mean of an increasingly limited window?
+This would mean using `df2100` as the data frame and taking the mean of all 2100 steps, and then 2099 steps, ...
+etc.
+down to 300 and seeing how the mean changes.
+
+```{r}
+int <- seq(300,2100)
+names(series2100) = speeds
+foo <- melt(series2100)
+names(foo) <- c("biomass","speed")
+means = vector("list",length(int))
+names(means) = int
+for(i in 1:length(int)){
+  means[[i]] = ddply(foo,.(speed),function(df) avg = mean(df$biomass[1:int[i]]))
+  }
+
+bar = melt(means)
+bar$variable = NULL
+names(bar) = c("speed", "mean", "interval")
+bar$interval = as.numeric(bar$interval)
+ggplot(bar, aes(x = interval , y = mean))  + geom_line() + facet_wrap( ~ speed)
+```
+
+New idea, find the maxima and minima and take the mean of those two numbers. See if that erases the variability in means.
+
+```{r}
+int <- seq(300,2100)
+names(series2100) = speeds
+foo <- melt(series2100)
+names(foo) <- c("biomass","speed")
+means_dif = vector("list",length(int))
+names(means) = int
+for(i in 1:length(int)){
+  means_dif[[i]] = ddply(foo,.(speed),function(df) avg=mean(max(df$biomass[1:int[i]]),min(df$biomass[1:int[i]])))
+  }
+
+bar = melt(means_dif)
+bar$variable = NULL
+names(bar) = c("speed", "mean", "interval")
+bar$interval = as.numeric(bar$interval)
+ggplot(bar, aes(x = interval , y = mean))  + geom_line() + facet_wrap( ~ speed)
+
+ggplot(bar, aes(x = speed, y = mean, group = interval)) + geom_line(aes(colour=interval))
+```
+
+And volia!
+It erases that variability.
+Looks like I can keep the 300 time steps and take the max and minimum and take the average.
+So should re-do that for the existing simulation data I have and plot a contour map
+
+```{r}
+
+```

### 2014-02-20 Writing/MS.tex
proofreading;

@@
 size is near (but above) $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$.
-Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$ and we use such a kernel in our analytical expressions.
+Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.
+In our analyses, as in \citep{Latore:1998fk}, we used the separable Gaussian kernel given by
+
+\[k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.\]
+
+\noindent To derive analytical expressions, we approximated the kernel, as described in the Appendix.
+Analytical results for a separable sinusoidal kernel are also described in the Appendix.
+We used
+simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
 
@@
 
-We would like to know the rate of environmental shift and the harvesting rate such that  the population will be able to persist as long as the 
-environment moves more slowly or we harvest less severely than those parameters. We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$. 
-We found these rates by finding the parameters that make the trivial pulse unstable.   In our analyses, as in \citep{Latore:1998fk}, we used the separable Gaussian kernel given by
-
-\[k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.\]
-
-\noindent To derive analytical expressions, we approximated the kernel, as described in the Appendix.
-Analytical results for a separable sinusoidal kernel are also described in the Appendix.
-We used
-simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
+If the population is harvested at low enough levels and the environment shifts slowly enough, the population will be able to persist.
+There are threshold values of the harvesting rate $h$ and the rate of environmental shift $c$ such that if the parameters are increased beyond these values, the population will be driven extinct.
+We found these critical parameters, $h^*$, and  $c^*$, by finding the parameters that make the trivial pulse unstable.
+Details are provided in Appendix \ref{stab}.
 
@@
 The population biomass at equilibrium depends on the function form of recruitment, but population persistence only depends on the intrinsic growth rate $R_0$.
-If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.
 For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that

### 2014-02-21 Writing/MS.tex
running new simulations

--- 2014-02-20 Writing/MS.tex
@@
 
-When the climate changes, the habitat with suitable conditions in which organisms can survive and reproduce moves. 
+When the climate changes, the areas with conditions in which organisms can survive and reproduce move. 
 This change does not occur in isolation but rather appears on a background of other disturbances. 
@@
 Positive synergy suggests that harvesting may aggravate the population's sensitivity to a shifting range.
-Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could mitigate the interaction between the two stressors.
+Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could mitigate the negative interaction of the two stressors.
 \hspace{10cm}
@@
 We studied the dynamics of a fish population constrained to a single, one-dimensional habitat patch by their inability to reproduce outside of the patch.
-This viable habitat patch (here after `patch') is shifting at a fixed velocity and the fish at each point in space can be harvested.
+This viable habitat patch (here after `patch') is shifting at a fixed velocity and fish at each point in space can be harvested.
 We first determined the climate velocity and harvesting rate that would drive the population extinct.
@@
 
-In the model of \cite{ZhouKot2011}, the adults from the current year produce offspring according to a recruitment function and these offspring disperse across the one-dimensional world according to a dispersal kernel to become the next generation's adults.
+In the model of \cite{ZhouKot2011}, the adults from the current year produce offspring according to a recruitment function, and these offspring disperse across the one-dimensional world according to a dispersal kernel to become the next generation's adults.
 We extend this model by additionally subjecting the adults to harvesting before they produce offspring so that only a proportion of the fish survive to reproduce.
@@
 offspring produced by a population of size $n$ (accounting for density dependence), $k(x-y)$ is the dispersal kernel giving the probability of a  larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it  shifts across space.
-We provide a list of variables and functions in Table \ref{variables}.
 We chose to use a Beverton-Holt recruitment function,
@@
 
-\noindent Regardless of the exact functional form of the recruitment function, the critical parameter in determining population persistence is how quickly recruitment increases when the population 
+\noindent Regardless of the exact functional form of the recruitment function, a critical parameter of the recruitment function in determining population persistence is how quickly recruitment increases when the population
 size is near (but above) $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$.
 Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$ and we use such a kernel in our analytical expressions.
+We provide a list of variables and functions in Table \ref{variables}.
 
@@
 One possible equilibrium traveling wave that solves Equation (\ref{traveling_pulse}) is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
-{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.  If a population is to persist, it must be able to avoid extinction and 
-grow even when it is small.   A small population can be thought of as a perturbation to the trivial traveling pulse. If the trivial 
-pulse is stable, the system will return to the trivial pulse even after the introduction 
-of a small population.
-If the trivial pulse is unstable, a small population may increase and form a stable population.
-Population persistence is therefore equivalent to the trivial traveling pulse being an unstable equilibrium.
+{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.  If a population becomes very small (or if we introduce a small 
+population), one of two things can happen. First, the population may crash and the trivial traveling pulse 
+without any fish may appear again. Second, those small numbers may increase and form a stable population. 
+In this sense, a small population can be thought of as a perturbation to the trivial traveling pulse. If the trivial 
+pulse is stable, the system will return to the trivial pulse even after a perturbation in the form of the introduction 
+of a small population. If a population is to persist, even when it is small it must be able to avoid extinction and 
+grow. For this to be the case, the trivial pulse must be unstable to small perturbations.
 
-We would like to know the rate of environmental shift and the harvesting rate such that  the population will be able to persist as long as the 
-environment moves more slowly or we harvest less severely than those parameters. We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$. 
+We would like to know the rate of environmental shift and the harvesting rate such that as long as the 
+environment moves more slowly or we harvest less severely than those parameters, then the population will be able to persist.
+We call these, respectively, the critical rate of environmental shift, $c^*$, and the critical harvesting rate, $h^*$.
 We found these rates by finding the parameters that make the trivial pulse unstable.   In our analyses, as in \citep{Latore:1998fk}, we used the separable Gaussian kernel given by
@@
 
-\noindent To derive analytical expressions, we approximated the kernel, as described in the Appendix.
+\noindent To find analytical results about the Gaussian kernel, we approximated the kernel, as described in the Appendix.
 Analytical results for a separable sinusoidal kernel are also described in the Appendix.
@@
 $).
-The population biomass at equilibrium depends on the function form of recruitment, but population persistence only depends on the intrinsic growth rate $R_0$.
+Whereas equilibrium biomass depends on the function form of recruitment, population persistence only depends on the intrinsic growth rate $R_0$.
 If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.
@@
 We also examined the effect of marine protected  areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and  climate change parameters where the fish population could survive.
-With MPAs in place, the population had a  slightly higher abundance along the edges of the patch where the population is limited by harvesting, which  translated into a slightly increased critical harvest rate (Figure \ref{management}).
+We found both MPA strategies examined (many small versus few large), allowed the population was able to withstand combinations of higher climate velocities and harvesting rates.
+However MPAs did not increase the critical climate velocity (Figure \ref{management}).
 
@@
 Our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
-As shown in Figure \ref{baseline}, species with a shorter generation time and a longer average dispersal distance will better track a high rate of environmental shift relative to a species that has a long generation time and short dispersal distance.
-This is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity) \citep{Perryetal2005}.
+As shown in Figure \ref{baseline}, species with a higher reproductive rate and a longer average dispersal distance will better track a high rate of environmental shift relative to a species that has a low reproductive rate and short dispersal distance.
+The finding that a higher reproductive rate can sustain higher climate velocities and harvesting rates is intuitive, especially because harvesting rate and reproductive rate cancel each other out.
+However it is worth pointing out that a higher reproductive rate can be generated either by shorter generation times or higher fecundity.
+Finding that species with shorter generation times is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity) \citep{Perryetal2005}.
 
@@
 
-\noindent {\bf Figure \ref{baseline}}: (a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-(b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+\noindent {\bf Figure \ref{baseline}}: (a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
 Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
@@
 These results are from an approximated Gaussian dispersal kernel with $L=1$.
+(b)
+The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
 

### 2014-02-21 Simluations/Aspatial_fast/Make.R

@@ -8,10 +8,10 @@ setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspa
 	
 # run analysis
 	# choose threshold or no threshold
-	model = "noThresh"	# "noThresh"; "Thresh"
+	model = "Thresh"	# "noThresh"; "Thresh"
 	
 	# if no threshold, choose MPA: "null", "cons", "fish"
-	MPA = "cons"
+	MPA = "null"
 
 # analysis
 	if(model=="noThresh") {sapply(c("Parameters_nothresh.R","Sim_noThresh.R"),source,.GlobalEnv)} else {

### 2014-02-21 Simluations/Aspatial_fast/Sim_thresh.R

@@ -1,6 +1,15 @@
 
 # added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
 
+if(MPA=="cons") {mpa.yes=cons.yes; mpa.no=cons.no} else {
+  if(MPA=="fish") {mpa.yes=fish.yes; mpa.no=fish.no} else {
+    if(MPA=="null") {mpa.yes=null.yes; mpa.no=null.no} else{
+      if(exists("MPA")) warning(paste("MPA needs to be 'cons', 'fish', or 'null'.",sep=""))
+    }
+  }
+}
+
+
 
 # initialize the world, set MPA (cons.yes/no; fish.yes/no; null.yes/no)
 	output <- startOut(w,maxt,mpa.yes,mpa.no,world)
@@ -11,7 +20,7 @@
 	biggestP <- max(init[,maxt])
 	thresholds <- thresholds * biggestP
 # add a buffer to last to make sure captures biggest population
-	tail(thresholds,1) <- tail(thresholds,1) + 0.001
+  thresholds[length(thresholds)]= thresholds[length(thresholds)] + 0.001
 
 # standard for equilibrium is the difference in the final step of the initialization 
 	init.diff <- diff(colSums(init))[149]

### 2014-02-21 Writing/MS.tex
edits

@@
 However MPAs did not increase the critical climate velocity (Figure \ref{management}).
+We also found that the spacing and size of the MPAs changed population dynamics.
+MPAs that were spaced more than 1 average dispersal distance apart resulted in large oscillations of population biomass at low climate velocities.
+As climate velocities increase, the mean population abundance declines but the population experiences less extreme oscillations in abundance, which results in the population being bounded further from possible extinction in a stochastic environment.
 
@@
 However it is worth pointing out that a higher reproductive rate can be generated either by shorter generation times or higher fecundity.
-Finding that species with shorter generation times is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity) \citep{Perryetal2005}.
+Finding that species with shorter generation times can better keep up with shifts in climate is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species \citep{Perryetal2005}.
+While higher reproductive rates improved a population's ability to persist, there was a tradeoff in increasing dispersal distances.
+At low speeds, we found that a short dispersal dispersal distance improved the maximum harvesting rate a population could sustain while at higher speeds a longer dispersal distance improved the maximum climate velocity in which the population could persist (Figure \ref{baseline}).
+This tradeoff is due to the proportion of dispersing  offspring at time step $t$ which lands within the patch at time step $t + 1$.
+When climate is shifting slowly, a large dispersal distance sends most offspring ahead of the patch, while with faster climate velocities a long dispersal distance allows the population to make it to the new patch (Figure \ref{baseline}).
+Thus climate velocity will selectively favor species with dispersal distances best matched to the rate of shift.
 
 We also examined whether frequently recommended management approaches ensure species persistence.
-We found increases in the population's biomass at equilibrium and an improved ability to persist.
+With these management strategies we found increases in the population's biomass at equilibrium and an improved ability to persist.
 Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
-Our results, that spatial management increased the maximum harvesting rate at which the population could survive, support the idea that MPAs could be used to reduce the impact of harvesting.
-However we did not find any evidence that MPAs increased the climate velocity under which the population could persist.
-The second management approach we investigated, harvesting thresholds, are already widely implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors.
-While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.
+Our results, that spatial management increased the maximum harvesting rate at which the population could survive  and improved the populations ability to persist at a slightly higher climate velocity conditional on a given harvesting rate [NEW FIGURE], support the idea that MPAs could be used to reduce the impact of harvesting and help to ameliorate the effects of climate velocity.
+However the spacing and size of the MPAs matter: few, large MPAs caused increased variability at low climate velocities while many smaller MPAs maintained a population that was bounded further from extinction.
+
+ When we applied harvesting thresholds, we found that it alleviates interactions between the two stressors.
+Thresholds have this effect as the management approach effectively prevents harvesting of the leading edge, which allows colonization to occur as if these individuals were moving into un-fished areas.
+It's interesting to note that novel, low abundance species are commonly unregulated in fisheries systems; in order to decouple the additive effects of harvest and climate change, this paradigm would have to be reversed: no new species would be allowed to harvest until they had become established.
+These results highlight that while management strategies only change harvesting practices, they can effect the way stressors interact and help to better implement harvesting rules and place protected areas.
 
@@
 We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
-However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect.
+However, we expect qualitatively similar results.
+An Allee effect would make it harder for populations to colonize new areas and add a threshold below which fishing drives the population to extinction.
+Thus an Allee effect would change lower the critical harvest rates and climate velocity, but we do not expect the additive nature of the interaction between climate and harvesting to change.
 We also did not include age structure in our model.
-The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work.
+The effects of both harvesting and climate change may be different across different age classes and may destabilize the system in complicated ways, including resonance \citep{Botsfordetal2011, Planqueetal2010}; including this level of complexity is left for future work.
 Similarly, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
@@
 Exploring how species interactions, age structure, and additional disturbances (e.g.
-pollution, disease, physiological response to temperature) affect population viability will improve our predictions and help us to understand whether species will persist under predicted climate and harvesting regimes.
-Finally, this work can help make general predictions as to whether specific life histories are likely to be selected over others as harvesting and range shifts increase.
+physiological response to temperature) affect population viability will improve our predictions and help us to understand whether species will persist under predicted climate and harvesting regimes.
+Finally, this work can help make general predictions as to whether specific life histories are likely to be selected over others as harvesting and range shifts increase and highlights the importance of considering stressors in combination as outcomes can deviate substantially from what would be predicted in isolation.
+This is especially true for management strategies which may result in unanticipated effects such as large fluctuations associated with big, distant MPAs shown here.
 

### 2014-02-21 Simluations/Aspatial_fast/Make.R

@@ -22,5 +22,7 @@ setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspa
 
 # 2014-02-10 
 # have run no-thresh cons
-
+# have run no-thresh fish
+# have run no-thresh null
+# have run thresh null
 

### 2014-02-21 Simluations/Aspatial_fast/Sim_thresh.R

@@ -32,7 +32,7 @@ for(q in 1:length(speeds)){
   
   for(j in 1:length(thresholds)){
     init.h<-array(0,c(w,1))
-    next.pop <- m(n=init[,maxt], s = 0, Fthresh = thresholds[j], Fharv = 1, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+    next.pop <- m(n=init[,maxt], s = 0, Fthresh = thresholds[j], Fharv = 1, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.start)
     init.h[,1]=next.pop[[1]]
     MPA.current = next.pop[[3]]
     h.diff <- abs(sum(init[,maxt]-init.h[,1]))

### 2014-02-21 Writing/MS.tex
formatting for Ecological Applications

@@
 \usepackage{geometry}
-\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
+\geometry{verbose,letterpaper,tmargin=2.4cm,bmargin=2.4cm,lmargin=2.4cm,rmargin=2.4cm}
 \usepackage[T1]{fontenc}
@@
 \title{Persistence of marine populations under climate and fishing}
-\author{Emma Fuller, Eleanor Brush, Malin Pinsky}
-\date{}
+\author{Emma Fuller$^1$, Eleanor Brush$^1$, Malin Pinsky$^{1,2}$}
+\date{\small (1): Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey 08544 USA \\
+(2): Department of Ecology, Evolution and Natural Resources, Rutgers University, New Brunswick, New Jersey 08901 USA}
 
@@
 \maketitle
-
+\pagebreak
 \begin{spacing}{1.9}
@@
 \end{flushleft}
-\end{spacing}
 
@@
 \pagebreak
+
+\section*{Figure Legends}
+
+\noindent {\bf Figure \ref{baseline}}: (a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
+Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
+These results are from an approximated Gaussian dispersal kernel with $L=1$.
+(b)
+The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+\hspace{6in}
+\hspace{6in}
+
+\noindent {\bf Figure \ref{Synergy}}: Positive synergy between the two stressors.
+The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
+This excess loss, on the order of $.001$, is small in comparison to the total biomass, which can be as large as $20$.
+These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+\hspace{6in}
+\hspace{6in}
+
+\noindent {\bf Figure \ref{management}}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+(a) No management.
+(b) Threshold harvesting levels.
+(c) MPAs.
+These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
+\pagebreak
+
 \section{Figures}
@@
 
-\noindent {\bf Figure \ref{baseline}}: (a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
-Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
-Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-These results are from an approximated Gaussian dispersal kernel with $L=1$.
-(b)
-The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-
-\pagebreak
-
 \begin{figure}[htbp]
@@
 \pagebreak
-
-\noindent {\bf Figure \ref{Synergy}}: Positive synergy between the two stressors.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-This excess loss, on the order of $.001$, is small in comparison to the total biomass, which can be as large as $20$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
 
@@
 
-\noindent {\bf Figure \ref{management}}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
-(a) No management.
-(b) Threshold harvesting levels.
-(c) MPAs.
-These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
-
-\pagebreak
-
 \section{Tables}
@@
 
-\appendix
-\section{Appendix}
-In Appendix \ref{stab}, we provide the details for assessing the persistence of a population with an integrodifference model and we discuss the effect of the harvesting function on population persistence.
-In Appendix \ref{sep}, we provide the details for assessing population persistence with separable dispersal kernels.
-In Appendix \ref{gausapp} and \ref{sinapp}, we derive expressions for the critical harvesting rate and rate of environmental shift for Gaussian and sinuosoidal dispersal kernels.
-In Appendix \ref{approxcrit}, we derive approximate expressions for these critical rates.
-
-\subsection{Determining stability \label{stab}}
-As in Zhou et al.
-\citep{ZhouKot2011}, let $k(x-y)$ be a dispersal kernel, let $f(n)$ be a recruitment function, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
-The integrodifference model describing the population over time is given by
-
-\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy. \label{integro} \end{equation}
-
-To find a traveling pulse, we are only interested in the population density as a function of the location within the patch rather than absolute position, $\overline{x}\equiv x-ct$.
-
-\begin{equation*}
-n^*(\overline{x})\equiv n^*(x-ct)=n_t(x).   \label{trav} 
-\end{equation*}
-
-The integrodifference equation (\ref{integro}) gives us an expression for $n^*$:
-
-\begin{align}
-n^*(\overline{x}-c)&=\int_{-L/2}^{L/2}k(\overline{x}-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y})))d\overline y \notag
-\\ \Rightarrow n^*(\overline{x})&=\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y}))d\overline y  \label{pulse}
-\end{align}
-As long as $f(0)=0$, there is a trivial solution to this problem where $n^*(\overline{x})\equiv 0$ for all $\overline{x}\in[-L/2,L/2]$, i.e.
-there is a trivial traveling pulse with no fish in it.
-If the trivial traveling pulse is unstable, even very small populations will persist or grow and avoid crashing back to the trivial pulse.
-To evaluate stability a traveling pulse, we introduce a small perturbation to the traveling pulse $n^*(\overline{x})$ and see if this perturbation grows or shrinks over time:
-
-\begin{align}
-n_t(x)&=n^*(\overline{x})+\xi_t(x) \notag
-\\ \Rightarrow \xi_{t+1}(x)&=n_{t+1}(x)-n^*(\overline{x}) \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(f(n_t(y)-g(n_t(y)))-f(n^*(\overline{y})-g(n^*(\overline{y}))\big)dy \text{ using (\ref{pulse})} \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(n^*(\overline{y})))f'(n^*(\overline{y})-g(n^*(\overline{y}))\xi_t(y)dy \notag
-\\& \text{ by linearizing around the traveling pulse} \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(0))f'(0)\xi_t(y)dy \text{ if $n^*(\overline{x})=0$} \label{perturb}
-\end{align}
-
-If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\R$ and $u:[-L/2,L/2]\to\R$, then the perturbation grows in time if and only if $\lambda >1$.
-Using Equation (\ref{perturb}), we can rewrite $\xi_t(x)$,
-\begin{align*}
-\lambda u(x-ct-c)&=(1-g'(0))f'(0)\int_{-L/2+ct}^{L/2+ct}k(x-y)u(y-ct)dy
-\\ \\\Rightarrow \lambda u(\overline{x})&=(1-g'(0))f'(0)\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})u(\overline{y})dy
-\end{align*}
-
-Define the integral operator
-$$ \psi_f(u)(x)=(1-g'(0))f'(0)\int_{-L/2}^{L/2}k(x+c-y)u(y)dy. $$
-Then the perturbation to the traveling pulse will satisfy 
-\begin{equation} \psi_f(u)(x)=\lambda u(x) \label{eigen} \end{equation}
-$\lambda$ and $u$ are thus an eigenvalue and eigenvector of the functional operator $\psi_f$.  The trivial traveling pulse is unstable when the dominant eigenvalue of $\psi_f$ is greater than $1$.
-
-The biomass in the equilibrium traveling wave depends on the specific functional form of the harvesting function $g(n)$.
-However, the persistence of the population only depends on $g'(0)$ so in this paper, we only considered a proportional harvesting function, i.e.
-the amount of fish harvested obeyed $g(n)=hn$.
-For this function, $g'(0)=h$.
-
-\subsection{Separable dispersal kernels \label{sep}}
-Jentzsch's theorem shows that there is an eigenfunction $u$, provided that the kernel $k$ satisfies some properties \citep{ZhouKot2011}.
-Finding the eigenfunctions and eigenvalues is in general a hard problem to solve.
-It becomes easier if the kernel $k$ is separable, i.e.
-there are functions $a_n,b_n$ such that $k(x-y)=\sum_{n=1}^\infty a_n(x)b_n(y)$.
-In that case, (\ref{eigen}) becomes
-\begin{align}
-\lambda u(x)&=f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right) \notag
-\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right) \notag
-\\ \Rightarrow \lambda d_k&=f'(0)\sum_{n=1}^\infty A_{nk}d_n  \label{problem}
-\end{align}
-where
-\begin{equation*}
-A_{nk}=\int_{-L/2}^{L/2}a_n(x)b_k(x-c)dx \text{ and } d_k=\int_{-L/2}^{L/2}b_k(x-c)u(x)dx
-\end{equation*}
-Finding the eigenvalues of (\ref{eigen}) then reduces to finding the eigenvalues of the matrix $(A_{nk})_{n,k=1}^\infty$.
-
-\subsection{Gaussian dispersal kernel \label{gausapp}}
-The Gaussian dispersal kernel is given by
-$$k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.$$
-As in \citep{Latore:1998fk}, this separable kernel can be written as
-$$k(x-y)=\sum_{n=0}^\infty a_n(x)b_n(y)$$
-where
-$$a_n(x)=b_n(x)=\frac{1}{\sqrt{2n!\sqrt{D\pi}}}e^{-x^2/4D}\left(\frac{x}{\sqrt{2D}}\right)^n.$$
-
-As a first approximation to $k$ we ignore all but the $0^{th}$ terms for $a_n$ and $b_n$ so that Equation (\ref{problem}) becomes
-\begin{align*}
-\lambda d_0(c)&=(1-h)f'(0)A_{00}(c)d_0(c)
-\\ \Rightarrow \lambda&=(1-h)R_0A_{00}(c)
-\\\text{ where } A_{00}(c)&=2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]
-\end{align*}
-where $\text{erf}$ is the error function.  The critical rate of environmental shift $c^*$ and the critical harvesting rate $h^*$ are those values of $c$ and $h$, respectively, that make $\lambda=1$.
-
-\subsection{Sinusoidal dispersal kernel \label{sinapp}}
-A sinusoidal dispersal kernel is given by 
-$$k(x-y)=\left\{\begin{array}{ccccc}
-\frac{w}{2}\cos(w(x-y)) & , & |x-y|\leq\frac{\pi}{2w}
-\\ 0 & , & |x-y|>\frac{\pi}{2w}
-\end{array}\right.
-$$
-where $L$ is the length of the patch and we assume $\frac{\pi}{2w}>L,c<\frac{\pi}{2w}-L$.
-
-In this case, $k(x-y)=\frac{w}{2}\cos(wx)\cos(w(y-c))+\frac{w}{2}\sin(wx)\sin(w(y-c))$ so that $A_{ij}$ and $d_i$ can be found for $i,j=1,2$ and (\ref{problem}) reduces to 
-$$\lambda^2-\left(\frac{R_0(1-h)wL}{2}\cos(wc)\right)\lambda+\frac{R_0^2(1-h)^2}{16}\left(w^2L^2-\sin^2(wL)\right)=0.$$
-
-If we solve for $\lambda$,we find
-\begin{equation*} \lambda=(1-h)R_0\left[\frac{wL\cos(wc)}{4}+\frac{1}{4}\sqrt{\sin^2(wL)-w^2L^2\sin^2(wc)}\right]. \label{cosine} \end{equation*}
-
-Zhou et al. \citep{ZhouKot2011} solve for the critical speed, $c^*$, at which the population will be driven extinct:
-$$c^*=c^*(R_0)=\frac{1}{w}\cos^{-1}\left[\frac{16+R_0^2(1-h)^2(w^2L^2-\sin^2(wL))}{8R_0(1-h)wL}\right].$$
-Similarly, we can solve for the critical harvesting rate, $h^*$, at which the population will be driven extinct:
-$$
-h^*=1-\frac{1}{R_0}\cdot\frac{4wL}{w^2L^2-\sin^2(wL)}\left[\cos(wc)-\sqrt{\cos^2(wc)-1+\frac{\sin^2(wL)}{w^2L^2}}\right] 
-$$
-
-\subsection{Approximate critical harvesting proportions \label{approxcrit}}
-~\\We will use the following Taylor series to make approximations of the critical harvesting proportions under the two dispersal kernels:
-\begin{align*}
-\cos(x)&=1-\frac{x^2}{2}
-\\ \cos^2(x)&=1-x^2
-\\ \sin^2(x)&=x^2-\frac{x^4}{3}
-\\ erf(x)&=\frac{2}{\sqrt{\pi}}(x-\frac{x^3}{3})
-\\ \exp(x)&=1+x+\frac{x^2}{2}
-\end{align*}
-For the sinusoidal kernel we found 
-\begin{equation}
-h^*=1-\frac{1}{R_0}\cdot\frac{4wL}{w^2L^2-\sin^2(wL)}\left[\cos(wc)-\sqrt{\cos^2(wc)-1+\frac{\sin^2(wL)}{w^2L^2}}\right] 
-\end{equation} 
-Using the Taylor series and the fact that $w=\frac{\sqrt{\frac{\pi^2}{4}-2}}{\sigma}$ where $\sigma^2$ is the variance of the sinusoidal kernel,
-\begin{align*}
-h^*&\sim 1-\frac{1}{R_0}\cdot\frac{12wL}{w^4L^4}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{12}{w^3L^3}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{12\sigma^3}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[1-\frac{(\frac{\pi^2}{4}-2)c^2}{2\sigma^2}-\sqrt{1-\frac{(\frac{\pi^2}{4}-2)c^2}{\sigma^2}-\frac{(\frac{\pi^2}{4}-2)L^2}{3\sigma^2}}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[2\sigma^2-(\frac{\pi^2}{4}-2)c^2-\frac{2}{\sqrt{3}}\sigma\sqrt{3\sigma^2-(\frac{\pi^2}{4}-2)3c^2-(\frac{\pi^2}{4}-2)L^2}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^34(\frac{\pi^2}{4}-2)^{3/2}}\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
-%\\&=1-\frac{12}{R_0L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
-\\&=1-\frac{1}{R_0}\cdot\frac{4\sqrt{3}}{L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sqrt{3}\sigma^2-(\pi^2-8)\sqrt{3}c^2-4\sigma\sqrt{12\sigma^2-(\pi^2-8)(3c^2+L^2)}\right]
-\end{align*}
-For the Gaussian kernel we found 
-\begin{equation}
-h^*=1-\frac{2\sqrt{2}\exp\left(\frac{c^{2}}{8D}\right)}{R_0\left[erf\left(\frac{L-c}{2\sqrt{2D}}\right)-erf\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]}
-\end{equation} 
-Using the Taylor series and the fact that $D=\frac{\sigma^2}{2}$ where $\sigma^2$ is the variance of the exponential kernel,
-\begin{align*}
-h^*&\sim 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\sqrt{\pi}\left[\frac{L-c}{2\sqrt{2D}}-\frac{(L-c)^3}{3(2\sqrt{2D})^3}-\frac{-L-c}{2\sqrt{2D}}+\frac{(-L-c)^3}{3(2\sqrt{2D})^3)}\right]}
-%\\&= 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\left[\frac{2L}{2\sqrt{2D}}+\frac{-L^3+3L^2c-3Lc^2+c^3-L^3-3L^2c-3Lc^2-c^3}{3(2\sqrt{2D})^3}\right]}
-%\\&= 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\left[\frac{L}{\sqrt{2D}}-2\frac{L^3+3Lc^2}{3(2\sqrt{2D})^3}\right]}
-%\\&= 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0L\left[\frac{48D-2(L^2+3c^2)}{3(2\sqrt{2D})^3}\right]}
-%\\&= 1-\frac{3*32D\sqrt{D\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0L\left(48D-2(L^2+3c^2)\right)}
-%\\&= 1-\frac{3*32D\sqrt{D\pi}(\frac{128D^2+16c^2D+c^4}{128D^2})}{R_0L\left(48D-2(L^2+3c^2)\right)}
-%\\&= 1-\frac{3\sqrt{\pi}(128D^2+16c^2D+c^4)}{4R_0L\sqrt{D}\left(48D-2(L^2+3c^2)\right)}
-%\\&= 1-\frac{3\sqrt{2\pi}(32\sigma^4+8c^2\sigma^2+c^4)}{4R_0L\sigma\left(24\sigma^2-2(L^2+3c^2)\right)}
-\\&= 1-\frac{1}{R_0}\cdot\frac{3\sqrt{2\pi}}{8L}\frac{(32\sigma^4+8c^2\sigma^2+c^4)}{\sigma\left(12\sigma^2-(L^2+3c^2)\right)}
-\end{align*}
-In the case of both kernels, the critical harvesting proportion can be approximated by a function that looks like 
-\begin{equation}
-h^*\sim1- \frac{1}{R_0}\cdot C(L)f(\sigma^2,c^2,L^2+3c^2)
-\end{equation}
-where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
-
+\end{spacing}
 \end{document}

### 2014-02-21 Simluations/Aspatial_fast/Make.R
running simulations, formatting paper

@@ -1,6 +1,6 @@
 rm(list=ls())
 setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
-
+require(zoo)
 
 # load parameters, functions
   source("Parameters.R")

### 2014-02-21 fish_figures.R

@@ -73,8 +73,10 @@ xdiff=.1
 ydiff=.2
 myaxes=list(arrows=FALSE,x=list(at=rev(seq(1,C,length.out=(cvals[C]-cvals[1])/xdiff+1)),labels=as.character(round(seq(cvals[1],cvals[C],by=xdiff),2))),y=list(at=rev(seq(1,H,length.out=(hvals[H]-hvals[1])/ydiff)),labels=as.character(round(seq(hvals[1],hvals[H],by=ydiff),2))))
 
+fig.width = 3	# for ecological applications, text needs to be at least 6 point font size
+
 quartz()
-pdf(file='eqbiomass.pdf',width=6.83)
+pdf(file='eqbiomass.pdf',width=fig.width)
 
 # wireframe(toplot_ebm[C:1,H:1],xlab=list("Rate of environmental shift",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
 par(oma=oma)
@@ -140,7 +142,7 @@ myaxes=list(arrows=FALSE,
 	z=list(at=seq(0,max(ebm),by=250),labels=as.character(seq(0,max(ebm),by=250))))
 
 quartz()
-pdf(file='eqbiomass_sim.pdf',width=6.83)
+pdf(file='eqbiomass_sim.pdf',width=fig.width)
 
 # wireframe(toplot_ebm[C:1,H:1],xlab=list("Rate of environmental shift",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
 par(oma=oma)
@@ -203,7 +205,7 @@ myaxes=list(arrows=FALSE,
 	z=list(at=seq(0,max(ebm),by=250),labels=as.character(seq(0,max(ebm),by=250))))
 
 quartz()
-pdf(file='eqbiomass_fishmpa.pdf',width=6.83)
+pdf(file='eqbiomass_fishmpa.pdf',width=fig.width)
 
[…]

### 2014-02-21 fish_figures (Autosaved).R
formatting figures, lots of work to do, cons simulation still not fixed

@@ -168,7 +168,7 @@ dev.off()
 	#mpas = read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv")
 # github copy
 	x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-02-20.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMi0yMC5jc3YiLCJleHBpcmVzIjoxMzkzNjIzMDUzfQ%3D%3D--d793f2abea78846c6b8535d73e76d5836207dea6")
-	maps = read.csv(textConnection(x))
+	mpas = read.csv(textConnection(x))
 cvals=unique(mpas$speed)
 hvals=unique(mpas$harvest)
 
@@ -210,7 +210,7 @@ myaxes=list(arrows=FALSE,
 	z=list(at=seq(0,max(ebm),by=250),labels=as.character(seq(0,max(ebm),by=250))))
 
 quartz()
-pdf(file='eqbiomass_fishmpa.pdf',width=fig.width)
+pdf(file='eqbiomass_fishmpa.pdf',width=fig.width,height=fig.width)
 
 # wireframe(toplot_ebm[C:1,H:1],xlab=list("Rate of environmental shift",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
 par(oma=oma)
@@ -220,7 +220,7 @@ image.plot(cvals,hvals,toplot_ebm,breaks=cuts,col=mycols,
 	yaxs=yaxs,xaxs=xaxs,axes=TRUE,
 	legend.shrink=1,legend.width=.1,zlim=range(cuts),
 	axis.args=list(at=myaxes$z$at,labels=myaxes$z$labels,cex.axis=cex.axis),
[…]

### 2014-02-27 Simluations/Aspatial_fast/Functions.R
working on simulation code, MPAs

@@ -23,6 +23,64 @@ f<-function(n,R0,K) return(R0*n/(1+(R0-1)/K*n))
 #Standard error removing NAs
 stderr <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
 
+
+moveMPA <- function(MPA.current = MPA.current, displaced = displaced, mpa.yes=mpa.yes, mpa.no=mpa.no, world){
+  
+  ##### Move MPAS
+  # move MPA forward by <displaced> amount
+  next_MPA = MPA.current[displaced:length(MPA.current)]
+  lost <- MPA.current[1:(displaced-1)] 
+  
+  # are there any MPAs in <next_MPA>?
+  #any(test[[i]]$next_MPA==1)
+  # IF FALSE, then need to figure out how many 0s were lost in move, and make sure to preserve interval of zeros == mpa.no, then fill in mpa.yes,mpa.no to length of world. But also this is only for when mpa.no exists on both sides of MPA_next. If exactly to the edge of one reserve is lost, should shift down to next if statement
+  if(any(next_MPA==1)==FALSE & any(lost==1)==FALSE){
+    # these are the intervals that are left behind as world moves forward
+    lost <- MPA.current[1:(displaced-1)] 
+    # this is the last continuous chunk of numbers at the end of <lost>  
+    length_last <- rep(tail(rle(lost)$value,1), tail(rle(lost)$lengths,1))
+    # want to know how long <length_last> is so can make sure to get correct interval
+    L_int <- length(length_last)
+    # how many more zeros do we need before we start with mpa.yes again?
+    zero_add <- sum(mpa.no==0) - L_int - sum(next_MPA==0)
+    # this is what needs to be appended to <next_MPA>
+    new_MPA <- c(rep(0,zero_add),rep(c(mpa.yes,mpa.no),length.out=length(world)))
+    
+  }else{
+    
+    # IF FALSE (there are some 1s) then we only care about the last interval of the <next_MPA>, is that protected or not?
+    last_step = tail(next_MPA,1)
+    # IF <last_step>==1
+    if(last_step ==1){ 
+      #then how many 1s are at the end of the <next_MPA> section?
[…]

### 2014-02-27 Simluations/Aspatial_fast/Make.R

@@ -8,10 +8,10 @@ require(zoo)
 	
 # run analysis
 	# choose threshold or no threshold
-	model = "Thresh"	# "noThresh"; "Thresh"
+	model = "noThresh"	# "noThresh"; "Thresh"
 	
 	# if no threshold, choose MPA: "null", "cons", "fish"
-	MPA = "null"
+	MPA = "cons"
 
 # analysis
 	if(model=="noThresh") {sapply(c("Parameters_nothresh.R","Sim_noThresh.R"),source,.GlobalEnv)} else {

### 2014-02-27 Simluations/Aspatial_fast/Parameters.R

@@ -7,12 +7,12 @@ b=.5 #parameter for Laplace dispersal kernel
 R0=5 #growth parameter for recruitment
 K=100 #carrying capacity parameter for juvenile density dependence
 maxt = 150 # number of generations to run for initializing of harvest
-speeds = seq(0,0.5,0.02)
-
+speeds = seq(0,.5,by=0.02)
+harvests = seq(0,.13,by=0.01)
 f_ind = 1 #per capita reproductive rate
 
 patch = seq(0,1,by=step_size)
-world = seq(-.51,1.5, by = step_size)
+world = seq(-.51,4.5, by = step_size)   # to run the MPA versions, world has to be at least 400 steps (max distance between MPAs in "cons" run)
 w = length(world)
 
 cons.yes = rep(1,4*b/step_size)

### 2014-02-27 fish_figures (Autosaved).R

@@ -1,523 +1,0 @@
-library(fields)
-library(lattice)
-library(colorspace)
-library(RCurl)
-setwd("/Users/eleanorbrush/Desktop")
-source("/Users/eleanorbrush/Desktop/biomass_synergy.R")
-source("/Users/eleanorbrush/Desktop/critical_parameters.R")
-
-
-### parameters from Zhou
-r=3
-w=pi/4
-L=1
-
-####
-#the world
-step_size=0.01
-L=1
-world=matrix(seq(-L/2,L/2,by=step_size),nrow=1)
-emmaR0=5
-emmaK=100
-
-#####
-#### plotting parameters
-cex.lab=3
-cex.axis=2
-line=5
-lwd=5
-yaxs='i'
-xaxs='i'
-cex.lab=1.5
-cex.axis=1
-line=1
-oma=c(0,0,0,2)
-fig.width = 3	# for ecological applications, text needs to be at least 6 point font size
-
-##########
-##########
-##### wireframe of the equilibrium biomass as a function of c and h
-r=emmaR0
-K=emmaK
-sig2=1
-tol=.00001
-diff=.01
-cvals=seq(0,1,by=diff)
-hvals=seq(0,1,by=diff)
-
-
-# ebm=eqbiomass(cvals,hvals,r,K,sig2,tol=tol)
-load("/Users/eleanorbrush/Desktop/ebm.rdata")
-
-#which params to work with
-H=sum(apply(ebm,2,max)>tol)+2
-hvals=hvals[1:H]
-C=length(cvals)
-toplot_ebm=ebm[1:C,1:H]
-
-#breaks and colors
-lower=.001
-cuts=c(0,lower,seq(.5,max(ebm)+.5,by=.5))
-# s=5
-# maxgray=length(cuts)-2+s-1
-# vecgray=(s:maxgray)/maxgray
-# mycols=c('black',gray(vecgray))
-# colorbar.labels=as.character(round(cuts[-1],digits=1))
-# mycolorkey=list(col=mycols,at=cuts,height=1,lty=1)
-# myalpha=.7
-mycols=(sequential_hcl(length(cuts)-2,c.=c(0,0),l=c(30,100)))
-mycols=c('black',mycols)
-
-#axis labels
-numxlabs=11
-numylabs=3
-xdiff=.1
-ydiff=.2
[…]

### 2014-02-27 fish_figures.R

@@ -168,7 +168,7 @@ dev.off()
 	#mpas = read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv")
 # github copy
 	x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-02-20.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMi0yMC5jc3YiLCJleHBpcmVzIjoxMzkzNjIzMDUzfQ%3D%3D--d793f2abea78846c6b8535d73e76d5836207dea6")
-	maps = read.csv(textConnection(x))
+	mpas = read.csv(textConnection(x))
 cvals=unique(mpas$speed)
 hvals=unique(mpas$harvest)
 
@@ -210,7 +210,7 @@ myaxes=list(arrows=FALSE,
 	z=list(at=seq(0,max(ebm),by=250),labels=as.character(seq(0,max(ebm),by=250))))
 
 quartz()
-pdf(file='eqbiomass_fishmpa.pdf',width=fig.width)
+pdf(file='eqbiomass_fishmpa.pdf',width=fig.width,height=fig.width)
 
 # wireframe(toplot_ebm[C:1,H:1],xlab=list("Rate of environmental shift",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
 par(oma=oma)
@@ -220,7 +220,7 @@ image.plot(cvals,hvals,toplot_ebm,breaks=cuts,col=mycols,
 	yaxs=yaxs,xaxs=xaxs,axes=TRUE,
 	legend.shrink=1,legend.width=.1,zlim=range(cuts),
 	axis.args=list(at=myaxes$z$at,labels=myaxes$z$labels,cex.axis=cex.axis),
-	legend.args=list(text="Equilibrium biomass",cex=cex.lab,side=2,line=0.5,las=0),
[…]

### 2014-02-27 Writing/MS.tex
Updated MS.tex
fine edits, still working on MPA sims

@@
 
-When the climate changes, the habitat with suitable conditions in which organisms can survive and reproduce moves. 
-This change does not occur in isolation but rather appears on a background of other disturbances. 
-In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we studied an integrodifference model that explicitly included the mechanisms of dispersal and reproduction.
-If the viable habitat moves too quickly or harvesting pressure is too great, the population is driven extinct.
-We found the rates of harvesting and environmental shift required to allow the population to persist and and studied how these critical parameters depend on the growth rate and dispersal behavior of the population.
-We then measured the interaction between the stressors.
-The stressors interact nearly additively: we found very low positive synergy at those levels of the stressors that almost drive the population extinct.
-Positive synergy suggests that harvesting may aggravate the population's sensitivity to a shifting range.
-Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could mitigate the interaction between the two stressors.
+When the climate changes, so does the location of habitats suitable for an organism's survival and reproduction.
+This change does not occur in isolation but rather appears on a background of other disturbances, making the study of interactions between stressors important.
+In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we analyzed an integrodifference model that explicitly included the mechanisms of dispersal and reproduction.
+We found the critical rates of harvesting and climate velocity depend on the growth rate and dispersal kernel of the population.
+We measured the interaction between the stressors and found that the disturbances interact nearly additively, with low positive synergy only at the greatest harvest rates and climate velocity that almost drive the population extinct.
+We also introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could be effective management tools sat they mitigate the interaction between the two stressors.
+
 \hspace{10cm}
@@
 
-There are many stressors that can disturb an ecosystem.
-Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}, but less work has been done to measure the effects of multiple stressors and the interactions between them.
+Many stressors can disturb an ecosystem, and ecologists have quantified the consequences of a great deal of these of perturbations \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
+Less work, however, has measured the effects of multiple stressors and the interactions between them.
 If disturbances interact synergistically, a perturbation that has little effect when it occurs individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008,Nyeetal2013,Gurevitchetal2000}.
@@
 If disturbances interact antagonistically, on the other hand, the effects of multiple stressors may be less than that predicted by the individual effects of the stressors.
-Since disturbances rarely occur in isolation, it is important to measure the synergy between  disturbances in order to understand how a system will be affected by their presence and to understand when multiple disturbances will drive a population extinct  \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
+Since disturbances rarely occur in isolation, measuring the effects of multiple distrubances gives a better understanding of the likely impacts to the system \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
 
-Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008}.
-They therefore provide an important case study of how disturbances interact in their effects on biological populations.
-Further, understanding these interactions will be crucial to managing populations subjected to both of these disturbances.
-Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and they are projected to continue moving in the future \citep{Kelletal2005, Mackenzieetal2007}.
-Species that are likely to undergo or already undergoing shifts in range are also subject to harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
-Synergistic interactions between overfishing and temperature-driven range shifts have been found in empirical case studies \citep{Lingetal2009} and synergistic interactions between warming temperatures, harvesting and connectivity have been identified in microcosm experiments\citep{Moraetal2007}.
+Climate change and fishing, two of the largest human impacts on the ocean \citep{Halpernetal2008}, provide an important case study of how disturbances interact in their effects on biological populations.
+Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and projected to continue in the future \citep{Kelletal2005, Mackenzieetal2007}.
+These shifting species, and those likely to move in the future, are also subject to harvesting, among other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
+Previous empirical work has found synergistic interactions between overfishing and temperature-driven range shifts \citep{Lingetal2009} and microcosm experiments have identified synergistic interactions between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
 This empirical work underscores the importance of understanding how range shifts and harvesting interact.
 
-A common approach to predicting how populations will be distributed in future after climate-driven range shifts has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
+A common approach to predicting future population distributions has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
 These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to explain and predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
-Despite these models' widespread adoption, SDMs have frequently been criticized as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
+Despite these models' widespread adoption, many papers have criticized SDMs as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
 Recent work on range shifts has addressed some of these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}.
@@
 Work on the joint impacts of climate and fishing often considers climate fluctuations (large anomalies around the mean) rather than directional changes in climate \citep{WaltersParma1996, KingMcFarlane2006}.
-When the effects of climate-driven range shifts on fishing are considered, the models are typically case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}.
-These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, however these models are so complex that understanding the relative importance of particular drivers, disturbances, and interactions is difficult (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances).
-The degree of detail and case-specificity in these studies makes it difficult to draw general conclusions.
+When studies consider the effects of climate-driven range shifts on fishing, the models are typically case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}.
+These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, however these models are so complex that it makes understanding the relative importance of particular drivers, disturbances, and interactions difficult (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances).
 
 Here we extended a previously studied model of a fish population subject to climate-driven range shift by also considering harvesting pressure.
-Reproduction and dispersal, two mechanistic processes central to species' responses to climate and fishing, are explicitly included.
+We explicitly include reproduction and dispersal, two mechanistic processes central to species' responses to climate and fishing.
 Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We found the rate of harvesting and the rate of environmental shift that drive the population extinct and how the threshold harvesting level of depends on how quickly the range is shifting.
-We also found that climate-driven range shifts and fishing interact nearly additively, with very low positive synergy at more extreme levels of the stressors.
+We found the critical harvesting rate and climate velocity that drives the population extinct and how these critical rates depend on one another.
+We also found that climate-driven range shifts and fishing interact nearly additively, with low positive synergy at more extreme levels of the stressors.
 
-We also examined the effect of threshold harvesting rules and marine protected areas (MPAs) on species persistence.
-Protected areas have been suggested as a key form of climate insurance and stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
-MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat is shifting rapidly.
-We found that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
+We also examined the efficacy of two different types of management strategies: threshold harvesting rules and marine protected areas (MPAs).
+Previous work has suggested protected areas as a key form of climate insurance and stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
+MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat shifts rapidly.
+We found that threshold harvesting rules erase the interaction between harvesting rates and climate velocity and that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
 
@@
 
-We studied the dynamics of a fish population constrained to a single, one-dimensional habitat patch by their inability to reproduce outside of the patch.
-This viable habitat patch (here after `patch') is shifting at a fixed velocity and the fish at each point in space can be harvested.
-We first determined the climate velocity and harvesting rate that would drive the population extinct.
-We then measured the drop in biomass caused by range shifts, harvesting, and both stressors together in order to determine whether they interact synergistically.
-We finally implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to determine how these management strategies affect population persistence.
+We studied the dynamics of a fish population constrained to a single, one-dimensional habitat patch by their inability to reproduce outside of that area as introduced by \cite{ZhouKot2011}.
+This viable habitat patch (here after `patch') shifts at a fixed velocity and harvest occurs at each point in space along the entire one-dimensional world.
+We first determined the climate velocity and harvesting rate that would drive the population extinct (hereafter the critical harvesting rate and climate velocity), and then measured synergy by calculating the drop in biomass caused by each stressor both individually and together.
+We finally add marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to determine how these management strategies affect population persistence.
 
@@
 We extend this model by additionally subjecting the adults to harvesting before they produce offspring so that only a proportion of the fish survive to reproduce.
-These processes-- recruitment, harvesting, and dispersal-- are incorporated into an integrodifference model to describe how the population changes over time.
+We incorporate these processes-- recruitment, harvesting, and dispersal-- into an integrodifference model to describe how the population changes over time.
 If $n_t(x)$
@@
 We provide a list of variables and functions in Table \ref{variables}.
-We chose to use a Beverton-Holt recruitment function,
+We use a Beverton-Holt recruitment function,
 
-\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}.\]
+\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
 
-\noindent Regardless of the exact functional form of the recruitment function, the critical parameter in determining population persistence is how quickly recruitment increases when the population 
-size is near (but above) $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$.
+\noindent but regardless of the exact functional form of the recruitment function, the critical parameter in determining population persistence is how quickly recruitment increases when the population size is near (but above) $0$, which is equivalent to the intrinsic growth rate $R_0=f'(0)$.
 Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.
-In our analyses, as in \citep{Latore:1998fk}, we used the separable Gaussian kernel given by
+In our analyses, as in \cite{Latore:1998fk}, we used the separable Gaussian kernel given by
 
@@
 
-\noindent To derive analytical expressions, we approximated the kernel, as described in the Appendix.
-Analytical results for a separable sinusoidal kernel are also described in the Appendix.
-We used
-simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
+\noindent To derive analytical expressions, we approximated the kernel, as described in the Appendix. %section number necessary?
+Analytical results for a separable sinusoidal kernel are also described in the Appendix.  %awkward
+We used simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.  % Why did we choose a Laplace dispersal kernel for simulations?
 
-At equilibrium, the population will be described by a traveling wave, 
-where the density of fish at a given point in space will change but the density of fish at a location relative to the 
-shifting  patch will not. We sought to describe how the population is distributed over the viable 
-patch as it shifts through the world in order to study the size of the population at equilibrium and whether or not the population could persist.  The traveling wave $n^*$ must 
-satisfy
+At equilibrium, a traveling wave will describe the population, where the density of fish at a given point in space will change but the density of fish at a location relative to the shifting patch will not.
+We sought to describe  the distribution of the population over the viable patch as it shifts through the world in order to study the size of the population at equilibrium and whether or not the population could persist.
+The traveling wave $n^*$ must satisfy
 
@@
 \subsection{Persistence }
-One possible equilibrium traveling wave that solves Equation (\ref{traveling_pulse}) is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
-{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.  If a population is to persist, it must be able to avoid extinction and 
-grow even when it is small.   A small population can be thought of as a perturbation to the trivial traveling pulse. If the trivial 
-pulse is stable, the system will return to the trivial pulse even after the introduction 
-of a small population.
-If the trivial pulse is unstable, a small population may increase and form a stable population.
+One possible equilibrium traveling wave that solves Equation (\ref{traveling_pulse}) is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.
+If a population persists, it must be able to avoid extinction and grow even when small.
+We can think of a small population as a perturbation to the trivial traveling pulse.
+If the trivial pulse is stable, the system will return to the trivial pulse even after the introduction of a small population.
+If the trivial pulse is unstable, a small population may increase and form a persistent population.
 Population persistence is therefore equivalent to the trivial traveling pulse being an unstable equilibrium.
 
-If the population is harvested at low enough levels and the environment shifts slowly enough, the population will be able to persist.
-There are threshold values of the harvesting rate $h$ and the rate of environmental shift $c$ such that if the parameters are increased beyond these values, the population will be driven extinct.
-We found these critical parameters, $h^*$, and  $c^*$, by finding the parameters that make the trivial pulse unstable.
-Details are provided in Appendix \ref{stab}.
+If we harvest the population at low enough levels and the climate velocity is slow enough, the population will be able to persist.
+There exists threshold values of the harvesting rate $h$ and a climate velocity $c$ such that if we increase parameters beyond these values, we drive the population extinct.
+We found these critical parameters, $h^*$, and  $c^*$, by finding the parameters that make the trivial pulse unstable (See Appendix \ref{stab}).
 
-For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment-- the length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
-$).
+For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses $\langle d \rangle$ and the intrinsic growth rate $R_0$; properties of the environment-- the length of the viable patch $L$ and how quickly the environment shifts $c$; and the harvesting rate $h$.
 The population biomass at equilibrium depends on the function form of recruitment, but population persistence only depends on the intrinsic growth rate $R_0$.
@@
 
-A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
+We derive a similar expression for a sinusoidal kernel in the Appendix [{\bf REF?}].  For both kernels, we can approximate the critical harvesting proportion by a function that looks like 
 \begin{equation*}
@@
 \citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct.
-In order to quantify whether the two stressors are interacting additively, synergistically, or antagonistically, we found the total biomass of the population when it reached an equilibrium traveling pulse and
-compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.
+In order to quantify whether the two stressors interact additively, synergistically, or antagonistically, we found the total biomass of the population when it reached an equilibrium traveling pulse and compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.
 For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
@@
 
-\noindent If the stressors aggravate each other, the effect of both stressors is worse than would be expected from 
-considering either stressor individually, and synergy is positive. If the stressors alleviate each other, the effect 
-of both stressors is better than would be expected from considering either stressor individually, and synergy is 
-negative. If the effect of both stressors is exactly as expected from considering either stressor individually, 
-there is no interaction and no synergy.
+\noindent If the stressors aggravate each other, synergy is positive, and the effect of both stressors is worse than would we expect from considering either stressor individually.
+If the stressors alleviate each other, the synergy is negative, and the effect of both stressors is better than we expecte from considering either stressor individually.
+If the effect of both stressors is exactly as expected from considering either stressor individually, there is no interaction and no synergy.
 
@@
 
-We used simulations to extend the basic integrodifference model in two ways that make it analytically 
-intractable. First, we examined the sensitivity of the model to choice of dispersal kernel by using the Laplace 
-dispersal kernel,
+We used simulations to extend the basic integrodifference model in two ways that make it analytically intractable.
+First, we examined the sensitivity of the model to choice of dispersal kernel by using the Laplace dispersal kernel,
 
@@
 Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
-Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
-specify a recruitment function.
+Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must specify a recruitment function.
 Again, we chose to use a Beverton-Holt function.
-In the first generation, we seeded the world with 50 individuals at a single point, as in \citep{ZhouKot2011}.
+In the first generation, we seeded the world with 50 individuals at a single point, as in \cite{ZhouKot2011}.
 We first ran through 150 generations in order for the population to reach equilibrium without harvesting or climate shift.
-We then added harvesting pressure, allowed the population to again reach equilibrium, and finally added climate change by moving the viable patch.
-Equilibrium biomass is calculated as the mean biomass of 300 time steps once the difference in biomass between time step $t$ and $t+1$ was no greater than $0.1$.
+We then added harvesting pressure, allowed the population to again reach equilibrium (150 generations), and finally added climate change by moving the viable patch.
+We calculate equilibrium biomass as the mean biomass of 300 time steps once the difference in biomass between time step $t$ and $t+1$ was no greater than $0.1$.
 
-We added harvesting pressure by harvesting a constant proportion of the population, in order to confirm our analytical results.
-We then evaluated the effect of a threshold harvest rule and marine protected
-areas (MPAs).  With a threshold rule, we evaluated the population at each point in space to determine how 
-much harvesting should occur. If the population abundance was below the designated threshold, no 
-harvesting occurred. If the population exceeded the threshold, then a proportion of the `surplus' individuals 
-were harvested.
+n, in order to confirm our analytical results, we first added harvesting pressure by harvesting a constant proportion of the population.
+We then evaluated the effect of a threshold harvest rule and marine protected areas (MPAs).
+With a threshold rule, we evaluated the population at each point in space to determine how much harvesting should occur.
+If the population abundance was below the designated threshold, no harvesting occurred.
+If the population exceeded the threshold, then we harvested all the `surplus' individuals.
 
@@
 
-Networks of MPAs were introduced into our simulations by designating segments of space in which harvesting was forbidden (i.e.
-harvesting rates were equal to $0$).
+We introduce networks of MPAs into our simulations by designating segments of space where the harvesting rate was equal to $0$.
 Conservation-oriented MPAs, are frequently large and rarely part of a larger network of reserves \citep{HastingsBotsford2003}.
@@
 \subsection{Interactions Between Stressors }
-We find the critical climate velocity and harvest  rate to be inversely related: as the harvesting rate $h$ increases, 
-the critical climate velocity $c^*$ decreases as the environment must move more slowly to accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
-environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}). This 
-means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
-the environment starts changing.
+We find the critical climate velocity and harvest  rate to be inversely related: as the harvesting rate $h$ increases, the critical climate velocity $c^*$ decreases, as the environment must move more slowly to accommodate the population growing more slowly (Figure \ref{baseline}).
+% this seems like a backwards way to say it.
+Instead of "as the environment must move more slowly to accommodate..." it seems like it should be "as the population can only survive in increasingly stationary habitats"..
+But "increasingly stationary" is not great...
+Ideas?
+Conversely, as the rate of environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}).
+This means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if the environment starts changing.
 When the climate velocity or harvesting pressure exceed their critical rates ($h^*, c^*$ respectively), the biomass of the population at equilibrium will be equal to $0$.
-Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure
-increases or the environmental shifts more quickly (Figure \ref{baseline}).
+Before the stressors reaches those thresholds, the equilibrium biomass of the population decreases as either the harvesting pressure increases or the environmental shifts more quickly (Figure \ref{baseline}).
 Our simulations confirm the analytical results with the critical speed $c^*$ declining as the critical harvest rate $h^*$ increases and vice versa (Figure \ref{nomang}).
 
-It is always the case that increasing the intrinsic growth rate, $R_0$, of the population increases the critical 
-speed $c^*$ and the critical harvesting rate $h^*$, since a population that grows more quickly can recover 
-more quickly from losses caused by these disturbances. However, whether or not dispersing farther is better 
-depends on how quickly the environment is shifting (Figure \ref{baseline}). When the environment is shifting 
-slowly, dispersing farther is detrimental since many larvae will disperse too far away from the viable patch. 
-When the environment is shifting quickly, on the other hand, dispersing farther can help the population persist 
-because some larvae will disperse into the space that will become viable shortly in the future.
-This affects the critical harvesting rate: at a low rate of environmental shift, populations that disperse less can be harvested more severely than those that disperse further, whereas at a high rate of environmental shift, populations that disperse further can be harvested more severely.
+It is always the case that increasing the intrinsic growth rate, $R_0$, of the population increases the critical climate velocity $c^*$ and the critical harvesting rate $h^*$, since a population that grows more quickly can recover more quickly from losses caused by these disturbances.
+However, whether or not dispersing farther is better depends on how quickly the environment is shifting (Figure \ref{baseline}).
+When the environment is shifting slowly, dispersing farther is detrimental since many larvae will disperse too far away from the viable patch.
+When the environment is shifting quickly, on the other hand, dispersing farther can help the population persist because some larvae will disperse into the space that will become viable shortly in the future.
+This affects the critical harvesting rate: at a low rate of environmental shift, we can more severely harvest populations that have a shorter dispersal distance than those that disperse farther, whereas at a high rate of environmental shift, we can more aggressively harvest populations that disperse farther.
 
-We found very low levels of positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
-{Synergy}).
-Where there is positive synergy, a doubly stressed population loses more biomass than would be predicted from either stressor individually.
-The stressors interact most strongly when they are both high, shortly before they drive the population extinct.
-However, the excess loss in biomass is extremely low, making it difficult to distinguish positive synergy from additive interactions.
+We found low levels of positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref{Synergy}).
+Where positive synergy exists, a doubly stressed population loses more biomass than we would predict from either stressor individually.
+The stressors interact most strongly at high values, shortly before they drive the population extinct.
+However, the excess loss in biomass is low, making it difficult to distinguish positive synergy from additive interactions.
+% this is my awkward sentence, any ideas?
 We found similar analytical results for a sinusoidal dispersal kernel, which indicates that this result is robust to changes in the dispersal kernel.
@@
 
-Without any management strategies, we found that when the population is harvested more severely, slower rates of environmental shift will suffice to drive the population extinct.
-However, when thresholds are in place, a small population can always escape harvesting pressure and the critical rate of environmental shift $c^*$ no longer depends on the harvesting rate (Figure \ref{management}).
+Without any management strategies, we found that the more severely we harvest the population, the slower the rate of environmental shift will suffice to drive the population extinct.
+However, when we put thresholds in place, a small population can always escape harvesting pressure and the critical rate of environmental shift $c^*$ no longer depends on the harvesting rate (Figure \ref{management}).
 In other words, as long as there is some threshold below which harvesting is not allowed, there is a constant critical rate of environmental shift that only depends on the growth rate, length of the viable patch, and average dispersal distance.
@@
 We also examined the effect of marine protected  areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and  climate change parameters where the fish population could survive.
-We found both MPA strategies examined (many small versus few large), allowed the population was able to withstand combinations of higher climate velocities and harvesting rates.
-However MPAs did not increase the critical climate velocity (Figure \ref{management}).
+We found that with either type of MPA strategies examined (many small versus few large), the population withstood combinations of higher climate velocities and harvesting rates, although the critical climate velocity $c^*$ was unchanged (Figure \ref{management}).
 We also found that the spacing and size of the MPAs changed population dynamics.
-MPAs that were spaced more than 1 average dispersal distance apart resulted in large oscillations of population biomass at low climate velocities.
-As climate velocities increase, the mean population abundance declines but the population experiences less extreme oscillations in abundance, which results in the population being bounded further from possible extinction in a stochastic environment.
+MPAs spaced more than one average dispersal distance apart resulted in large oscillations of population biomass at low climate velocities relative to small, closely spaced, MPAs.
+For both of these MPA strategies we find that as climate velocities increase, the mean population abundance declines but the population experiences less extreme oscillations in abundance, which results in a population bounded farther from possible extinction in a stochastic environment.
+Additionally, large MPAs were able to increase equilibrium biomass under relatively high harvest and intermediate speeds relative to a population harvested at the same rate, but at a slower speed.
+This effect disappears at faster climate velocities and is not present in the other, many-small MPA, strategy we tested.
 
@@
 
-Knowing whether two disturbances interact in their impacts for a given population is important for management.
+Understanding interactions among disturbances will help to design management for populations subjected to these disturbances.
 The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.
@@
 For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
-More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
+We also found a negative relationship between the critical harvesting rate and the rate of environmental shift.
 That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
@@
 This indicates that this result is robust to changes in the dispersal kernel.
-We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.
+We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \cite{Crainetal2008, DarlingCote2008,Nyeetal2013}.
 We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
@@
 
-The fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
-This means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
-However, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
+Worryingly, we find the highest synergy in those populations whose persistence is most tenuous.
+This means that harvesting levels or climate velocity that are sustainable individually together can drive a population to extinction.
+However the drop in biomass caused by both stressors was never much higher than the null prediction, i.e.
 synergistic effects were quite small.
-Synergy between harvesting and climate changes has been identified in experimental populations \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
-Additionally, in the experimental populations, synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
-While we did find (very) low levels of positive synergy, we did not find as much as might be predicted from these empirical studies.
+%we need a transition here, feels abrupt
+Experimental work has identified synergy between harvesting and climate changes \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
+Additionally, experimental populations have identified synergy between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
+While we did find (very) low levels of positive synergy, we did not find as much as predicted from these empirical studies.
 However, these previous results are not directly comparable to ours because they focus on different aspects of climate change, e.g.
@@
 
-We also examined whether frequently recommended management approaches ensure species persistence.
+We also examined whether frequently recommended management approaches, MPAs and harvest control rules, ensure species persistence.
 With these management strategies we found increases in the population's biomass at equilibrium and an improved ability to persist.
-Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
-Our results, that spatial management increased the maximum harvesting rate at which the population could survive  and improved the populations ability to persist at a slightly higher climate velocity conditional on a given harvesting rate [NEW FIGURE], support the idea that MPAs could be used to reduce the impact of harvesting and help to ameliorate the effects of climate velocity.
-However the spacing and size of the MPAs matter: few, large MPAs caused increased variability at low climate velocities while many smaller MPAs maintained a population that was bounded further from extinction.
+Previous work has advanced protected areas as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
+Our results show that both threshold and MPAs increase the equilibrium biomass at a given climate velocity, which support their use as a tool to ameliorate the effect of climate velocity.
+However, for MPAs, the details mater: few, large MPAs caused increased variability at low climate velocities while many smaller MPAs maintained a population bounded farther from extinction.
+Finally, with sufficiently high harvesting pressure, few, large MPAs rescued populations at intermediate speeds.
+With intermediate speeds, the population was able to reach a protected area fast enough to avoid extinction, and the protected area was large enough to allow a partial rebuilding of the population before it moved out the other side.
+However this effect disappears as speed continues to increase, suggesting that understanding the relationship between climate velocity, dispersal distance and reproductive rate are important parameters in designing management strategies effective under both climate change and harvesting pressure.
 
- When we applied harvesting thresholds, we found that it alleviates interactions between the two stressors.
+Unlike MPAs, thresholds are not explicitly spatial, and we found that this harvest control rule alleviates interactions between the two stressors.
 Thresholds have this effect as the management approach effectively prevents harvesting of the leading edge, which allows colonization to occur as if these individuals were moving into un-fished areas.
-It's interesting to note that novel, low abundance species are commonly unregulated in fisheries systems; in order to decouple the additive effects of harvest and climate change, this paradigm would have to be reversed: no new species would be allowed to harvest until they had become established.
-These results highlight that while management strategies only change harvesting practices, they can effect the way stressors interact and help to better implement harvesting rules and place protected areas.
+It's interesting to note that novel, low abundance species are commonly unregulated in fisheries systems; so in order to decouple the additive effects of harvest and climate change, management would have to reverse this paradigm by allowing no harvest of new species until they had become established.
+These results highlight that while management strategies only change harvesting practices, they can effect the way stressors interact and so understanding these interacts help to better design harvesting rules and place protected areas.
 
 The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
-However, it  ignores many of the complexities present in marine fisheries.
-We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
-However, we expect qualitatively similar results.
+However, this simplistic approach requires that we ignore complexities known to be present in marine fisheries.
+For example, we do not include Allee effects, so that even if the population shrank to low levels it was possible for it to persist over time.
+However, with Alee effects we expect qualitatively similar results.
 An Allee effect would make it harder for populations to colonize new areas and add a threshold below which fishing drives the population to extinction.
@@
 We also did not include age structure in our model.
-The effects of both harvesting and climate change may be different across different age classes and may destabilize the system in complicated ways, including resonance \citep{Botsfordetal2011, Planqueetal2010}; including this level of complexity is left for future work.
+The effects of both harvesting and climate change may be different across different age classes and may destabilize the system in complicated ways, including resonance \citep{Botsfordetal2011, Planqueetal2010}; and we leave this additional complexity for future work.
 Similarly, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
 Besides these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
-By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
+By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000} and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
 
 Using a simple mechanistic model like the one we present here provides a useful framework for incorporating additional ecological complexities which can mediate species persistence under multiple disturbances.
-Exploring how species interactions, age structure, and additional disturbances (e.g.
+Using this modeling framework as a starting point, we believe exploring how species interactions, age structure, and additional disturbances (e.g.
 physiological response to temperature) affect population viability will improve our predictions and help us to understand whether species will persist under predicted climate and harvesting regimes.
-Finally, this work can help make general predictions as to whether specific life histories are likely to be selected over others as harvesting and range shifts increase and highlights the importance of considering stressors in combination as outcomes can deviate substantially from what would be predicted in isolation.
+Finally, this work can help make general predictions as to whether specific life histories offer selective advantages over others as harvesting and range shifts increase and highlights the importance of considering stressors in combination as outcomes can deviate from what we would predict in isolation.
 This is especially true for management strategies which may result in unanticipated effects such as large fluctuations associated with big, distant MPAs shown here.

### 2014-02-28 Writing/MS.tex
last few of malins comments in the discussion

--- 2014-02-21 Writing/MS.tex
@@
 the critical climate velocity $c^*$ decreases as the environment must move more slowly to accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
-environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}). This 
-means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
-the environment starts changing.
+environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}).
+This means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if the environment starts changing.
 When the climate velocity or harvesting pressure exceed their critical rates ($h^*, c^*$ respectively), the biomass of the population at equilibrium will be equal to $0$.
-Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure
-increases or the environmental shifts more quickly (Figure \ref{baseline}).
+Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure increases or the environmental shifts more quickly (Figure \ref{baseline}).
 Our simulations confirm the analytical results with the critical speed $c^*$ declining as the critical harvest rate $h^*$ increases and vice versa (Figure \ref{nomang}).
 
-It is always the case that increasing the intrinsic growth rate, $R_0$, of the population increases the critical 
-speed $c^*$ and the critical harvesting rate $h^*$, since a population that grows more quickly can recover 
-more quickly from losses caused by these disturbances. However, whether or not dispersing farther is better 
-depends on how quickly the environment is shifting (Figure \ref{baseline}). When the environment is shifting 
-slowly, dispersing farther is detrimental since many larvae will disperse too far away from the viable patch. 
-When the environment is shifting quickly, on the other hand, dispersing farther can help the population persist 
-because some larvae will disperse into the space that will become viable shortly in the future.
+It is always the case that increasing the intrinsic growth rate, $R_0$, of the population increases the critical speed $c^*$ and the critical harvesting rate $h^*$, since a population that grows more quickly can recover more quickly from losses caused by these disturbances.
+However, whether or not dispersing farther is better depends on how quickly the environment is shifting (Figure \ref{baseline}).
+When the environment is shifting slowly, dispersing farther is detrimental since many larvae will disperse too far away from the viable patch.
+When the environment is shifting quickly, on the other hand, dispersing farther can help the population persist because some larvae will disperse into the space that will become viable shortly in the future.
 This affects the critical harvesting rate: at a low rate of environmental shift, populations that disperse less can be harvested more severely than those that disperse further, whereas at a high rate of environmental shift, populations that disperse further can be harvested more severely.
@@
 Without any management strategies, we found that when the population is harvested more severely, slower rates of environmental shift will suffice to drive the population extinct.
-However, when thresholds are in place, a small population can always escape harvesting pressure and the critical rate of environmental shift $c^*$ no longer depends on the harvesting rate (Figure \ref{management}).
+However, when thresholds are in place, the critical rate of environmental shift $c^*$ no longer depends on the harvesting rate (Figure \ref{management}).
 In other words, as long as there is some threshold below which harvesting is not allowed, there is a constant critical rate of environmental shift that only depends on the growth rate, length of the viable patch, and average dispersal distance.
+This is because, with thresholds in place, a small portion of the population always escapes harvesting pressure and only the size, not the growth rate, of this small population depends on the level of threshold implemented.
+This small population can then track the shifting environment as well as a larger one.
 
 We also examined the effect of marine protected  areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and  climate change parameters where the fish population could survive.
-We found both MPA strategies examined (many small versus few large), allowed the population was able to withstand combinations of higher climate velocities and harvesting rates.
-However MPAs did not increase the critical climate velocity (Figure \ref{management}).
-We also found that the spacing and size of the MPAs changed population dynamics.
-MPAs that were spaced more than 1 average dispersal distance apart resulted in large oscillations of population biomass at low climate velocities.
-As climate velocities increase, the mean population abundance declines but the population experiences less extreme oscillations in abundance, which results in the population being bounded further from possible extinction in a stochastic environment.
+With MPAs in place, the population can withstand a slightly higher harvesting rate in those parts of the world where harvesting is allowed (Figure \ref{management}).
+More surprisingly, perhaps, implementing MPAs slightly increased the critical rate of environmental shift (Figure \ref{management}).
+Even though MPAs are meant to protect marine populations from harvesting pressure, by providing refuges in which their intrinsic growth rates can increase back to the normal level, MPAs allow marine populations to withstand higher rates of environmental shift.
 
@@
 That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
-The curved line separating parameters that will allow the population to persist from those that won't is an indication of an interaction between the stressors.
-%say more about this line of evidence for an interaction
+This is an indication of an interaction between the stressors.
 
@@
 The fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
-This means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
+This is further evidence that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
 However, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
 synergistic effects were quite small.
-Synergy between harvesting and climate changes has been identified in experimental populations \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
-Additionally, in the experimental populations, synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
-While we did find (very) low levels of positive synergy, we did not find as much as might be predicted from these empirical studies.
+Synergy between harvesting and the effects of climate change has been identified in experimental populations \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
+In the experimental populations, synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
+While we did find (very low levels of) positive synergy, we did not find as much as might be predicted from these empirical studies.
 However, these previous results are not directly comparable to ours because they focus on different aspects of climate change, e.g.
-warming temperature \citep{Moraetal2007,Kirby:2009fk} or a more variable climate \citep{Planque:2010uq}.
+warming temperature \citep{Moraetal2007,Kirby:2009fk} or a more variable climate \citep{Planque:2010uq}, rather than range shifts.
 Additionally, while we can isolate the affects of climate shift and harvesting in our simple analytical model, there are other forces acting on real populations that may produce the observed synergistic effects.
 
-%e.g.?!?!
-
-%Nevertheless, our results generally agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
-
 Our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
-As shown in Figure \ref{baseline}, species with a higher reproductive rate and a longer average dispersal distance will better track a high rate of environmental shift relative to a species that has a low reproductive rate and short dispersal distance.
-The finding that a higher reproductive rate can sustain higher climate velocities and harvesting rates is intuitive, especially because harvesting rate and reproductive rate cancel each other out.
-However it is worth pointing out that a higher reproductive rate can be generated either by shorter generation times or higher fecundity.
-Finding that species with shorter generation times can better keep up with shifts in climate is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species \citep{Perryetal2005}.
-While higher reproductive rates improved a population's ability to persist, there was a tradeoff in increasing dispersal distances.
-At low speeds, we found that a short dispersal dispersal distance improved the maximum harvesting rate a population could sustain while at higher speeds a longer dispersal distance improved the maximum climate velocity in which the population could persist (Figure \ref{baseline}).
-This tradeoff is due to the proportion of dispersing  offspring at time step $t$ which lands within the patch at time step $t + 1$.
-When climate is shifting slowly, a large dispersal distance sends most offspring ahead of the patch, while with faster climate velocities a long dispersal distance allows the population to make it to the new patch (Figure \ref{baseline}).
-Thus climate velocity will selectively favor species with dispersal distances best matched to the rate of shift.
+Species with higher intrinsic growth rates will always better track a shifting environment than species with lower intrinsic growth rates.
+This is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non-shifting species (smaller body sizes, faster maturation, smaller sizes at maturity) \citep{Perryetal2005}.
+However, as shown in Figure \ref{baseline}, having a longer average dispersal distance translates into a better ability to track environmental shift only if the environment is shifting quickly.
 
 We also examined whether frequently recommended management approaches ensure species persistence.
-With these management strategies we found increases in the population's biomass at equilibrium and an improved ability to persist.
+We found increases in the population's biomass at equilibrium and an improved ability to persist.
+Harvesting thresholds, are already implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors by allowing a small portion of the population to escape harvesting pressure and successfully track the shifting environment.
 Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
-Our results, that spatial management increased the maximum harvesting rate at which the population could survive  and improved the populations ability to persist at a slightly higher climate velocity conditional on a given harvesting rate [NEW FIGURE], support the idea that MPAs could be used to reduce the impact of harvesting and help to ameliorate the effects of climate velocity.
-However the spacing and size of the MPAs matter: few, large MPAs caused increased variability at low climate velocities while many smaller MPAs maintained a population that was bounded further from extinction.
-
- When we applied harvesting thresholds, we found that it alleviates interactions between the two stressors.
-Thresholds have this effect as the management approach effectively prevents harvesting of the leading edge, which allows colonization to occur as if these individuals were moving into un-fished areas.
-It's interesting to note that novel, low abundance species are commonly unregulated in fisheries systems; in order to decouple the additive effects of harvest and climate change, this paradigm would have to be reversed: no new species would be allowed to harvest until they had become established.
-These results highlight that while management strategies only change harvesting practices, they can effect the way stressors interact and help to better implement harvesting rules and place protected areas.
+Our results, that spatial management increased the maximum harvesting rate at which the population could survive, support the idea that MPAs could be used to reduce the impact of harvesting.
+We also found evidence that MPAs increased the climate velocity under which the population could persist.
+MPAs help populations to persist by providing regions of space in which to grow and recover before being re-subjected to harvesting.
+While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.
+This is encouraging evidence that a single set of of management practices may help to protect marine populations from both harvesting and climate change.
 
@@
 However, it  ignores many of the complexities present in marine fisheries.
-We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
-However, we expect qualitatively similar results.
-An Allee effect would make it harder for populations to colonize new areas and add a threshold below which fishing drives the population to extinction.
-Thus an Allee effect would change lower the critical harvest rates and climate velocity, but we do not expect the additive nature of the interaction between climate and harvesting to change.
-We also did not include age structure in our model.
-The effects of both harvesting and climate change may be different across different age classes and may destabilize the system in complicated ways, including resonance \citep{Botsfordetal2011, Planqueetal2010}; including this level of complexity is left for future work.
+First, we could consider additional disturbances, like pollution, disease, physiological response to temperature.
+However, if the main effect of these disturbances is to increase mortality rates than we would expect that they would interact with climate shift in a way similar to harvesting pressure.
+We did not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
+However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect.
+We also did not include age structure in our model, which has previously been shown to affect a population's resistance to harvesting and climate change \citep{Botsfordetal2001}.
+The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work.
 Similarly, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
 Besides these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
-By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000}and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
+By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000} and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
 
-Using a simple mechanistic model like the one we present here provides a useful framework for incorporating additional ecological complexities which can mediate species persistence under multiple disturbances.
-Exploring how species interactions, age structure, and additional disturbances (e.g.
-physiological response to temperature) affect population viability will improve our predictions and help us to understand whether species will persist under predicted climate and harvesting regimes.
-Finally, this work can help make general predictions as to whether specific life histories are likely to be selected over others as harvesting and range shifts increase and highlights the importance of considering stressors in combination as outcomes can deviate substantially from what would be predicted in isolation.
-This is especially true for management strategies which may result in unanticipated effects such as large fluctuations associated with big, distant MPAs shown here.
+Using a simple mechanistic model like the one we present here provides a useful framework for understanding when it is or is not necessary to consider the interactions between multiple disturbances and for incorporating additional ecological complexities which can mediate species persistence under those disturbances.
+Exploring how species interactions, age structure, and additional disturbances affect population viability will improve our predictions and help us to understand whether species will persist under predicted climate and harvesting regimes.
+Finally, this work can help to make general predictions as to whether specific life histories are likely to be selected over others as harvesting and range shifts increase.
 
@@
 These results are from an approximated Gaussian dispersal kernel with $L=1$.
-(b)
-The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+(b) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
 These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.

### 2014-02-28 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd

@@
 # for synced online github data
-   x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxMy0xMi0wOS5jc3YiLCJleHBpcmVzIjoxMzkyODI5NjEzfQ%3D%3D--9fbc7c170a93836981fa63448bf378fbbdb13905")
+   x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2014-02-20.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxNC0wMi0yMC5jc3YiLCJleHBpcmVzIjoxMzkzNjIzMjk4fQ%3D%3D--b299cd8c20eb28a2d626a3d8d467685b75986fbf")
   sim = read.csv(textConnection(x))
 
-  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxMy0xMi0wOS5jc3YiLCJleHBpcmVzIjoxMzkyODI5NjQzfQ%3D%3D--a432e4025f1b9042246c8a3d6b49af3464a03546")
+  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-02-20.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMi0yMC5jc3YiLCJleHBpcmVzIjoxMzkzNjIzMDUzfQ%3D%3D--d793f2abea78846c6b8535d73e76d5836207dea6")
   fishmpa = read.csv(textConnection(fish))
 
-  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-02-10.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMi0xMC5jc3YiLCJleHBpcmVzIjoxMzkyODI5NjYwfQ%3D%3D--00004e8d0186aa3a6f7f52bb264a406f147d1abf")
+  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-02-20.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMi0yMC5jc3YiLCJleHBpcmVzIjoxMzkzNjIzMDEzfQ%3D%3D--9f521e92c8abdd1c4a8a547576e661ea41d342d6")
   consmpa = read.csv(textConnection(cons))
@@
 
-```{r}
-
-```
+UPDATE, the problem persists.
+Malin's idea is to take a time step where the MPA and the range line up at the 'northern' edge.
+Because the world is only a patch length long, need to figure out when the MPA is at the northern-most edge.

### 2014-02-28 Writing/Appendix.tex

--- 2014-02-28 Writing/MS.tex
@@
 \usepackage{lineno}
-\linenumbers
-\title{Persistence of marine populations under climate and fishing}
-\author{Emma Fuller, Eleanor Brush, Malin Pinsky}
-\date{}
-
 \begin{document}
-\maketitle
-
-\begin{spacing}{1.9}
-\begin{flushleft}
-
-\section{Abstract}
-
-When the climate changes, the habitat with suitable conditions in which organisms can survive and reproduce moves. 
-This change does not occur in isolation but rather appears on a background of other disturbances. 
-In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we studied an integrodifference model that explicitly included the mechanisms of dispersal and reproduction.
-If the viable habitat moves too quickly or harvesting pressure is too great, the population is driven extinct.
-We found the rates of harvesting and environmental shift required to allow the population to persist and and studied how these critical parameters depend on the growth rate and dispersal behavior of the population.
-We then measured the interaction between the stressors.
-The stressors interact nearly additively: we found very low positive synergy at those levels of the stressors that almost drive the population extinct.
-Positive synergy suggests that harvesting may aggravate the population's sensitivity to a shifting range.
-Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could mitigate the interaction between the two stressors.
-\hspace{10cm}
-
-\noindent {\bf Keywords:} Climate change, fishing, integrodifference model, synergy, multiple disturbances
-
-\section{Introduction}
-
-There are many stressors that can disturb an ecosystem.
-Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}, but less work has been done to measure the effects of multiple stressors and the interactions between them.
-If disturbances interact synergistically, a perturbation that has little effect when it occurs individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008,Nyeetal2013,Gurevitchetal2000}.
-In the most extreme (and worrying) cases, synergistic interactions between multiple stressors will drive a population extinct even though it could persist in the face of any single stressor (e.g.
-\citet{Pelletieretal2006}).
-If disturbances interact antagonistically, on the other hand, the effects of multiple stressors may be less than that predicted by the individual effects of the stressors.
-Since disturbances rarely occur in isolation, it is important to measure the synergy between  disturbances in order to understand how a system will be affected by their presence and to understand when multiple disturbances will drive a population extinct  \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
-
-Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008}.
-They therefore provide an important case study of how disturbances interact in their effects on biological populations.
-Further, understanding these interactions will be crucial to managing populations subjected to both of these disturbances.
-Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and they are projected to continue moving in the future \citep{Kelletal2005, Mackenzieetal2007}.
-Species that are likely to undergo or already undergoing shifts in range are also subject to harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
-Synergistic interactions between overfishing and temperature-driven range shifts have been found in empirical case studies \citep{Lingetal2009} and synergistic interactions between warming temperatures, harvesting and connectivity have been identified in microcosm experiments\citep{Moraetal2007}.
-This empirical work underscores the importance of understanding how range shifts and harvesting interact.
-
-A common approach to predicting how populations will be distributed in future after climate-driven range shifts has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
-These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to explain and predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
-Despite these models' widespread adoption, SDMs have frequently been criticized as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
-Recent work on range shifts has addressed some of these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}.
-However these models only address one disturbance, climate-driven range shifts.
-
-Work on the joint impacts of climate and fishing often considers climate fluctuations (large anomalies around the mean) rather than directional changes in climate \citep{WaltersParma1996, KingMcFarlane2006}.
-When the effects of climate-driven range shifts on fishing are considered, the models are typically case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}.
-These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, however these models are so complex that understanding the relative importance of particular drivers, disturbances, and interactions is difficult (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances).
-The degree of detail and case-specificity in these studies makes it difficult to draw general conclusions.
-
-Here we extended a previously studied model of a fish population subject to climate-driven range shift by also considering harvesting pressure.
-Reproduction and dispersal, two mechanistic processes central to species' responses to climate and fishing, are explicitly included.
-Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We found the rate of harvesting and the rate of environmental shift that drive the population extinct and how the threshold harvesting level of depends on how quickly the range is shifting.
-We also found that climate-driven range shifts and fishing interact nearly additively, with very low positive synergy at more extreme levels of the stressors.
-
-We also examined the effect of threshold harvesting rules and marine protected areas (MPAs) on species persistence.
-Protected areas have been suggested as a key form of climate insurance and stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
-MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat is shifting rapidly.
-We found that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
-
-\section{Methods}
-
-We studied the dynamics of a fish population constrained to a single, one-dimensional habitat patch by their inability to reproduce outside of the patch.
-This viable habitat patch (here after `patch') is shifting at a fixed velocity and the fish at each point in space can be harvested.
-We first determined the climate velocity and harvesting rate that would drive the population extinct.
-We then measured the drop in biomass caused by range shifts, harvesting, and both stressors together in order to determine whether they interact synergistically.
-We finally implemented marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to determine how these management strategies affect population persistence.
-
-\subsection{The Model }
-
-In the model of \cite{ZhouKot2011}, the adults from the current year produce offspring according to a recruitment function and these offspring disperse across the one-dimensional world according to a dispersal kernel to become the next generation's adults.
-We extend this model by additionally subjecting the adults to harvesting before they produce offspring so that only a proportion of the fish survive to reproduce.
-These processes-- recruitment, harvesting, and dispersal-- are incorporated into an integrodifference model to describe how the population changes over time.
-If $n_t(x)$
-is the density of fish at position $x$ at time $t$, then the density of fish at the next generation is given by
-
-\begin{equation*}
-n_{t+1}(x)=\int^{\frac{L}{2}+ct}_{-\frac{L}{2}+ct}k(x-y)f((1-h)n_t(y))dy \label{integrodifference},
-\end{equation*}
-
-\noindent where $h$ is the proportion of adults harvested, $f(n)$ is the recruitment function giving the number of 
-offspring produced by a population of size $n$ (accounting for density dependence), $k(x-y)$ is the dispersal kernel giving the probability of a  larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it  shifts across space.
-We provide a list of variables and functions in Table \ref{variables}.
-We chose to use a Beverton-Holt recruitment function,
-
-\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}.\]
-
-\noindent Regardless of the exact functional form of the recruitment function, the critical parameter in determining population persistence is how quickly recruitment increases when the population 
-size is near (but above) $0$, which is equivalent to the intrinsic growth rate, $R_0=f'(0)$.
-Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.
-In our analyses, as in \citep{Latore:1998fk}, we used the separable Gaussian kernel given by
-
-\[k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.\]
-
-\noindent To derive analytical expressions, we approximated the kernel, as described in the Appendix.
-Analytical results for a separable sinusoidal kernel are also described in the Appendix.
-We used
-simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
-
-At equilibrium, the population will be described by a traveling wave, 
-where the density of fish at a given point in space will change but the density of fish at a location relative to the 
-shifting  patch will not. We sought to describe how the population is distributed over the viable 
-patch as it shifts through the world in order to study the size of the population at equilibrium and whether or not the population could persist.  The traveling wave $n^*$ must 
-satisfy
-
-\begin{equation}
-n^*(\bar{x})=\int^{\frac{L}{2}}_{-\frac{L}{2}}k(\bar{x}+c-\bar{y})f((1-h))n^*(\bar{y}))d
-\bar{y}, \label{traveling_pulse}
-\end{equation}
-
-\noindent where $\bar{x}\in\left[-\frac{L}{2}, \frac{L}{2}\right]$ describes the position within the patch \citep{ZhouKot2011}.
-
-\subsection{Persistence }
-One possible equilibrium traveling wave that solves Equation (\ref{traveling_pulse}) is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}
-{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.  If a population is to persist, it must be able to avoid extinction and 
-grow even when it is small.   A small population can be thought of as a perturbation to the trivial traveling pulse. If the trivial 
-pulse is stable, the system will return to the trivial pulse even after the introduction 
-of a small population.
-If the trivial pulse is unstable, a small population may increase and form a stable population.
-Population persistence is therefore equivalent to the trivial traveling pulse being an unstable equilibrium.
-
-If the population is harvested at low enough levels and the environment shifts slowly enough, the population will be able to persist.
-There are threshold values of the harvesting rate $h$ and the rate of environmental shift $c$ such that if the parameters are increased beyond these values, the population will be driven extinct.
-We found these critical parameters, $h^*$, and  $c^*$, by finding the parameters that make the trivial pulse unstable.
-Details are provided in Appendix \ref{stab}.
-
-For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment-- the length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
-$).
-The population biomass at equilibrium depends on the function form of recruitment, but population persistence only depends on the intrinsic growth rate $R_0$.
-For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
-
-\[R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.\]
-
-A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
-\begin{equation*}
-h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\langle d \rangle,c^2,L^2+3c^2),
-\end{equation*}
-where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
-
-\subsection{Calculating synergy }
-
-\citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct.
-In order to quantify whether the two stressors are interacting additively, synergistically, or antagonistically, we found the total biomass of the population when it reached an equilibrium traveling pulse and
-compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.
-For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
-
-\begin{equation}
-n^*(x)=\sum^\infty_{i=1}
-a_i(x)\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f((1-h)n^*(y))dy=\sum^\infty_{i=1}m_ia_i(x), \label{separable_integrodifference}
-\end{equation}
-
-\noindent where the $m_i$ satisfy the recursive equations
-
-\begin{equation}
-m_i=\int^{\frac{L}{2}}_{-\frac{L}{2}}b_i(y-c)f\left((1-h)\sum^\infty_{j=1}m_ja_j(x)\right)
-dy. \label{recursive_m}
-\end{equation}
-
-\noindent \citep{Latore:1998fk}. Equation (\ref{recursive_m}) allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
-equilibrium traveling pulse by using these $m_i$ and integrating Equation (\ref{separable_integrodifference}).
-
-We used $B_0$ to denote the equilibrium biomass 
-without either stressor, $B_\text{h}$ the equilibrium biomass with harvesting but a constant environment, $B_\text{c}$ the 
-equilibrium biomass with a shifting environment but no harvesting, and $B_\text{hc}$ the equilibrium biomass with 
-both stressors. For each stressor or combination of stressors, we found the drop in  biomass caused 
-by stressor $s$,
-
-\[E_\text{s}=B_0-B_\text{s}.\]
-
-\noindent If the stressors do not interact, the drop caused by both stressors would be the sum of the drops caused by 
-either individually. The synergy is therefore defined as
-
-\[S = E_\text{hc}-\left(E_\text{h}+E_\text{c}\right).\]
-
-\noindent If the stressors aggravate each other, the effect of both stressors is worse than would be expected from 
-considering either stressor individually, and synergy is positive. If the stressors alleviate each other, the effect 
-of both stressors is better than would be expected from considering either stressor individually, and synergy is 
-negative. If the effect of both stressors is exactly as expected from considering either stressor individually, 
-there is no interaction and no synergy.
-
-\subsection{Simulations }
-
-We used simulations to extend the basic integrodifference model in two ways that make it analytically 
-intractable. First, we examined the sensitivity of the model to choice of dispersal kernel by using the Laplace 
-dispersal kernel,
-
-\[ k(x-y)=\frac{1}{2}be^{-b|x-y|},\]
-
-\noindent a commonly used model of larval dispersal \citep{Pinsky:2011fk}.
-Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
-Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must
-specify a recruitment function.
-Again, we chose to use a Beverton-Holt function.
-In the first generation, we seeded the world with 50 individuals at a single point, as in \citep{ZhouKot2011}.
-We first ran through 150 generations in order for the population to reach equilibrium without harvesting or climate shift.
-We then added harvesting pressure, allowed the population to again reach equilibrium, and finally added climate change by moving the viable patch.
-Equilibrium biomass is calculated as the mean biomass of 300 time steps once the difference in biomass between time step $t$ and $t+1$ was no greater than $0.1$.
-
-We added harvesting pressure by harvesting a constant proportion of the population, in order to confirm our analytical results.
-We then evaluated the effect of a threshold harvest rule and marine protected
-areas (MPAs).  With a threshold rule, we evaluated the population at each point in space to determine how 
-much harvesting should occur. If the population abundance was below the designated threshold, no 
-harvesting occurred. If the population exceeded the threshold, then a proportion of the `surplus' individuals 
-were harvested.
-
-MPAs are a form of management designed to check the impact of fishing on targeted populations and are typically designed to meet either conservation of fishery management goals \citep{Agardy1994, HollandBrazee1996, Gainesetal2010a}.
-To implement an MPA management strategy in our model, we examine the effect of both of these commonly advocated approaches.
-While both conservation and fisheries oriented MPA schemes align in their goal of maintaining a sustainable fished population, they differ in desired level of adult spillover.
-Fisheries-oriented MPAs are often designed such that they maximize adult spillover into fishable areas by creating many small reserves closely spaced \citep{HastingsBotsford2003}.
-The converse of this is the goal of conservation-oriented MPAs which seek to reduce adult spillover by minizing the ratio between the reserve edge length relative to area protected \citep{Gainesetal2010a}.
-
-Networks of MPAs were introduced into our simulations by designating segments of space in which harvesting was forbidden (i.e.
-harvesting rates were equal to $0$).
-Conservation-oriented MPAs, are frequently large and rarely part of a larger network of reserves \citep{HastingsBotsford2003}.
-For solitary reserves to be successful at protecting target species, they must encompass self-sustaining fish populations \citep{HastingsBotsford2006, Gainesetal2010a}.
-As such modeling studies estimate that isolated reserves must be at least as large as the average dispersal distance for the targeted fish species \citep{Lockwoodetal2002, HastingsBotsford2003, Botsfordetal2001, Gainesetal2010}.
-To implement conservation MPAs we created reserves with a length of 4 times the average dispersal distance and had a distance of 8 times the average dispersal distance between them to ensure that populations would be self sustiaining and not dependent on other dispersal for other reserves \citep{Lockwoodetal2002}.
-
-Previous work has shown that if MPAs are to benefit fisheries, the reserves should be broken into a network, closely spaced to maximize adult spillover into fishable areas and export of larvae from reserve to reserve \citep{ HastingsBotsford2003, Gaylordetal2005, Gainesetal2010a}.
-To mimic this management scheme, MPAs had a length of $\tfrac{1}{3}$ of the average dispersal distance and had a distance of $\tfrac{2}{3}$ of the average dispersal distance between them.
-
-\section{Results}
-
-\subsection{Interactions Between Stressors }
-We find the critical climate velocity and harvest  rate to be inversely related: as the harvesting rate $h$ increases, 
-the critical climate velocity $c^*$ decreases as the environment must move more slowly to accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
-environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}).
-This means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if the environment starts changing.
-When the climate velocity or harvesting pressure exceed their critical rates ($h^*, c^*$ respectively), the biomass of the population at equilibrium will be equal to $0$.
-Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure increases or the environmental shifts more quickly (Figure \ref{baseline}).
-Our simulations confirm the analytical results with the critical speed $c^*$ declining as the critical harvest rate $h^*$ increases and vice versa (Figure \ref{nomang}).
-
-It is always the case that increasing the intrinsic growth rate, $R_0$, of the population increases the critical speed $c^*$ and the critical harvesting rate $h^*$, since a population that grows more quickly can recover more quickly from losses caused by these disturbances.
-However, whether or not dispersing farther is better depends on how quickly the environment is shifting (Figure \ref{baseline}).
-When the environment is shifting slowly, dispersing farther is detrimental since many larvae will disperse too far away from the viable patch.
-When the environment is shifting quickly, on the other hand, dispersing farther can help the population persist because some larvae will disperse into the space that will become viable shortly in the future.
-This affects the critical harvesting rate: at a low rate of environmental shift, populations that disperse less can be harvested more severely than those that disperse further, whereas at a high rate of environmental shift, populations that disperse further can be harvested more severely.
-
-We found very low levels of positive synergy between the two stressors in our analysis of the Gaussian kernel (Figure \ref
-{Synergy}).
-Where there is positive synergy, a doubly stressed population loses more biomass than would be predicted from either stressor individually.
-The stressors interact most strongly when they are both high, shortly before they drive the population extinct.
-However, the excess loss in biomass is extremely low, making it difficult to distinguish positive synergy from additive interactions.
-We found similar analytical results for a sinusoidal dispersal kernel, which indicates that this result is robust to changes in the dispersal kernel.
-
-\subsection{Management strategies }
-
-Without any management strategies, we found that when the population is harvested more severely, slower rates of environmental shift will suffice to drive the population extinct.
-However, when thresholds are in place, the critical rate of environmental shift $c^*$ no longer depends on the harvesting rate (Figure \ref{management}).
-In other words, as long as there is some threshold below which harvesting is not allowed, there is a constant critical rate of environmental shift that only depends on the growth rate, length of the viable patch, and average dispersal distance.
-This is because, with thresholds in place, a small portion of the population always escapes harvesting pressure and only the size, not the growth rate, of this small population depends on the level of threshold implemented.
-This small population can then track the shifting environment as well as a larger one.
-
-We also examined the effect of marine protected  areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and  climate change parameters where the fish population could survive.
-With MPAs in place, the population can withstand a slightly higher harvesting rate in those parts of the world where harvesting is allowed (Figure \ref{management}).
-More surprisingly, perhaps, implementing MPAs slightly increased the critical rate of environmental shift (Figure \ref{management}).
-Even though MPAs are meant to protect marine populations from harvesting pressure, by providing refuges in which their intrinsic growth rates can increase back to the normal level, MPAs allow marine populations to withstand higher rates of environmental shift.
-
-\section{Discussion}
-
-Knowing whether two disturbances interact in their impacts for a given population is important for management.
-The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.
-Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.
-
-For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
-More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
-That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
-This is an indication of an interaction between the stressors.
-
-To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.
-We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.
-We found similar results from the analytically derived biomass and the simulation derived biomass.
-This indicates that this result is robust to changes in the dispersal kernel.
-We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \citep{Crainetal2008, DarlingCote2008,Nyeetal2013}.
-We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
-In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.
-Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
-
-The fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.
-This is further evidence that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.
-However, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e.
-synergistic effects were quite small.
-Synergy between harvesting and the effects of climate change has been identified in experimental populations \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
-In the experimental populations, synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
-While we did find (very low levels of) positive synergy, we did not find as much as might be predicted from these empirical studies.
-However, these previous results are not directly comparable to ours because they focus on different aspects of climate change, e.g.
-warming temperature \citep{Moraetal2007,Kirby:2009fk} or a more variable climate \citep{Planque:2010uq}, rather than range shifts.
-Additionally, while we can isolate the affects of climate shift and harvesting in our simple analytical model, there are other forces acting on real populations that may produce the observed synergistic effects.
-
-Our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
-Species with higher intrinsic growth rates will always better track a shifting environment than species with lower intrinsic growth rates.
-This is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non-shifting species (smaller body sizes, faster maturation, smaller sizes at maturity) \citep{Perryetal2005}.
-However, as shown in Figure \ref{baseline}, having a longer average dispersal distance translates into a better ability to track environmental shift only if the environment is shifting quickly.
-
-We also examined whether frequently recommended management approaches ensure species persistence.
-We found increases in the population's biomass at equilibrium and an improved ability to persist.
-Harvesting thresholds, are already implemented in fisheries management, and we found that this management tactic alleviates interactions between the two stressors by allowing a small portion of the population to escape harvesting pressure and successfully track the shifting environment.
-Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
-Our results, that spatial management increased the maximum harvesting rate at which the population could survive, support the idea that MPAs could be used to reduce the impact of harvesting.
-We also found evidence that MPAs increased the climate velocity under which the population could persist.
-MPAs help populations to persist by providing regions of space in which to grow and recover before being re-subjected to harvesting.
-While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.
-This is encouraging evidence that a single set of of management practices may help to protect marine populations from both harvesting and climate change.
-
-The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.
-However, it  ignores many of the complexities present in marine fisheries.
-First, we could consider additional disturbances, like pollution, disease, physiological response to temperature.
-However, if the main effect of these disturbances is to increase mortality rates than we would expect that they would interact with climate shift in a way similar to harvesting pressure.
-We did not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time.
-However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect.
-We also did not include age structure in our model, which has previously been shown to affect a population's resistance to harvesting and climate change \citep{Botsfordetal2001}.
-The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work.
-Similarly, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.
-Besides these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.
-By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000} and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
-
-Using a simple mechanistic model like the one we present here provides a useful framework for understanding when it is or is not necessary to consider the interactions between multiple disturbances and for incorporating additional ecological complexities which can mediate species persistence under those disturbances.
-Exploring how species interactions, age structure, and additional disturbances affect population viability will improve our predictions and help us to understand whether species will persist under predicted climate and harvesting regimes.
-Finally, this work can help to make general predictions as to whether specific life histories are likely to be selected over others as harvesting and range shifts increase.
-
-\section*{ Acknowledgements}
-We thank Catherine Offord and Will Scott for their contributions to an earlier version of this manuscript.
-
-\end{flushleft}
-\end{spacing}
-
-\bibliography{fish}
-\pagebreak
-\section{Figures}
-
-\begin{figure}[htbp]
-\begin{subfigure}{3in}
-\subcaption{\label{rates}}
-\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
-\end{subfigure}
-\begin{subfigure}{3in}
-\subcaption{\label{biomass}}
-\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
-\end{subfigure}
-
-\caption{%(a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
-Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
-Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-(b) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-These results are from an approximated Gaussian dispersal kernel with $L=1$.
-}
-
-\label{baseline}
-\end{figure}
-
-\pagebreak
-
-\noindent {\bf Figure \ref{baseline}}: (a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
-Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
-Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-These results are from an approximated Gaussian dispersal kernel with $L=1$.
-(b) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-
-\pagebreak
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=3in]{plots/synergy.pdf}
-\caption{
-%Positive synergy between the two stressors.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-}
-\label{Synergy}
-\end{center}
-\end{figure}
-
-\pagebreak
-
-\noindent {\bf Figure \ref{Synergy}}: Positive synergy between the two stressors.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-This excess loss, on the order of $.001$, is small in comparison to the total biomass, which can be as large as $20$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
-
-\pagebreak
-
-\begin{figure}[htbp]
-
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
-\label{nomang}
-\end{subfigure}
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
-\end{subfigure}
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
-\end{subfigure}
-\caption{
-%The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
-(a) No management.
-(b) Threshold harvesting levels.
-(c) MPAs.
-These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
-}
-\label{management}
-\end{figure}
-
-\pagebreak
-
-\noindent {\bf Figure \ref{management}}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
-(a) No management.
-(b) Threshold harvesting levels.
-(c) MPAs.
-These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
-
-\pagebreak
-
-\section{Tables}
-\begin{table}[h]
-\caption{Table of variables used in the text}
-\begin{tabular}{@{}lllllll@{}}
-  Variable & Definition
-\\\cmidrule{1-1} \cmidrule{2-2}   
-$n_t(x)$ & density of fish at position $x$ at time $t$
-\\ $n^*(\overline{x})$ & density of fish at equilibrium at position $\overline{x}$ relative to the patch 
-\\ $k(x-y)$ & dispersal kernel, the probability of larva traveling from position $y$ to position $x$
-\\ $\langle d \rangle $ & expected distance traveled by larva
-\\ $f(n)$ & recruitment function, the number of offspring produced by a population of size $n$
-\\ $R_0$ & intrinsic growth rate, $R_0=f'(0)$
-\\ $h$ & proportion of adults harvested
-\\ $L$ & patch length
-\\ $c$ & rate of environmental shift
-\end{tabular}
-\label{variables}
-\end{table}
-
-\pagebreak
 

### 2014-02-28 Writing/MS.tex

--- 2014-02-27 Writing/MS.tex
@@
 We also introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could be effective management tools sat they mitigate the interaction between the two stressors.
-
 \hspace{10cm}
@@
 For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
-We also found a negative relationship between the critical harvesting rate and the rate of environmental shift.
+More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
 That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
-The curved line separating parameters that will allow the population to persist from those that won't is an indication of an interaction between the stressors.
-%say more about this line of evidence for an interaction
+This is an indication of an interaction between the stressors.
 
@@
 %we need a transition here, feels abrupt
-Experimental work has identified synergy between harvesting and climate changes \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
-Additionally, experimental populations have identified synergy between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
+Synergy between harvesting and the effects of climate change has been identified in experimental populations \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
+In the experimental populations, synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
 While we did find (very) low levels of positive synergy, we did not find as much as predicted from these empirical studies.
@@
 Additionally, while we can isolate the affects of climate shift and harvesting in our simple analytical model, there are other forces acting on real populations that may produce the observed synergistic effects.
-
-%e.g.?!?!
-
-%Nevertheless, our results generally agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
 
@@
 With these management strategies we found increases in the population's biomass at equilibrium and an improved ability to persist.
+We found that a threshold harvesting rule alleviates interactions between the two stressors.
+Thresholds have this effect as the management approach effectively prevents harvesting of the leading edge, which allows colonization to occur as if these individuals were moving into un-fished areas.
+It's interesting to note that novel, low abundance species are commonly unregulated in fisheries systems; so in order to decouple the additive effects of harvest and climate change, management would have to reverse this paradigm by allowing no harvest of new species until they had become established.
+
+Unlike thresholds, MPAs are explicitly spatial.
 Previous work has advanced protected areas as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation \citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.
@@
 
-Unlike MPAs, thresholds are not explicitly spatial, and we found that this harvest control rule alleviates interactions between the two stressors.
-Thresholds have this effect as the management approach effectively prevents harvesting of the leading edge, which allows colonization to occur as if these individuals were moving into un-fished areas.
-It's interesting to note that novel, low abundance species are commonly unregulated in fisheries systems; so in order to decouple the additive effects of harvest and climate change, management would have to reverse this paradigm by allowing no harvest of new species until they had become established.
-These results highlight that while management strategies only change harvesting practices, they can effect the way stressors interact and so understanding these interacts help to better design harvesting rules and place protected areas.
+While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.
+This is encouraging evidence that a single set of of management practices may help to protect marine populations from both harvesting and climate change.
 

### 2014-02-28 Writing/MS.tex
fine tooth edits in abstract and intro

@@
 \title{Persistence of marine populations under climate and fishing}
-\author{Emma Fuller$^1$, Eleanor Brush$^1$, Malin Pinsky$^{1,2}$}
+\author{Emma Fuller$^1$, Eleanor Brush$^2$, Malin Pinsky$^{1,3}$}
 \date{\small (1): Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey 08544 USA \\
-(2): Department of Ecology, Evolution and Natural Resources, Rutgers University, New Brunswick, New Jersey 08901 USA}
+(2): Program in Quantitative and Computational Biology, Princeton University, Princeton, New Jersey 08544 USA \\
+(3): Department of Ecology, Evolution and Natural Resources, Rutgers University, New Brunswick, New Jersey 08901 USA}
 
@@
 In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we analyzed an integrodifference model that explicitly included the mechanisms of dispersal and reproduction.
-We found the critical rates of harvesting and climate velocity depend on the growth rate and dispersal kernel of the population.
+We found how the critical rates of harvesting and climate velocity depend on the growth rate and dispersal kernel of the population.
 We measured the interaction between the stressors and found that the disturbances interact nearly additively, with low positive synergy only at the greatest harvest rates and climate velocity that almost drive the population extinct.
-We also introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that under some circumstances these approaches could be effective management tools sat they mitigate the interaction between the two stressors.
+We also introduced two conservation techniques into simulations of the population model --- threshold harvest rules and marine protected areas (MPAs) --- and found that under some circumstances these approaches could be effective management tools as they mitigate the interaction between the two stressors.
 \hspace{10cm}
@@
 Many stressors can disturb an ecosystem, and ecologists have quantified the consequences of a great deal of these of perturbations \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
-Less work, however, has measured the effects of multiple stressors and the interactions between them.
+Less work, however, has been done to measure the effects of multiple stressors and the interactions between them.
 If disturbances interact synergistically, a perturbation that has little effect when it occurs individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008,Nyeetal2013,Gurevitchetal2000}.
@@
 If disturbances interact antagonistically, on the other hand, the effects of multiple stressors may be less than that predicted by the individual effects of the stressors.
-Since disturbances rarely occur in isolation, measuring the effects of multiple distrubances gives a better understanding of the likely impacts to the system \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
+Since disturbances rarely occur in isolation, measuring the effects of multiple disturbances gives a better understanding of the likely impacts to the system \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}.
 
 Climate change and fishing, two of the largest human impacts on the ocean \citep{Halpernetal2008}, provide an important case study of how disturbances interact in their effects on biological populations.
-Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and projected to continue in the future \citep{Kelletal2005, Mackenzieetal2007}.
+Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue in the future \citep{Kelletal2005, Mackenzieetal2007}.
 These shifting species, and those likely to move in the future, are also subject to harvesting, among other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
@@
 A common approach to predicting future population distributions has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
-These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to explain and predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
+These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
 Despite these models' widespread adoption, many papers have criticized SDMs as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
@@
 Here we extended a previously studied model of a fish population subject to climate-driven range shift by also considering harvesting pressure.
-We explicitly include reproduction and dispersal, two mechanistic processes central to species' responses to climate and fishing.
+The model explicitly included reproduction and dispersal, two mechanistic processes central to species' responses to climate and fishing.
 Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We found the critical harvesting rate and climate velocity that drives the population extinct and how these critical rates depend on one another.
+We found the critical harvesting rate and climate velocity that drive the population extinct and how these critical rates depend on one another.
 We also found that climate-driven range shifts and fishing interact nearly additively, with low positive synergy at more extreme levels of the stressors.
@@
 We also examined the efficacy of two different types of management strategies: threshold harvesting rules and marine protected areas (MPAs).
-Previous work has suggested protected areas as a key form of climate insurance and stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
 MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat shifts rapidly.
-We found that threshold harvesting rules erase the interaction between harvesting rates and climate velocity and that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.
+Previous work has suggested protected areas can be a key form of climate insurance and can provide stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
+We found that threshold harvesting rules remove the interaction between harvesting rates and climate velocity and that MPAs can help a species persist with higher harvesting pressure and slightly increase the maximum climate velocity with which a species can keep up.
 

### 2014-03-01 Simluations/Aspatial_fast/Data/troubleshoot_moveMPA.R
finished re-running simulations

@@ -1,0 +1,110 @@
+# just doing the MPA moving thing
+reps = 450
+test = vector("list",length=reps)
+
+displaced = ifelse(s>0,s/step_size,1)
+
+test[[1]]$MPA_finish = MPA.start
+
+for(i in 2:reps){  
+# move MPA forward by <displaced> amount
+  test[[i]]$next_MPA = test[[i-1]]$MPA_finish[displaced:length(test[[i-1]]$MPA_finish)]
+  lost <- test[[i-1]]$MPA_finish[1:(displaced-1)] 
+ # are there any MPAs in <next_MPA>?
+  #any(test[[i]]$next_MPA==1)
+  # IF FALSE, then need to figure out how many 0s were lost in move, and make sure to preserve interval of zeros == mpa.no, then fill in mpa.yes,mpa.no to length of world
+  if(any(test[[i]]$next_MPA==1)==FALSE & any(lost==1)==FALSE){
+    # these are the intervals that are left behind as world moves forward
+      lost <- test[[i-1]]$MPA_finish[1:(displaced-1)] 
+    # this is the last continuous chunk of numbers at the end of <lost>  
+      length_last <- rep(tail(rle(lost)$value,1), tail(rle(lost)$lengths,1))
+    # want to know how long <length_last> is so can make sure to get correct interval
+      L_int <- length(length_last)
+    # how many more zeros do we need before we start with mpa.yes again?
+      zero_add <- sum(mpa.no==0) - L_int - sum(test[[i]]$next_MPA==0)
+    # this is what needs to be appended to <next_MPA>
+      test[[i]]$new_MPA <- c(rep(0,zero_add),rep(c(mpa.yes,mpa.no),length.out=w))
+      
+  }else{
+  
+  # IF FALSE (there are some 1s) then we only care about the last interval of the <next_MPA>, is that protected or not?
+    last_step = tail(test[[i]]$next_MPA,1)
+   # IF <last_step>==1
+    if(last_step ==1){ 
+      #then how many 1s are at the end of the <next_MPA> section?
+        end_step = rep(tail(rle(test[[i]]$next_MPA)$value,1), tail(rle(test[[i]]$next_MPA)$lengths,1)) # number of 1s at very end of lost interval
+      # length of <end_step>
+      end_int = length(end_step)
[…]

### 2014-03-02 fish_figures.R
plot labels

@@ -79,10 +79,10 @@ myaxes=list(arrows=FALSE,x=list(at=rev(seq(1,C,length.out=(cvals[C]-cvals[1])/xd
 quartz()
 pdf(file='eqbiomass.pdf',width=fig.width)
 
-# wireframe(toplot_ebm[C:1,H:1],xlab=list("Rate of environmental shift",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
+# wireframe(toplot_ebm[C:1,H:1],xlab=list("Climate velocity",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
 par(oma=oma)
 image.plot(cvals,hvals[1:H],toplot_ebm[,1:H],breaks=cuts,col=mycols,
-	xlab="Rate of environmental shift",ylab="Harvesting rate",
+	xlab="Climate velocity",ylab="Harvesting rate",
 	cex.lab=cex.lab,cex.axis=cex.axis, 
 	yaxs=yaxs,xaxs=xaxs,axes=TRUE,
 	legend.shrink=1,legend.width=.1,zlim=range(cuts),
@@ -147,10 +147,10 @@ myaxes=list(arrows=FALSE,
 quartz()
 pdf(file='eqbiomass_sim.pdf',width=fig.width, height=fig.width)
 
-# wireframe(toplot_ebm[C:1,H:1],xlab=list("Rate of environmental shift",rot=10),ylab=list("Harvesting rate",rot=-35),zlab=list('Equilibrium biomass',rot=90),at=cuts,col.regions=mycols,alpha.regions=myalpha,scales=myaxes,drape=FALSE,screen=list(z=30,x=-80),par.settings = list(axis.line = list(col = "transparent")))
[…]

### 2014-03-03 Writing/MS.tex

@@
 In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we analyzed an integrodifference model that explicitly included the mechanisms of dispersal and reproduction.
-We found how the critical rates of harvesting and climate velocity depend on the growth rate and dispersal kernel of the population.
+We found how the critical rates of harvesting and climate velocity that suffice to drive the population extinct depend on the growth rate and dispersal kernel of the population.
 We measured the interaction between the stressors and found that the disturbances interact nearly additively, with low positive synergy only at the greatest harvest rates and climate velocity that almost drive the population extinct.
-We also introduced two conservation techniques into simulations of the population model --- threshold harvest rules and marine protected areas (MPAs) --- and found that under some circumstances these approaches could be effective management tools as they mitigate the interaction between the two stressors.
+We also introduced two conservation techniques into simulations of the population model, threshold harvest rules and marine protected areas (MPAs), and found that under some circumstances these approaches could be effective management tools as they mitigate the interaction between the two stressors.
 \hspace{10cm}
@@
 
-Many stressors can disturb an ecosystem, and ecologists have quantified the consequences of a great deal of these of perturbations \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
+Many stressors can disturb an ecosystem, and ecologists have quantified the consequences of many of these of perturbations \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
 Less work, however, has been done to measure the effects of multiple stressors and the interactions between them.
@@
 These shifting species, and those likely to move in the future, are also subject to harvesting, among other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
-Previous empirical work has found synergistic interactions between overfishing and temperature-driven range shifts \citep{Lingetal2009} and microcosm experiments have identified synergistic interactions between warming temperatures, harvesting and connectivity \citep{Moraetal2007}.
+Previous empirical work has found synergistic interactions between overfishing and temperature-driven range shifts \citep{Lingetal2009} and synergistic interactions between warming temperatures, harvesting and connectivity have been identified in microcosm experiments \citep{Moraetal2007}.
 This empirical work underscores the importance of understanding how range shifts and harvesting interact.
@@
 A common approach to predicting future population distributions has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs).
-These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
+These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperature, rainfall, or salinity, to predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
 Despite these models' widespread adoption, many papers have criticized SDMs as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
@@
 
-We studied the dynamics of a fish population constrained to a single, one-dimensional habitat patch by their inability to reproduce outside of that area as introduced by \cite{ZhouKot2011}.
+We studied a model of the dynamics of a fish population constrained to a single, one-dimensional habitat patch by their inability to reproduce outside of that area, as introduced by \cite{ZhouKot2011}.
 This viable habitat patch (here after `patch') shifts at a fixed velocity and harvest occurs at each point in space along the entire one-dimensional world.
-We first determined the climate velocity and harvesting rate that would drive the population extinct (hereafter the critical harvesting rate and climate velocity), and then measured synergy by calculating the drop in biomass caused by each stressor both individually and together.
-We finally add marine protected areas (MPAs) and threshold harvesting rules in numerical simulations of the model to determine how these management strategies affect population persistence.
+We first determined the harvesting rate climate velocity that would drive the population extinct (hereafter the critical harvesting rate and critical climate velocity), and then measured synergy by calculating the drop in biomass caused by each stressor both individually and together.
+We finally added threshold harvesting rules and marine protected areas (MPAs) in numerical simulations of the model to determine how these management strategies affect population persistence.
 
@@
 We extend this model by additionally subjecting the adults to harvesting before they produce offspring so that only a proportion of the fish survive to reproduce.
-We incorporate these processes-- recruitment, harvesting, and dispersal-- into an integrodifference model to describe how the population changes over time.
+These processes-- recruitment, harvesting, and dispersal-- are incorporated into an integrodifference model to describe how the population changes over time.
 If $n_t(x)$
@@
 We provide a list of variables and functions in Table \ref{variables}.
-We use a Beverton-Holt recruitment function,
+We used a Beverton-Holt recruitment function,
 
-\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
+\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}.\]
 
-\noindent but regardless of the exact functional form of the recruitment function, the critical parameter in determining population persistence is how quickly recruitment increases when the population size is near (but above) $0$, which is equivalent to the intrinsic growth rate $R_0=f'(0)$.
-Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.
+\noindent  Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.
 In our analyses, as in \cite{Latore:1998fk}, we used the separable Gaussian kernel given by
@@
 
-\noindent To derive analytical expressions, we approximated the kernel, as described in the Appendix. %section number necessary?
-Analytical results for a separable sinusoidal kernel are also described in the Appendix.  %awkward
-We used simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.  % Why did we choose a Laplace dispersal kernel for simulations?
+\noindent To derive analytical expressions, we approximated the kernel, as described Appendix A.3.
+Analytical results for a separable sinusoidal kernel are also described in Appendix A.4.  
+We used simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
 
-At equilibrium, a traveling wave will describe the population, where the density of fish at a given point in space will change but the density of fish at a location relative to the shifting patch will not.
-We sought to describe  the distribution of the population over the viable patch as it shifts through the world in order to study the size of the population at equilibrium and whether or not the population could persist.
+At equilibrium, the population will move in a traveling wave, where the density of fish at a given point in space will change but the density of fish at a location relative to the shifting patch will not.
 The traveling wave $n^*$ must satisfy
@@
 \noindent where $\bar{x}\in\left[-\frac{L}{2}, \frac{L}{2}\right]$ describes the position within the patch \citep{ZhouKot2011}.
-
-\subsection{Persistence }
-One possible equilibrium traveling wave that solves Equation (\ref{traveling_pulse}) is the `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}{2}, \frac{L}{2}\right]$, i.e.~a patch with no fish in it.
-If a population persists, it must be able to avoid extinction and grow even when small.
-We can think of a small population as a perturbation to the trivial traveling pulse.
-If the trivial pulse is stable, the system will return to the trivial pulse even after the introduction of a small population.
-If the trivial pulse is unstable, a small population may increase and form a persistent population.
-Population persistence is therefore equivalent to the trivial traveling pulse being an unstable equilibrium.
-
-If we harvest the population at low enough levels and the climate velocity is slow enough, the population will be able to persist.
-There exists threshold values of the harvesting rate $h$ and a climate velocity $c$ such that if we increase parameters beyond these values, we drive the population extinct.
-We found these critical parameters, $h^*$, and  $c^*$, by finding the parameters that make the trivial pulse unstable (See Appendix \ref{stab}).
-
-For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses $\langle d \rangle$ and the intrinsic growth rate $R_0$; properties of the environment-- the length of the viable patch $L$ and how quickly the environment shifts $c$; and the harvesting rate $h$.
-The population biomass at equilibrium depends on the function form of recruitment, but population persistence only depends on the intrinsic growth rate $R_0$.
-For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
-
-\[R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.\]
-
-We derive a similar expression for a sinusoidal kernel in the Appendix [{\bf REF?}].  For both kernels, we can approximate the critical harvesting proportion by a function that looks like 
-\begin{equation*}
-h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\langle d \rangle,c^2,L^2+3c^2),
-\end{equation*}
-where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
-
-\subsection{Calculating synergy }
-
-\citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct.
-In order to quantify whether the two stressors interact additively, synergistically, or antagonistically, we found the total biomass of the population when it reached an equilibrium traveling pulse and compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.
 For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy
@@
 
-\noindent \citep{Latore:1998fk}. Equation (\ref{recursive_m}) allowed us to find the values of $m_i$ numerically. We then found the total biomass in the 
-equilibrium traveling pulse by using these $m_i$ and integrating Equation (\ref{separable_integrodifference}).
+\noindent \citep{Latore:1998fk}.
+
+\subsection{Persistence }
+If the population is harvested at low enough levels and the climate velocity is slow enough, the population will be able to persist.
+There are threshold values of the harvesting rate $h$ and the climate velocity $c$ such that if we increase the parameters beyond these values, the population will be driven extinct.
+When the population is extinct, the system is in equilibrium, i.e.
+there is a `trivial' traveling pulse, $n^*(\bar{x}) = 0$ for all $x \in \left[-\frac{L}{2}, \frac{L}{2}\right]$, which satisfies Equation (\ref{traveling_pulse}).
+If a population persists, it must be able to avoid extinction and grow even when small.
+If the trivial pulse is stable, the system will return to extinction even after the introduction of a small population.
+If the trivial pulse is unstable, a small population may increase and form a persistent population.
+Population persistence is therefore equivalent to the trivial traveling pulse being an unstable equilibrium.
+We found the critical parameters, $h^*$ and  $c^*$, by finding the parameters that make the trivial pulse unstable.
+See Appendix A.1 for details.
+
+Regardless of the exact functional form of the recruitment function, the critical parameter in determining population persistence is how quickly recruitment increases when the population size is near (but above) $0$, which is equivalent to the intrinsic growth rate $R_0=f'(0)$.
+For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses $\langle d \rangle$ and the intrinsic growth rate $R_0$; properties of the environment-- the length of the viable patch $L$ and how quickly the environment shifts $c$; and the harvesting rate $h$.
+For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
+
+\[R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.\]
+
+We derive a similar expression for a sinusoidal kernel in the Appendix A.4.  For both kernels, we can approximate the critical harvesting proportion by a function that looks like 
+\begin{equation*}
+h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\langle d \rangle,c^2,L^2+3c^2),
+\end{equation*}
+where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
+
+\subsection{Calculating synergy }
+
+\citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct.
+In order to quantify whether the two stressors interact additively, synergistically, or antagonistically, we found the total biomass of the population when it reached an equilibrium traveling pulse and compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.
+Equations (\ref{recursive_m}) and (\ref{separable_integrodifference}) allowed us to numerically find the  total biomass in the equilibrium traveling pulse.
 
@@
 
-\noindent If the stressors aggravate each other, synergy is positive, and the effect of both stressors is worse than would we expect from considering either stressor individually.
-If the stressors alleviate each other, the synergy is negative, and the effect of both stressors is better than we expecte from considering either stressor individually.
+\noindent If the stressors aggravate each other, the effect of both stressors is greater than  we would expect from considering either stressor individually and synergy is positive.
+If the stressors alleviate each other, the effect of both stressors is less than we would expect from considering either stressor individually and synergy is negative.
 If the effect of both stressors is exactly as expected from considering either stressor individually, there is no interaction and no synergy.
@@
 \noindent a commonly used model of larval dispersal \citep{Pinsky:2011fk}.
-Second, we examined harvesting rules more complex than harvesting a constant proportion of the population.
-Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must specify a recruitment function.
-Again, we chose to use a Beverton-Holt function.
-In the first generation, we seeded the world with 50 individuals at a single point, as in \cite{ZhouKot2011}.
+Second, we implemented two management strategies, threshold rules and MPAs, to understand their effect on population persistence and on the synergistic interactions between stressors.
+For every simulation, in the first generation, we seeded the world with 50 individuals at a single point, as in \cite{ZhouKot2011}.
 We first ran through 150 generations in order for the population to reach equilibrium without harvesting or climate shift.
 We then added harvesting pressure, allowed the population to again reach equilibrium (150 generations), and finally added climate change by moving the viable patch.
-We calculate equilibrium biomass as the mean biomass of 300 time steps once the difference in biomass between time step $t$ and $t+1$ was no greater than $0.1$.
+We calculated equilibrium biomass as the mean biomass of 300 time steps once the difference in biomass between successive generations was no greater than $0.1$.
 
-n, in order to confirm our analytical results, we first added harvesting pressure by harvesting a constant proportion of the population.
-We then evaluated the effect of a threshold harvest rule and marine protected areas (MPAs).
+Under the two management strategies, harvesting pressure was implemented differently.
 With a threshold rule, we evaluated the population at each point in space to determine how much harvesting should occur.
@@
 
-MPAs are a form of management designed to check the impact of fishing on targeted populations and are typically designed to meet either conservation of fishery management goals \citep{Agardy1994, HollandBrazee1996, Gainesetal2010a}.
-To implement an MPA management strategy in our model, we examine the effect of both of these commonly advocated approaches.
-While both conservation and fisheries oriented MPA schemes align in their goal of maintaining a sustainable fished population, they differ in desired level of adult spillover.
-Fisheries-oriented MPAs are often designed such that they maximize adult spillover into fishable areas by creating many small reserves closely spaced \citep{HastingsBotsford2003}.
-The converse of this is the goal of conservation-oriented MPAs which seek to reduce adult spillover by minizing the ratio between the reserve edge length relative to area protected \citep{Gainesetal2010a}.
-
 We introduce networks of MPAs into our simulations by designating segments of space where the harvesting rate was equal to $0$.
-Conservation-oriented MPAs, are frequently large and rarely part of a larger network of reserves \citep{HastingsBotsford2003}.
-For solitary reserves to be successful at protecting target species, they must encompass self-sustaining fish populations \citep{HastingsBotsford2006, Gainesetal2010a}.
-As such modeling studies estimate that isolated reserves must be at least as large as the average dispersal distance for the targeted fish species \citep{Lockwoodetal2002, HastingsBotsford2003, Botsfordetal2001, Gainesetal2010}.
-To implement conservation MPAs we created reserves with a length of 4 times the average dispersal distance and had a distance of 8 times the average dispersal distance between them to ensure that populations would be self sustiaining and not dependent on other dispersal for other reserves \citep{Lockwoodetal2002}.
-
-Previous work has shown that if MPAs are to benefit fisheries, the reserves should be broken into a network, closely spaced to maximize adult spillover into fishable areas and export of larvae from reserve to reserve \citep{ HastingsBotsford2003, Gaylordetal2005, Gainesetal2010a}.
-To mimic this management scheme, MPAs had a length of $\tfrac{1}{3}$ of the average dispersal distance and had a distance of $\tfrac{2}{3}$ of the average dispersal distance between them.
+MPAs are typically designed to meet either fishery management  or conservation goals \citep{Agardy1994, HollandBrazee1996, Gainesetal2010a}.
+Fisheries-oriented MPAs are often designed such that they maximize adult spillover into fishable areas by creating many small reserves closely spaced \citep{HastingsBotsford2003,Gaylordetal2005, Gainesetal2010a}.
+To mimic this management scheme, we implemented MPAs with a length of $\tfrac{1}{3}$ of the average dispersal distance and a distance of $\tfrac{2}{3}$ of the average dispersal distance between them.
+Conservation-oriented MPAs seek to reduce adult spillover by creating fewer larger protected areas \citep{HastingsBotsford2006, Gainesetal2010a}.
+To mimic this scheme.
+we implemented MPAs with a length of $4$ times the average dispersal distance and a distance of $8$ times the average dispersal distance between them \citep{Lockwoodetal2002}.
 
@@
 \subsection{Interactions Between Stressors }
-We find the critical climate velocity and harvest  rate to be inversely related: as the harvesting rate $h$ increases, the critical climate velocity $c^*$ decreases, as the environment must move more slowly to accommodate the population growing more slowly (Figure \ref{baseline}).
-% this seems like a backwards way to say it.
-Instead of "as the environment must move more slowly to accommodate..." it seems like it should be "as the population can only survive in increasingly stationary habitats"..
-But "increasingly stationary" is not great...
-Ideas?
-Conversely, as the rate of environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}).
+The critical climate velocity and harvest  rate are inversely related.
+As the climate velocity shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}).
 This means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if the environment starts changing.
-When the climate velocity or harvesting pressure exceed their critical rates ($h^*, c^*$ respectively), the biomass of the population at equilibrium will be equal to $0$.
+Conversely, as the harvesting rate $h$ increases, the critical climate velocity $c^*$ decreases (Figure \ref{baseline}).
+This means that as harvesting pressure increases, it becomes increasingly easy for a shifting environment to drive the population extinct.
+
+When the climate velocity or harvesting pressure exceed their critical rates ($c^*, h^*$ respectively), the biomass of the population at equilibrium will be equal to $0$.
 Before the stressors reaches those thresholds, the equilibrium biomass of the population decreases as either the harvesting pressure increases or the environmental shifts more quickly (Figure \ref{baseline}).
@@
 
-It is always the case that increasing the intrinsic growth rate, $R_0$, of the population increases the critical climate velocity $c^*$ and the critical harvesting rate $h^*$, since a population that grows more quickly can recover more quickly from losses caused by these disturbances.
+It is always the case that increasing the intrinsic growth rate, $R_0$, increases the critical climate velocity $c^*$ and the critical harvesting rate $h^*$, since a population that grows more quickly can recover more quickly from losses caused by these disturbances.
 However, whether or not dispersing farther is better depends on how quickly the environment is shifting (Figure \ref{baseline}).
@@
 When the environment is shifting quickly, on the other hand, dispersing farther can help the population persist because some larvae will disperse into the space that will become viable shortly in the future.
-This affects the critical harvesting rate: at a low rate of environmental shift, we can more severely harvest populations that have a shorter dispersal distance than those that disperse farther, whereas at a high rate of environmental shift, we can more aggressively harvest populations that disperse farther.
+This affects the critical harvesting rate: at a low climate velocity, we can more severely harvest populations that have a shorter dispersal distance than those that disperse farther, whereas at a high climate velocity, we can more aggressively harvest populations that disperse farther.
 
@@
 The stressors interact most strongly at high values, shortly before they drive the population extinct.
-However, the excess loss in biomass is low, making it difficult to distinguish positive synergy from additive interactions.
-% this is my awkward sentence, any ideas?
+However, the excess loss in biomass is very low, meaning the stressors interact more or less additively.
 We found similar analytical results for a sinusoidal dispersal kernel, which indicates that this result is robust to changes in the dispersal kernel.
@@
 
-Without any management strategies, we found that the more severely we harvest the population, the slower the rate of environmental shift will suffice to drive the population extinct.
-However, when we put thresholds in place, a small population can always escape harvesting pressure and the critical rate of environmental shift $c^*$ no longer depends on the harvesting rate (Figure \ref{management}).
-In other words, as long as there is some threshold below which harvesting is not allowed, there is a constant critical rate of environmental shift that only depends on the growth rate, length of the viable patch, and average dispersal distance.
+Without any management strategies, we found that the more severely we harvest the population, a slower climate velocity will suffice to drive the population extinct.
+However, when we put thresholds in place, a small population can always escape harvesting pressure and the critical climate velocity $c^*$ no longer depends on the harvesting rate (Figure \ref{management}).
+In other words, as long as there is some threshold below which harvesting is not allowed, there is a constant critical climate velocity that only depends on the growth rate, length of the viable patch, and average dispersal distance.
 
-We also examined the effect of marine protected  areas (MPAs) on the population's persistence to see whether it might extend the range of harvesting and  climate change parameters where the fish population could survive.
-We found that with either type of MPA strategies examined (many small versus few large), the population withstood combinations of higher climate velocities and harvesting rates, although the critical climate velocity $c^*$ was unchanged (Figure \ref{management}).
-We also found that the spacing and size of the MPAs changed population dynamics.
-MPAs spaced more than one average dispersal distance apart resulted in large oscillations of population biomass at low climate velocities relative to small, closely spaced, MPAs.
-For both of these MPA strategies we find that as climate velocities increase, the mean population abundance declines but the population experiences less extreme oscillations in abundance, which results in a population bounded farther from possible extinction in a stochastic environment.
-Additionally, large MPAs were able to increase equilibrium biomass under relatively high harvest and intermediate speeds relative to a population harvested at the same rate, but at a slower speed.
+With either type of MPA strategies examined (many small versus few large), the population withstood combinations of higher climate velocities and harvesting rates (Figure \ref{management}).
+At lower climate velocities, MPAs spaced more than one average dispersal distance apart resulted in larger fluctuations of population biomass relative to small, closely spaced, MPAs.
+As climate velocities increase, for both MPA strategies, the mean population abundance declines but the population experiences less extreme oscillations in abundance.
+Since minimum population biomass is increased, the population would be more buffered from possible extinction in a stochastic environment.
+%Additionally, large MPAs were able to increase equilibrium biomass under relatively high harvest and intermediate speeds relative to a population harvested at the same rate, but at a slower speed.
 This effect disappears at faster climate velocities and is not present in the other, many-small MPA, strategy we tested.
+%I don't quite follow what I commented out above. -Eleanor
 
@@
 
-For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.
-More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.
+For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the climate velocity, the better a population can adjust to harvest and climate change.
+More interestingly, we found a negative relationship between the critical harvesting rate and the climate velocity.
 That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
@@
 
-Our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others.
-As shown in Figure \ref{baseline}, species with a higher reproductive rate and a longer average dispersal distance will better track a high rate of environmental shift relative to a species that has a low reproductive rate and short dispersal distance.
+Our results suggest that particular combinations of harvesting and climate velocity will affect some species more than others.
+As shown in Figure \ref{baseline}, species with a higher reproductive rate and a longer average dispersal distance will better track a high climate velocity relative to a species that has a low reproductive rate and short dispersal distance.
 The finding that a higher reproductive rate can sustain higher climate velocities and harvesting rates is intuitive, especially because harvesting rate and reproductive rate cancel each other out.
@@
 
-\noindent {\bf Figure \ref{baseline}}: (a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+\noindent {\bf Figure \ref{baseline}}: (a) The critical harvesting rate on the y-axis as a function of the climate velocity on the x-axis.
 Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
@@
 (b)
-The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis.
 These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
@@
 \noindent {\bf Figure \ref{Synergy}}: Positive synergy between the two stressors.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
+The x-axis shows the climate velocity, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
 This excess loss, on the order of $.001$, is small in comparison to the total biomass, which can be as large as $20$.
@@
 
-\noindent {\bf Figure \ref{management}}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+\noindent {\bf Figure \ref{management}}: The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis with and without management strategies.
 (a) No management.
@@
 
-\caption{%(a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.
+\caption{%(a) The critical harvesting rate on the y-axis as a function of the climate velocity on the x-axis.
 Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.
 Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.
-(b) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis.
+(b) The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis.
 These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
@@
 %Positive synergy between the two stressors.
-The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
+The x-axis shows the climate velocity, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
 These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
@@
 \caption{
-%The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.
+%The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis with and without management strategies.
 (a) No management.
@@
 \\ $L$ & patch length
-\\ $c$ & rate of environmental shift
+\\ $c$ & climate velocity
 \end{tabular}

### 2014-03-03 Writing/MS.tex
proofread through discussion

@@
 The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.
-Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.
+Here we have studied
+ a general model that incorporates dispersal and reproduction to examine how climate and harvesting interact in their effects on species persistence and biomass.
 
 For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the climate velocity, the better a population can adjust to harvest and climate change.
-More interestingly, we found a negative relationship between the critical harvesting rate and the climate velocity.
+Further, we found a negative relationship between the critical harvesting rate and the climate velocity.
 That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.
-This is an indication of an interaction between the stressors.
+This is evidence that the stressors interact since each stressor's ability to drive the population extinct depends on the severity of the other stressor.
 
@@
 We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.
-We found similar results from the analytically derived biomass and the simulation derived biomass.
-This indicates that this result is robust to changes in the dispersal kernel.
 We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \cite{Crainetal2008, DarlingCote2008,Nyeetal2013}.
-We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.
 In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.
@@
 Our results suggest that particular combinations of harvesting and climate velocity will affect some species more than others.
-As shown in Figure \ref{baseline}, species with a higher reproductive rate and a longer average dispersal distance will better track a high climate velocity relative to a species that has a low reproductive rate and short dispersal distance.
+Species with a higher reproductive rate and a longer average dispersal distance will better track a high climate velocity relative to a species that has a low reproductive rate and short dispersal distance (Figure \ref{baseline}).
 The finding that a higher reproductive rate can sustain higher climate velocities and harvesting rates is intuitive, especially because harvesting rate and reproductive rate cancel each other out.
@@
 Finding that species with shorter generation times can better keep up with shifts in climate is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species \citep{Perryetal2005}.
-While higher reproductive rates improved a population's ability to persist, there was a tradeoff in increasing dispersal distances.
+While higher reproductive rates improved a population's ability to persist, increasing dispersal distances did not necessarily.
 At low speeds, we found that a short dispersal dispersal distance improved the maximum harvesting rate a population could sustain while at higher speeds a longer dispersal distance improved the maximum climate velocity in which the population could persist (Figure \ref{baseline}).
-This tradeoff is due to the proportion of dispersing  offspring at time step $t$ which lands within the patch at time step $t + 1$.
-When climate is shifting slowly, a large dispersal distance sends most offspring ahead of the patch, while with faster climate velocities a long dispersal distance allows the population to make it to the new patch (Figure \ref{baseline}).
+This is because when climate is shifting slowly, a large dispersal distance sends most offspring ahead of the patch, while with faster climate velocities a long dispersal distance allows the population to make it to the new patch (Figure \ref{baseline}).
 Thus climate velocity will selectively favor species with dispersal distances best matched to the rate of shift.

### 2014-03-03 Writing/Appendix.tex
proofread appendix;

@@
 If the trivial traveling pulse is unstable, even very small populations will persist or grow and avoid crashing back to the trivial pulse.
-To evaluate stability a traveling pulse, we introduce a small perturbation to the traveling pulse $n^*(\overline{x})$ and see if this perturbation grows or shrinks over time:
+To evaluate the stability of a traveling pulse, we introduce a small perturbation to the traveling pulse $n^*(\overline{x})$ and see if this perturbation grows or shrinks over time:
 
@@
 If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\R$ and $u:[-L/2,L/2]\to\R$, then the perturbation grows in time if and only if $\lambda >1$.
-Using Equation (\ref{perturb}), we can rewrite $\xi_t(x)$,
+Using Equation (\ref{perturb}), we can rewrite $\xi_{t+1}(x)$,
 \begin{align*}
@@
 
+\bibliography{fish}
+
 \end{document}

### 2014-03-03 Writing/MS.tex
fine edits

@@
 When the climate changes, so does the location of habitats suitable for an organism's survival and reproduction.
-This change does not occur in isolation but rather appears on a background of other disturbances, making the study of interactions between stressors important.
-In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we analyzed an integrodifference model that explicitly included the mechanisms of dispersal and reproduction.
-We found how the critical rates of harvesting and climate velocity that suffice to drive the population extinct depend on the growth rate and dispersal kernel of the population.
-We measured the interaction between the stressors and found that the disturbances interact nearly additively, with low positive synergy only at the greatest harvest rates and climate velocity that almost drive the population extinct.
-We also introduced two conservation techniques into simulations of the population model, threshold harvest rules and marine protected areas (MPAs), and found that under some circumstances these approaches could be effective management tools as they mitigate the interaction between the two stressors.
+This change does not occur in isolation but appears on a background of other disturbances, making the study of interactions between stressors important.
+In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we analyze an integrodifference model that explicitly includes the mechanisms of dispersal and reproduction.
+We show how the critical rates of harvesting and climate velocity that suffice to drive the population extinct depend on the growth rate and dispersal kernel of the population.
+We measure the interaction between the stressors and find the disturbances interact nearly additively in the parameter space that results in a stable population, with low positive synergy present only at the greatest harvest rates and climate velocity.
+Using simulations, we introduce two conservation techniques, threshold harvest rules and marine protected areas (MPAs), and show that these approaches can be effective management tools as they can mitigate the interaction between the two stressors.
+
 \hspace{10cm}
@@
 
-Many stressors can disturb an ecosystem, and ecologists have quantified the consequences of many of these of perturbations \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
+A number of stressors can disturb an ecosystem, and ecologists have quantified the consequences of many of these of perturbations \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}.
 Less work, however, has been done to measure the effects of multiple stressors and the interactions between them.
-If disturbances interact synergistically, a perturbation that has little effect when it occurs individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008,Nyeetal2013,Gurevitchetal2000}.
-In the most extreme (and worrying) cases, synergistic interactions between multiple stressors will drive a population extinct even though it could persist in the face of any single stressor (e.g.
+If disturbances interact synergistically, a perturbation that has little effect when occurring individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008,Nyeetal2013,Gurevitchetal2000}.
+In the most extreme (and worrying) cases, synergistic interactions between multiple stressors could drive a population extinct even though assessments of impacts individually predict the population to be robust (e.g.
 \citet{Pelletieretal2006}).
@@
 
-Climate change and fishing, two of the largest human impacts on the ocean \citep{Halpernetal2008}, provide an important case study of how disturbances interact in their effects on biological populations.
-Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue in the future \citep{Kelletal2005, Mackenzieetal2007}.
-These shifting species, and those likely to move in the future, are also subject to harvesting, among other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
+Climate change and fishing, two of the largest human impacts on the ocean \citep{Halpernetal2008}, provide an important example of ecological disturbances occurring in unison.
+Marine fish are already moving in response to climate change \citep{ Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011, Pinskyetal2013} and are projected to continue in the future \citep{Kelletal2005, Mackenzieetal2007}.
+These shifting species are also subject to harvesting, among other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.
 Previous empirical work has found synergistic interactions between overfishing and temperature-driven range shifts \citep{Lingetal2009} and synergistic interactions between warming temperatures, harvesting and connectivity have been identified in microcosm experiments \citep{Moraetal2007}.
@@
 These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperature, rainfall, or salinity, to predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}.
-Despite these models' widespread adoption, many papers have criticized SDMs as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
+Despite these models' widespread adoption, many authors have criticized SDMs as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
 Recent work on range shifts has addressed some of these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}.
@@
 
-Work on the joint impacts of climate and fishing often considers climate fluctuations (large anomalies around the mean) rather than directional changes in climate \citep{WaltersParma1996, KingMcFarlane2006}.
+Previous work has considered the joint impacts of climate and fishing, however these studies consider climate fluctuations (large anomalies around the mean) rather than directional shifts in temperature \citep{WaltersParma1996, KingMcFarlane2006}.
 When studies consider the effects of climate-driven range shifts on fishing, the models are typically case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}.
-These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, however these models are so complex that it makes understanding the relative importance of particular drivers, disturbances, and interactions difficult (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances).
+These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, but the complexity of these models makes it difficult to  understand the relative importance of particular drivers, disturbances, and interactions (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances).
 
-Here we extended a previously studied model of a fish population subject to climate-driven range shift by also considering harvesting pressure.
+Here we extend a previously studied model \citep{ZhouKot2011} to a fish population subject to climate-driven range shift and harvesting pressure.
 The model explicitly included reproduction and dispersal, two mechanistic processes central to species' responses to climate and fishing.
 Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.
-We found the critical harvesting rate and climate velocity that drive the population extinct and how these critical rates depend on one another.
-We also found that climate-driven range shifts and fishing interact nearly additively, with low positive synergy at more extreme levels of the stressors.
+We find the critical harvesting rate and climate velocity that drive the population extinct and how these critical rates depend on one another.
+We also show that climate-driven range shifts and fishing interact nearly additively, with low positive synergy at more extreme levels of the stressors.
 
-We also examined the efficacy of two different types of management strategies: threshold harvesting rules and marine protected areas (MPAs).
+We also examine the efficacy of two different types of management strategies: threshold harvesting rules and marine protected areas (MPAs).
 MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat shifts rapidly.
 Previous work has suggested protected areas can be a key form of climate insurance and can provide stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}.
-We found that threshold harvesting rules remove the interaction between harvesting rates and climate velocity and that MPAs can help a species persist with higher harvesting pressure and slightly increase the maximum climate velocity with which a species can keep up.
+We find that threshold harvesting rules remove the interaction between harvesting rates and climate velocity and that MPAs can help a species persist with higher harvesting pressure and slightly increase the maximum climate velocity with which a species can keep up.
 
@@
 We studied a model of the dynamics of a fish population constrained to a single, one-dimensional habitat patch by their inability to reproduce outside of that area, as introduced by \cite{ZhouKot2011}.
-This viable habitat patch (here after `patch') shifts at a fixed velocity and harvest occurs at each point in space along the entire one-dimensional world.
-We first determined the harvesting rate climate velocity that would drive the population extinct (hereafter the critical harvesting rate and critical climate velocity), and then measured synergy by calculating the drop in biomass caused by each stressor both individually and together.
-We finally added threshold harvesting rules and marine protected areas (MPAs) in numerical simulations of the model to determine how these management strategies affect population persistence.
+This viable habitat patch (hereafter `patch') shifts at a fixed velocity and harvest occurs at each point in space along the entire one-dimensional world.
+We first analytically determined the harvesting rate climate velocity that would drive the population extinct (hereafter the critical harvesting rate and critical climate velocity), and then measured synergy by calculating the drop in biomass caused by each stressor both individually and together.
+We then added threshold harvesting rules and marine protected areas (MPAs) in numerical simulations of the model to determine how these management strategies affect population persistence.
 
@@
 We extend this model by additionally subjecting the adults to harvesting before they produce offspring so that only a proportion of the fish survive to reproduce.
-These processes-- recruitment, harvesting, and dispersal-- are incorporated into an integrodifference model to describe how the population changes over time.
+These processes -- recruitment, harvesting, and dispersal -- are incorporated into an integrodifference model to describe how the population changes over time.
 If $n_t(x)$
@@
 
-\noindent where $h$ is the proportion of adults harvested, $f(n)$ is the recruitment function giving the number of 
-offspring produced by a population of size $n$ (accounting for density dependence), $k(x-y)$ is the dispersal kernel giving the probability of a  larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it  shifts across space.
-We provide a list of variables and functions in Table \ref{variables}.
-We used a Beverton-Holt recruitment function,
+\noindent where $h$ is the proportion of adults harvested, $k(x-y)$ is the dispersal kernel giving the probability of a  larva traveling from position $y$ to position $x$, $L$ is the length of the patch, and $c$ is the rate at which it  shifts across space.
+We used a Beverton-Holt stock-recruitment function for $f(n)$,
 
-\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}.\]
+\[f(n_t)=\frac{R_0n_t}{1+\left(\frac{R_0-1}{K}\right)n_t}\]
 
-\noindent  Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.
+\noindent  which gives the number of offspring produced by a population of size $n$ (accounting for density dependence).
+Here $R_0$ is the intrinsic growth rate and $K$ is carrying capacity (see table 1 for a full description of parameters and functions).
+
+Analyzing this kind of model becomes easier if the dispersal kernel is separable into its dependence on the source of larvae and its dependence on the destination of the larvae, i.e.~if there are functions $a_i, b_i$ such that $k(x- y) = \sum^\infty_{i=1} a_i(x)b_i(y)$.
 In our analyses, as in \cite{Latore:1998fk}, we used the separable Gaussian kernel given by
@@
 
-\noindent To derive analytical expressions, we approximated the kernel, as described Appendix A.3.
-Analytical results for a separable sinusoidal kernel are also described in Appendix A.4.  
+\noindent To derive analytical expressions, we approximated the kernel, as described Appendix A.3, and analytical results for a separable sinusoidal kernel are also described in Appendix A.4.
 We used simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
+%% Why did we choose a Laplace for simulations??
 
@@
 Regardless of the exact functional form of the recruitment function, the critical parameter in determining population persistence is how quickly recruitment increases when the population size is near (but above) $0$, which is equivalent to the intrinsic growth rate $R_0=f'(0)$.
-For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses $\langle d \rangle$ and the intrinsic growth rate $R_0$; properties of the environment-- the length of the viable patch $L$ and how quickly the environment shifts $c$; and the harvesting rate $h$.
+For each kernel, the population's ability to persist depends on properties of the population itself -- the expected distance a larva disperses $\langle d \rangle$ and the intrinsic growth rate $R_0$; properties of the environment -- the length of the viable patch $L$ and how quickly the environment shifts $c$; and the harvesting rate $h$.
+%% A little confusing.
+First you say that the critical parameter for persistence is equivalent to the intrinsic growth rate, and then list a bunch of other parameters that are also important for determining it.
 For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that
@@
 \noindent a commonly used model of larval dispersal \citep{Pinsky:2011fk}.
-Second, we implemented two management strategies, threshold rules and MPAs, to understand their effect on population persistence and on the synergistic interactions between stressors.
-For every simulation, in the first generation, we seeded the world with 50 individuals at a single point, as in \cite{ZhouKot2011}.
+Second, we implemented two management strategies, threshold rules and MPAs, to examine their effect on population persistence and on the interactions between stressors.
+For every simulation we seeded the world with 50 individuals at a single point, as in \cite{ZhouKot2011}.
 We first ran through 150 generations in order for the population to reach equilibrium without harvesting or climate shift.
@@
 If the population exceeded the threshold, then we harvested all the `surplus' individuals.
-
 We introduce networks of MPAs into our simulations by designating segments of space where the harvesting rate was equal to $0$.
-MPAs are typically designed to meet either fishery management  or conservation goals \citep{Agardy1994, HollandBrazee1996, Gainesetal2010a}.
+MPAs are typically designed to meet either fishery management  or conservation goals \citep{Agardy1994, HollandBrazee1996, Gainesetal2010a}, thus their spacing and size differ.
 Fisheries-oriented MPAs are often designed such that they maximize adult spillover into fishable areas by creating many small reserves closely spaced \citep{HastingsBotsford2003,Gaylordetal2005, Gainesetal2010a}.
@@
 Conservation-oriented MPAs seek to reduce adult spillover by creating fewer larger protected areas \citep{HastingsBotsford2006, Gainesetal2010a}.
-To mimic this scheme.
-we implemented MPAs with a length of $4$ times the average dispersal distance and a distance of $8$ times the average dispersal distance between them \citep{Lockwoodetal2002}.
+To mimic this scheme we implemented MPAs with a length of $4$ times the average dispersal distance and a distance of $8$ times the average dispersal distance between them \citep{Lockwoodetal2002}.
 
@@
 Conversely, as the harvesting rate $h$ increases, the critical climate velocity $c^*$ decreases (Figure \ref{baseline}).
-This means that as harvesting pressure increases, it becomes increasingly easy for a shifting environment to drive the population extinct.
+Thus as harvesting pressure increases, it becomes increasingly easy for a shifting environment to drive the population extinct.
 
@@
 Where positive synergy exists, a doubly stressed population loses more biomass than we would predict from either stressor individually.
-The stressors interact most strongly at high values, shortly before they drive the population extinct.
-However, the excess loss in biomass is very low, meaning the stressors interact more or less additively.
+The stressors interact most strongly at high harvest and climate velocity rates, shortly before they drive the population extinct.
+However, the synergistic  loss in biomass is very low, meaning that these stressors interact more or less additively.
 We found similar analytical results for a sinusoidal dispersal kernel, which indicates that this result is robust to changes in the dispersal kernel.
@@
 As climate velocities increase, for both MPA strategies, the mean population abundance declines but the population experiences less extreme oscillations in abundance.
-Since minimum population biomass is increased, the population would be more buffered from possible extinction in a stochastic environment.
-%Additionally, large MPAs were able to increase equilibrium biomass under relatively high harvest and intermediate speeds relative to a population harvested at the same rate, but at a slower speed.
-This effect disappears at faster climate velocities and is not present in the other, many-small MPA, strategy we tested.
-%I don't quite follow what I commented out above. -Eleanor
+Since minimum population biomass is increased, the population is a larger buffer to possible extinction in a stochastic environment.
 
@@
 
-Understanding interactions among disturbances will help to design management for populations subjected to these disturbances.
+Understanding interactions among disturbances will help to design management for populations subjected to these stressors.
 The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.
-Here we have studied
+Here we have analyzed
  a general model that incorporates dispersal and reproduction to examine how climate and harvesting interact in their effects on species persistence and biomass.
 
-For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the climate velocity, the better a population can adjust to harvest and climate change.
+For each dispersal kernel we studied, we found that the higher the growth rate and the more the mean dispersal distance matches the climate velocity, the better a population can persist under harvesting and climate change.
 Further, we found a negative relationship between the critical harvesting rate and the climate velocity.
@@
 Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.
-
 Worryingly, we find the highest synergy in those populations whose persistence is most tenuous.
@@
 synergistic effects were quite small.
-%we need a transition here, feels abrupt
-Synergy between harvesting and the effects of climate change has been identified in experimental populations \citep{Moraetal2007}, in specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
-In the experimental populations, synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}.
+
+Despite the absence of synergy in our analysis, whether or not we should assume that synergy is unlikely to exist between climate velocity and harvesting remains to be seen.
+Synergy between harvesting and the effects of climate change has been identified in experimental populations \citep{Moraetal2007}, and observationally at both the population \citep{Planque:2010uq}, and ecosystem level \citep{Kirby:2009fk,Planque:2010uq}.
+Some of the discrepancies may be due to the ways in which climate was measured.
+In the experimental populations, effects of climate were mimicked by increased temperatures, and organisms were unable to relocate to thermal optima.
+Synergy was identified between warming and harvesting but not between habitat fragmentation \citep{Moraetal2007}, which may be more similar to the range shift we analyzed in our theoretical model.
 While we did find (very) low levels of positive synergy, we did not find as much as predicted from these empirical studies.
@@
 
-Our results suggest that particular combinations of harvesting and climate velocity will affect some species more than others.
+Absence of synergy does not mean absence of effect, and our results suggest that particular combinations of harvesting and climate velocity will affect some species more than others.
 Species with a higher reproductive rate and a longer average dispersal distance will better track a high climate velocity relative to a species that has a low reproductive rate and short dispersal distance (Figure \ref{baseline}).
@@
 Thresholds have this effect as the management approach effectively prevents harvesting of the leading edge, which allows colonization to occur as if these individuals were moving into un-fished areas.
-It's interesting to note that novel, low abundance species are commonly unregulated in fisheries systems; so in order to decouple the additive effects of harvest and climate change, management would have to reverse this paradigm by allowing no harvest of new species until they had become established.
+It's interesting to note that novel, low abundance species are commonly unregulated in fisheries systems; so in order to decouple the additive effects of harvest and climate change, management would have to reverse this paradigm by allowing no harvest of shifting species until they had become established in new areas.
 
@@
 Our results show that both threshold and MPAs increase the equilibrium biomass at a given climate velocity, which support their use as a tool to ameliorate the effect of climate velocity.
-However, for MPAs, the details mater: few, large MPAs caused increased variability at low climate velocities while many smaller MPAs maintained a population bounded farther from extinction.
+However for MPAs the details mater: few, large MPAs caused increased variability at low climate velocities while many smaller MPAs maintained a population bounded farther from extinction.
 Finally, with sufficiently high harvesting pressure, few, large MPAs rescued populations at intermediate speeds.
@@
 However this effect disappears as speed continues to increase, suggesting that understanding the relationship between climate velocity, dispersal distance and reproductive rate are important parameters in designing management strategies effective under both climate change and harvesting pressure.
-
-While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.
-This is encouraging evidence that a single set of of management practices may help to protect marine populations from both harvesting and climate change.
 
@@
 This is especially true for management strategies which may result in unanticipated effects such as large fluctuations associated with big, distant MPAs shown here.
+While the management strategies only change harvesting practices and do not directly address the effects of climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement harvesting rules and place protected areas.
+This is encouraging evidence that a single set of of management practices may help to protect marine populations from both harvesting and climate change.
 

### 2014-03-03 fish_figures.R
eqbiomass data and working on figures

@@ -28,11 +28,60 @@ line=5
 lwd=5
 yaxs='i'
 xaxs='i'
-cex.lab=1.5
+cex.lab=15
 cex.axis=1
 line=1
 oma=c(0,0,0,2)
 fig.width = 3	# for ecological applications, text needs to be at least 6 point font size
+
+############
+#############
+#relationship between critical harvesting rate and critical climate velocity
+
+cvals=matrix(seq(0,1,by=0.01),nrow=1)
+
+nr=3
+ns=2
+rvals=c(rep(3,ns),rep(5,ns),rep(10,ns))
+rcols=c(rep('black',ns),rep('red',ns),rep('blue',ns))
+xvals=c(.1,.25)
+sigvals=rep(pi/2*xvals,nr)
+sigltys=rep(c(1,2),nr)
+l=length(rvals)
+
+lwd=1
+cex.lab=.75
+cex.axis=.5
+cex.legend=.5
+tck=-.02
+mar=c(3,3,1,1)
+mgp=c(1.5,.25,0)
+oma=c(0,0,0,0)
+
+quartz()
+postscript("r2plots_rev.eps", horizontal = FALSE, onefile = FALSE, paper = "special", width = fig.width,height=fig.width,bg="white")
+
+par(mar=mar,mgp=mgp,oma=oma)
+plot(cvals,cvals,col='white',ylim=c(0,1),xlab="Climate velocity",ylab="Critical harvesting rate",bty='l',axes=FALSE,xaxs='i',yaxs='i',cex.lab=cex.lab)
+
+for(i in 1:l){
+	r=rvals[i]
+	sig2=sigvals[i]
+	h=apply(cvals,2,hstar_g,r=r,sig2=sig2)
+	lines(cvals,h,col=rcols[i],lty=sigltys[i],lwd=lwd)
+}
+
+urvals=unique(rvals)
+urcols=unique(rcols)
+usigvals=unique(sigvals)
+usigltys=unique(sigltys)
+
+axis(1,lwd=lwd,cex.axis=cex.axis,tck=tck)
+axis(2,lwd=lwd,cex.axis=cex.axis,tck=tck)
+
[…]

### 2014-03-03 fish_figures.R
eqbiomass data

@@ -54,12 +54,12 @@ cex.lab=.75
 cex.axis=.5
 cex.legend=.5
 tck=-.02
-mar=c(3,3,1,1)
-mgp=c(1.5,.25,0)
+mar=c(2,2,1,1)
+mgp=c(1,.25,0)
 oma=c(0,0,0,0)
 
 quartz()
-postscript("r2plots_rev.eps", horizontal = FALSE, onefile = FALSE, paper = "special", width = fig.width,height=fig.width,bg="white")
+postscript("critical_rates.eps", horizontal = FALSE, onefile = FALSE, paper = "special", width = fig.width,height=fig.width,bg="white")
 
 par(mar=mar,mgp=mgp,oma=oma)
 plot(cvals,cvals,col='white',ylim=c(0,1),xlab="Climate velocity",ylab="Critical harvesting rate",bty='l',axes=FALSE,xaxs='i',yaxs='i',cex.lab=cex.lab)
@@ -79,7 +79,7 @@ usigltys=unique(sigltys)
 axis(1,lwd=lwd,cex.axis=cex.axis,tck=tck)
 axis(2,lwd=lwd,cex.axis=cex.axis,tck=tck)
 
-legend(.63,1.03,col=c(urcols,rep('black',3),'white'),lty=c(rep(1,3),usigltys,1),lwd=lwd,legend=c(bquote(R[0]==.(urvals[1])),bquote(R[0]==.(urvals[2])), bquote(R[0]==.(urvals[3])),expression(symbol("\341")*d*symbol("\361")== .1),expression(symbol("\341")*d*symbol("\361")== .25),expression()),bty='n',cex=cex.legend)
+legend(.7,.97,col=c(urcols,rep('black',3),'white'),lty=c(rep(1,3),usigltys,1),lwd=lwd,legend=c(bquote(R[0]==.(urvals[1])),bquote(R[0]==.(urvals[2])), bquote(R[0]==.(urvals[3])),expression(symbol("\341")*d*symbol("\361")== .1),expression(symbol("\341")*d*symbol("\361")== .25),expression()),bty='n',cex=cex.legend)
 
 graphics.off()
 

### 2014-03-03 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd
messing with figures, fine edits

@@
 require(reshape2)
+require(RColorBrewer)
 ```
@@
 # for synced online github data
-   x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2014-02-20.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxNC0wMi0yMC5jc3YiLCJleHBpcmVzIjoxMzkzNjIzMjk4fQ%3D%3D--b299cd8c20eb28a2d626a3d8d467685b75986fbf")
+x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzYwfQ%3D%3D--5fa73caaafffae7c95d930f61def04ee9b96930a")
 sim = read.csv(textConnection(x))
 
-  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-02-20.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMi0yMC5jc3YiLCJleHBpcmVzIjoxMzkzNjIzMDUzfQ%3D%3D--d793f2abea78846c6b8535d73e76d5836207dea6")
+fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzQ2fQ%3D%3D--9b6ceb0092173222543c95381dfbb0f42f0c884f")
 fishmpa = read.csv(textConnection(fish))
 
-  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-02-20.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMi0yMC5jc3YiLCJleHBpcmVzIjoxMzkzNjIzMDEzfQ%3D%3D--9f521e92c8abdd1c4a8a547576e661ea41d342d6")
+cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-03-01.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMy0wMS5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzIwfQ%3D%3D--5f4f584256a4175eebe2ef5465e83c92ebe65cc0")
 consmpa = read.csv(textConnection(cons))
 
+thresh = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/Thresh_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL1RocmVzaF8yMDE0LTAzLTAyLmNzdiIsImV4cGlyZXMiOjEzOTQ0ODIyMTV9--13d78257cda683d18eb8ecc71dcf0f4e6101b41d")
+threshz = read.csv(textConnection(thresh))
+
 # combine all data into one dataframe
-sim$management = rep("noManagement",nrow(sim))
-consmpa$management = rep("consMPA", nrow(consmpa))
-fishmpa$management = rep("fishMPA", nrow(fishmpa))
-
+sim$management = rep("No MPAs",nrow(sim))
+consmpa$management = rep("Few Large MPAs", nrow(consmpa))
+fishmpa$management = rep("Many Small MPAs", nrow(fishmpa))
+threshz$management = rep("Thresholds", nrow(threshz))
 allData <- rbind(sim,consmpa,fishmpa)
@@
 ```{r Compare_Management,echo = FALSE, fig.width=10, fig.height=10}
-contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,3), drape=TRUE)
+sb <- trellis.par.get("strip.background") 
+sb[["col"]][1] <- "lightgray"
+trellis.par.set("strip.background", sb)
+
+ltheme <- canonical.theme(color = FALSE) ## in-built B&W theme 
+ltheme$strip.background$col <- "transparent" ## change strip bg 
+lattice.options(default.theme = ltheme) ## set as default
+
+lw <- list(left.padding = list(x = 0, units = "inches")) 
+lw$right.padding <- list(x = -0.1, units = "inches") 
+lh <- list(bottom.padding = list(x = 0, units = "inches")) 
+lh$top.padding <- list(x = -0.2, units = "inches")
+
+lattice.options(layout.widths = lw, layout.heights = lh)
+
+pdf(file="foo.pdf",width=3,height=6)
+contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,3), drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"))
+dev.off()
+
+contourplot(Equil.pop ~ speed * -thresh | factor(management), data = threshz, region=TRUE, drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"),scales=list( y=list(at=seq(-8,0,2),labels=round(rev(unique(threshz$thresh)/max(threshz$thresh)),3))),ylab=list(label="proportion of unharvested population below threshold"))
+
+# replicating our figure plots
+
+# build dataframe
 

### 2014-03-03 alt_fish_figures.R

@@ -1,0 +1,49 @@
+# for synced online github data
+x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzYwfQ%3D%3D--5fa73caaafffae7c95d930f61def04ee9b96930a")
+sim = read.csv(textConnection(x))
+
+fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzQ2fQ%3D%3D--9b6ceb0092173222543c95381dfbb0f42f0c884f")
+fishmpa = read.csv(textConnection(fish))
+
+cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-03-01.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMy0wMS5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzIwfQ%3D%3D--5f4f584256a4175eebe2ef5465e83c92ebe65cc0")
+consmpa = read.csv(textConnection(cons))
+
[…]

### 2014-03-03 Writing/MS.tex
messing with figures, fine edits
fixing tense

@@
 This change does not occur in isolation but appears on a background of other disturbances, making the study of interactions between stressors important.
-In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we analyze an integrodifference model that explicitly includes the mechanisms of dispersal and reproduction.
-We show how the critical rates of harvesting and climate velocity that suffice to drive the population extinct depend on the growth rate and dispersal kernel of the population.
-We measure the interaction between the stressors and find the disturbances interact nearly additively in the parameter space that results in a stable population, with low positive synergy present only at the greatest harvest rates and climate velocity.
-Using simulations, we introduce two conservation techniques, threshold harvest rules and marine protected areas (MPAs), and show that these approaches can be effective management tools as they can mitigate the interaction between the two stressors.
+In order to understand how two disturbances, range shift and harvesting, interact and affect population persistence, we analyzed an integrodifference model that explicitly includes the mechanisms of dispersal and reproduction.
+We have shown how the critical rates of harvesting and climate velocity that suffice to drive the population extinct depend on the growth rate and dispersal kernel of the population.
+We measured the interaction between the stressors and find the disturbances interact nearly additively in the parameter space that results in a stable population, with low positive synergy present only at the greatest harvest rates and climate velocity.
+Using simulations, we introduced two conservation techniques, threshold harvest rules and marine protected areas (MPAs), and have shown that these approaches can be effective management tools as they can mitigate the interaction between the two stressors.
 
@@
 \noindent {\bf Figure \ref{management}}: The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis with and without management strategies.
-(a) No management.
-(b) Threshold harvesting levels.
-(c) MPAs.
+(a) MPAs (b) Threshold harvesting levels.
 These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
@@
 \subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
+\includegraphics[width=1\textwidth]{plots/MPAs.pdf}
 \label{nomang}
@@
 \subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
-\end{subfigure}
-\begin{subfigure}{.33\textwidth}
-\subcaption{}
-\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
+\includegraphics[width=1\textwidth]{plots/Threshold.pdf}
 \end{subfigure}
@@
 %The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis with and without management strategies.
-(a) No management.
-(b) Threshold harvesting levels.
-(c) MPAs.
+(a) MPAs  (b) Threshold harvesting levels.
 These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.

### 2014-03-03 Writing/MS.tex
emmas comment on line 104

--- 2014-03-03 Writing/MS.tex
@@
 
-Regardless of the exact functional form of the recruitment function, the critical parameter in determining population persistence is how quickly recruitment increases when the population size is near (but above) $0$, which is equivalent to the intrinsic growth rate $R_0=f'(0)$.
-For each kernel, the population's ability to persist depends on properties of the population itself -- the expected distance a larva disperses $\langle d \rangle$ and the intrinsic growth rate $R_0$; properties of the environment -- the length of the viable patch $L$ and how quickly the environment shifts $c$; and the harvesting rate $h$.
-%% A little confusing.
-First you say that the critical parameter for persistence is equivalent to the intrinsic growth rate, and then list a bunch of other parameters that are also important for determining it.
+Regardless of its exact functional form, the only property of the recruitment function that determines whether or not a population can persist is how quickly recruitment increases when the population size is near (but above) $0$, which is equivalent to the intrinsic growth rate $R_0=f'(0)$.
+For each kernel, therefore, the population's ability to persist depends on properties of the population itself -- the expected distance a larva disperses $\langle d \rangle$ and the intrinsic growth rate $R_0$; properties of the environment -- the length of the viable patch $L$ and how quickly the environment shifts $c$; and the harvesting rate $h$.
 For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that

### 2014-03-03 fish_figures.R
emmas comment on line 104

@@ -22,16 +22,26 @@ emmaK=100
 
 #####
 #### plotting parameters
-cex.lab=3
-cex.axis=2
-line=5
-lwd=5
-yaxs='i'
-xaxs='i'
-cex.lab=15
-cex.axis=1
-line=1
-oma=c(0,0,0,2)
+# cex.lab=3
+# cex.axis=2
+# line=5
+# lwd=5
+# yaxs='i'
+# xaxs='i'
+# cex.lab=15
+# cex.axis=1
+# line=1
+# oma=c(0,0,0,2)
+
+lwd=1
+cex.lab=.75
+cex.axis=.5
+cex.legend=.5
+tck=-.02
+mar=c(2,2,1,1)
+mgp=c(1,.25,0)
+oma=c(0,0,0,0)
+
 fig.width = 3	# for ecological applications, text needs to be at least 6 point font size
 
 ############
@@ -48,15 +58,6 @@ xvals=c(.1,.25)
 sigvals=rep(pi/2*xvals,nr)
 sigltys=rep(c(1,2),nr)
 l=length(rvals)
-
-lwd=1
-cex.lab=.75
-cex.axis=.5
-cex.legend=.5
-tck=-.02
-mar=c(2,2,1,1)
-mgp=c(1,.25,0)
-oma=c(0,0,0,0)
 
 quartz()
 postscript("critical_rates.eps", horizontal = FALSE, onefile = FALSE, paper = "special", width = fig.width,height=fig.width,bg="white")
@@ -90,13 +91,13 @@ r=emmaR0
 K=emmaK
 sig2=1
 tol=.00001
-diff=.01
+diff=.25
 cvals=seq(0,1,by=diff)
 hvals=seq(0,1,by=diff)
 
 
-# ebm=eqbiomass(cvals,hvals,r,K,sig2,tol=tol)
-load("/Users/eleanorbrush/Desktop/ebm.rdata")
[…]

### 2014-03-03 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd

@@
 threshz$management = rep("Thresholds", nrow(threshz))
-allData <- rbind(sim,consmpa,fishmpa)
+allData <- rbind(consmpa,fishmpa)
 ```
@@
 ```{r Compare_Management,echo = FALSE, fig.width=10, fig.height=10}
+
+# lattice plot settings
+
 sb <- trellis.par.get("strip.background") 
@@
 
-pdf(file="foo.pdf",width=3,height=6)
-contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,3), drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"))
+# mpa plots
+
+pdf(file="MPAs.pdf",width=3,height=5)
+contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,2), drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"))
 dev.off()
 
-contourplot(Equil.pop ~ speed * -thresh | factor(management), data = threshz, region=TRUE, drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"),scales=list( y=list(at=seq(-8,0,2),labels=round(rev(unique(threshz$thresh)/max(threshz$thresh)),3))),ylab=list(label="proportion of unharvested population below threshold"))
+#threshold plot
 
-# replicating our figure plots
-
-# build dataframe
+pdf(file="Threshold.pdf",width=3,height=3)
+contourplot(Equil.pop ~ speed * -thresh | factor(management), data = threshz, region=TRUE, drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"),scales=list( y=list(at=seq(-8,0,2),labels=round(rev(unique(threshz$thresh)/max(threshz$thresh)),3))),ylab=list(label="proportion below threshold"))
+dev.off()
 

### 2014-03-04 Writing/MS.tex
changing comments
converting to a doc file
finished turing figures into eps and put them into the tex file

@@
 
-\begin{equation*}
-n_{t+1}(x)=\int^{\frac{L}{2}+ct}_{-\frac{L}{2}+ct}k(x-y)f((1-h)n_t(y))dy \label{integrodifference},
-\end{equation*}
+\[n_{t+1}(x)=\int^{\frac{L}{2}+ct}_{-\frac{L}{2}+ct}k(x-y)f((1-h)n_t(y))dy \label{integrodifference},\]
 
@@
 We used simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.
-%% Why did we choose a Laplace for simulations??
+%% Why did we choose a Laplace for simulations?
+Why not just do the same for both analytics and simulations?
 
@@
 
+\section{Tables}
+\begin{table}[h]
+\caption{Table of variables used in the text}
+\begin{tabular}{@{}lllllll@{}}
+  Variable & Definition
+\\\cmidrule{1-1} \cmidrule{2-2}   
+$n_t(x)$ & density of fish at position $x$ at time $t$
+\\ $n^*(\overline{x})$ & density of fish at equilibrium at position $\overline{x}$ relative to the patch 
+\\ $k(x-y)$ & dispersal kernel, the probability of larva traveling from position $y$ to position $x$
+\\ $\langle d \rangle $ & expected distance traveled by larva
+\\ $f(n)$ & recruitment function, the number of offspring produced by a population of size $n$
+\\ $R_0$ & intrinsic growth rate, $R_0=f'(0)$
+\\ $h$ & proportion of adults harvested
+\\ $L$ & patch length
+\\ $c$ & climate velocity
+\end{tabular}
+\label{variables}
+\end{table}
+
+\pagebreak
+
 \section*{Figure Legends}
@@
 The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.5$.
 \hspace{6in}
@@
 The x-axis shows the climate velocity, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-This excess loss, on the order of $.001$, is small in comparison to the total biomass, which can be as large as $20$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+This excess loss, on the order of $.05$, is small in comparison to the total biomass, which can be greater than $40$.
+These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.5$.
 \hspace{6in}
+
 \hspace{6in}
@@
 \subcaption{\label{rates}}
-\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
+\includegraphics[width=\textwidth]{plots/critical_rates.eps}
 \end{subfigure}
@@
 \subcaption{\label{biomass}}
-\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
+\includegraphics[width=\textwidth]{plots/eqbiomass.eps}
 \end{subfigure}
@@
 (b) The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.5$.
 These results are from an approximated Gaussian dispersal kernel with $L=1$.
@@
 \begin{center}
-\includegraphics[width=3in]{plots/synergy.pdf}
+\includegraphics[width=3in]{plots/synergy.eps}
 \caption{
@@
 The x-axis shows the climate velocity, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
+These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.5$.
 }
@@
 
-\section{Tables}
-\begin{table}[h]
-\caption{Table of variables used in the text}
-\begin{tabular}{@{}lllllll@{}}
-  Variable & Definition
-\\\cmidrule{1-1} \cmidrule{2-2}   
-$n_t(x)$ & density of fish at position $x$ at time $t$
-\\ $n^*(\overline{x})$ & density of fish at equilibrium at position $\overline{x}$ relative to the patch 
-\\ $k(x-y)$ & dispersal kernel, the probability of larva traveling from position $y$ to position $x$
-\\ $\langle d \rangle $ & expected distance traveled by larva
-\\ $f(n)$ & recruitment function, the number of offspring produced by a population of size $n$
-\\ $R_0$ & intrinsic growth rate, $R_0=f'(0)$
-\\ $h$ & proportion of adults harvested
-\\ $L$ & patch length
-\\ $c$ & climate velocity
-\end{tabular}
-\label{variables}
-\end{table}
-
-\pagebreak
-
 \end{spacing}

### 2014-03-04 Simluations/Aspatial_fast/Figures/variable_steps.R
examining consmpa, small fixes to sim

@@ -1,9 +1,7 @@
 # MPA simulation with variable numbers of steps for 'equilibrium' step. Set steps outside of this file
 
 # adding in subset of parameters, so instead of running Parameters_nothresh.R, am manually entering
-harvests = 0.1
-speeds = seq(0.02,0.1,.01)
-MPA = "cons"
+
 
 sims = vector("list", length(speeds))
 
@@ -71,6 +69,19 @@ for(q in 1:length(speeds)){
         roll.mean <- diff.rolls[length(diff.rolls)]
         print(T)
       }
+    }
+    
+    T = ncol(move)
+    # then run until the MPA.current matches the MPA.start so we compare populations at the same point in spatial cycle
+    req = 0   # this is the condition for stopping, can be modified if count gets too high
+    while(length(which(MPA.current!=MPA.start)) > req){
+      T = T + 1
+      next.pop <- m(n=move[,T-1], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
+      
+      move <- cbind(move, next.pop[[1]])
+      MPA.current <- next.pop[[3]]
+      print(T)
+      print(length(which(MPA.current!=MPA.start)))		  
     }
     
     # simulate for <steps> steps after equilibrium and take average

### 2014-03-04 Simluations/Aspatial_fast/load_data.R
updating data

@@ -1,0 +1,20 @@
+# Load data for figures
+
+# analytics
+
+x=getURL("https://raw.github.com/emfuller/MovingFish/38c6f3bb51dae267fe4056159609de100fb34826/eqbiomass_gaus.csv?token=6224444__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC8zOGM2ZjNiYjUxZGFlMjY3ZmU0MDU2MTU5NjA5ZGUxMDBmYjM0ODI2L2VxYmlvbWFzc19nYXVzLmNzdiIsImV4cGlyZXMiOjEzOTQ0ODQ5OTh9--d95944353679e18e44def40cd49ca41a22502039")
+ebm=read.csv(textConnection(x))
+
+# simulations
+
+x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDA4NjYyfQ%3D%3D--7f380bfb17f1cd58ff4f131134ed434700767def")
+	sim = read.csv(textConnection(x))
+
+	x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDA4NjQxfQ%3D%3D--2dba33c2046114811967e01914fc795e1771f425")
+	mpas = read.csv(textConnection(x))
+
[…]

### 2014-03-04 fish_figures.R

@@ -5,7 +5,7 @@ library(RCurl)
 setwd("/Users/eleanorbrush/Desktop")
 source("/Users/eleanorbrush/Desktop/biomass_synergy.R")
 source("/Users/eleanorbrush/Desktop/critical_parameters.R")
-
+source("load_data.R")
 
 ### parameters from Zhou
 r=3
@@ -96,8 +96,8 @@ cvals=seq(0,1,by=diff)
 hvals=seq(0,1,by=diff)
 
 
-x=getURL("https://raw.github.com/emfuller/MovingFish/38c6f3bb51dae267fe4056159609de100fb34826/eqbiomass_gaus.csv?token=6224444__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC8zOGM2ZjNiYjUxZGFlMjY3ZmU0MDU2MTU5NjA5ZGUxMDBmYjM0ODI2L2VxYmlvbWFzc19nYXVzLmNzdiIsImV4cGlyZXMiOjEzOTQ0ODQ5OTh9--d95944353679e18e44def40cd49ca41a22502039")
-ebm=read.csv(textConnection(x))
+# x=getURL("https://raw.github.com/emfuller/MovingFish/38c6f3bb51dae267fe4056159609de100fb34826/eqbiomass_gaus.csv?token=6224444__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC8zOGM2ZjNiYjUxZGFlMjY3ZmU0MDU2MTU5NjA5ZGUxMDBmYjM0ODI2L2VxYmlvbWFzc19nYXVzLmNzdiIsImV4cGlyZXMiOjEzOTQ0ODQ5OTh9--d95944353679e18e44def40cd49ca41a22502039")
+# ebm=read.csv(textConnection(x))
 
 #which params to work with
 H=sum(apply(ebm,2,max)>tol)+2
@@ -140,8 +140,8 @@ graphics.off()
 ##### wireframe of the equilibrium biomass as a function of c and h : SIMULATIONS
 
 # github copy
[…]

### 2014-03-04 fish_figures.R
tweaks to fish_figures

@@ -2,9 +2,8 @@ library(fields)
 library(lattice)
 library(colorspace)
 library(RCurl)
-setwd("/Users/eleanorbrush/Desktop")
-source("/Users/eleanorbrush/Desktop/biomass_synergy.R")
-source("/Users/eleanorbrush/Desktop/critical_parameters.R")
+source("biomass_synergy.R")
+source("critical_parameters.R")
 
 
 ### parameters from Zhou
@@ -34,13 +33,14 @@ emmaK=100
 # oma=c(0,0,0,2)
 
 lwd=1
+yaxs='i'
+xaxs='i'
 cex.lab=.75
 cex.axis=.5
 cex.legend=.5
 tck=-.02
 mar=c(2,2,1,1)
 mgp=c(1,.25,0)
-oma=c(0,0,0,0)
 
 fig.width = 3	# for ecological applications, text needs to be at least 6 point font size
 
@@ -62,7 +62,7 @@ l=length(rvals)
 quartz()
 postscript("critical_rates.eps", horizontal = FALSE, onefile = FALSE, paper = "special", width = fig.width,height=fig.width,bg="white")
 
-par(mar=mar,mgp=mgp,oma=oma)
+par(mar=mar,mgp=mgp)
 plot(cvals,cvals,col='white',ylim=c(0,1),xlab="Climate velocity",ylab="Critical harvesting rate",bty='l',axes=FALSE,xaxs='i',yaxs='i',cex.lab=cex.lab)
 
 for(i in 1:l){
@@ -86,7 +86,7 @@ graphics.off()
 
 ##########
 ##########
-##### wireframe of the equilibrium biomass as a function of c and h
+##### heatmap of the equilibrium biomass as a function of c and h
 r=emmaR0
 K=emmaK
 sig2=1
@@ -97,13 +97,13 @@ hvals=seq(0,1,by=diff)
 
 
[…]

### 2014-03-04 Simluations/Aspatial_fast/Figures/diagnose_consmpa.Rmd

--- 2014-03-03 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd
@@
+# Figuring out the weird contour plot for `consmpa`
+
 # Effect of MPAs
@@
 
-pdf(file="MPAs.pdf",width=3,height=5)
 contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,2), drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"))
-dev.off()
 
@@
 
-pdf(file="Threshold.pdf",width=3,height=3)
 contourplot(Equil.pop ~ speed * -thresh | factor(management), data = threshz, region=TRUE, drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"),scales=list( y=list(at=seq(-8,0,2),labels=round(rev(unique(threshz$thresh)/max(threshz$thresh)),3))),ylab=list(label="proportion below threshold"))
-dev.off()
 
@@
 
-Looks like conservation MPAs do the best, but there are some weird fluctuations at higher harvest rates, why is that? Can take slices to look more closely
+Looking at the weird variance in some of the `consmpa` simulations, try to target
 
-```{r variance, echo = FALSE, fig.width=10, fig.height=5}
-subData <- subset(allData,harvest == 0.04 )
-pd <- position_dodge(.1)
-ggplot(subData, aes(x = speed, y = Equil.pop))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()
+```{r}
+ ggplot(consmpa, aes(x=speed, y = Equil.pop)) + geom_line() + facet_wrap(~harvest)
 ```
 
-Trying the same plot but a higher rate of catch: $h = 0.10$ I get,
-
-```{r variance_high, echo = FALSE, fig.width=10, fig.height=5}
-subData <- subset(allData,harvest == 0.10 )
-pd <- position_dodge(.1)
-ggplot(subData, aes(x = speed, y = Equil.pop))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()
-```
+Looks like it's in `h=0.08,0.09, 0.1, 0.12`
 
@@
 
-```{r sub, echo=FALSE,fig.width=15, fig.height=5}
-subData <- subset(allData,harvest > 0.05 & harvest < 0.13 & management == "consMPA" & speed < 0.3)
-contourplot(Equil.pop ~ speed * harvest, data = subData, region = TRUE, drape = TRUE)
+```{r sub, echo=FALSE}
+subData <- subset(consmpa,harvest > 0.07 & harvest < 0.13 & speed < 0.3 & speed > 0.1)
+ggplot(subData, aes(x=speed, y = Equil.pop)) + geom_line() + facet_wrap(~harvest) + geom_point()
 ```
+And it's in different places depending on the harvest.. So the simulations to look at are
 
-The next step is to run a few of these simulations and save entire outputs (biomass over space/time). Will do $h = 0.10, s = 0.02-0.07$
++ `h = 0.08; s = 0.24-0.28`
++ `h = 0.09; s = 0.24-0.28`
++ `h = 0.10; s = 0.14-0.16`
++ `h = 0.12; s = 0.14-0.16`
+
+The next step is to run a few of these simulations and save entire outputs (biomass over space/time). Will do `h = 0.08, s = 0.24-0.28`
 
@@
 steps = 300   # setting number of steps over which to take equilibrium 
+harvests = 0.08
+speeds = seq(0.24,0.28,0.02)
+MPA = "cons"
 source("Figures/variable_steps.R")
+```
+What it looks like is that `h=0.08, s = 0.28` failed to converge. I finally killed it after 3821 generations.
+
+```{r}
 sims300 = sims
@@
 series300 = series
-notify("Simulation is finished!")
 ```
-After re-running the model, can see that it replicates the pattern of fluctuating total biomas
+After re-running the model, can see that it replicates the pattern of fluctuating total biomass and huge error bars around `s = 0.26`
 ```{r final-plot,echo=FALSE,message=FALSE,results='hide'}
@@
 ```
-And when you look at the time-series population trends, it's as we expected.
-All the populations have the same pattern, that they oscillate up and down depending on whether or not they're in a reserve, but the slowest speeds go through the biggest oscillations which means that the average population is much lower.
-However I wonder if just extending the number of time steps over which I run the average would reduce this effect.
-```{r time-series,echo=FALSE,message=FALSE,results='hide'}
+And can see crazy fluctuations in `s = 0.26` that's not present in either of the other two simulations
+```{r}
   xyplot(sums ~ time | speed, data = df300, type='l',xlab="time",ylab="Population biomass",lwd=2, main = "300 steps")
 ```
-After re-running the simulation (this time with 600 time steps for equilibrium)
-```{r second-sim,echo=FALSE,fig.keep='none',message=FALSE,results='hide'}
-# pre-allocating list to save simulations
 
-steps = 600   # setting number of steps over which to take equilibrium 
-source("Figures/variable_steps.R")
-sims600 = sims
-df600 = df
-series600 = series
-notify("Simulation is finished!")
+Re-running on `Sim_noThresh.R` for `s = 0.28` also doens't get convergence.
+So not sure why that simulation was able to run overnight..
+Unless it takes just thousands of runs.
+But I think I could check that in the dataframe.
+
+Switching to `s=0.26` issue, I think the issue may be that the `move` matrix is so small: it didn't run for many generations.
+So reaching back for 300 time steps involves a lot of the decline from when only harvest occurred, so it artificially inflates population biomass.
+So to check, look at the high `Equil.pop` and see how big the `move` matrix is.
+
+```{r}
+consmpa$steps.move[consmpa$harvest==0.08 & consmpa$speed==0.26]
+consmpa$steps.move[consmpa$harvest==0.09 & consmpa$speed==0.26]
+consmpa$steps.move[consmpa$harvest==0.1 & consmpa$speed==0.16]
+consmpa$steps.move[consmpa$harvest==0.12 & consmpa$speed==0.16]
+```
+This still doesn't help me with the fact that `s = 0.28` now doesn't want to converge.
+But will leave that for now.
+Pretty sure that if I make the simulation run for a minimum of 500 generations.
+So need to target the parameter combinations that need to be re-run.
+
+Look for differences within each speed simulation where the number of `steps.move` drops abruptly.
+
+```{r}
+ggplot(consmpa, aes(x=harvest, y = steps.move)) + geom_line() + geom_point() + facet_wrap(~speed)
 ```
 
-```{r plots,echo=FALSE,message=FALSE,results='hide'}
-  xyplot(sums ~ time | factor(speed), data = df600, type='l',xlab="time",ylab="Population biomass",lwd=2, main = "600 steps")
-
-finals600 <- llply(series600,mean)
-df_finals600 <- melt(finals600)
-names(df_finals600) <- c("totalPop","Speed")
-df_finals600$Speed = speeds
-xyplot(totalPop ~ Speed, data=df_finals600,type='l',lwd=2,xlab="Speed")
-```
-
-This seems to matter a little bit
-
-```{r compare,echo=FALSE,message=FALSE,results='hide'}
-plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean")
-lines(speeds,df_finals600$totalPop,col="red",type="o",pch=19)
-```
-
-This is weird, shouldn't be like this, not sure why it is. Needs more work. Next step is to try 900 steps to see if the equilibrium varies
-
-```{r thirdSim, echo=FALSE,fig.keep='none'}
-steps = 900   # setting number of steps over which to take equilibrium 
-source("Figures/variable_steps.R")
-sims900 = sims
-df900 = df
-series900 = series
-
-xyplot(sums ~ time | factor(speed), data = df900, type='l',xlab="time",ylab="Population biomass",lwd=2, main="900 steps")
-
-finals900 <- llply(series900,mean)
-df_finals900 <- melt(finals900)
-names(df_finals900) <- c("totalPop","Speed")
-df_finals900$Speed = speeds
-xyplot(totalPop ~ Speed, data=df_finals900,type='l',lwd=2,xlab="Speed")
-
-plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean")
-lines(speeds,df_finals600$totalPop,col="red",type="o",pch=19)
-lines(speeds,df_finals900$totalPop,col="blue", type="o",pch=19)
-
-```
-
-```{r fourthSim, echo=FALSE,fig.keep='none'}
-steps = 1200   # setting number of steps over which to take equilibrium 
-source("Figures/variable_steps.R")
-sims1200 = sims
-df1200 = df
-series1200 = series
-
-xyplot(sums ~ time | factor(speed), data = df1200, type='l',xlab="time",ylab="Population biomass",lwd=2, main="1200 steps")
-
-finals1200 <- llply(series1200,mean)
-df_finals1200 <- melt(finals1200)
-names(df_finals1200) <- c("totalPop","Speed")
-df_finals1200$Speed = speeds
-xyplot(totalPop ~ Speed, data=df_finals1200,type='l',lwd=2,xlab="Speed")
-
-plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean")
-lines(speeds,df_finals600$totalPop,col="red",type="o",pch=19)
-lines(speeds,df_finals900$totalPop,col="blue", type="o",pch=19)
-lines(speeds,df_finals1200$totalPop,col="green", type="o",pch=19)
-
-notify("simluation is done!")
-```
-
-```{r fifthSim, echo=FALSE,fig.keep='none'}
-steps = 1500   # setting number of steps over which to take equilibrium 
-source("Figures/variable_steps.R")
-sims1500 = sims
-df1500 = df
-series1500 = series
-
-xyplot(sums ~ time | factor(speed), data = df1500, type='l',xlab="time",ylab="Population biomass",lwd=2, main="1500 steps")
-
-finals1500 <- llply(series1500,mean)
-df_finals1500 <- melt(finals1500)
-names(df_finals1500) <- c("totalPop","Speed")
-df_finals1500$Speed = speeds
-xyplot(totalPop ~ Speed, data=df_finals1500,type='l',lwd=2,xlab="Speed")
-
-plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean")
-lines(speeds,df_finals600$totalPop,col="red",type="o",pch=19)
-lines(speeds,df_finals900$totalPop,col="blue", type="o",pch=19)
-lines(speeds,df_finals1200$totalPop,col="green", type="o",pch=19)
-lines(speeds,df_finals1500$totalPop,col="purple", type="o",pch=19)
-
-notify("simluation is done!")
-```
-
-```{r sixthSim, echo=FALSE,fig.keep='none'}
-steps = 1800   # setting number of steps over which to take equilibrium 
-source("Figures/variable_steps.R")
-sims1800 = sims
-df1800 = df
-series1800 = series
-
-xyplot(sums ~ time | factor(speed), data = df1800, type='l',xlab="time",ylab="Population biomass",lwd=2, main="1800 steps")
-
-finals1800 <- llply(series1800,mean)
-df_finals1800 <- melt(finals1800)
-names(df_finals1800) <- c("totalPop","Speed")
-df_finals1800$Speed = speeds
-xyplot(totalPop ~ Speed, data=df_finals1800,type='l',lwd=2,xlab="Speed")
-
-require(RColorBrewer)
-colz <- colorRampPalette(brewer.pal(6,"Dark2"))(6)
-
-plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean", col=colz[1], lwd=3, bty="n")
-lines(speeds,df_finals600$totalPop,col=colz[2],type="o",pch=19,lwd=3)
-lines(speeds,df_finals900$totalPop,col=colz[3], type="o",pch=19,lwd=3)
-lines(speeds,df_finals1200$totalPop,col=colz[4], type="o",pch=19,lwd=3)
-lines(speeds,df_finals1500$totalPop,col=colz[5], type="o",pch=19,lwd=3)
-lines(speeds,df_finals1800$totalPop,col=colz[6], type="o",pch=19,lwd=3)
-
-legend("topright",legend=seq(300,1800,300), lwd=3, pch=19,col=colz, title="Steps", bty="n")
-
-notify("simluation is done!")
-```
-
-```{r sixthSim, echo=FALSE,fig.keep='none'}
-steps = 2100   # setting number of steps over which to take equilibrium 
-source("Figures/variable_steps.R")
-sims2100 = sims
-df2100 = df
-series2100 = series
-
-xyplot(sums ~ time | factor(speed), data = df2100, type='l',xlab="time",ylab="Population biomass",lwd=2, main="2100 steps")
-
-finals2100 <- llply(series2100,mean)
-df_finals2100 <- melt(finals2100)
-names(df_finals2100) <- c("totalPop","Speed")
-df_finals2100$Speed = speeds
-xyplot(totalPop ~ Speed, data=df_finals2100,type='l',lwd=2,xlab="Speed")
-
-require(RColorBrewer)
-colz <- colorRampPalette(brewer.pal(6,"Dark2"))(7)
-
-plot(speeds,df_finals300$totalPop,pch=19,type="o",main="Mean", col=colz[1], lwd=3, bty="n")
-lines(speeds,df_finals600$totalPop,col=colz[2],type="o",pch=19,lwd=3)
-lines(speeds,df_finals900$totalPop,col=colz[3], type="o",pch=19,lwd=3)
-lines(speeds,df_finals1200$totalPop,col=colz[4], type="o",pch=19,lwd=3)
-lines(speeds,df_finals1500$totalPop,col=colz[5], type="o",pch=19,lwd=3)
-lines(speeds,df_finals1800$totalPop,col=colz[6], type="o",pch=19,lwd=3)
-lines(speeds,df_finals2100$totalPop,col=colz[7], type="o",pch=19,lwd=3)
-
-legend("topright",legend=seq(300,2100,300), lwd=3, pch=19,col=colz, title="Steps", bty="n")
-
-notify("simluation is done!")
-```
-
-Idea, can I sub-sample the 2100 step run and take the mean of an increasingly limited window?
-This would mean using `df2100` as the data frame and taking the mean of all 2100 steps, and then 2099 steps, ...
-etc.
-down to 300 and seeing how the mean changes.
-
-```{r}
-int <- seq(300,2100)
-names(series2100) = speeds
-foo <- melt(series2100)
-names(foo) <- c("biomass","speed")
-means = vector("list",length(int))
-names(means) = int
-for(i in 1:length(int)){
-  means[[i]] = ddply(foo,.(speed),function(df) avg = mean(df$biomass[1:int[i]]))
-  }
-
-bar = melt(means)
-bar$variable = NULL
-names(bar) = c("speed", "mean", "interval")
-bar$interval = as.numeric(bar$interval)
-ggplot(bar, aes(x = interval , y = mean))  + geom_line() + facet_wrap( ~ speed)
-```
-
-New idea, find the maxima and minima and take the mean of those two numbers. See if that erases the variability in means.
-
-```{r}
-int <- seq(300,2100)
-names(series2100) = speeds
-foo <- melt(series2100)
-names(foo) <- c("biomass","speed")
-means_dif = vector("list",length(int))
-names(means) = int
-for(i in 1:length(int)){
-  means_dif[[i]] = ddply(foo,.(speed),function(df) avg=mean(max(df$biomass[1:int[i]]),min(df$biomass[1:int[i]])))
-  }
-
-bar = melt(means_dif)
-bar$variable = NULL
-names(bar) = c("speed", "mean", "interval")
-bar$interval = as.numeric(bar$interval)
-ggplot(bar, aes(x = interval , y = mean))  + geom_line() + facet_wrap( ~ speed)
-
-ggplot(bar, aes(x = speed, y = mean, group = interval)) + geom_line(aes(colour=interval))
-```
-
-And volia!
-It erases that variability.
-Looks like I can keep the 300 time steps and take the max and minimum and take the average.
-So should re-do that for the existing simulation data I have and plot a contour map
-
-UPDATE, the problem persists.
-Malin's idea is to take a time step where the MPA and the range line up at the 'northern' edge.
-Because the world is only a patch length long, need to figure out when the MPA is at the northern-most edge.
+Looks like all of `s=0.16` should be re-run, also all of `s=0.26`. This is because the number of generations dips below 400. The rest seems ok.

### 2014-03-04 Writing/MS.tex
finishing up with figures

@@
 Our results show that both threshold and MPAs increase the equilibrium biomass at a given climate velocity, which support their use as a tool to ameliorate the effect of climate velocity.
-However for MPAs the details mater: few, large MPAs caused increased variability at low climate velocities while many smaller MPAs maintained a population bounded farther from extinction.
+However for MPAs the details mater: few, large MPAs caused increased variability at low climate velocities while many smaller MPAs maintained a population bounded farther from extinction (see Figure 1, Appendix A.6).
 Finally, with sufficiently high harvesting pressure, few, large MPAs rescued populations at intermediate speeds.
@@
 \subcaption{}
-\includegraphics[width=1\textwidth]{plots/MPAs.pdf}
+\includegraphics[width=1\textwidth]{plots/eqbiomass_sim.pdf}
 \label{nomang}
@@
 \subcaption{}
-\includegraphics[width=1\textwidth]{plots/Threshold.pdf}
+\includegraphics[width=1\textwidth]{plots/eqbiomass_consmpa.eps}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=1\textwidth]{plots/eqbiomass_fishmpa.eps}
+\end{subfigure}
+\begin{subfigure}{.33\textwidth}
+\subcaption{}
+\includegraphics[width=1\textwidth]{plots/eqbiomass_thresh.eps}
 \end{subfigure}

### 2014-03-04 Simluations/Aspatial_fast/Figures/bounded_flux.R

@@ -1,0 +1,74 @@
+# show that variability of population size is different between consmpa and fishmpa
+
+setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
+source("Parameters.R")
+source("Functions.R")
+require(zoo)
+require(plyr)
+require(lattice)
+require(dplyr)
+require(reshape)
+
+########
+# Cons #
+########
+
+steps = 300   # setting number of steps over which to take equilibrium 
+harvests = 0.08
+speeds = c(0.1,0.3)
+MPA = "cons"
+source("Figures/variable_steps.R")
+
+
+simsCons = sims
+dfCons = df
+seriesCons = series
+
+finalsCons <- llply(seriesCons,mean)
+sdCons <- llply(seriesCons,sd)
+sdCons <- melt(sdCons)
+df_finalsCons <- melt(finalsCons)
+df_finalsCons$sd = sdCons$value
+names(df_finalsCons) <- c("totalPop","Speed","sd")
+df_finalsCons$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finalsCons,type='l',lwd=2,xlab="Speed", main = "Cons steps")
+ggplot(df_finalsCons,aes(x = Speed, y = totalPop)) + geom_errorbar(aes(ymin=totalPop - sd, ymax = totalPop + sd)) + geom_line() + geom_point()
+
+xyplot(sums ~ time | speed, data = dfCons, type='l',xlab="time",ylab="Population biomass",lwd=2, main = "Cons steps")
+
+
+########
+# Fish #
+########
+
+steps = 300   # setting number of steps over which to take equilibrium 
+MPA = "fish"
+source("Figures/variable_steps.R")
+
+
+simsfish = sims
+dffish = df
+seriesfish = series
+
+finalsfish <- llply(seriesfish,mean)
+sdfish <- llply(seriesfish,sd)
+sdfish <- melt(sdfish)
+df_finalsfish <- melt(finalsfish)
+df_finalsfish$sd = sdfish$value
+names(df_finalsfish) <- c("totalPop","Speed","sd")
+df_finalsfish$Speed = speeds
+xyplot(totalPop ~ Speed, data=df_finalsfish,type='l',lwd=2,xlab="Speed", main = "fish steps")
[…]

### 2014-03-04 fish_figures.R

@@ -304,6 +304,21 @@ image.plot(cvals,hvals,toplot_ebm2,breaks=cuts,col=mycols,
 box()
 dev.off()
 
+postscript("eqbiomass_consmpa.eps", horizontal = FALSE, onefile = FALSE, paper = "special", width = fig.width,height=fig.width,bg="white")
+
+par(oma=oma)
+image.plot(cvals,hvals,toplot_ebm2,breaks=cuts,col=mycols,
+	xlab="Climate velocity",ylab="Harvesting rate",
+	cex.lab=cex.lab,cex.axis=cex.axis, 
+	yaxs=yaxs,xaxs=xaxs,axes=TRUE,
+	legend.shrink=1,legend.width=.1,zlim=range(cuts),
+	axis.args=list(at=myaxes$z$at,labels=myaxes$z$labels,cex.axis=cex.axis),
+	#legend.args=list(text="Equilibrium biomass",cex=cex.lab,side=2,line=0.5,las=0),
+	bigplot=c(.13,.85,.15,.95),smallplot=c(.91,.95,.15,.95),horizontal=FALSE)
+box()
+
+graphics.off()
+
 ##########
 ########## 
 ##### wireframe of the equilibrium biomass as a function of c and h : THRESHOLD

### 2014-03-04 load_data.R

@@ -13,7 +13,7 @@ x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspati
 	x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDA4NjQxfQ%3D%3D--2dba33c2046114811967e01914fc795e1771f425")
 	mpas = read.csv(textConnection(x))
 
-x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-03-04.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMy0wNC5jc3YiLCJleHBpcmVzIjoxMzk0NTUzMjQxfQ%3D%3D--9ddfd59a719d0410fa72a83cca37a0e789ea0d34")
+x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-03-04.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMy0wNC5jc3YiLCJleHBpcmVzIjoxMzk0NTU5MTY0fQ%3D%3D--dbd27e32e8632468f8c1113dbd4b470bba84a535")
 	mpas2 = read.csv(textConnection(x))
 
[…]

### 2014-03-04 Writing/MS.tex
last tweaks to figures for now

--- 2014-03-04 Writing/MS.tex
@@
 The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis.
-These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.5$.
+These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=10$, $\langle d \rangle = 0.5$.
 \hspace{6in}
@@
 This excess loss, on the order of $.05$, is small in comparison to the total biomass, which can be greater than $40$.
-These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.5$.
+These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=10$, $\langle d \rangle = 0.5$.
 \hspace{6in}

### 2014-03-04 fish_figures.R
tweaks to figures

@@ -4,20 +4,9 @@ library(colorspace)
 library(RCurl)
 source("biomass_synergy.R")
 source("critical_parameters.R")
-source("load_data.R")
+source("Simluations/Aspatial_fast/load_data.R")
 
-### parameters from Zhou
-r=3
-w=pi/4
-L=1
-
-####
-#the world
-step_size=0.01
-L=1
-world=matrix(seq(-L/2,L/2,by=step_size),nrow=1)
-emmaR0=5
-emmaK=100
+analytical_ebm=ebm
 
 #####
 #### plotting parameters
@@ -87,24 +76,17 @@ graphics.off()
 ##########
 ##########
 ##### heatmap of the equilibrium biomass as a function of c and h
-r=emmaR0
-K=emmaK
-sig2=1
-tol=.00001
-diff=.25
-cvals=seq(0,1,by=diff)
-hvals=seq(0,1,by=diff)
 
-
-# x=getURL("https://raw.github.com/emfuller/MovingFish/38c6f3bb51dae267fe4056159609de100fb34826/eqbiomass_gaus.csv?token=6224444__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC8zOGM2ZjNiYjUxZGFlMjY3ZmU0MDU2MTU5NjA5ZGUxMDBmYjM0ODI2L2VxYmlvbWFzc19nYXVzLmNzdiIsImV4cGlyZXMiOjEzOTQ0ODQ5OTh9--d95944353679e18e44def40cd49ca41a22502039")
-# ebm=read.csv(textConnection(x))
-
+cvals=as.numeric(str_sub(row.names(ebm),start=3))
+hvals=as.numeric(str_sub(dimnames(ebm)[[2]],start=3))
 
 #which params to work with
+tol=0.0001
 H=sum(apply(ebm,2,max)>tol)+1
-hvals=hvals[1:H]
+# hvals=hvals[1:H]
 C=length(cvals)
-toplot_ebm=as.matrix(ebm[1:C,1:H],nrow=C)
+toplot_ebm=as.matrix(ebm,nrow=C)
+toplot_ebm[toplot_ebm<0.01]=0
 
 #breaks and colors
 lower=.001
@@ -126,7 +108,7 @@ quartz()
[…]

### 2014-03-04 biomass_synergy.R
last tweaks to figures for now

@@ -1,0 +1,79 @@
+library(fields)
+library(lattice)
+
+#Beverton-Holt recruitment 
+
+f<-function(n,R0,K){
+	next_gen<-R0*n/(1+(R0-1)/K*n)
+	return(next_gen)
+	}
+	
+int<-function(v){
+	l=length(v)
+	step_size/2*(sum(2*v)-v[1]-v[l])
+}
+
+finda<-function(c,h,r,K,sig2,L=1,tol=NA){
+	if(is.na(tol)){tol=.001}
+	a=1
+	fish=a*exp(-world^2/(2*sig2))
+	fish=(1-h)*fish
+	v=exp(-(world-c)^2/(2*sig2))*apply(fish,2,f,R0=r,K=K)
+	newa=1/sqrt(2*sig2*pi)*int(v)
+	while(abs(a-newa)>tol){
+		a=newa
+		fish=a*exp(-world^2/(2*sig2))
+	fish=(1-h)*fish
+	v=exp(-(world-c)^2/(2*sig2))*apply(fish,2,f,R0=r,K=K)
+	newa=1/sqrt(2*sig2*pi)*int(v)
+	}
+	return(newa)
+}
+
+biomass<-function(c,h,r,K,sig2,L=1,tol=NA){
+	if(is.na(tol)){tol=.001}
+	a=finda(c,h,r,K,sig2,L,tol)
+	d=a*sqrt(2*sig2*pi)*(pnorm(L/2/sqrt(sig2))-pnorm(-L/2/sqrt(sig2)))
+	return(d)
+}
+
+eqbiomass<-function(cvals,hvals,r,K,sig2,L=1,tol=NA){
+	if(is.na(tol)){tol=.001}
+	M=array(,c(length(cvals),length(hvals)))
+	for(i in 1:length(cvals)){
+		for(j in 1:length(hvals)){
+			M[i,j]<-biomass(cvals[i],hvals[j],r,K,sig2,L,tol)
+	}
+}
+	rownames(M)=paste("c",cvals)
+	colnames(M)=paste("h",hvals)
+	M=round(M,digits=4)
+	return(M)
+}
+
+synergy<-function(ebm,digs=NA){
+	if(is.na(digs)){digs=3}
+	cvals=as.numeric(substring(rownames(ebm),2))
+	hvals=as.numeric(substring(colnames(ebm),2))
+	M=array(,c(length(cvals),length(hvals)))
[…]

### 2014-03-12 Writing/Appendix.tex
finishing up with figures
editing the appendix, particularly adding to the math

@@
 \subsection{Determining stability \label{stab}}
-As in Zhou et al.
-\citep{ZhouKot2011}, let $k(x-y)$ be a dispersal kernel, let $f(n)$ be a recruitment function, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
-The integrodifference model describing the population over time is given by
+Let $k(x)$ be a dispersal kernel describing the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring produced by a population of size $n$, let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$, and let $n_t(x)$ be the number of adults at position $x$ at time $t$.
+In the absence of harvesting, the integrodifference model describing the population over time is given by
+
+\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y))dy. \label{integro} \end{equation}
+as described in \citep{ZhouKot2011}.  With the addition of harvesting, the model becomes
 
@@
 
-To find a traveling pulse, we are only interested in the population density as a function of the location within the patch rather than absolute position, $\overline{x}\equiv x-ct$.
+In evaluating persistence, we follow the methods of \citep{ZhouKot2011}.
+A traveling pulse is a solution such that population size relative to location within the patch (rather than absolute position) is constant over time, i.e.
 
 \begin{equation*}
-n^*(\overline{x})\equiv n^*(x-ct)=n_t(x).   \label{trav} 
+n^*(\overline{x}_t)\equiv n^*(x-ct)=n_t(x),   \label{trav} 
 \end{equation*}
+where $\overline{x}_t\equiv x-ct$ gives position relative to the patch.
 
@@
 \begin{align}
-n^*(\overline{x}-c)&=\int_{-L/2}^{L/2}k(\overline{x}-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y})))d\overline y \notag
-\\ \Rightarrow n^*(\overline{x})&=\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y}))d\overline y  \label{pulse}
+n^*(\overline{x}_{t+1})&=n_{t+1}(x) \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-\overline{y}_t-ct)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))dy \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(\overline{x}_t-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))dy \notag
+\\ \Rightarrow n^*(\overline{x}_t-c)&=\int_{-L/2+ct}^{L/2+ct}k(\overline{x}_t-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))dy \notag
+%\\n^*(\overline{x}-c)&=\int_{-L/2}^{L/2}k(\overline{x}-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y})))d\overline y \notag
+\\ \Rightarrow n^*(\overline{x}_t)&=\int_{-L/2}^{L/2}k(\overline{x}_t+c-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t  \label{pulse}
 \end{align}
-As long as $f(0)=0$, there is a trivial solution to this problem where $n^*(\overline{x})\equiv 0$ for all $\overline{x}\in[-L/2,L/2]$, i.e.
-there is a trivial traveling pulse with no fish in it.
+As long as $f(0)=0$, there is a trivial solution to this problem where $n^*(\overline{x})\equiv 0$ for all $\overline{x}$, i.e., there is a trivial traveling pulse with no fish in it.
 If the trivial traveling pulse is unstable, even very small populations will persist or grow and avoid crashing back to the trivial pulse.
@@
 \begin{align}
-n_t(x)&=n^*(\overline{x})+\xi_t(x) \notag
-\\ \Rightarrow \xi_{t+1}(x)&=n_{t+1}(x)-n^*(\overline{x}) \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(f(n_t(y)-g(n_t(y)))-f(n^*(\overline{y})-g(n^*(\overline{y}))\big)dy \text{ using (\ref{pulse})} \notag
+n_t(x)&=n^*(\overline{x}_t)+\xi_t(x) \notag
+\\ \Rightarrow \xi_{t+1}(x)&=n_{t+1}(x)-n^*(\overline{x}_{t+1}) \notag
+\\&=n_{t+1}(x)-n^*(\overline{x}_{t}-c) \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2}^{L/2}k(\overline{x}_{t}-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \text{ using (\ref{pulse})} \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2}^{L/2}k(\overline{x}_{t}-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2+ct}^{L/2+ct}k(x-ct-(y-ct))f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(f(n_t(y)-g(n_t(y)))-f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))\big)dy \notag
+%%
+%\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2}^{L/2}k(x+ct-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y}))d\overline y \notag
+%\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(f(n_t(y)-g(n_t(y)))-f(n^*(\overline{y})-g(n^*(\overline{y}))\big)dy \notag
 \\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(n^*(\overline{y})))f'(n^*(\overline{y})-g(n^*(\overline{y}))\xi_t(y)dy \notag
@@
 \begin{equation} \psi_f(u)(x)=\lambda u(x) \label{eigen} \end{equation}
-$\lambda$ and $u$ are thus an eigenvalue and eigenvector of the functional operator $\psi_f$.  The trivial traveling pulse is unstable when the dominant eigenvalue of $\psi_f$ is greater than $1$.
+$\lambda$ and $u$ are thus an eigenvalue and eigenfunction of the functional operator $\psi_f$.  The trivial traveling pulse is unstable when the dominant eigenvalue of $\psi_f$ is greater than $1$.
 
@@
 \subsection{Separable dispersal kernels \label{sep}}
-Jentzsch's theorem shows that there is an eigenfunction $u$, provided that the kernel $k$ satisfies some properties \citep{ZhouKot2011}.
+It is not immediately obvious that the operator $\psi$ will have any eigenfunctions.
+However, Jentzsch's theorem guarantees that there is an eigenfunction $u$, provided that the kernel $k$ satisfies some properties \citep{ZhouKot2011}.
 Finding the eigenfunctions and eigenvalues is in general a hard problem to solve.
@@
 \lambda u(x)&=f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right) \notag
-\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right) \notag
+\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right) \notag \text{for any $k$}
 \\ \Rightarrow \lambda d_k&=f'(0)\sum_{n=1}^\infty A_{nk}d_n  \label{problem}
@@
 $$k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.$$
-As in \citep{Latore:1998fk}, this separable kernel can be written as
-$$k(x-y)=\sum_{n=0}^\infty a_n(x)b_n(y)$$
-where
-$$a_n(x)=b_n(x)=\frac{1}{\sqrt{2n!\sqrt{D\pi}}}e^{-x^2/4D}\left(\frac{x}{\sqrt{2D}}\right)^n.$$
+This is a separable kernel with
+$a_n(x)=b_n(x)=\frac{1}{\sqrt{2n!\sqrt{D\pi}}}e^{-x^2/4D}\left(\frac{x}{\sqrt{2D}}\right)^n$ \citep{Latore:1998fk}.
 
@@
 \subsection{Sinusoidal dispersal kernel \label{sinapp}}
-A sinusoidal dispersal kernel is given by 
+The sinusoidal dispersal kernel is given by 
 $$k(x-y)=\left\{\begin{array}{ccccc}
@@
 $$c^*=c^*(R_0)=\frac{1}{w}\cos^{-1}\left[\frac{16+R_0^2(1-h)^2(w^2L^2-\sin^2(wL))}{8R_0(1-h)wL}\right].$$
-Similarly, we can solve for the critical harvesting rate, $h^*$, at which the population will be driven extinct:
+In our model, we can additionally solve for the critical harvesting rate, $h^*$, at which the population will be driven extinct:
 $$
@@
 \\ \exp(x)&=1+x+\frac{x^2}{2}
-\end{align*}
-For the sinusoidal kernel we found 
-\begin{equation}
-h^*=1-\frac{1}{R_0}\cdot\frac{4wL}{w^2L^2-\sin^2(wL)}\left[\cos(wc)-\sqrt{\cos^2(wc)-1+\frac{\sin^2(wL)}{w^2L^2}}\right] 
-\end{equation} 
-Using the Taylor series and the fact that $w=\frac{\sqrt{\frac{\pi^2}{4}-2}}{\sigma}$ where $\sigma^2$ is the variance of the sinusoidal kernel,
-\begin{align*}
-h^*&\sim 1-\frac{1}{R_0}\cdot\frac{12wL}{w^4L^4}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{12}{w^3L^3}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{12\sigma^3}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[1-\frac{(\frac{\pi^2}{4}-2)c^2}{2\sigma^2}-\sqrt{1-\frac{(\frac{\pi^2}{4}-2)c^2}{\sigma^2}-\frac{(\frac{\pi^2}{4}-2)L^2}{3\sigma^2}}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[2\sigma^2-(\frac{\pi^2}{4}-2)c^2-\frac{2}{\sqrt{3}}\sigma\sqrt{3\sigma^2-(\frac{\pi^2}{4}-2)3c^2-(\frac{\pi^2}{4}-2)L^2}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^34(\frac{\pi^2}{4}-2)^{3/2}}\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
-%\\&=1-\frac{12}{R_0L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
-\\&=1-\frac{1}{R_0}\cdot\frac{4\sqrt{3}}{L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sqrt{3}\sigma^2-(\pi^2-8)\sqrt{3}c^2-4\sigma\sqrt{12\sigma^2-(\pi^2-8)(3c^2+L^2)}\right]
 \end{align*}
@@
 \end{align*}
+For the sinusoidal kernel we found 
+\begin{equation}
+h^*=1-\frac{1}{R_0}\cdot\frac{4wL}{w^2L^2-\sin^2(wL)}\left[\cos(wc)-\sqrt{\cos^2(wc)-1+\frac{\sin^2(wL)}{w^2L^2}}\right] 
+\end{equation} 
+Using the Taylor series and the fact that $w=\frac{\sqrt{\frac{\pi^2}{4}-2}}{\sigma}$ where $\sigma^2$ is the variance of the sinusoidal kernel,
+\begin{align*}
+h^*&\sim 1-\frac{1}{R_0}\cdot\frac{12wL}{w^4L^4}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
+%\\&=1-\frac{1}{R_0}\cdot\frac{12}{w^3L^3}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
+%\\&=1-\frac{1}{R_0}\cdot\frac{12\sigma^3}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[1-\frac{(\frac{\pi^2}{4}-2)c^2}{2\sigma^2}-\sqrt{1-\frac{(\frac{\pi^2}{4}-2)c^2}{\sigma^2}-\frac{(\frac{\pi^2}{4}-2)L^2}{3\sigma^2}}\right]
+%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[2\sigma^2-(\frac{\pi^2}{4}-2)c^2-\frac{2}{\sqrt{3}}\sigma\sqrt{3\sigma^2-(\frac{\pi^2}{4}-2)3c^2-(\frac{\pi^2}{4}-2)L^2}\right]
+%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^34(\frac{\pi^2}{4}-2)^{3/2}}\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
+%\\&=1-\frac{12}{R_0L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
+\\&=1-\frac{1}{R_0}\cdot\frac{4\sqrt{3}}{L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sqrt{3}\sigma^2-(\pi^2-8)\sqrt{3}c^2-4\sigma\sqrt{12\sigma^2-(\pi^2-8)(3c^2+L^2)}\right]
+\end{align*}
 In the case of both kernels, the critical harvesting proportion can be approximated by a function that looks like 
 \begin{equation}
-h^*\sim1- \frac{1}{R_0}\cdot C(L)f(\sigma^2,c^2,L^2+3c^2)
+h^*\sim1- \frac{1}{R_0}\cdot g(L)q(\sigma^2,c^2,L^2+3c^2)
 \end{equation}
-where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.
+where $g(L)$ is a decreasing function of the length of the viable patch $L$.
+
+\subsection{MPA fluctuations}
+
+\begin{figure}
+\includegraphics[width=1\textwidth]{plots/bounded_flux.pdf}
+\caption{Comparing population over time where $s = 0.1, h = 0.08$, see bounded\_flux.R for code that made this plot. }
+\end{figure}
 

### 2014-03-12 alt_fish_figures.R
Incorporating Malin's edits

@@ -1,22 +1,17 @@
 # for synced online github data
-x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzYwfQ%3D%3D--5fa73caaafffae7c95d930f61def04ee9b96930a")
-sim = read.csv(textConnection(x))
-
-fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzQ2fQ%3D%3D--9b6ceb0092173222543c95381dfbb0f42f0c884f")
-fishmpa = read.csv(textConnection(fish))
-
-cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-03-01.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMy0wMS5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzIwfQ%3D%3D--5f4f584256a4175eebe2ef5465e83c92ebe65cc0")
-consmpa = read.csv(textConnection(cons))
-
[…]

### 2014-03-12 load_data.R

@@ -3,20 +3,20 @@ library(RCurl)
 
 # analytics
 
-x=getURL("https://raw.github.com/emfuller/MovingFish/d6b73bca90c71f0aedb616a241df24b536e0f80b/eqbiomass_gaus.csv?token=6224444__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9kNmI3M2JjYTkwYzcxZjBhZWRiNjE2YTI0MWRmMjRiNTM2ZTBmODBiL2VxYmlvbWFzc19nYXVzLmNzdiIsImV4cGlyZXMiOjEzOTQ1NTkzNjF9--9f57d5392b49fe11ef00599873ecc6b66926ba10")
+x=getURL("https://raw.github.com/emfuller/MovingFish/master/eqbiomass_gaus.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvZXFiaW9tYXNzX2dhdXMuY3N2IiwiZXhwaXJlcyI6MTM5NTE4NTQ4N30%3D--65033562cc7e8c52f04dbf0ea06cae97e4008776")
 ebm=read.csv(textConnection(x),row.names=1)
 ebm=as.matrix(ebm,nrow=length(ebm[,1]))
 
 # simulations
 
-x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDA4NjYyfQ%3D%3D--7f380bfb17f1cd58ff4f131134ed434700767def")
[…]

### 2014-03-13 alt_fish_figures.R
Finishing my round of edits, have redone the plots

@@ -1,51 +1,0 @@
-# for synced online github data
-source("load_data.R")
-require(lattice)
-require(RColorBrewer)
-threshz = thresh
-fishmpa = mpas
-consmpa = mpas2
-
-# combine all data into one dataframe
-sim$management = rep("No MPAs",nrow(sim))
-consmpa$management = rep("Few Large MPAs", nrow(consmpa))
-fishmpa$management = rep("Many Small MPAs", nrow(fishmpa))
-threshz$management = rep("Thresholds", nrow(threshz))
-allData <- rbind(consmpa,fishmpa, sim)
-
-
-# lattice plot settings
-
-sb <- trellis.par.get("strip.background") 
-sb[["col"]][1] <- "lightgray"
-trellis.par.set("strip.background", sb) 
-
-ltheme <- canonical.theme(color = FALSE) ## in-built B&W theme 
-ltheme$strip.background$col <- "transparent" ## change strip bg 
-lattice.options(default.theme = ltheme) ## set as default 
-
-lw <- list(left.padding = list(x = 0, units = "inches")) 
-lw$right.padding <- list(x = -0.1, units = "inches") 
-lh <- list(bottom.padding = list(x = 0, units = "inches")) 
-lh$top.padding <- list(x = -0.2, units = "inches")
-
-lattice.options(layout.widths = lw, layout.heights = lh) 
-
-# mpa plots
-
-pdf(file="Writing/plots/MPAs.pdf",width=3,height=5)
-contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(2,2), drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"))
-dev.off()
-
-#threshold plot
-
-pdf(file="Writing/plots/Threshold.pdf",width=3,height=3)
-contourplot(Equil.pop ~ speed * -thresh | factor(management), data = threshz, region=TRUE, drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"),scales=list( y=list(at=seq(-8,0,2),labels=round(rev(unique(threshz$thresh)/max(threshz$thresh)),3))),ylab=list(label="proportion below threshold"))
-dev.off()
-
-### ggplots
-require(ggplot2)
-require(ggthemes)
-
[…]

### 2014-03-13 plots/Fish_Figures.R

@@ -1,0 +1,127 @@
+library(fields)
+library(lattice)
+library(colorspace)
+library(RCurl)
+source("biomass_synergy.R")
+source("critical_parameters.R")
+source("load_data.R")
+require(stringr)
+require(colorspace)
+require(reshape2)
+require(gridExtra)
+require(ggplot2)
+require(ggthemes)
+
+###################################### Figure 1
+
+### A
+	analytical_ebm=ebm
+	cvals=matrix(seq(0,1,by=0.01),nrow=1)
+
+	nr=3
+	ns=2
+	rvals=c(rep(3,ns),rep(5,ns),rep(10,ns))
+	rcols=c(rep('black',ns),rep('red',ns),rep('blue',ns))
+	xvals=c(.1,.25)
+	sigvals=rep(pi/2*xvals,nr)
+	sigltys=rep(c(1,2),nr)
+	l=length(rvals)
+
+# save data as list
+	h = vector("list",length=l)
+
+	for(i in 1:l){
+		r=rvals[i]
+		sig2=sigvals[i]
+		h[[i]]=apply(cvals,2,hstar_g,r=r,sig2=sig2)
+	}
+	
+	melted_h <- melt(h)
+	names(melted_h) <- c("h", "Param_combo")
+	melted_h$speed <- rep(cvals, 6)
+	melted_h$R <- rep(rvals, each = 101)
+	melted_h$d <- round(rep(sigvals, each =101),digits=2)
+
+	plot1 <- ggplot(melted_h, aes(x=speed, y = h, group=Param_combo)) + 
+		geom_line(aes(color=factor(R), linetype=factor(d)),size=1) + 
+		theme_tufte() + 
+		scale_color_manual(values=c("grey", "dark grey", "black")) + 
+		theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(1,0,0,0),"cm")) + 
+		xlab("") + ylab("Harvesting rate") + 
+		labs(linetype=expression(symbol("\341")*d*symbol("\361")), color=expression("R"[0]), title="A")
+	
+
+#### B
+	melted_ebm <- melt(ebm)
+	speeds <- rep(seq(0,1, 0.01), 101)
+	harvests <- rep(seq(0,1,0.01), each = 101)
+	melted_ebm$speed = speeds
+	melted_ebm$harvest = harvests
+
+	plot2 <- ggplot(melted_ebm, aes(x=speed, y = harvest, fill=value)) + 
[…]

### 2014-03-13 plots/old/fish_figures.R

@@ -5,7 +5,8 @@ library(RCurl)
 source("biomass_synergy.R")
 source("critical_parameters.R")
 source("Simluations/Aspatial_fast/load_data.R")
-
+require(stringr)
+require(colorspace)
 analytical_ebm=ebm
 
 #####

### 2014-03-14 Writing/Appendix.tex
editing

@@
 \subsection{Determining stability \label{stab}}
-Let $k(x)$ be a dispersal kernel describing the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring produced by a population of size $n$, let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$, and let $n_t(x)$ be the number of adults at position $x$ at time $t$.
+Let $n_t(x)$ be the number of adults at position $x$ at time $t$, let $k(x)$ be a dispersal kernel describing the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring produced by a population of size $n$, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
 In the absence of harvesting, the integrodifference model describing the population over time is given by
@@
 
+\pagebreak
 \subsection{MPA fluctuations}
 
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=1\textwidth]{plots/bounded_flux.pdf}
@@
 
+\pagebreak
 \bibliography{fish}

### 2014-03-14 Writing/Appendix.tex
edits

@@
 In Appendix \ref{approxcrit}, we derive approximate expressions for these critical rates.
+In Appendix \ref{MPA} we provide details on differences between small and large MPA simulations.
 
@@
 \pagebreak
-\subsection{MPA fluctuations}
-
+\subsection{MPA fluctuations \label{MPA}}
+~\\After the simulations come to equilibrium, the fluctuations in total biomass per generation fluctuate more in MPAs that are larger and spaced farther apart than simulations in which the MPAs that are smaller and more closely spaced.
+The large MPAs have a slightly larger average population, however large MPAs here can induce fluctuations of biomass even in deterministic simulations.
+Thus we expect if reproduction was stochastic, large MPAs spaced far apart would be more likely to result in extinction of the population than more closely spaced, smaller MPAs.
 \begin{figure}[h]
-\includegraphics[width=1\textwidth]{plots/bounded_flux.pdf}
-\caption{Comparing population over time where $s = 0.1, h = 0.08$, see bounded\_flux.R for code that made this plot. }
+\includegraphics[width=.85\textwidth]{plots/bounded_flux.pdf}
+\caption{Total population biomass at is on the y axis, and generation is on the x axis. These simulations were run with climate velocity $ = 0.1$, and a proportional harvest rate $= 0.08$. }
+% see bounded\_flux.R for code that made this plot
 \end{figure}

### 2014-03-14 Simluations/Aspatial_fast/Figures/bounded_flux.R

@@ -72,3 +72,6 @@ subnewdata <- subset(newdata, speed == 1)
 pdf(file="bounded_flux.pdf")
 xyplot(sums ~ time | MPA, data = subnewdata, type='l',lwd=2,ylab="Population biomass")
 dev.off()
+
+mean(subnewdata$sums[subnewdata$MPA=="many small MPAs"])
+mean(subnewdata$sums[subnewdata$MPA=="few large MPAs"])

### 2014-03-14 Simluations/Aspatial_fast/Figures/variable_steps.R

@@ -58,7 +58,7 @@ for(q in 1:length(speeds)){
       next.pop <- m(n=move[,i-1], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
       move <- cbind(move, next.pop[[1]])
       MPA.current = next.pop[[3]]
-      harv[(i-T)] <- sum(next.pop[[2]])
+      #harv[(i-T)] <- sum(next.pop[[2]])
     }  
     
     T = ncol(move)
@@ -97,7 +97,7 @@ for(q in 1:length(speeds)){
       next.pop <- m(n=move[,i-1], s = speeds[q], Fharv=harvests[j], mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
       move <- cbind(move, next.pop[[1]])
       MPA.current = next.pop[[3]]
-      harv[(i-T)] <- sum(next.pop[[2]])
+      #harv[(i-T)] <- sum(next.pop[[2]])
       
     }
   }

### 2014-03-28 plots/Fish_Figures.R
updating figures

@@ -112,20 +112,11 @@ ggplot(melt_syn, aes(x=Speed, y = Harvest, fill=Synergy)) +
 	consmpa$management = rep("Few Large MPAs", nrow(consmpa))
 	fishmpa$management = rep("Many Small MPAs", nrow(fishmpa))
 	threshz$management = rep("Thresholds", nrow(threshz))
-	allData <- rbind(consmpa,fishmpa, sim)
-
-	ggplot(allData, aes(x = speed, y = harvest, fill = Equil.pop)) + 
-		geom_raster(interpolate=TRUE) + 
-		facet_wrap(~management, scales="free") + 
-		theme_tufte() + 
-		scale_fill_gradient(low="black", high="white") + 
-		theme(legend.position="bottom", text=element_text(family="Helvetica", size = 11)) + 
-		guides(fill = guide_colorbar(barwidth = 18, barheight = 1, title.position="top", title="Equilibrium Biomass"))
-		
+	threshz$ord_thresh <- threshz$thresh/max(threshz$thresh)
 	plotA <- ggplot(sim, aes(x=speed, y = harvest, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="white")  + labs(title="A") + xlab("") + ylab("") + theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm"), legend.position="none") 
 
 
-	plotB <- ggplot(threshz, aes(x=speed, y = -thresh, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="white") + labs(title="B")+ xlab("") + ylab("")+ theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm"),legend.position="none") 
+	plotB <- ggplot(threshz, aes(x=speed, y = ord_thresh, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="white") + labs(title="B")+ xlab("") + ylab("")+ theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm"),legend.position="none") + scale_y_reverse()
 
 
 	

### 2014-03-28 Writing/Appendix.tex
small edits to appendix

@@
 \subsection{Determining stability \label{stab}}
-Let $n_t(x)$ be the number of adults at position $x$ at time $t$, let $k(x)$ be a dispersal kernel describing the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring produced by a population of size $n$, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
+Let $n_t(x)$ be the number of adults at position $x$ at time $t$, let $k(x)$ be a dispersal kernel describingt the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring produced by a population of size $n$, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
 In the absence of harvesting, the integrodifference model describing the population over time is given by
 
-\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y))dy. \label{integro} \end{equation}
-as described in \citep{ZhouKot2011}.  With the addition of harvesting, the model becomes
+\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y))dy \label{integro} \end{equation}
+as described in \citet{ZhouKot2011}.  With the addition of harvesting, the model becomes
 
@@
 
-In evaluating persistence, we follow the methods of \citep{ZhouKot2011}.
+In evaluating persistence, we apply the methods of \citet{ZhouKot2011} to the new model, Equation 2.
 A traveling pulse is a solution such that population size relative to location within the patch (rather than absolute position) is constant over time, i.e.
@@
 Finding the eigenfunctions and eigenvalues is in general a hard problem to solve.
-It becomes easier if the kernel $k$ is separable, i.e.
-there are functions $a_n,b_n$ such that $k(x-y)=\sum_{n=1}^\infty a_n(x)b_n(y)$.
+It becomes easier if the kernel $k$ is separable, i.e., there are functions $a_n,b_n$ such that $k(x-y)=\sum_{n=1}^\infty a_n(x)b_n(y)$.
 In that case, (\ref{eigen}) becomes
@@
 The Gaussian dispersal kernel is given by
-$$k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}}.$$
+$$k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}},$$
+where $D=\frac{\pi}{4}\langle d\rangle$ and $\langle d \rangle$ is the expected distance a larva disperses.
 This is a separable kernel with
@@
 
-Zhou et al. \citep{ZhouKot2011} solve for the critical speed, $c^*$, at which the population will be driven extinct:
+\citet{ZhouKot2011} solve for the critical speed, $c^*$, at which the population will be driven extinct:
 $$c^*=c^*(R_0)=\frac{1}{w}\cos^{-1}\left[\frac{16+R_0^2(1-h)^2(w^2L^2-\sin^2(wL))}{8R_0(1-h)wL}\right].$$
@@
 \begin{figure}[h]
-\includegraphics[width=.85\textwidth]{plots/bounded_flux.pdf}
-\caption{Total population biomass at is on the y axis, and generation is on the x axis. These simulations were run with climate velocity $ = 0.1$, and a proportional harvest rate $= 0.08$. }
+\centering
+\includegraphics[width=.5\textwidth]{plots/bounded_flux.pdf}
+\caption{Total population biomass is on the y axis, and generation is on the x axis. These simulations were run with climate velocity $ = 0.1$, and a proportional harvest rate $= 0.08$. }
 % see bounded\_flux.R for code that made this plot

### 2014-03-28 Writing/Appendix.tex
clarifying math in appendix

@@
 \\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2}^{L/2}k(\overline{x}_{t}-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \text{ using (\ref{pulse})} \notag
-\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2}^{L/2}k(\overline{x}_{t}-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \notag
+%\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2}^{L/2}k(\overline{x}_{t}-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \notag
 \\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2+ct}^{L/2+ct}k(x-ct-(y-ct))f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \notag
@@
 %\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(f(n_t(y)-g(n_t(y)))-f(n^*(\overline{y})-g(n^*(\overline{y}))\big)dy \notag
+\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(n^*(\overline{y})))(n_t(y)-n^*(\overline{y}_t))f'(n^*(\overline{y})-g(n^*(\overline{y}))dy \notag
+\\& \text{ by linearizing around the traveling pulse} \notag
 \\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(n^*(\overline{y})))f'(n^*(\overline{y})-g(n^*(\overline{y}))\xi_t(y)dy \notag
-\\& \text{ by linearizing around the traveling pulse} \notag
 \\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(0))f'(0)\xi_t(y)dy \text{ if $n^*(\overline{x})=0$} \label{perturb}
@@
 \begin{align}
-\lambda u(x)&=f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right) \notag
-\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right) \notag \text{for any $k$}
-\\ \Rightarrow \lambda d_k&=f'(0)\sum_{n=1}^\infty A_{nk}d_n  \label{problem}
+\lambda u(x)&=(1-g'(0))f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right) \notag
+\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=(1-g'(0))f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right) \notag
+\\& \text{for any $k$} \notag
+\\ \Rightarrow \lambda d_k&=(1-g'(0))f'(0)\sum_{n=1}^\infty A_{nk}d_n  \label{problem}
 \end{align}
@@
 \end{equation*}
-Finding the eigenvalues of (\ref{eigen}) then reduces to finding the eigenvalues of the matrix $(A_{nk})_{n,k=1}^\infty$.
+Finding the eigenvalues of (\ref{eigen}) then reduces to finding the eigenvalues of the matrix comprised of entires $(A_{nk})_{n,k=1}^\infty$.
 
@@
 $$k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}},$$
-where $D=\frac{\pi}{4}\langle d\rangle$ and $\langle d \rangle$ is the expected distance a larva disperses.
+where $D$ is one half the variance of the kernel.
 This is a separable kernel with

### 2014-06-09 Simluations/Figures/MPA_figs.Rmd
finished simulations

--- 2014-02-11 Writing/plots/figureCode/MPA_figs.Rmd
@@
 ```{r loadData, echo=FALSE, message=FALSE}
-#sim=read.csv("~/Desktop/noMPAnotThresh_add_2013-12-09.csv")
-# for emma's local copy
-#sim=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv")
-
 # for synced online github data
-x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL25vTVBBbm90VGhyZXNoX2FkZF8yMDEzLTEyLTA5LmNzdiIsImV4cGlyZXMiOjEzOTI2NDYxMDl9--46e45a3eb2faef7fc22b6124253c925bd2290de9")
+x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzYwfQ%3D%3D--5fa73caaafffae7c95d930f61def04ee9b96930a")
 sim = read.csv(textConnection(x))
 
-fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2Zpc2hNUEFub3RUaHJlc2hfYWRkXzIwMTMtMTItMDkuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE0NH0%3D--8eff4ee16d08efcf6b266c3e4e42adde7aeff969")
+fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzQ2fQ%3D%3D--9b6ceb0092173222543c95381dfbb0f42f0c884f")
 fishmpa = read.csv(textConnection(fish))
 
-cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2NvbnNNUEFub3RUaHJlc2hfYWRkXzIwMTQtMDEtMjMuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE2Nn0%3D--586d3cb31ac0f1fe3c49f48886e3796b5132bb77")
+cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-03-01.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMy0wMS5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzIwfQ%3D%3D--5f4f584256a4175eebe2ef5465e83c92ebe65cc0")
 consmpa = read.csv(textConnection(cons))
@@
 ```{r results}
-par(mfrow=c(1,3))
+
+ggplot(consmpa, aes(speed, harvest)) + geom_tile(aes(fill=Equil.pop, alpha = cut(Equil.pop, breaks=c(0,1, Inf)))) + scale_alpha_manual(values=c(0,1),guide="none") + scale_fill_gradient(low="gray99",high="black") + theme_minimal()
+
 noMang <- plotfig(sim); plotParams(noMang)

### 2014-06-09 Simluations/Figures/compareCatch.Rmd
finished simulations

--- 2014-01-29 Writing/plots/figureCode/compareCatch.Rmd
@@
 # for synced online github data
-  x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL25vTVBBbm90VGhyZXNoX2FkZF8yMDEzLTEyLTA5LmNzdiIsImV4cGlyZXMiOjEzOTEyNjg5OTB9--e041d4ef68764b1fc49140c128bda685c6f740f3")
+   x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL25vTVBBbm90VGhyZXNoX2FkZF8yMDEzLTEyLTA5LmNzdiIsImV4cGlyZXMiOjEzOTI2NDYxMDl9--46e45a3eb2faef7fc22b6124253c925bd2290de9")
   sim = read.csv(textConnection(x))
 
-  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2Zpc2hNUEFub3RUaHJlc2hfYWRkXzIwMTMtMTItMDkuY3N2IiwiZXhwaXJlcyI6MTM5MTI2OTAzM30%3D--f35f2593a7e901c900188fadca920fdd68a57e35")
+  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2Zpc2hNUEFub3RUaHJlc2hfYWRkXzIwMTMtMTItMDkuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE0NH0%3D--8eff4ee16d08efcf6b266c3e4e42adde7aeff969")
   fishmpa = read.csv(textConnection(fish))
 
-  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2NvbnNNUEFub3RUaHJlc2hfYWRkXzIwMTQtMDEtMjMuY3N2IiwiZXhwaXJlcyI6MTM5MTI2OTA3Mn0%3D--ffb8451d33daaeab48f8b3be3357cacbda0aadbb")
+  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2NvbnNNUEFub3RUaHJlc2hfYWRkXzIwMTQtMDEtMjMuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE2Nn0%3D--586d3cb31ac0f1fe3c49f48886e3796b5132bb77")
   consmpa = read.csv(textConnection(cons))

### 2014-06-09 Simluations/Aspatial_fast/Functions.R

@@ -11,7 +11,6 @@ f<-function(n,R0,K) return(R0*n/(1+(R0-1)/K*n))
 #Standard error removing NAs
 stderr <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
 
-
 moveMPA <- function(MPA.current = MPA.current, displaced = displaced, mpa.yes=mpa.yes, mpa.no=mpa.no, world){
   
   ##### Move MPAS
@@ -77,19 +76,14 @@ m <- function(n, s, Fthresh = NA, Fharv = NA, mpa.yes = NA, mpa.no = NA, MPA.cur
 	# harvesting occurs first - check to see how should re-allocate effort
 	
 	if(!is.na(effort_re_allocate)){	
-		total_catch = sum(n*Fharv)
+		total_catch = sum(n)*Fharv
 		available_total_pop = sum(n[which(MPA.current==0)]) # pop with no MPA coverage
-		available_fish <- n[which(MPA.current==0)] # available_total_pop # proportion at each point
+		available_fish = rep(0,length(n))
+		available_fish[MPA.current==0] <- n[MPA.current==0]/available_total_pop # available_total_pop 
+		# proportion at each point
 		catch_in_space <- total_catch*available_fish # allocate catch
 		
-		# NEED check that total catch is preserved
-		#sum(catch_in_space == total_catch)
-		
-		harvest <- MPA.current
-		harvest[harvest==0]<-catch_in_space
-		harvest[harvest==1]<-0
-		
-		next_gen = n-harvest
+		next_gen = n-catch_in_space
 		next_gen[next_gen<0] = 0
 	}else{
 		if(!is.na(Fthresh)) { # if thresholds
@@ -135,133 +129,41 @@ m <- function(n, s, Fthresh = NA, Fharv = NA, mpa.yes = NA, mpa.no = NA, MPA.cur
  	return(list(n2,harv,MPA))
 }
 
-# goes until the slope is less than 0.001 as measured with linear regression. Returns final population 
-startOut <- function(s, mpa.yes, mpa.no, burn_in, Fharv, Fthresh, init, MPA.start){
-	# burn in:
+longRun <- function(s, mpa.yes, mpa.no, Fthresh, Fharv, init, MPA.start, generations_total, generations_av, effort_re_allocate=NA){
 	MPA.current <- MPA.start
-	cat("burning in...\n")
-		for(t in 1:burn_in){
[…]

### 2014-06-09 Simluations/Aspatial_fast/Make.R

@@ -1,25 +1,38 @@
 rm(list=ls())
 setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
-require(zoo)
+require(plyr)
+require(lattice)
 
 # load parameters, functions
-  source("Parameters.R")
+	source("Parameters.R")
 	source("Functions.R")
 	
+# set up different parameter combinations for all simulations
+
+sims <- data.frame(model = c("noThresh","noThresh","noThresh","noThresh","noThresh","Thresh"),
+					MPA = c("cons","cons","fish","fish","null","null"),
+					effort_allocate = c(NA, TRUE, NA, TRUE, NA, NA), stringsAsFactors=FALSE)
+					
+for(run in 1:nrow(sims)){
+
 # run analysis
 	# choose threshold or no threshold
-	model = "noThresh"	# "noThresh"; "Thresh"
-	
+	model = sims$model[run]	# "noThresh"; "Thresh"
+
 	# if no threshold, choose MPA: "null", "cons", "fish"
-	MPA = "cons"
+	MPA = sims$MPA[run]
+		# choose how effort should be allocated if MPAs present. 
+		effort_allocate = sims$effort_allocate[run]
+		effort_allocate = ifelse(MPA!="null",effort_allocate,NA)
 
 # analysis
+timed <- system.time(
 	if(model=="noThresh") {sapply(c("Parameters_nothresh.R","Sim_noThresh.R"),source,.GlobalEnv)} else {
 		if(model=="Thresh" & MPA == "null") {sapply(c("Parameters_thresh.R","Sim_thresh.R"),source,.GlobalEnv)} else {
 			warning("model needs to be 'noThresh' or 'Thresh', MPA needs to be 'null'")
 			}
 		}
-
-# 2014-02-27 
-
-
+	)
+		
+cat(paste("Time elapsed: ",timed[1]/360," hours\n","Finished running a ", model, " simulation with ", MPA, " MPAs", " and effort re_allocate set to ", effort_allocate,".\n",nrow(sims)-run, " simulations left to go...",sep=""))
+}

### 2014-06-09 Simluations/Aspatial_fast/Parameters.R

@@ -1,4 +1,4 @@
-######################################
+ ######################################
 ## Parameters & Building Structures ##
 ######################################
 

### 2014-06-09 Simluations/Aspatial_fast/Parameters_thresh.R

@@ -5,16 +5,10 @@
 # build dataframes
 	summaries <- data.frame(
 		Equil.pop = rep(NA,length=length(speeds)*length(thresholds)), 
-		Equil.pop.sd = rep(NA, length=length(speeds)*length(thresholds)), 
-		Equil.pop.se = rep(NA, length=length(speeds)*length(thresholds)), 
-		Equil.harvest = rep(NA,length=length(speeds)*length(thresholds)),
-		Equil.harvest.sd = rep(NA,length=length(speeds)*length(thresholds)), 
-		Equil.harvest.se = rep(NA,length=length(speeds)*length(thresholds)), 
+		Equil.sd = rep(NA, length=length(speeds)*length(thresholds)), 
 		speed = rep(NA,length=length(speeds)*length(thresholds)), 
 		harvest = rep(NA,length=length(speeds)*length(thresholds)), 
-		thresh=rep(NA,length=length(speeds)*length(thresholds)), 
-		steps.harv = rep(NA,length=length(speeds)*length(thresholds)),
-		steps.move = rep(NA,length=length(speeds)*length(thresholds)))
+		thresh=rep(NA,length=length(speeds)*length(thresholds)))
 
 
 # index for row number

### 2014-06-09 Simluations/Aspatial_fast/Sim_thresh.R

@@ -1,109 +1,27 @@
-
-# added a stipulation for equilibrium: either the difference between two time steps is as small as the initializing step or the difference in the rolling mean with a window of 100 values [move_window below] of the total population abundance is 0. With MPAs getting oscillations with perfect oscillations such that it's never 'at equilibrium' but is not trending. so is at equilibrim
-
-if(MPA=="cons") {mpa.yes=cons.yes; mpa.no=cons.no} else {
-  if(MPA=="fish") {mpa.yes=fish.yes; mpa.no=fish.no} else {
-    if(MPA=="null") {mpa.yes=null.yes; mpa.no=null.no} else{
-      if(exists("MPA")) warning(paste("MPA needs to be 'cons', 'fish', or 'null'.",sep=""))
-    }
-  }
-}
-
-
-
-# initialize the world, set MPA (cons.yes/no; fish.yes/no; null.yes/no)
-	output <- startOut(w,maxt,mpa.yes,mpa.no,world)
-	init <- output[[1]]
+# initializing the population with no pressure (no harvesting, no climate)
+	init<-rep(0,w) # rows are world, columns are time
+	init[which(patch==0.55)]=50
+	MPA.start = rep(c(mpa.yes,mpa.no),length.out=length(world))
+	
+	output <- startUp(s=0,mpa.yes=mpa.yes,mpa.no=mpa.no,burn_in=burn_in, Fharv=NA, Fthresh=NA, init=init, MPA.start = MPA.start)
+	init.s <- output[[1]]
 	MPA.start <- output[[2]]
-
-# thresholds need to be relative to maximum population along 1d world
-	biggestP <- max(init[,maxt])
-	thresholds <- thresholds * biggestP
-# add a buffer to last to make sure captures biggest population
-  thresholds[length(thresholds)]= thresholds[length(thresholds)] + 0.001
-
-# standard for equilibrium is the difference in the final step of the initialization 
-	init.diff <- diff(colSums(init))[149]
-
-# Loop through various speeds and harvest intensities as definited in the "Parameters.R" file
-# using while loops. Only saving two time steps 
-
+	
 for(q in 1:length(speeds)){
-  
[…]

### 2014-06-10 Simluations/Aspatial_fast/Functions.R
working on figuring out effort

@@ -118,7 +118,6 @@ m <- function(n, s, Fthresh = NA, Fharv = NA, mpa.yes = NA, mpa.no = NA, MPA.cur
 	# keep individuals still in patch + those now in it due to move
 	next_patch[1:length(patch)] = next_n[1:length(patch)]
 	
-    harv = n-next_gen
     babies = next_patch*f_ind
     n2 = babies %*% d *step_size
     n2 = sapply(n2,f,R0,K)
@@ -140,7 +139,6 @@ longRun <- function(s, mpa.yes, mpa.no, Fthresh, Fharv, init, MPA.start, generat
 	
 	# make dataframe for simulation average
 	pop <- rep(0,generations_av)
-	sd <- rep(0,generations_av)
 	for(keep in 1:generations_av){
 		output = m(n=init, s = s, Fthresh=Fthresh,Fharv=Fharv, mpa.yes = mpa.yes, mpa.no = mpa.no, MPA.current = MPA.current)
 		init = output[[1]]

### 2014-06-10 plots/Fish_Figures.R

@@ -13,8 +13,6 @@ require(ggplot2)
 require(ggthemes)
 
 ###################################### Figure 1
-
-### A
 	analytical_ebm=ebm
 	cvals=matrix(seq(0,1,by=0.01),nrow=1)
 
@@ -46,31 +44,34 @@ require(ggthemes)
 		geom_line(aes(color=factor(R), linetype=factor(d)),size=1) + 
 		theme_tufte() + 
 		scale_color_manual(values=c("grey", "dark grey", "black")) + 
-		theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(1,0,0,0),"cm")) + 
-		xlab("") + ylab("Harvesting rate") + 
-		labs(linetype=expression(symbol("\341")*d*symbol("\361")), color=expression("R"[0]), title="A")
+		theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm")) + 
+		xlab("Climate Velocity") + ylab("Harvesting rate") + 
+		labs(linetype=expression(symbol("\341")*d*symbol("\361")), color=expression("R"[0]), title="")
 	
+	png("Fig1.png",height=3,width=4,units="in",res=300)
+	print(plot1)
+	dev.off() 
+	
+################################################################## FIGURE 2
 
-#### B
+#### A
 	melted_ebm <- melt(ebm)
-	speeds <- rep(seq(0,1, 0.01), 101)
-	harvests <- rep(seq(0,1,0.01), each = 101)
+	speeds <- rep(seq(0,1, 0.01), 100)
+	harvests <- rep(seq(0.01,1,0.01), each = 101)
 	melted_ebm$speed = speeds
 	melted_ebm$harvest = harvests
 
-	plot2 <- ggplot(melted_ebm, aes(x=speed, y = harvest, fill=value)) + 
+	plot2a <- ggplot(melted_ebm, aes(x=speed, y = harvest, fill=value)) + 
 		geom_tile() + theme_tufte() + 
 		scale_fill_gradient(low="black", high="white") + 
-		theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(1,0,0,0),"cm")) + 
+		theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm")) + 
 		guides(fill = guide_colorbar(barwidth=1, barheight=10, title.position="top", title="Equilibrium\nBiomass")) + 
[…]

### 2014-06-10 Simluations/Figures/diagnose_consmpa.Rmd

--- 2014-03-04 Simluations/Aspatial_fast/Figures/diagnose_consmpa.Rmd
@@
 ```{r}
- ggplot(consmpa, aes(x=speed, y = Equil.pop)) + geom_line() + facet_wrap(~harvest)
+ ggplot(consmpa, aes(x=speed, y = Equil.pop)) + geom_line() + facet_wrap(~harvest) + geom_point()
 ```
@@
 + `h = 0.09; s = 0.24-0.28`
-+ `h = 0.10; s = 0.14-0.16`
-+ `h = 0.12; s = 0.14-0.16`
++ `h = 0.10; s = 0.14-0.18`
++ `h = 0.12; s = 0.14-0.18`
 
@@
 steps = 300   # setting number of steps over which to take equilibrium 
-harvests = 0.08
-speeds = seq(0.24,0.28,0.02)
+harvests = seq(0,.15,0.01)
+speeds = c(0.1,0.3)
 MPA = "cons"

### 2014-06-11 Writing/Appendix.tex
working on figuring out effort
reordering events in appendix

@@
 
-\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y))dy \label{integro} \end{equation}
+\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0f(n_t(y))dy \label{integro} \end{equation}
 as described in \citet{ZhouKot2011}.  With the addition of harvesting, the model becomes
 
-\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy. \label{integro} \end{equation}
+\begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy. \label{integro} \end{equation}
 
@@
 n^*(\overline{x}_{t+1})&=n_{t+1}(x) \notag
-\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy \notag
-\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-\overline{y}_t-ct)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))dy \notag
-\\ &=\int_{-L/2+ct}^{L/2+ct}k(\overline{x}_t-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))dy \notag
-\\ \Rightarrow n^*(\overline{x}_t-c)&=\int_{-L/2+ct}^{L/2+ct}k(\overline{x}_t-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))dy \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-\overline{y}_t-ct)R_0g(f(n^*(\overline{y}_t)))dy \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(\overline{x}_t-\overline{y}_t)R_0g(f(n^*(\overline{y}_t)))dy \notag
+\\ \Rightarrow n^*(\overline{x}_t-c)&=\int_{-L/2+ct}^{L/2+ct}k(\overline{x}_t-\overline{y}_t)R_0g(f(n^*(\overline{y}_t)))dy \notag
 %\\n^*(\overline{x}-c)&=\int_{-L/2}^{L/2}k(\overline{x}-\overline{y})f(n^*(\overline{y})-g(n^*(\overline{y})))d\overline y \notag
-\\ \Rightarrow n^*(\overline{x}_t)&=\int_{-L/2}^{L/2}k(\overline{x}_t+c-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t  \label{pulse}
+\\ \Rightarrow n^*(\overline{x}_t)&=\int_{-L/2}^{L/2}k(\overline{x}_t+c-\overline{y}_t)R_0g(f(n^*(\overline{y}_t)))d\overline y_t  \label{pulse}
 \end{align}
@@
 \\&=n_{t+1}(x)-n^*(\overline{x}_{t}-c) \notag
-\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2}^{L/2}k(\overline{x}_{t}-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \text{ using (\ref{pulse})} \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy-\int_{-L/2}^{L/2}k(\overline{x}_{t}-\overline{y}_t)R_0g(f(n^*(\overline{y}_t)))d\overline y_t \text{ using (\ref{pulse})} \notag
 %\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2}^{L/2}k(\overline{x}_{t}-\overline{y}_t)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \notag
-\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2+ct}^{L/2+ct}k(x-ct-(y-ct))f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \notag
-\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n_t(y)-g(n_t(y)))dy-\int_{-L/2+ct}^{L/2+ct}k(x-y)f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))d\overline y_t \notag
-\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(f(n_t(y)-g(n_t(y)))-f(n^*(\overline{y}_t)-g(n^*(\overline{y}_t))\big)dy \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy-\int_{-L/2+ct}^{L/2+ct}k(x-ct-(y-ct))R_0g(f(n^*(\overline{y}_t)))d\overline y_t \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy-\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n^*(\overline{y}_t)))d\overline y_t \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(R_0g(f(n_t(y)))-R_0g(f(n^*(\overline{y}_t)))\big)dy \notag
+\\ &=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0\big(g(f(n_t(y)))-g(f(n^*(\overline{y}_t)))\big)dy \notag
 %%
@@
 %\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(f(n_t(y)-g(n_t(y)))-f(n^*(\overline{y})-g(n^*(\overline{y}))\big)dy \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(n^*(\overline{y})))(n_t(y)-n^*(\overline{y}_t))f'(n^*(\overline{y})-g(n^*(\overline{y}))dy \notag
+\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g'(f(n^*(\overline{y})))f'(n^*(\overline{y}))(n_t(y)-n^*(\overline{y}_t))dy \notag
 \\& \text{ by linearizing around the traveling pulse} \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(n^*(\overline{y})))f'(n^*(\overline{y})-g(n^*(\overline{y}))\xi_t(y)dy \notag
-\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)(1-g'(0))f'(0)\xi_t(y)dy \text{ if $n^*(\overline{x})=0$} \label{perturb}
+\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g'(f(n^*(\overline{y})))f'(n^*(\overline{y}))\xi_t(y)dy \notag
+\\ \Rightarrow \xi_{t+1}(x)&=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g'(0)f'(0)\xi_t(y)dy \text{ if $n^*(\overline{x})=0$ and $f(0)=0$} \label{perturb}
 \end{align}
@@
 \begin{align*}
-\lambda u(x-ct-c)&=(1-g'(0))f'(0)\int_{-L/2+ct}^{L/2+ct}k(x-y)u(y-ct)dy
-\\ \\\Rightarrow \lambda u(\overline{x})&=(1-g'(0))f'(0)\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})u(\overline{y})dy
+\lambda u(x-ct-c)&=R_0g'(0)f'(0)\int_{-L/2+ct}^{L/2+ct}k(x-y)u(y-ct)dy
+\\ \\\Rightarrow \lambda u(\overline{x})&=R_0g'(0)f'(0)\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})u(\overline{y})dy
 \end{align*}
@@
 Define the integral operator
-$$ \psi_f(u)(x)=(1-g'(0))f'(0)\int_{-L/2}^{L/2}k(x+c-y)u(y)dy. $$
+$$ \psi_f(u)(x)=R_0g'(0)f'(0)\int_{-L/2}^{L/2}k(x+c-y)u(y)dy. $$
 Then the perturbation to the traveling pulse will satisfy 
@@
 
-The biomass in the equilibrium traveling wave depends on the specific functional form of the harvesting function $g(n)$.
-However, the persistence of the population only depends on $g'(0)$ so in this paper, we only considered a proportional harvesting function, i.e.
-the amount of fish harvested obeyed $g(n)=hn$.
-For this function, $g'(0)=h$.
+The biomass in the equilibrium traveling wave depends on the specific functional forms of the harvesting function $g(n)$ and the recruitment function $f(n)$.
+However, the persistence of the population only depends on $g'(0)$ and $f'(0)$.
+In this paper, we only considered a proportional harvesting function, i.e.
+the amount of fish harvested obeyed $g(n)=(1-h)n$.
+For this function, $g'(0)=1-h$.
+For the recruitment function we considered, $f'(0)=1$.
 
@@
 \begin{align}
-\lambda u(x)&=(1-g'(0))f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right) \notag
-\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=(1-g'(0))f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right) \notag
+\lambda u(x)&=R_0g'(0)f'(0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right) \notag
+\\ \Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx&=R_0g'(0)f'(0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right) \notag
 \\& \text{for any $k$} \notag
-\\ \Rightarrow \lambda d_k&=(1-g'(0))f'(0)\sum_{n=1}^\infty A_{nk}d_n  \label{problem}
+\\ \Rightarrow \lambda d_k&=R_0g'(0)f'(0)\sum_{n=1}^\infty A_{nk}d_n  \label{problem}
 \end{align}
@@
 \begin{align*}
-\lambda d_0(c)&=(1-h)f'(0)A_{00}(c)d_0(c)
-\\ \Rightarrow \lambda&=(1-h)R_0A_{00}(c)
+\lambda d_0(c)&=R_0(1-h)A_{00}(c)d_0(c)
+\\ \Rightarrow \lambda&=R_0(1-h)A_{00}(c)
 \\\text{ where } A_{00}(c)&=2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]
@@
 \centering
-\includegraphics[width=.5\textwidth]{plots/bounded_flux.pdf}
+\includegraphics[width=.75\textwidth]{plots/bounded_flux.pdf}
 \caption{Total population biomass is on the y axis, and generation is on the x axis. These simulations were run with climate velocity $ = 0.1$, and a proportional harvest rate $= 0.08$. }

### 2014-06-11 Writing/Appendix.tex
adding recursive eq biomass equations to appendix

@@
 
+To find the equilibrium biomass, we rewrite (\ref{pulse}) using the separable kernel as in \cite{Latore:1998fk}:
+
+\begin{align*}
+ n^*(x)&=\int_{-L/2}^{L/2}k(x+c-y)R_0g(f(n^*(y)))dy
+ \\&= \int_{-L/2}^{L/2}\left(\sum_{n=1}^\infty a_n(x)b_n(y-c)\right)R_0g(f(n^*(y)))dy
+  \\&=\sum_{n=1}^\infty a_n(x) \int_{-L/2}^{L/2}b_n(y-c)R_0g(f(n^*(y)))dy
+\end{align*}
+If we define $m_n=\int_{-L/2}^{L/2}b_n(y-c)R_0g(f(n^*(y)))dy$ then we find that 
+\begin{align}
+n^*(x)&=\sum_{n=1}^\infty m_na_n(x) \text { and } \notag
+\\m_n&=\int_{-L/2}^{L/2}b_n(y-c)R_0g\left(f\left(\sum_{n=1}^\infty m_na_n(y)\right)\right)dy \label{recursive}
+\end{align}
+The recursive equation (\ref{recursive}) allows us to find the $m_n$ numerically and we then find the total equilibrium biomass by integrating $n^*(x)$ over space.
+
 \subsection{Gaussian dispersal kernel \label{gausapp}}
@@
 \begin{equation}
-h^*\sim1- \frac{1}{R_0}\cdot g(L)q(\sigma^2,c^2,L^2+3c^2)
+h^*\sim1- \frac{1}{R_0}\cdot p(L)q(\sigma^2,c^2,L^2+3c^2)
 \end{equation}
-where $g(L)$ is a decreasing function of the length of the viable patch $L$.
+where $p(L)$ is a decreasing function of the length of the viable patch $L$.
 

### 2014-06-30 Writing/Appendix.tex
removed the word recursive from appendix

@@
 \end{align}
-The recursive equation (\ref{recursive}) allows us to find the $m_n$ numerically and we then find the total equilibrium biomass by integrating $n^*(x)$ over space.
+The equations (\ref{recursive}) allows us to find the $m_n$ numerically and we then find the total equilibrium biomass by integrating $n^*(x)$ over space.
 

### 2014-07-09 Writing/Appendix.tex
small rewording of text in appendix

@@
 \subsection{Determining stability \label{stab}}
-Let $n_t(x)$ be the number of adults at position $x$ at time $t$, let $k(x)$ be a dispersal kernel describingt the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring produced by a population of size $n$, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
+Let $n_t(x)$ be the number of adults at position $x$ at time $t$, let $k(x)$ be a dispersal kernel describingt the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring that settle and survive in juvenile population of size $n$, let $R_0$ be the intrinsic growth rate of the population, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
 In the absence of harvesting, the integrodifference model describing the population over time is given by
@@
 \end{align}
-As long as $f(0)=0$, there is a trivial solution to this problem where $n^*(\overline{x})\equiv 0$ for all $\overline{x}$, i.e., there is a trivial traveling pulse with no fish in it.
+As long as $f(0)=0$, there is a trivial solution to this problem where $n^*(\overline{x})\equiv 0$ for all $\overline{x}$, i.e., there is a trivial traveling pulse with no adults in it.
 If the trivial traveling pulse is unstable, even very small populations will persist or grow and avoid crashing back to the trivial pulse.
@@
 The biomass in the equilibrium traveling wave depends on the specific functional forms of the harvesting function $g(n)$ and the recruitment function $f(n)$.
-However, the persistence of the population only depends on $g'(0)$ and $f'(0)$.
+However, the persistence of the population only depends on $R_0$, $g'(0)$ and $f'(0)$.
 In this paper, we only considered a proportional harvesting function, i.e.

### 2014-07-09 Simluations/Aspatial_fast/Functions.R
adding edits for ms

@@ -2,7 +2,8 @@
 k<-function(x,y,b) return(1/2*b*exp(-b*abs(x-y)))    
 
 # dispersal matrix
-d<-array(,c(w,w)) #dispersal probabilities to point i from every point j
+d<-array(,c(w,w)) 
+#dispersal probabilities to point i from every point j
 for(i in 1:w) d[i,]=k(world[i],world,b)
 
 #Beverton-Holt recruitment
@@ -20,102 +21,136 @@ moveMPA <- function(MPA.current = MPA.current, displaced = displaced, mpa.yes=mp
   
   # are there any MPAs in <next_MPA>?
   #any(test[[i]]$next_MPA==1)
-  # IF FALSE, then need to figure out how many 0s were lost in move, and make sure to preserve interval of zeros == mpa.no, then fill in mpa.yes,mpa.no to length of world. But also this is only for when mpa.no exists on both sides of MPA_next. If exactly to the edge of one reserve is lost, should shift down to next if statement
+  # IF FALSE, then need to figure out how many 0s were lost in
+  # move, and make sure to preserve interval of zeros == mpa.no,
+  # then fill in mpa.yes,mpa.no to length of world. But also this
+  # is only for when mpa.no exists on both sides of MPA_next. If
+  # exactly to the edge of one reserve is lost, should shift down
+  # to next if statement
   if(any(next_MPA==1)==FALSE & any(lost==1)==FALSE){
-    # these are the intervals that are left behind as world moves forward
+    # these are the intervals that are left behind as world moves
+    # forward
     lost <- MPA.current[1:(displaced-1)] 
-    # this is the last continuous chunk of numbers at the end of <lost>  
-    length_last <- rep(tail(rle(lost)$value,1), tail(rle(lost)$lengths,1))
-    # want to know how long <length_last> is so can make sure to get correct interval
+    # this is the last continuous chunk of numbers at the end of
+    # <lost>  
+    length_last <- rep(tail(rle(lost)$value,1), 
+    	tail(rle(lost)$lengths,1))
+    # want to know how long <length_last> is so can make sure to
+    # get correct interval
     L_int <- length(length_last)
[…]

### 2014-07-09 Simluations/Aspatial_fast/Make.R

@@ -13,8 +13,8 @@ sims <- data.frame(model = c("noThresh","noThresh","noThresh","noThresh","noThre
 					MPA = c("cons","cons","fish","fish","null","null"),
 					effort_allocate = c(NA, TRUE, NA, TRUE, NA, NA), stringsAsFactors=FALSE)
 					
-for(run in 1:nrow(sims)){
-
+#for(run in 1:nrow(sims)){
+for(run in c(2,4)){ # just re-do effort reallocation
 # run analysis
 	# choose threshold or no threshold
 	model = sims$model[run]	# "noThresh"; "Thresh"

### 2014-07-11 Simluations/Aspatial_fast/Functions.R
adding new simulation data, final touches on MS

@@ -98,7 +98,7 @@ m <- function(n, s, Fthresh = NA, Fharv = NA, mpa.yes = NA,
 	# harvesting occurs first - check to see how should 
 	# re-allocate effort
 	
-	if(!is.na(effort_re_allocate)){	
+	if(!is.na(effort_re_allocate) & !is.na(Fharv)){	# harvesting non-zero and effort reallocate
 		#total_catch = sum(n)*Fharv
 		total_catch = sum(n)*Fharv * 1.5 # increasing effort by 50% to account for effort-reallocation
 		available_total_pop = sum(n[which(MPA.current==0)]) 
@@ -163,13 +163,13 @@ m <- function(n, s, Fthresh = NA, Fharv = NA, mpa.yes = NA,
 
 longRun <- function(s, mpa.yes, mpa.no, Fthresh, Fharv, init,
 	 MPA.start, generations_total, generations_av, 	
-	 effort_re_allocate=NA){
+	 effort_re_allocate=effort_allocate){
 	MPA.current <- MPA.start
 	burn_in <- generations_total - generations_av
 	for(t in 1:(burn_in)){
 		output = m(n=init, s = s, Fthresh=Fthresh,Fharv=Fharv,
 		 mpa.yes = mpa.yes, mpa.no = mpa.no, 
-		 MPA.current = MPA.current)
+		 MPA.current = MPA.current, effort_re_allocate=effort_allocate)
 		init= output[[1]]
 		MPA.current = output[[2]]
 	}
@@ -179,7 +179,7 @@ longRun <- function(s, mpa.yes, mpa.no, Fthresh, Fharv, init,
 	for(keep in 1:generations_av){
 		output = m(n=init, s = s, Fthresh=Fthresh,Fharv=Fharv,
 		 mpa.yes = mpa.yes, mpa.no = mpa.no, 
-		 MPA.current = MPA.current)
+		 MPA.current = MPA.current, effort_re_allocate = effort_allocate)
 		init = output[[1]]
 		MPA.current = output[[2]]
 		pop[keep] = sum(output[[1]])
@@ -197,12 +197,13 @@ longRun <- function(s, mpa.yes, mpa.no, Fthresh, Fharv, init,
 #  adding speed treatment
 
 startUp <- function(s, mpa.yes, mpa.no, Fthresh, Fharv, init,
-	 MPA.start, burn_in,effort_re_allocate=NA){
[…]

### 2014-07-11 Simluations/Aspatial_fast/Make.R

@@ -11,7 +11,7 @@ require(lattice)
 
 sims <- data.frame(model = c("noThresh","noThresh","noThresh","noThresh","noThresh","Thresh"),
 					MPA = c("cons","cons","fish","fish","null","null"),
-					effort_allocate = c(NA, TRUE, NA, TRUE, NA, NA), stringsAsFactors=FALSE)
+					effort_allocate = c(NA, "yes", NA, "yes", NA, NA), stringsAsFactors=FALSE)
 					
 #for(run in 1:nrow(sims)){
 for(run in c(2,4)){ # just re-do effort reallocation
@@ -34,5 +34,5 @@ timed <- system.time(
 		}
 	)
 		
-cat(paste("Time elapsed: ",timed[1]/360," hours\n","Finished running a ", model, " simulation with ", MPA, " MPAs", " and effort re_allocate set to ", effort_allocate,".\n",nrow(sims)-run, " simulations left to go...",sep=""))
+cat(paste("Time elapsed: ",round(timed[1]/3600,3)," hours\n","Finished running a ", model, " simulation with ", MPA, " MPAs", " and effort re_allocate set to ", effort_allocate,".\n",nrow(sims)-run, " simulations left to go...",sep=""))
 }

### 2014-07-11 Simluations/Aspatial_fast/Sim_noThresh.R

@@ -12,7 +12,7 @@ if(MPA=="cons") {mpa.yes=cons.yes; mpa.no=cons.no} else {
 	init[which(patch==0.55)]=50
 	MPA.start = rep(c(mpa.yes,mpa.no),length.out=length(world))
 	
-	output <- startUp(s=0,mpa.yes=mpa.yes,mpa.no=mpa.no,burn_in=burn_in, Fharv=NA, Fthresh=NA, init=init, MPA.start = MPA.start)
+	output <- startUp(s=0,mpa.yes=mpa.yes,mpa.no=mpa.no,burn_in=burn_in, Fharv=NA, Fthresh=NA, init=init, MPA.start = MPA.start, effort_re_allocate=NA)
 	init.s <- output[[1]]
 	MPA.start <- output[[2]]
 	
@@ -34,4 +34,4 @@ for(q in 1:length(speeds)){
 }	
 
 
-write.csv(summaries,file = paste("Data/MPA",MPA,"_",Sys.Date(),".csv",sep=""))
+write.csv(summaries,file = paste("Data/MPA",MPA,"_",effort_allocate,"_",Sys.Date(),".csv",sep=""))

### 2014-07-11 plots/Fish_Figures (Autosaved).R

@@ -220,7 +220,8 @@ ggplot(diff_TRUE, aes(x=speed, y = harvest, fill=Equil.pop)) + geom_tile() # loo
 
 # what about between harvest re-compensated
 
-consTrue <- data[[5]]
+consTrue <- data[[5]]# not sure why there are NAs, making them 0
+consTrue$Equil.pop[is.na(consTrue$Equil.pop)] <- 0
 fishTrue <- data[[6]]
 
 #compare cons without effort reallocated and effort removed
@@ -235,7 +236,8 @@ ggplot(diff_fish, aes(x=speed, y=harvest, fill=Equil.pop)) + geom_tile()
 
 # generate plot for appendix using reallocated effort
 
-
+	consTrue$management = rep("Few Large MPAs", nrow(consTrue))
+	fishTrue$management = rep("Many Small MPAs", nrow(fishTrue))
 
 plotA <- ggplot(sim, aes(x=speed, y = harvest, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="gray95")  + labs(title="A") + xlab("") + ylab("") + theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm"), legend.position="none") 
 

### 2014-07-11 plots/load_data.R

@@ -3,7 +3,11 @@ library(RCurl)
 
 # analytics
 
-x=getURL("https://raw.githubusercontent.com/emfuller/MovingFish/master/plots/eqbiomass_gaus.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvcGxvdHMvZXFiaW9tYXNzX2dhdXMuY3N2IiwiZXhwaXJlcyI6MTQwMjk0OTk1OX0%3D--7f3a49902fd6f47b2b8cda6b306da6a86d348bd0")
+where_load <- function(where_load){
+
+if(where_load=="github"){
+	
+	x=getURL("https://raw.githubusercontent.com/emfuller/MovingFish/master/plots/eqbiomass_gaus.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvcGxvdHMvZXFiaW9tYXNzX2dhdXMuY3N2IiwiZXhwaXJlcyI6MTQwMjk0OTk1OX0%3D--7f3a49902fd6f47b2b8cda6b306da6a86d348bd0")
 ebm=read.csv(textConnection(x),row.names=1)
 ebm=as.matrix(ebm,nrow=length(ebm[,1]))
 
@@ -20,3 +24,19 @@ x = getURL("https://raw.githubusercontent.com/emfuller/MovingFish/master/Simluat
 
 x = getURL("https://raw.githubusercontent.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/Thresh_2014-06-06.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL1RocmVzaF8yMDE0LTA2LTA2LmNzdiIsImV4cGlyZXMiOjE0MDI5NDk5Mjl9--0b46cc290706e13e37f2db47568f5752f1eb09ef")
 	thresh = read.csv(textConnection(x), row.names=1)
+}else{
+	if(where_load=="local"){
+		setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/")
[…]

### 2014-07-11 Writing/Appendix.tex
adding edits for ms
adding appendix figures

@@
 \usepackage{lineno}
+\usepackage{listings}
 \begin{document}
@@
 In Appendix \ref{MPA} we provide details on differences between small and large MPA simulations.
+In Appendix \ref{effort} we show how reallocation of effort does not qualitatively change our results in our MPA simulations.
 
@@
 In this paper, we only considered a proportional harvesting function, i.e.
-the amount of fish harvested obeyed $g(n)=(1-h)n$.
+the amount of adults harvested obeyed $g(n)=(1-h)n$.
 For this function, $g'(0)=1-h$.
@@
 \pagebreak
+
+\subsection{Re-allocation of Effort \label{effort}}
+~\\Spatial shifts of harvesters in response to MPA creation can affect sustainability of the system.
+We model the two extremes of response: either that harvesting pressure that occurred in now-MPAs is reallocated outside the reservers, or harvesting inside MPAs disappears from the system with the creation of MPAs and MPAs are effectively an effort control management technique.
+To reallocate effort we increased quota by 50\% in outside of MPAs.
+
+We found no qualitative change in our results.
+Reallocation of effort was effectively like increasing the harvesting rate, and thus the critical harvest rate was lower than when effort was removed from the system.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.75\textwidth]{plots/fig3a.pdf}
+\caption{Climate velocity is on the x axis, harvest rate is on the y axis. These simulations were run with effort reallocated to areas outside of MPAs. }
+% see bounded\_flux.R for code that made this plot
+\end{figure}
+
 \bibliography{fish}

### 2014-07-11 plots/Fish_Figures (Autosaved).R

@@ -1,279 +1,0 @@
-library(fields)
-library(lattice)
-library(colorspace)
-library(RCurl)
-source("biomass_synergy.R")
-source("critical_parameters.R")
-source("load_data.R")
-require(stringr)
-require(colorspace)
-require(reshape2)
-require(gridExtra)
-require(ggplot2)
-require(ggthemes)
-
-where_load = "local"
-data <- where_load("local")
-
-sim <- data[[1]]
-mpas <- data[[2]]
-mpas2 <- data[[3]]
-thresh <- data[[4]]
-
-###################################### Figure 1
-	analytical_ebm=ebm
-	cvals=matrix(seq(0,1,by=0.01),nrow=1)
-
-	nr=3
-	ns=2
-	rvals=c(rep(3,ns),rep(5,ns),rep(10,ns))
-	rcols=c(rep('black',ns),rep('red',ns),rep('blue',ns))
-	xvals=c(.1,.5)
-	sigvals=rep(pi/2*xvals^2,nr)
-	sigltys=rep(c(1,2),nr)
-	l=length(rvals)
-
-# save data as list
-	h = vector("list",length=l)
-
-	for(i in 1:l){
-		r=rvals[i]
-		sig2=sigvals[i]
-		h[[i]]=apply(cvals,2,hstar_g,r=r,sig2=sig2)
-	}
-	
-	melted_h <- melt(h)
-	names(melted_h) <- c("h", "Param_combo")
-	melted_h$speed <- rep(cvals, 6)
-	melted_h$R <- rep(rvals, each = 101)
-	melted_h$d <- round(rep(xvals, each =101),digits=2)
-
-	plot1 <- ggplot(melted_h, aes(x=speed, y = h, group=Param_combo)) + 
-		geom_line(aes(color=factor(R), linetype=factor(d)),size=1) + 
-		theme_tufte() + 
-		scale_color_manual(values=c("grey", "dark grey", "black")) + 
-		theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm")) + 
-		xlab("Climate Velocity") + ylab("Harvesting rate") + 
-		labs(linetype=expression(symbol("\341")*d*symbol("\361")), color=expression("R"[0]), title="")
-	
-	png("Fig1.png",height=3,width=4,units="in",res=300)
-	print(plot1)
-	dev.off() 
-	
-################################################################## FIGURE 2
-
[…]

### 2014-07-11 plots/Fish_Figures.R

@@ -220,7 +220,8 @@ ggplot(diff_TRUE, aes(x=speed, y = harvest, fill=Equil.pop)) + geom_tile() # loo
 
 # what about between harvest re-compensated
 
-consTrue <- data[[5]]
+consTrue <- data[[5]]# not sure why there are NAs, making them 0
+consTrue$Equil.pop[is.na(consTrue$Equil.pop)] <- 0
 fishTrue <- data[[6]]
 
 #compare cons without effort reallocated and effort removed
@@ -235,7 +236,8 @@ ggplot(diff_fish, aes(x=speed, y=harvest, fill=Equil.pop)) + geom_tile()
 
 # generate plot for appendix using reallocated effort
 
-
+	consTrue$management = rep("Few Large MPAs", nrow(consTrue))
+	fishTrue$management = rep("Many Small MPAs", nrow(fishTrue))
 
 plotA <- ggplot(sim, aes(x=speed, y = harvest, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="gray95")  + labs(title="A") + xlab("") + ylab("") + theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm"), legend.position="none") 
 
@@ -272,6 +274,6 @@ return(legend)}
 legend <- g_legend(plotD)
 lwidth <- sum(legend$width)
 
-png(file="fig3.png",width=8,height=6,res=300,units="in")
+png(file="fig3a.png",width=8,height=6,res=300,units="in")
 		grid.arrange(arrangeGrob(plotA, plotB , plotC,plotD + theme(legend.position="none")), legend, left ="\nHarvest", sub="Climate velocity\n", widths=unit.c(unit(1, "npc") - lwidth,lwidth),ncol=2)
 		dev.off()

### 2014-07-11 Writing/Appendix.tex
adding a bit about MPAs in methods/results

@@
 \subsection{Re-allocation of Effort \label{effort}}
-~\\Spatial shifts of harvesters in response to MPA creation can affect sustainability of the system.
+~\\Spatial shifts of harvesters in response to MPA creation can affect sustainability of the system \citep{Botsfordetal2009}.
 We model the two extremes of response: either that harvesting pressure that occurred in now-MPAs is reallocated outside the reservers, or harvesting inside MPAs disappears from the system with the creation of MPAs and MPAs are effectively an effort control management technique.
-To reallocate effort we increased quota by 50\% in outside of MPAs.
+To reallocate effort we increased quota by 50\% in outside of MPAs (see \cite{Kaplanetal2005} for similar approach).
 

### 2014-07-11 plots/Fish_Figures.R

@@ -274,6 +274,6 @@ return(legend)}
 legend <- g_legend(plotD)
 lwidth <- sum(legend$width)
 
-png(file="fig3a.png",width=8,height=6,res=300,units="in")
+pdf(file="fig3a.pdf",width=8,height=6)
 		grid.arrange(arrangeGrob(plotA, plotB , plotC,plotD + theme(legend.position="none")), legend, left ="\nHarvest", sub="Climate velocity\n", widths=unit.c(unit(1, "npc") - lwidth,lwidth),ncol=2)
 		dev.off()

### 2015-01-28 Writing/older_stuff/Appendix.tex
accepted all of Malin's comments, and reorganized our writing folder for clarity

--- 2014-07-09 Writing/Appendix.tex
@@
 \pagebreak
-\subsection{MPA fluctuations \label{MPA}}
-~\\After the simulations come to equilibrium, the fluctuations in total biomass per generation fluctuate more in MPAs that are larger and spaced farther apart than simulations in which the MPAs that are smaller and more closely spaced.
-The large MPAs have a slightly larger average population, however large MPAs here can induce fluctuations of biomass even in deterministic simulations.
-Thus we expect if reproduction was stochastic, large MPAs spaced far apart would be more likely to result in extinction of the population than more closely spaced, smaller MPAs.
+\subsection{Protected Area fluctuations \label{MPA}}
+~\\After the simulations come to equilibrium, the fluctuations in total biomass per generation fluctuate more in reserves that are larger and spaced farther apart than simulations in which the reserves that are smaller and more closely spaced.
+The large reserves have a slightly larger average population, however large reserves here can induce fluctuations of biomass even in deterministic simulations.
+Thus we expect if reproduction was stochastic, large reserves spaced far apart would be more likely to result in extinction of the population than more closely spaced, smaller reserves.
+We find the same effect regardless of whether or not effort remains constant.
 \begin{figure}[h]
 \centering
-\includegraphics[width=.75\textwidth]{plots/bounded_flux.pdf}
-\caption{Total population biomass is on the y axis, and generation is on the x axis. These simulations were run with climate velocity $ = 0.1$, and a proportional harvest rate $= 0.08$. }
+\includegraphics[width=.75\textwidth]{plots/compare_flucts.png}
+\caption{Total population biomass is on the y axis, and generation is on the x axis. These simulations were run with climate velocity $ = 0.1$, and a proportional harvest rate $= 0.02$. }
 % see bounded\_flux.R for code that made this plot
@@
 \pagebreak
+
 \bibliography{fish}

### 2015-01-28 Simluations/Aspatial_fast/Parameters.R
starting to adjust parameters for rockfish parameterization

@@ -3,14 +3,24 @@
 ## Parameters & Building Structures ##
 ######################################
 
+# adding in parameterization for rockfish
+# currently need to figure out how to change the dispersal kernal. and re-scale so that interval of 1 is equal to 1000 km. So can re-run with proper scaling. 
+
+# - Mean dispersal distance: 73km (Gaussian)
+# - R0 = 2.86
+# - K = 1 (for simplicity)
+# - h = 0 to 1
+# - c = 0 to 200 km/decade
+# - L = 1000 km
+
 step_size=0.01 #distance between points in space
 b=.5 #parameter for Laplace dispersal kernel
-R0=5 #growth parameter for recruitment
-K=100 #carrying capacity parameter for juvenile density dependence
+R0=2.86 #growth parameter for recruitment
+K=1 #carrying capacity parameter for juvenile density dependence
 threshold = 0.001 #difference between generation populations. 
 burn_in = 2000 # number of generations to run simulations before checking for equilibrium conditions
 speeds = seq(0,.5,by=0.02)
-harvests = seq(0,.2,by=0.01)
+harvests = seq(0,1,by=0.1)
 f_ind = 1 #per capita reproductive rate
 generations_total = 8000
 generations_av = 2000

### 2015-02-10 Writing/older_stuff/Appendix.tex
generating plots of rockfish

@@
 In Appendix \ref{effort} we show how reallocation of effort does not qualitatively change our results in our MPA simulations.
+In Appendix \ref{rock} we parameterize our model for black rockfish in the California Current and demonstrate the qualitatively similarities of the results.
 
@@
 
+\pagebreak
+\subsection{Effort reallocation \label{effort}}
+Effort reallocation does not affect our results. 
+\pagebreak
+
+\pagebreak
+\subsection{California Current Black Rockfish Parameterization \label{rock}}
+We parameterize our model for black rockfish in the California Current, with MPAs of spacing and width that is qualitatively similar to the MLP Act reserves, and with a maximum climate velocity equal to that observed empirically.
+The parameters are as follows
+
+We find qualitatively similar results with this parameterization to what is presented in the main text. 
+\begin{figure}[h]
+\centering
+\includegraphics[width=.75\textwidth]{plots/rockfish_sims.pdf}
+\caption{The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis under alternative management strategies as parameterized for black rockfish.
+(a) The equilibrium biomass for simulations with constant harvest rates.
+(b) Equilibrium biomass for simulations with threshold management.
+For threshold management, the maximum threshold below which no harvesting is allowed is set to be the largest population size observed at a given time step before harvesting.
+For a less sever threshold, we use a proportion of this maximum threshold, so that a lower proportion gives a lower threshold and allows for more harvesting.
+We show this proportion on the y-axis.
+(c) Equilibrium biomass for simulations with protected areas where harvesting pressure outside reserves unchanged.
+(d) Equilibrium biomass for simulations with protected areas with harvesting pressure reallocated outside reserves.
+These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=2.83$, $K=1$, and $b=73$.}
+\end{figure}
+
+\pagebreak
 \bibliography{fish}

### 2015-02-10 plots/Fish_Figures.R

@@ -252,6 +252,7 @@ polygon(fN.p[[1]][fN.p[[2]],], col=alpha("grey",0.5),bor=F)
 
 cY.p <- poly(consTrue)
 polygon(cY.p[[1]][cY.p[[2]],],col=alpha("grey",0.5),bor=F)
+
 ############## Appendix
 
 #### Additional harvest simulations: compensation
@@ -375,3 +376,51 @@ ggplot(diff_cons, aes(x=speed, y=harvest, fill=Equil.pop)) + geom_tile() # no di
 diff_fish <- sim
 diff_fish$Equil.pop <- fishTrue$Equil.pop - sim$Equil.pop
 ggplot(diff_fish, aes(x=speed, y=harvest, fill=Equil.pop)) + geom_tile()
+
+## plot rockfish simulations
+setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/rockfish_sim/Data/")
+noThresh <- read.csv('MPAnull_NA_2015-02-09.csv')
+rockMPA_eff <- read.csv('MPArock_yes_2015-02-09.csv')
+rockMPA_na <- read.csv('MPArock_NA_2015-02-09.csv')
+thresh_noMPA <- read.csv('Thresh_2015-02-09.csv')
+
+
+plotA <- ggplot(noThresh, aes(x=speed, y = harvest, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="gray95")  + labs(title="A") + xlab("") + ylab("") + theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm"), legend.position="none") 
+
+
+thresh_noMPA$ord_thresh <- thresh_noMPA$thresh/max(thresh_noMPA$thresh)
+
+
+plotB <- ggplot(thresh_noMPA, aes(x=speed, y = ord_thresh, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="gray95") + labs(title="B")+ xlab("") + ylab("")+ theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm"),legend.position="none") + scale_y_reverse()
+
+plotC <-ggplot(rockMPA_na, aes(x=speed, y = harvest, fill = Equil.pop)) + 
+  geom_raster(interpolate=TRUE) + theme_tufte() + 
[…]

### 2015-02-10 Writing/older_stuff/Appendix.tex

@@
 In Appendix \ref{MPA} we provide details on differences between small and large MPA simulations.
-In Appendix \ref{effort} we show how reallocation of effort does not qualitatively change our results in our MPA simulations.
 In Appendix \ref{rock} we parameterize our model for black rockfish in the California Current and demonstrate the qualitatively similarities of the results.
@@
 \pagebreak
-
-\pagebreak
-\subsection{Effort reallocation \label{effort}}
-Effort reallocation does not affect our results. 
-\pagebreak
-
-\pagebreak
 \subsection{California Current Black Rockfish Parameterization \label{rock}}
 We parameterize our model for black rockfish in the California Current, with MPAs of spacing and width that is qualitatively similar to the MLP Act reserves, and with a maximum climate velocity equal to that observed empirically.
-The parameters are as follows
+The parameters and references are as follows
 
-We find qualitatively similar results with this parameterization to what is presented in the main text. 
+\begin{table}[htdp]
+\caption{Rockfish simulation parameters}
+\begin{center}
+\begin{tabular}{c l l}
+\hline
+Parameter & Value & Source\\
+\hline
+$R_0$ & 2.86 & \citet{white2010decision}, equivalent to $\left(\frac{1}{CRT}\right)$ \\
+$K$ & 1 & \citet{white2010decision} \\
+$h$ & 0-100\% &\\
+$c$ & 0-200km/decade & \citet{burrows2011pace}\\
+$L$ & 1000km & \\
+MPA width & 20km & \citet{Gainesetal2010}\\
+Space between MPAs & 76km & \citet{Gainesetal2010}\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \begin{figure}[h]
@@
 (c) Equilibrium biomass for simulations with protected areas where harvesting pressure outside reserves unchanged.
-(d) Equilibrium biomass for simulations with protected areas with harvesting pressure reallocated outside reserves.
-These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=2.83$, $K=1$, and $b=73$.}
+(d) Equilibrium biomass for simulations with protected areas with harvesting pressure reallocated outside reserves.}
 \end{figure}
 
-\pagebreak
+We find qualitatively similar results with this parameterization to what is presented in the main text.
+In particular, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse under some rates of range shift and harvesting rates, than no MPA at all (i.e.
+comparing Fig.
+2a and 2e).
+
 \bibliography{fish}

### 2015-02-11 Writing/older_stuff/Appendix.tex
working on finalizing discussion and incorporation of black rockfish results

@@
 In Appendix \ref{MPA} we provide details on differences between small and large MPA simulations.
-In Appendix \ref{rock} we parameterize our model for black rockfish in the California Current and demonstrate the qualitatively similarities of the results.
+In Appendix \ref{rock} we parameterize our model for black rockfish in the California Current and demonstrate results for those parameters that are qualitatively similar to those presented in the main text.
 
@@
 
-\pagebreak
+%\pagebreak
 \subsection{Protected Area fluctuations \label{MPA}}
-~\\After the simulations come to equilibrium, the fluctuations in total biomass per generation fluctuate more in reserves that are larger and spaced farther apart than simulations in which the reserves that are smaller and more closely spaced.
+~\\After the simulations come to equilibrium, the fluctuations in total biomass per generation fluctuate more in reserves that are larger and spaced farther apart than simulations in which the reserves that are smaller and more closely spaced (Figure \ref{fluctuations}).
 The large reserves have a slightly larger average population, however large reserves here can induce fluctuations of biomass even in deterministic simulations.
 Thus we expect if reproduction was stochastic, large reserves spaced far apart would be more likely to result in extinction of the population than more closely spaced, smaller reserves.
-We find the same effect regardless of whether or not effort remains constant.
-\begin{figure}[h]
-\centering
-\includegraphics[width=.75\textwidth]{plots/compare_flucts.png}
-\caption{Total population biomass is on the y axis, and generation is on the x axis. These simulations were run with climate velocity $ = 0.1$, and a proportional harvest rate $= 0.02$. }
-% see bounded\_flux.R for code that made this plot
-\end{figure}
+We find the same effect regardless of whether or not effort remains constant or is removed from the system.
 
-\pagebreak
+%\pagebreak
 \subsection{California Current Black Rockfish Parameterization \label{rock}}
 We parameterize our model for black rockfish in the California Current, with MPAs of spacing and width that is qualitatively similar to the MLP Act reserves, and with a maximum climate velocity equal to that observed empirically.
-The parameters and references are as follows
+The parameters and references are provided in Table \ref{rockfish_params}.
+
+Our results with this parameterization are qualitatively similar to the results presented in the main text.
+In particular, we find the same negative relationship between critical harvesting rate $h^*$ and the climate velocity $c$ (Figure \ref{crith_rockfish}) and an essentially additive interaction between the effects of the two stressors on biomass (Figure \ref{synergy_rockfish}).
+Additionally, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse under some rates of range shift and harvesting rates, than no MPA at all (compare Fig.
+\ref{simulation_rockfish}a and \ref{simulation_rockfish}d).
 
 \begin{table}[htdp]
-\caption{Rockfish simulation parameters}
+\caption{\label{rockfish_params} Rockfish simulation parameters}
 \begin{center}
@@
 \centering
+\includegraphics[width=.75\textwidth]{plots/compare_flucts.png}
+\caption{\label{fluctuations} Total population biomass is on the y axis, and generation is on the x axis.
+These simulations were run with climate velocity $ = 0.1$, and a proportional harvest rate $= 0.02$.
+}
+% see bounded\_flux.R for code that made this plot
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\caption{ \label{crith_rockfish}
+ Lines indicate the critical threshold for persistence as a function of harvesting rate on the y-axis and climate velocity as a fraction of the length of the viable patch on the x-axis from an approximated Gaussian dispersal kernel as parameterized for black rockfish.
+}
+ \end{figure}
+
+ \begin{figure}[h]
+ \centering
+ \caption{ \label{synergy_rockfish}
+ (a) The equilibrium biomass of the population as a function of the climate velocity relative to patch length on the x-axis and the proportional harvesting rate on the y-axis.
+(b) Interaction between the two stressors as a function of climate velocity and harvesting rate.
+The heat map indicates the interaction measure $S$, i.e., the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually ($E_{hc}-E_h-E_c$).
+$S$ of $0$ indicates additive interaction of the stressors.
+The excess loss, on the order of $0.001$, is small in comparison to the total biomass, which can be as large as $20$.
+These results are from an approximated Gaussian dispersal kernel as parameterized for  black rockfish.
+ }
+ \end{figure}
+
+\begin{figure}[h]
+\centering
 \includegraphics[width=.75\textwidth]{plots/rockfish_sims.pdf}
-\caption{The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvesting rate on the y-axis under alternative management strategies as parameterized for black rockfish.
+\caption{\label{simulation_rockfish} The equilibrium biomass of the population as a function of the climate velocity relative to patch length on the x-axis and the harvesting rate on the y-axis under alternative management strategies as parameterized for black rockfish.
 (a) The equilibrium biomass for simulations with constant harvest rates.
@@
 
-We find qualitatively similar results with this parameterization to what is presented in the main text.
-In particular, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse under some rates of range shift and harvesting rates, than no MPA at all (i.e.
-comparing Fig.
-2a and 2e).
-
 \bibliography{fish}

### 2015-02-19 Writing/final_draft/Appendix.tex
add axes and moved tick marks in figures and added all rockfish figures to appendix
updating Appendix to add figures
starting Malin's edits. Appendix is unfinished

--- 2015-02-11 Writing/older_stuff/Appendix.tex
@@
 In Appendix \ref{MPA} we provide details on differences between small and large MPA simulations.
-In Appendix \ref{rock} we parameterize our model for black rockfish in the California Current and demonstrate results for those parameters that are qualitatively similar to those presented in the main text.
+In Appendix \ref{rock} we parameterize our model for black rockfish in the California Current and demonstrate results for  parameters  which are qualitatively similar to those  presented in the main text.
 
@@
 \subsection{California Current Black Rockfish Parameterization \label{rock}}
-We parameterize our model for black rockfish in the California Current, with MPAs of spacing and width that is qualitatively similar to the MLP Act reserves, and with a maximum climate velocity equal to that observed empirically.
+~\\We parameterize our model for black rockfish in the California Current, with MPAs of spacing and width that is qualitatively similar to the MLP Act reserves, and with a maximum climate velocity equal to that observed empirically.
 The parameters and references are provided in Table \ref{rockfish_params}.
@@
 $c$ & 0-200km/decade & \citet{burrows2011pace}\\
-$L$ & 1000km & \\
+$L$ & 1000km  & \citet{fishbase} \\
 MPA width & 20km & \citet{Gainesetal2010}\\
 Space between MPAs & 76km & \citet{Gainesetal2010}\\
+Dispersal kernal &  Gaussian & \citet{white2010decision}\\
 \hline
@@
 \centering
-\includegraphics[width=.75\textwidth]{plots/compare_flucts.png}
+\includegraphics[width=.75\textwidth]{../../plots/compare_flucts.png}
 \caption{\label{fluctuations} Total population biomass is on the y axis, and generation is on the x axis.
@@
 \centering
+\includegraphics[width=.75\textwidth]{../../plots/rockfish_criticalh.pdf}
 \caption{ \label{crith_rockfish}
@@
  \centering
+ \includegraphics[width=.75\textwidth]{../../plots/rockfisheqbiomass.pdf}
  \caption{ \label{synergy_rockfish}
@@
 \centering
-\includegraphics[width=.75\textwidth]{plots/rockfish_sims.pdf}
+\includegraphics[width=.75\textwidth]{../../plots/rockfish_sims.pdf}
 \caption{\label{simulation_rockfish} The equilibrium biomass of the population as a function of the climate velocity relative to patch length on the x-axis and the harvesting rate on the y-axis under alternative management strategies as parameterized for black rockfish.
@@
 
+\cleardoublepage
+
 \bibliography{fish}

### 2015-02-19 Simluations/Aspatial_fast/rockfish_sim/Parameters.R

@@ -23,7 +23,7 @@ w = length(world)
 # MPAs are 20km wide and 75km apart. Step size is 2 km (1000*step_size). 
 
 rock.yes = rep(1,.02/step_size) 
-rock.no = rep(0,.076/step_size) # 10*8 = 80 (more than 75. so conservative)
+rock.no = rep(0,.076/step_size) # (more than 75. so conservative)
 
 null.yes = rep(0,length(world))
 null.no = rep(0, length(world))

### 2015-02-19 Writing/final_draft/Appendix.tex
updating Appendix

@@
 
+\renewcommand{\thefigure}{S\arabic{figure}}
+ \renewcommand{\thetable}{S\arabic{table}}
+
 \appendix
@@
 In Appendix \ref{MPA} we provide details on differences between small and large MPA simulations.
-In Appendix \ref{rock} we parameterize our model for black rockfish in the California Current and demonstrate results for  parameters  which are qualitatively similar to those  presented in the main text.
+In Appendix \ref{rock}, we parameterize our model for black rockfish ($Sebastes melanops$)in the California Current and demonstrate that results for  parameters  are qualitatively similar to results  presented in the main text.
 
@@
 \subsection{California Current Black Rockfish Parameterization \label{rock}}
-~\\We parameterize our model for black rockfish in the California Current, with MPAs of spacing and width that is qualitatively similar to the MLP Act reserves, and with a maximum climate velocity equal to that observed empirically.
+~\\We parameterize our model for black rockfish ($Sebastes melanops$) in the California Current, with MPAs of spacing and width qualitatively similar to those in the Marine Life Projection Act (MLPA), and with a maximum climate velocity equal to that observed empirically.
 The parameters and references are provided in Table \ref{rockfish_params}.
@@
 In particular, we find the same negative relationship between critical harvesting rate $h^*$ and the climate velocity $c$ (Figure \ref{crith_rockfish}) and an essentially additive interaction between the effects of the two stressors on biomass (Figure \ref{synergy_rockfish}).
-Additionally, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse under some rates of range shift and harvesting rates, than no MPA at all (compare Fig.
-\ref{simulation_rockfish}a and \ref{simulation_rockfish}d).
+Additionally, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse than no MPA at all (compare Figure \ref{simulation_rockfish}a and \ref{simulation_rockfish}d).
 

### 2015-02-19 Writing/final_draft/Appendix.tex
updating appendix, fixing bibs, adding some new figures

@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{natbib}
+\usepackage[round]{natbib}
 \bibliographystyle{plainnat}
@@
 $R_0$ & 2.86 & \citet{white2010decision}, equivalent to $\left(\frac{1}{CRT}\right)$ \\
-$K$ & 1 & \citet{white2010decision} \\
+$K$ & 1 & \cite{white2010decision} \\
 $h$ & 0-100\% &\\
-$c$ & 0-200km/decade & \citet{burrows2011pace}\\
-$L$ & 1000km  & \citet{fishbase} \\
-MPA width & 20km & \citet{Gainesetal2010}\\
-Space between MPAs & 76km & \citet{Gainesetal2010}\\
-Dispersal kernal &  Gaussian & \citet{white2010decision}\\
+$c$ & 0-200km/decade & \cite{burrows2011pace}\\
+$L$ & 1000km  & \cite{fishbase} \\
+MPA width & 20km & \cite{Gainesetal2010}\\
+Space between MPAs & 76km & \cite{Gainesetal2010}\\
+Dispersal kernal &  Gaussian & \cite{white2010decision}\\
 \hline
@@
 \caption{ \label{crith_rockfish}
-Lines indicate the critical threshold for persistence as a function of harvesting rate on the y-axis and climate velocity as a fraction of the length of the viable patch on the x-axis from an approximated Gaussian dispersal kernel as parameterized for black rockfish.
-}
+Lines indicate the critical threshold for persistence as a function of harvest fraction on the y-axis and climate velocity on the x-axis. Model parameterized for black rockfish. }
  \end{figure}
@@
  \caption{ \label{synergy_rockfish}
- (a) The equilibrium biomass of the population as a function of the climate velocity relative to patch length on the x-axis and the proportional harvesting rate on the y-axis.
-(b) Interaction between the two stressors as a function of climate velocity and harvesting rate.
+ (a) Results from a model parameterized for black rockfish.
+The equilibrium biomass of a black rockfish population as a function of the climate velocity on the x-axis and the harvest fraction on the y-axis.
+(b) Interaction between the two stressors as a function of climate velocity and harvest fraction.
 The heat map indicates the interaction measure $S$, i.e., the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually ($E_{hc}-E_h-E_c$).
 $S$ of $0$ indicates additive interaction of the stressors.
-The excess loss, on the order of $0.001$, is small in comparison to the total biomass, which can be as large as $20$.
-These results are from an approximated Gaussian dispersal kernel as parameterized for  black rockfish.
+The excess loss, on the order of $4$, is small in comparison to the total biomass, which can be as large as $160$.
  }
@@
 \includegraphics[width=.75\textwidth]{../../plots/rockfish_sims.pdf}
-\caption{\label{simulation_rockfish} The equilibrium biomass of the population as a function of the climate velocity relative to patch length on the x-axis and the harvesting rate on the y-axis under alternative management strategies as parameterized for black rockfish.
+\caption{\label{simulation_rockfish} The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvest fraction on the y-axis under alternative management strategies as parameterized for black rockfish.
 (a) The equilibrium biomass for simulations with constant harvest rates.
 (b) Equilibrium biomass for simulations with threshold management.
-For threshold management, the maximum threshold below which no harvesting is allowed is set to be the largest population size observed at a given time step before harvesting.
-For a less sever threshold, we use a proportion of this maximum threshold, so that a lower proportion gives a lower threshold and allows for more harvesting.
-We show this proportion on the y-axis.
-(c) Equilibrium biomass for simulations with protected areas where harvesting pressure outside reserves unchanged.
-(d) Equilibrium biomass for simulations with protected areas with harvesting pressure reallocated outside reserves.}
+For threshold management, we set a threshold density below which no fishing is allowed.
+The threshold ranges between 0 (no fishing allowed) and 1 (all fish taken), with intermediate density thresholds determined as fractions of the maximum population density observed at a given time step before harvesting.
+We show this on the y-axis.
+(c) Equilibrium biomass for simulations with protected areas where harvesting pressure outside reserves is unchanged  (i.e., harvest effort inside reserves is eliminated).
+(d) Equilibrium biomass for simulations with protected areas in which harvesting pressure is reallocated outside reserves.}
 \end{figure}
@@
 
+\bibstyle{cbe}
 \bibliography{fish}

### 2015-02-24 Writing/final_draft/Appendix.tex

@@
 \section{Appendix}
-In Appendix \ref{stab}, we provide the details for assessing the persistence of a population with an integrodifference model and we discuss the effect of the harvesting function on population persistence.
+~\\In Appendix \ref{stab}, we provide the details for assessing the persistence of a population with an integrodifference model and we discuss the effect of the harvesting function on population persistence.
 In Appendix \ref{sep}, we provide the details for assessing population persistence with separable dispersal kernels.
@@
 \subsection{Determining stability \label{stab}}
-Let $n_t(x)$ be the number of adults at position $x$ at time $t$, let $k(x)$ be a dispersal kernel describingt the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring that settle and survive in juvenile population of size $n$, let $R_0$ be the intrinsic growth rate of the population, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
+~\\Let $n_t(x)$ be the number of adults at position $x$ at time $t$, let $k(x)$ be a dispersal kernel describingt the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring that settle and survive in juvenile population of size $n$, let $R_0$ be the intrinsic growth rate of the population, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$.
 In the absence of harvesting, the integrodifference model describing the population over time is given by
@@
 \subsection{Separable dispersal kernels \label{sep}}
-It is not immediately obvious that the operator $\psi$ will have any eigenfunctions.
+~\\It is not immediately obvious that the operator $\psi$ will have any eigenfunctions.
 However, Jentzsch's theorem guarantees that there is an eigenfunction $u$, provided that the kernel $k$ satisfies some properties \citep{ZhouKot2011}.
@@
 \subsection{Gaussian dispersal kernel \label{gausapp}}
-The Gaussian dispersal kernel is given by
+~\\The Gaussian dispersal kernel is given by
 $$k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}},$$
@@
 \subsection{Sinusoidal dispersal kernel \label{sinapp}}
-The sinusoidal dispersal kernel is given by 
+~\\The sinusoidal dispersal kernel is given by 
 $$k(x-y)=\left\{\begin{array}{ccccc}

### 2015-02-24 plots/Fish_Figures.R
manually changing figures, updating appendix

@@ -377,13 +377,21 @@ diff_fish <- sim
 diff_fish$Equil.pop <- fishTrue$Equil.pop - sim$Equil.pop
 ggplot(diff_fish, aes(x=speed, y=harvest, fill=Equil.pop)) + geom_tile()
 
-## plot rockfish simulations
+## plot rockfish simulations ----
+library(fields)
+library(ggplot2)
+library(ggthemes)
 setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/rockfish_sim/Data/")
 noThresh <- read.csv('MPAnull_NA_2015-02-09.csv')
 rockMPA_eff <- read.csv('MPArock_yes_2015-02-09.csv')
 rockMPA_na <- read.csv('MPArock_NA_2015-02-09.csv')
 thresh_noMPA <- read.csv('Thresh_2015-02-09.csv')
 
+# change speeds units
+noThresh$speed <- noThresh$speed * 1000
+rockMPA_eff$speed <- rockMPA_eff$speed * 1000
+rockMPA_na$speed <- rockMPA_na$speed * 1000
+thresh_noMPA$speed <- thresh_noMPA$speed * 1000
 
 plotA <- ggplot(noThresh, aes(x=speed, y = harvest, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="gray95")  + labs(title="A") + xlab("") + ylab("") + theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm"), legend.position="none") 
 
@@ -422,5 +430,5 @@ legend <- g_legend(plotD)
 lwidth <- sum(legend$width)
 
 pdf(file="rockfish_sims.pdf",width=8,height=6)
-grid.arrange(arrangeGrob(plotA, plotB , plotC,plotD + theme(legend.position="none")), legend, left ="\nHarvest Fraction", sub="Climate velocity (km/decade)\n", widths=unit.c(unit(1, "npc") - lwidth,lwidth),ncol=2)
+grid.arrange(arrangeGrob(plotA, plotB , plotC,plotD + theme(legend.position="none")), legend, left ="\nHarvest fraction", sub="Climate velocity (km/decade)\n", widths=unit.c(unit(1, "npc") - lwidth,lwidth),ncol=2)
 dev.off()

### 2015-03-16 Writing/submitted/appendix/Appendix.tex
finished ecosphere formatting
changes to conform to ecosphere requirements

--- 2015-02-24 Writing/final_draft/Appendix.tex
@@
 
-\renewcommand{\thefigure}{S\arabic{figure}}
- \renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\thefigure}{A\arabic{figure}}
+ \renewcommand{\thetable}{A\arabic{table}}
+ \renewcommand{\thesubsection}{\Alph{subsection}}
 
-\appendix
-\section{Appendix}
+\section*{Appendix}
 ~\\In Appendix \ref{stab}, we provide the details for assessing the persistence of a population with an integrodifference model and we discuss the effect of the harvesting function on population persistence.
@@
 In Appendix \ref{MPA} we provide details on differences between small and large MPA simulations.
-In Appendix \ref{rock}, we parameterize our model for black rockfish ($Sebastes melanops$)in the California Current and demonstrate that results for  parameters  are qualitatively similar to results  presented in the main text.
+In Appendix \ref{rock}, we parameterize our model for black rockfish (\emph{Sebastes melanops})in the California Current and demonstrate that results for  parameters  are qualitatively similar to results  presented in the main text.
 
@@
 \begin{equation} n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0f(n_t(y))dy \label{integro} \end{equation}
-as described in \citet{ZhouKot2011}.  With the addition of harvesting, the model becomes
+as described in (Zhou and Kot 2011).  With the addition of harvesting, the model becomes
 
@@
 
-In evaluating persistence, we apply the methods of \citet{ZhouKot2011} to the new model, Equation 2.
+In evaluating persistence, we apply the methods of Zhou and Kot (2011) to the new model, Equation 2.
 A traveling pulse is a solution such that population size relative to location within the patch (rather than absolute position) is constant over time, i.e.
@@
 ~\\It is not immediately obvious that the operator $\psi$ will have any eigenfunctions.
-However, Jentzsch's theorem guarantees that there is an eigenfunction $u$, provided that the kernel $k$ satisfies some properties \citep{ZhouKot2011}.
+However, Jentzsch's theorem guarantees that there is an eigenfunction $u$, provided that the kernel $k$ satisfies some properties (Zhou and Kot 2011).
 Finding the eigenfunctions and eigenvalues is in general a hard problem to solve.
@@
 
-To find the equilibrium biomass, we rewrite (\ref{pulse}) using the separable kernel as in \cite{Latore:1998fk}:
+To find the equilibrium biomass, we rewrite (\ref{pulse}) using the separable kernel as in (Latore et al.1998):
 
@@
 This is a separable kernel with
-$a_n(x)=b_n(x)=\frac{1}{\sqrt{2n!\sqrt{D\pi}}}e^{-x^2/4D}\left(\frac{x}{\sqrt{2D}}\right)^n$ \citep{Latore:1998fk}.
+$a_n(x)=b_n(x)=\frac{1}{\sqrt{2n!\sqrt{D\pi}}}e^{-x^2/4D}\left(\frac{x}{\sqrt{2D}}\right)^n$ (Latore et al. 1998).
 
@@
 
-\citet{ZhouKot2011} solve for the critical speed, $c^*$, at which the population will be driven extinct:
+Zhou and Kot (2011) solve for the critical speed, $c^*$, at which the population will be driven extinct:
 $$c^*=c^*(R_0)=\frac{1}{w}\cos^{-1}\left[\frac{16+R_0^2(1-h)^2(w^2L^2-\sin^2(wL))}{8R_0(1-h)wL}\right].$$
@@
 \subsection{California Current Black Rockfish Parameterization \label{rock}}
-~\\We parameterize our model for black rockfish ($Sebastes melanops$) in the California Current, with MPAs of spacing and width qualitatively similar to those in the Marine Life Projection Act (MLPA), and with a maximum climate velocity equal to that observed empirically.
+~\\We parameterize our model for black rockfish (\emph{Sebastes melanops}) in the California Current, with MPAs of spacing and width qualitatively similar to those in the Marine Life Projection Act (MLPA), and with a maximum climate velocity equal to that observed empirically.
 The parameters and references are provided in Table \ref{rockfish_params}.
@@
 In particular, we find the same negative relationship between critical harvesting rate $h^*$ and the climate velocity $c$ (Figure \ref{crith_rockfish}) and an essentially additive interaction between the effects of the two stressors on biomass (Figure \ref{synergy_rockfish}).
-Additionally, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse than no MPA at all (compare Figure \ref{simulation_rockfish}a and \ref{simulation_rockfish}d).
+Additionally, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse than no MPA at all (compare Figure \ref{simulation_rockfish}A and \ref{simulation_rockfish}D).
 
@@
 \hline
-$R_0$ & 2.86 & \citet{white2010decision}, equivalent to $\left(\frac{1}{CRT}\right)$ \\
-$K$ & 1 & \cite{white2010decision} \\
+$R_0$ & 2.86 & , equivalent to $\left(\frac{1}{CRT}\right)$ \\
+$K$ & 1 & White et al. (2010) \\
 $h$ & 0-100\% &\\
-$c$ & 0-200km/decade & \cite{burrows2011pace}\\
-$L$ & 1000km  & \cite{fishbase} \\
-MPA width & 20km & \cite{Gainesetal2010}\\
-Space between MPAs & 76km & \cite{Gainesetal2010}\\
-Dispersal kernal &  Gaussian & \cite{white2010decision}\\
+$c$ & 0-200km/decade & Burrows et al. (2011)\\
+$L$ & 1000km  & Froese (2014) \\
+MPA width & 20km & Gaines et al. (2010b) \\
+Space between MPAs & 76km & Gaines et al. (2010b) \\
+Dispersal kernal &  Gaussian & White et al. (2010)\\
 \hline
@@
 \centering
-\includegraphics[width=.75\textwidth]{../../plots/compare_flucts.png}
+\includegraphics[width=.45\textwidth]{../../../plots/rockfish_criticalh2.png}
+\caption{ \label{crith_rockfish}
+The line indicate the critical harvest fraction as a function of climate velocity on the x-axis. Model parameterized for black rockfish. }
+ \end{figure}
+
+  \begin{figure}[h]
+ \centering
+ \includegraphics[width=.75\textwidth]{../../../plots/rockfisheqbiomass3.pdf}
+ \caption{ \label{synergy_rockfish}
+ (A) Results from a model parameterized for black rockfish.
+The equilibrium biomass of a black rockfish population as a function of the climate velocity on the x-axis and the harvest fraction on the y-axis.
+(B) Interaction between the two stressors as a function of climate velocity and harvest fraction.
+The heat map indicates the interaction measure $S$, i.e., the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually ($E_{hc}-E_h-E_c$).
+$S$ of $0$ indicates additive interaction of the stressors.
+The excess loss is small in comparison to the total biomass.
+ }
+ \end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.75\textwidth]{../../../plots/compare_flucts.png}
 \caption{\label{fluctuations} Total population biomass is on the y axis, and generation is on the x axis.
@@
 \centering
-\includegraphics[width=.75\textwidth]{../../plots/rockfish_criticalh.pdf}
-\caption{ \label{crith_rockfish}
-Lines indicate the critical threshold for persistence as a function of harvest fraction on the y-axis and climate velocity on the x-axis. Model parameterized for black rockfish. }
- \end{figure}
-
- \begin{figure}[h]
- \centering
- \includegraphics[width=.75\textwidth]{../../plots/rockfisheqbiomass.pdf}
- \caption{ \label{synergy_rockfish}
- (a) Results from a model parameterized for black rockfish.
-The equilibrium biomass of a black rockfish population as a function of the climate velocity on the x-axis and the harvest fraction on the y-axis.
-(b) Interaction between the two stressors as a function of climate velocity and harvest fraction.
-The heat map indicates the interaction measure $S$, i.e., the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually ($E_{hc}-E_h-E_c$).
-$S$ of $0$ indicates additive interaction of the stressors.
-The excess loss, on the order of $4$, is small in comparison to the total biomass, which can be as large as $160$.
- }
- \end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=.75\textwidth]{../../plots/rockfish_sims.pdf}
+\includegraphics[width=.75\textwidth]{../../../plots/rockfish_sims.pdf}
 \caption{\label{simulation_rockfish} The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvest fraction on the y-axis under alternative management strategies as parameterized for black rockfish.
-(a) The equilibrium biomass for simulations with constant harvest rates.
-(b) Equilibrium biomass for simulations with threshold management.
+(A) The equilibrium biomass for simulations with constant harvest rates.
+(B) Equilibrium biomass for simulations with threshold management.
 For threshold management, we set a threshold density below which no fishing is allowed.
@@
 We show this on the y-axis.
-(c) Equilibrium biomass for simulations with protected areas where harvesting pressure outside reserves is unchanged  (i.e., harvest effort inside reserves is eliminated).
-(d) Equilibrium biomass for simulations with protected areas in which harvesting pressure is reallocated outside reserves.}
+(C) Equilibrium biomass for simulations with protected areas where harvesting pressure outside reserves is unchanged  (i.e., harvest effort inside reserves is eliminated).
+(D) Equilibrium biomass for simulations with protected areas in which harvesting pressure is reallocated outside reserves.}
 \end{figure}
@@
 \bibstyle{cbe}
-\bibliography{fish}
 

### 2015-03-16 plots/Fish_Figures.R
small changes to font sizes, axis labels in plots

@@ -52,13 +52,19 @@ ebm <- data[[7]]
 		geom_line(aes(color=factor(R), linetype=factor(d)),size=1) + 
 		theme_tufte() + 
 		scale_color_manual(values=c("grey", "dark grey", "black")) + 
-		theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm")) + 
-		xlab("Climate Velocity") + ylab("Harvesting rate") + 
+		theme(text=element_text(family="Helvetica", size=10), plot.margin=unit(c(0,0,0,0),"cm")) + 
+		xlab("Climate velocity") + ylab("Harvested fraction") + 
+		theme(axis.line = element_line(colour = 'black')) +
+		scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
+		 expand_limits(y = c(0,1), x=c(0,1)) +
 		labs(linetype=expression(symbol("\341")*d*symbol("\361")), color=expression("R"[0]), title="")
 	
-	png("Fig1.png",height=3,width=4,units="in",res=300)
-	print(plot1)
-	dev.off() 
+# png("Fig1.png",height=3,width=4,units="in",res=300)
+	# print(plot1)
+	# dev.off() 
+	pdf(file="Fig1.pdf",width=6.1,height=3)
+print(plot1)
+dev.off()
 	
 ################################################################## FIGURE 2
 
@@ -72,10 +78,11 @@ ebm <- data[[7]]
 	plot2a <- ggplot(melted_ebm, aes(x=speed, y = harvest, fill=value)) + 
 		geom_tile() + theme_tufte() + 
 		scale_fill_gradient(low="black", high="white") + 
-		theme(text=element_text(family="Helvetica", size=14), plot.margin=unit(c(0,0,0,0),"cm")) + 
-		guides(fill = guide_colorbar(barwidth=1, barheight=10, title.position="top", title="Equilibrium\nBiomass")) + 
+		theme(text=element_text(family="Helvetica", size=10), plot.margin=unit(c(0,0,0,0),"cm"),plot.title=element_text(size=10)) + 
+		scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
+		guides(fill = guide_colorbar(barwidth=1, barheight=10, title.position="top", title="Equilibrium\nbiomass")) + 
 		xlab("") + 
-		ylab("Harvesting Rate") + 
[…]

### 2015-03-16 plots/rockfish_analytical.R
i accidentally removed some things from my git branch, adding them back on

@@ -1,0 +1,114 @@
+library(fields)
+library(lattice)
+library(colorspace)
+library(RCurl)
+source("biomass_synergy.r")
+source("critical_parameters.r")
+require(stringr)
+require(colorspace)
+require(reshape2)
+require(gridExtra)
+require(ggplot2)
+require(ggthemes)
+
+d=73
+sig2=pi/2*d^2
+R=2.86
+K=1
+L=1000
+c=1
+
+lambdag(10,R,sig2,L)
+# cstar=cstar_g(R,sig2,L)
+cstar=154
+
+cvals=seq(0,cstar,length.out=26)
+Nc=length(cvals)
+hvals=seq(0,1,length.out=26)
+crit_hvals=array(,Nc)
+
+for(i in 1:Nc){
+	c=cvals[i]
+		hstar=hstar_g(c,R,sig2,L)
+		crit_hvals[i]=hstar
+}
+	
+htoplot=data.frame(h=crit_hvals,speed=cvals)
+
+	
+# plot(cvals,hvals)	
+
+rockfishplot1 <- ggplot(htoplot, aes(x=speed, y = h)) + 
+		geom_line(size=1) + 
+		theme_tufte() + 
+		# scale_color_manual(values=c("grey", "dark grey", "black")) + 
+		theme(text=element_text(family="Helvetica", size=10)) + 	
+		theme(axis.line = element_line(colour = 'black')) +
+		scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
+		 expand_limits(y = c(0,1), x=c(0,L)) +
+		xlab("Climate velocity (km/year)") + ylab("Harvested fraction") 
+
+# png("rockfish_criticalh.png",height=3,width=4,units="in",res=300)	
+pdf(file="rockfish_criticalh.pdf",width=2.9,height=3)
+print(rockfishplot1)
+dev.off()
+# png("rockfish_criticalh.png",height=3, width=4,units='in',res=300)
+# print(rockfishplot1)
+# dev.off()
+	
+# rockfish_analytical=eqbiomass(cvals,hvals,R,K,sig2,L,tol=NA)
+rockfish_analytical=read.csv('rockfish_analytical_ebm.csv',sep='\t')
+# making data
+	ebm=rockfish_analytical
+	ebm=as.matrix(ebm,nrow=length(hvals))
+	melted_ebm <- melt(ebm)
+	num=dim(ebm)[1]
[…]

### 2015-03-17 Writing/submitted/appendix/Appendix.tex
continued to tweak figures and appendix
more small changes to figures and putting them into the ms and appendix

@@
 \usepackage{subcaption}
-\usepackage{brush_style}
 \usepackage{setspace}
@@
 
-If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\R$ and $u:[-L/2,L/2]\to\R$, then the perturbation grows in time if and only if $\lambda >1$.
+If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\mathbb{R}$ and $u:[-L/2,L/2]\to\mathbb{R}$, then the perturbation grows in time if and only if $\lambda >1$.
 Using Equation (\ref{perturb}), we can rewrite $\xi_{t+1}(x)$,
@@
 \centering
-\includegraphics[width=.45\textwidth]{../../../plots/rockfish_criticalh2.png}
+\includegraphics[width=2.9in]{../figures/rockfish_criticalh.pdf}
 \caption{ \label{crith_rockfish}
@@
  \centering
- \includegraphics[width=.75\textwidth]{../../../plots/rockfisheqbiomass3.pdf}
+ \includegraphics[width=6.1in]{../figures/rockfisheqbiomass.pdf}
  \caption{ \label{synergy_rockfish}
@@
 \centering
-\includegraphics[width=.75\textwidth]{../../../plots/compare_flucts.png}
+\includegraphics[width=6.1in]{../figures/compare_flucts.png}
 \caption{\label{fluctuations} Total population biomass is on the y axis, and generation is on the x axis.
@@
 \centering
-\includegraphics[width=.75\textwidth]{../../../plots/rockfish_sims.pdf}
+\includegraphics[width=6.1in]{../figures/rockfish_sims.pdf}
 \caption{\label{simulation_rockfish} The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvest fraction on the y-axis under alternative management strategies as parameterized for black rockfish.

### 2015-03-17 plots/Fish_Figures.R

@@ -62,7 +62,7 @@ ebm <- data[[7]]
 # png("Fig1.png",height=3,width=4,units="in",res=300)
 	# print(plot1)
 	# dev.off() 
-	pdf(file="Fig1.pdf",width=6.1,height=3)
+	pdf(file="Fig1.pdf",width=2.9,height=2)
 print(plot1)
 dev.off()
 	
@@ -124,8 +124,8 @@ plot2b <- ggplot(melt_syn, aes(x=Speed, y = Harvest, fill=Synergy)) +
 #### Together
 # png("Fig2.png",height=3, width=8,units="in",res=300)
 # grid.arrange(plot2a, plot2b, ncol=2, sub = textGrob("Climate velocity", gp=gpar(fontsize=10),just="bottom"))
-dev.off() 
-pdf(file="Fig2.pdf",width=6.1,height=3)
+# dev.off() 
+pdf(file="Fig2.pdf",width=6.1,height=2.6)
 grid.arrange(plot2a, plot2b, ncol=2, sub=textGrob("Climate velocity",gp=gpar(fontsize=10),just='bottom'))
 dev.off() 
 ################################################################### Figure 3
@@ -451,5 +451,5 @@ legend <- g_legend(plotD)
 lwidth <- sum(legend$width)
 
 pdf(file="rockfish_sims.pdf",width=8,height=6)
-grid.arrange(arrangeGrob(plotA, plotB , plotC,plotD + theme(legend.position="none")), legend,  sub=textGrob("Climate velocity (km/decade)\n",gp=gpar(size=10)), widths=unit.c(unit(1, "npc") - lwidth,lwidth),ncol=2)
+grid.arrange(arrangeGrob(plotA, plotB , plotC,plotD + theme(legend.position="none")), legend,  sub=textGrob("Climate velocity (km/year)\n",gp=gpar(fontsize=10)), widths=unit.c(unit(1, "npc") - lwidth,lwidth),ncol=2)
 dev.off()

### 2015-03-17 plots/plotting_fluctuations.R

@@ -6,7 +6,8 @@ pfNA <- read.csv("popfishnoThreshNA.csv")
 pfEf <- read.csv("popfishnoThreshyes.csv")
 pn <- read.csv("popnullnoThreshNA.csv")
 
-png(file="compare_flucts.png", height=6,width=6,units="in",res=300)
+# png(file="compare_flucts.png", height=6,width=6,units="in",res=300)
+pdf(file='compare_flucts.pdf',height=6,width=6.1)
 plot(pcNA[1:200,], type='l',ylim=c(1000,2200),lwd=2,bty="n",xlab="Generation", ylab="Total population size")
 lines(pcEf,type='l',col="black",lwd=2,lty=9)
 

### 2015-03-17 plots/rockfish_analytical.R

@@ -49,7 +49,7 @@ rockfishplot1 <- ggplot(htoplot, aes(x=speed, y = h)) +
 		xlab("Climate velocity (km/year)") + ylab("Harvested fraction") 
 
 # png("rockfish_criticalh.png",height=3,width=4,units="in",res=300)	
-pdf(file="rockfish_criticalh.pdf",width=2.9,height=3)
+pdf(file="rockfish_criticalh.pdf",width=2.9,height=2)
 print(rockfishplot1)
 dev.off()
 # png("rockfish_criticalh.png",height=3, width=4,units='in',res=300)
@@ -106,8 +106,8 @@ rockfishsynplot <- ggplot(melt_syn, aes(x=Speed, y = Harvest, fill=Synergy)) +
 	theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank()) + labs(title="B")
 
 # #### Together
-pdf("rockfisheqbiomass.pdf",height=3, width=6.1)
-grid.arrange(rockfishebmplot, rockfishsynplot, ncol=2, sub = textGrob("Climate velocity (km/year)", just="bottom"))
+pdf("rockfisheqbiomass.pdf",height=2.6, width=6.1)
+grid.arrange(rockfishebmplot, rockfishsynplot, ncol=2, sub = textGrob("Climate velocity (km/year)", just="bottom",gp=gpar(fontsize=10)))
 dev.off() 
 # png("rockfisheqbiomass.png",height=3, width=8,units='in',res=300)
 # grid.arrange(rockfishebmplot, rockfishsynplot, ncol=2, sub = textGrob("Climate velocity  (km/year)", just="bottom"))

### 2015-03-17 Writing/submitted/appendix/Appendix.tex
hopefully final changes to figures, moved figures and table from appendix to manuscript and removed them from appendix

@@
 
-\renewcommand{\thefigure}{A\arabic{figure}}
- \renewcommand{\thetable}{A\arabic{table}}
  \renewcommand{\thesubsection}{\Alph{subsection}}
@@
 h^*&\sim 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\sqrt{\pi}\left[\frac{L-c}{2\sqrt{2D}}-\frac{(L-c)^3}{3(2\sqrt{2D})^3}-\frac{-L-c}{2\sqrt{2D}}+\frac{(-L-c)^3}{3(2\sqrt{2D})^3)}\right]}
-%\\&= 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\left[\frac{2L}{2\sqrt{2D}}+\frac{-L^3+3L^2c-3Lc^2+c^3-L^3-3L^2c-3Lc^2-c^3}{3(2\sqrt{2D})^3}\right]}
-%\\&= 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\left[\frac{L}{\sqrt{2D}}-2\frac{L^3+3Lc^2}{3(2\sqrt{2D})^3}\right]}
-%\\&= 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0L\left[\frac{48D-2(L^2+3c^2)}{3(2\sqrt{2D})^3}\right]}
-%\\&= 1-\frac{3*32D\sqrt{D\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0L\left(48D-2(L^2+3c^2)\right)}
-%\\&= 1-\frac{3*32D\sqrt{D\pi}(\frac{128D^2+16c^2D+c^4}{128D^2})}{R_0L\left(48D-2(L^2+3c^2)\right)}
-%\\&= 1-\frac{3\sqrt{\pi}(128D^2+16c^2D+c^4)}{4R_0L\sqrt{D}\left(48D-2(L^2+3c^2)\right)}
-%\\&= 1-\frac{3\sqrt{2\pi}(32\sigma^4+8c^2\sigma^2+c^4)}{4R_0L\sigma\left(24\sigma^2-2(L^2+3c^2)\right)}
 \\&= 1-\frac{1}{R_0}\cdot\frac{3\sqrt{2\pi}}{8L}\frac{(32\sigma^4+8c^2\sigma^2+c^4)}{\sigma\left(12\sigma^2-(L^2+3c^2)\right)}
@@
 h^*&\sim 1-\frac{1}{R_0}\cdot\frac{12wL}{w^4L^4}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{12}{w^3L^3}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{12\sigma^3}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[1-\frac{(\frac{\pi^2}{4}-2)c^2}{2\sigma^2}-\sqrt{1-\frac{(\frac{\pi^2}{4}-2)c^2}{\sigma^2}-\frac{(\frac{\pi^2}{4}-2)L^2}{3\sigma^2}}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^3(\frac{\pi^2}{4}-2)^{3/2}}\left[2\sigma^2-(\frac{\pi^2}{4}-2)c^2-\frac{2}{\sqrt{3}}\sigma\sqrt{3\sigma^2-(\frac{\pi^2}{4}-2)3c^2-(\frac{\pi^2}{4}-2)L^2}\right]
-%\\&=1-\frac{1}{R_0}\cdot\frac{6\sigma}{L^34(\frac{\pi^2}{4}-2)^{3/2}}\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
-%\\&=1-\frac{12}{R_0L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sigma^2-(\pi^2-8)c^2-\frac{4}{\sqrt{3}}\sigma\sqrt{12\sigma^2-(\pi^2-8)3c^2-(\pi^2-8)L^2}\right]
 \\&=1-\frac{1}{R_0}\cdot\frac{4\sqrt{3}}{L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sqrt{3}\sigma^2-(\pi^2-8)\sqrt{3}c^2-4\sigma\sqrt{12\sigma^2-(\pi^2-8)(3c^2+L^2)}\right]
@@
 %\pagebreak
+\subsection{California Current Black Rockfish Parameterization \label{rock}}
+~\\We parameterize our model for black rockfish (\emph{Sebastes melanops}) in the California Current, with MPAs of spacing and width qualitatively similar to those in the Marine Life Projection Act (MLPA), and with a maximum climate velocity equal to that observed empirically.
+The parameters and references are provided in Table A1.
+
+Our results with this parameterization are qualitatively similar to the results presented in the main text.
+In particular, we find the same negative relationship between critical harvesting rate $h^*$ and the climate velocity $c$ (Figure A1) and an essentially additive interaction between the effects of the two stressors on biomass (Figure A2).
+Additionally, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse than no MPA at all (compare Figure A4A and A4D).
+
+%\pagebreak
 \subsection{Protected Area fluctuations \label{MPA}}
-~\\After the simulations come to equilibrium, the fluctuations in total biomass per generation fluctuate more in reserves that are larger and spaced farther apart than simulations in which the reserves that are smaller and more closely spaced (Figure \ref{fluctuations}).
+~\\After the simulations come to equilibrium, the fluctuations in total biomass per generation fluctuate more in reserves that are larger and spaced farther apart than simulations in which the reserves that are smaller and more closely spaced (Figure A3).
 The large reserves have a slightly larger average population, however large reserves here can induce fluctuations of biomass even in deterministic simulations.
@@
 
-%\pagebreak
-\subsection{California Current Black Rockfish Parameterization \label{rock}}
-~\\We parameterize our model for black rockfish (\emph{Sebastes melanops}) in the California Current, with MPAs of spacing and width qualitatively similar to those in the Marine Life Projection Act (MLPA), and with a maximum climate velocity equal to that observed empirically.
-The parameters and references are provided in Table \ref{rockfish_params}.
-
-Our results with this parameterization are qualitatively similar to the results presented in the main text.
-In particular, we find the same negative relationship between critical harvesting rate $h^*$ and the climate velocity $c$ (Figure \ref{crith_rockfish}) and an essentially additive interaction between the effects of the two stressors on biomass (Figure \ref{synergy_rockfish}).
-Additionally, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse than no MPA at all (compare Figure \ref{simulation_rockfish}A and \ref{simulation_rockfish}D).
-
-\begin{table}[htdp]
-\caption{\label{rockfish_params} Rockfish simulation parameters}
-\begin{center}
-\begin{tabular}{c l l}
-\hline
-Parameter & Value & Source\\
-\hline
-$R_0$ & 2.86 & , equivalent to $\left(\frac{1}{CRT}\right)$ \\
-$K$ & 1 & White et al. (2010) \\
-$h$ & 0-100\% &\\
-$c$ & 0-200km/decade & Burrows et al. (2011)\\
-$L$ & 1000km  & Froese (2014) \\
-MPA width & 20km & Gaines et al. (2010b) \\
-Space between MPAs & 76km & Gaines et al. (2010b) \\
-Dispersal kernal &  Gaussian & White et al. (2010)\\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=2.9in]{../figures/rockfish_criticalh.pdf}
-\caption{ \label{crith_rockfish}
-The line indicate the critical harvest fraction as a function of climate velocity on the x-axis. Model parameterized for black rockfish. }
- \end{figure}
-
-  \begin{figure}[h]
- \centering
- \includegraphics[width=6.1in]{../figures/rockfisheqbiomass.pdf}
- \caption{ \label{synergy_rockfish}
- (A) Results from a model parameterized for black rockfish.
-The equilibrium biomass of a black rockfish population as a function of the climate velocity on the x-axis and the harvest fraction on the y-axis.
-(B) Interaction between the two stressors as a function of climate velocity and harvest fraction.
-The heat map indicates the interaction measure $S$, i.e., the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually ($E_{hc}-E_h-E_c$).
-$S$ of $0$ indicates additive interaction of the stressors.
-The excess loss is small in comparison to the total biomass.
- }
- \end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=6.1in]{../figures/compare_flucts.png}
-\caption{\label{fluctuations} Total population biomass is on the y axis, and generation is on the x axis.
-These simulations were run with climate velocity $ = 0.1$, and a proportional harvest rate $= 0.02$.
-}
-% see bounded\_flux.R for code that made this plot
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=6.1in]{../figures/rockfish_sims.pdf}
-\caption{\label{simulation_rockfish} The equilibrium biomass of the population as a function of the climate velocity on the x-axis and the harvest fraction on the y-axis under alternative management strategies as parameterized for black rockfish.
-(A) The equilibrium biomass for simulations with constant harvest rates.
-(B) Equilibrium biomass for simulations with threshold management.
-For threshold management, we set a threshold density below which no fishing is allowed.
-The threshold ranges between 0 (no fishing allowed) and 1 (all fish taken), with intermediate density thresholds determined as fractions of the maximum population density observed at a given time step before harvesting.
-We show this on the y-axis.
-(C) Equilibrium biomass for simulations with protected areas where harvesting pressure outside reserves is unchanged  (i.e., harvest effort inside reserves is eliminated).
-(D) Equilibrium biomass for simulations with protected areas in which harvesting pressure is reallocated outside reserves.}
-\end{figure}
-
-\cleardoublepage
-
-\bibstyle{cbe}
-
 \end{document}

### 2015-03-17 plots/Fish_Figures.R

@@ -53,7 +53,7 @@ ebm <- data[[7]]
 		theme_tufte() + 
 		scale_color_manual(values=c("grey", "dark grey", "black")) + 
 		theme(text=element_text(family="Helvetica", size=10), plot.margin=unit(c(0,0,0,0),"cm")) + 
-		xlab("Climate velocity") + ylab("Harvested fraction") + 
+		xlab("Climate velocity") + ylab("Harvesting rate") + 
 		theme(axis.line = element_line(colour = 'black')) +
 		scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
 		 expand_limits(y = c(0,1), x=c(0,1)) +
@@ -82,7 +82,7 @@ dev.off()
 		scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
 		guides(fill = guide_colorbar(barwidth=1, barheight=10, title.position="top", title="Equilibrium\nbiomass")) + 
 		xlab("") + 
-		ylab("Harvested fraction") + 
+		ylab("Harvesting rate") + 
 		labs(title="A")
 
 #### B
@@ -145,7 +145,7 @@ dev.off()
 	consTrue$management = "Few Large Reserves- Constant Effort"
 	fishTrue$managemetn = "Many Small Reserves - Constant Effort"
 	
-	plotA <- ggplot(sim, aes(x=speed, y = harvest, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="gray95")  + labs(title="A") + xlab("") + ylab("Harvested fraction") + theme(text=element_text(family="Helvetica", size=10),axis.title=element_text(size=10),plot.title=element_text(size=10),plot.margin=unit(c(0,0,0,0),"cm"), legend.position="none") +
+	plotA <- ggplot(sim, aes(x=speed, y = harvest, fill = Equil.pop)) + geom_raster(interpolate=TRUE) + theme_tufte() + scale_fill_gradient(low="black", high="gray95")  + labs(title="A") + xlab("") + ylab("Harvesting rate") + theme(text=element_text(family="Helvetica", size=10),axis.title=element_text(size=10),plot.title=element_text(size=10),plot.margin=unit(c(0,0,0,0),"cm"), legend.position="none") +
[…]

### 2015-03-17 plots/critical_parameters.R

@@ -77,6 +77,15 @@ cstar_g<-function(r,sig2,L=1){
 	return(x)
 }
 
+
+cstar_g_big<-function(r,sig2,L=1){
+	if(lambdag(0,r,sig2,L)<1){ #if eigenvalue is less than 1 for c=0, c=0 is the critical speed
+		x=0
+	} else{ #if eigenvalue is greater than 1 for c=0, look for c where eigenvalue = 1
+	x=uniroot(fg,interval=c(0,200),r=r,sig2=sig2,L=L)$root}
+	return(x)
+}
+
 #### analytical expression for critical harvesting level for sinusoidal kernel
 hstar_s<-function(c,r,sig2,L=1){
 	if(is.na(L)){L=1}
@@ -93,6 +102,13 @@ hstar_s<-function(c,r,sig2,L=1){
 ###### numerically find critical harvesting proportion for gaussian kernel
 hstar_g<-function(c,r,sig2,L=1){
 	y=cstar_g(r,sig2,L)
+	if(y<=c){h=0}
+	else{h=1-1/lambdag(c,r,sig2,L)}
+	return(h)
+}
+
+hstar_g_big<-function(c,r,sig2,L=1){
+	y=cstar_g_big(r,sig2,L)
 	if(y<=c){h=0}
 	else{h=1-1/lambdag(c,r,sig2,L)}
 	return(h)

### 2015-03-17 plots/rockfish_analytical.R

@@ -29,11 +29,12 @@ crit_hvals=array(,Nc)
 
 for(i in 1:Nc){
 	c=cvals[i]
-		hstar=hstar_g(c,R,sig2,L)
+		hstar=hstar_g_big(c,R,sig2,L)
 		crit_hvals[i]=hstar
 }
-	
-htoplot=data.frame(h=crit_hvals,speed=cvals)
+
+gen_time=7	
+htoplot=data.frame(h=crit_hvals,speed=cvals/gen_time*10)
 
 	
 # plot(cvals,hvals)	
@@ -45,11 +46,11 @@ rockfishplot1 <- ggplot(htoplot, aes(x=speed, y = h)) +
 		theme(text=element_text(family="Helvetica", size=10)) + 	
 		theme(axis.line = element_line(colour = 'black')) +
 		scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
-		 expand_limits(y = c(0,1), x=c(0,L)) +
-		xlab("Climate velocity (km/year)") + ylab("Harvested fraction") 
+		 expand_limits(y = c(0,1), x=c(0,300)) +
+		xlab("Climate velocity (km/decade)") + ylab("Harvesting rate") 
 
 # png("rockfish_criticalh.png",height=3,width=4,units="in",res=300)	
-pdf(file="rockfish_criticalh.pdf",width=2.9,height=2)
+pdf(file="FigA1.pdf",width=2.9,height=2)
 print(rockfishplot1)
 dev.off()
 # png("rockfish_criticalh.png",height=3, width=4,units='in',res=300)
@@ -65,7 +66,7 @@ rockfish_analytical=read.csv('rockfish_analytical_ebm.csv',sep='\t')
 	num=dim(ebm)[1]
 	speeds <- rep(cvals, num)
 	harvests <- rep(hvals, each = num)
-	melted_ebm$Speed = speeds
+	melted_ebm$Speed = speeds/gen_time*10
 	melted_ebm$Harvest = harvests
 
 	syn=synergy(ebm,2)
@@ -75,7 +76,7 @@ rockfish_analytical=read.csv('rockfish_analytical_ebm.csv',sep='\t')
 	
 	melt_syn <- melt(syn)
 	names(melt_syn) <- c("Speed","Harvest","Synergy")	
-	speeds <- rep(cvals[2:num], num-1)
+	speeds <- rep(cvals[2:num], num-1)/gen_time*10
 	harvests <- rep(hvals[2:num], each = num-1)
[…]

### 2015-04-04 Writing/submitted/appendix/component_parts/Appendix.Rmd
making formatting changes required by EcoSphere

---
title: "Appendix"
author: ' '
date: ' '
output: word_document
---

In Appendix A, we provide the details for assessing the persistence
of a population with an integrodifference model and we discuss the
effect of the harvesting function on population persistence. In Appendix B, we provide the details for assessing population persistence with separable dispersal kernels. In Appendix C and D, we derive expressions for the critical harvesting rate and rate of environmental shift for Gaussian and sinuosoidal dispersal kernels. In Appendix E, we derive approximate expressions for these critical rates. In Appendix F, we parameterize our model for black rockfish (*Sebastes melanops*)in the California Current and demonstrate that results for parameters are qualitatively similar to results presented in the main text. In Appendix G we provide details on differences between small and large MPA simulations.

A) Determining stability
----------------------------
Let $n_t(x)$ be the number of adults at position $x$ at time $t$, let $k(x)$ be a dispersal kernel describingt the probability of a larva traveling a distance $x$, let $f(n)$ be the recruitment function describing the number of offspring that settle and survive in juvenile population of size $n$, let $R_0$ be the intrinsic growth rate of the population, and let $g(n)$ be the harvesting function describing the number of adults harvested from a population of size $n$. In the absence of harvesting, the integrodifference model describing the population over time is given by

$$n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0f(n_t(y))dy \label{integro}$$

as described in (Zhou and Kot 2011). With the addition of harvesting, the model becomes

$$n_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy.$$

In evaluating persistence, we apply the methods of Zhou and Kot (2011) to the new model, Equation 2. A traveling pulse is a solution such that population size relative to location within the patch (rather than absolute position) is constant over time, i.e.

$$n^*(\overline{x}_t)\equiv n^*(x-ct)=n_t(x)$$

where $\overline{x}_t\equiv x-ct$ gives position relative to the patch.

The integrodifference Eq. 2 gives us an expression for $n^*$:

$$n^*(\overline{x}_{t+1})=n_{t+1}(x)$$

$$n^*(\overline{x}_{t+1})=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy$$

$$n^*(\overline{x}_{t+1})=\int_{-L/2+ct}^{L/2+ct}k(x-\overline{y}_t-ct)R_0g(f(n^*(\overline{y}_t)))dy$$

$$n^*(\overline{x}_{t+1})=\int_{-L/2+ct}^{L/2+ct}k(\overline{x}_t-\overline{y}_t)R_0g(f(n^*(\overline{y}_t)))dy$$

$$\Rightarrow n^*(\overline{x}_t-c)=\int_{-L/2+ct}^{L/2+ct}k(\overline{x}_t-\overline{y}_t)R_0g(f(n^*(\overline{y}_t)))dy$$

$$\Rightarrow n^*(\overline{x}_t)=\int_{-L/2}^{L/2}k(\overline{x}_t+c-\overline{y}_t)R_0g(f(n^*(\overline{y}_t)))d\overline y_t$$

As long as $f(0)=0$, there is a trivial solution to this problem where $n^*(\overline{x})\equiv 0$ for all $\overline{x}$, i.e., there is a trivial traveling pulse with no adults in it. If the trivial traveling pulse is unstable, even very small populations will persist or grow and avoid crashing back to the trivial pulse. To evaluate the stability of a traveling pulse, we introduce a small perturbation to the traveling pulse $n^*(\overline{x})$ and see if this perturbation grows or shrinks over time:

$$n_t(x)=n^*(\overline{x}_t)+\xi_t(x)$$

$$\Rightarrow \xi_{t+1}(x)=n_{t+1}(x)-n^*(\overline{x}_{t+1})$$

$$ \xi_{t+1}(x)=n_{t+1}(x)-n^*(\overline{x}_{t}-c)$$

$$\xi_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy-\int_{-L/2}^{L/2}k(\overline{x}_{t}-\overline{y}_t)R_0g(f(n^*(\overline{y}_t)))d\overline y_t \text{ using Eq. 3}$$

$$\xi_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy-\int_{-L/2+ct}^{L/2+ct}k(x-ct-(y-ct))R_0g(f(n^*(\overline{y}_t)))d\overline y_t$$

$$\xi_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n_t(y)))dy-\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g(f(n^*(\overline{y}_t)))d\overline y_t$$

$$\xi_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)\big(R_0g(f(n_t(y)))-R_0g(f(n^*(\overline{y}_t)))\big)dy$$

$$\xi_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0\big(g(f(n_t(y)))-g(f(n^*(\overline{y}_t)))\big)dy$$

$$\Rightarrow \xi_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g\prime (f(n^*(\overline{y})))f\prime (n^*(\overline{y}))(n_t(y)-n^*(\overline{y}_t))dy$$

by linearizing around the traveling pulse

$$\Rightarrow \xi_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g\prime (f(n^*(\overline{y})))f\prime(n^*(\overline{y}))\xi_t(y)dy$$

$$\Rightarrow \xi_{t+1}(x)=\int_{-L/2+ct}^{L/2+ct}k(x-y)R_0g\prime(0)f\prime(0)\xi_t(y)dy$$

if $n^*(\overline{x})=0$ and $f(0)=0$.

If we assume $\xi_t(x)=\lambda^tu(x-ct)$ for some $\lambda\in\mathbb{R}$ and $u:[-L/2,L/2]\to\mathbb{R}$, then the perturbation grows in time if and only if $\lambda >1$. Using Eq. 4, we can rewrite $\xi_{t+1}(x)$,

$$\lambda u(x-ct-c)=R_0g\prime(0)f\prime(0)\int_{-L/2+ct}^{L/2+ct}k(x-y)u(y-ct)dy$$

$$\Rightarrow \lambda u(\overline{x})=R_0g\prime (0)f\prime (0)\int_{-L/2}^{L/2}k(\overline{x}+c-\overline{y})u(\overline{y})dy$$

Define the integral operator

$$\psi_f(u)(x)=R_0g\prime (0)f\prime (0)\int_{-L/2}^{L/2}k(x+c-y)u(y)dy.$$

Then the perturbation to the traveling pulse will satisfy

$$\psi_f(u)(x)=\lambda u(x)$$

$\lambda$ and $u$ are thus an eigenvalue and eigenfunction of the functional operator $\psi_f$. The trivial traveling pulse is unstable when the dominant eigenvalue of $\psi_f$ is greater than $1$.

The biomass in the equilibrium traveling wave depends on the specific functional forms of the harvesting function $g(n)$ and the recruitment function $f(n)$. However, the persistence of the population only depends on $R_0$, $g\prime (0)$ and $f\prime (0)$. In this paper, we only considered a proportional harvesting function, i.e. the amount of adults harvested obeyed $g(n)=(1-h)n$. For this function, $g\prime (0)=1-h$. For the recruitment function we considered, $f\prime(0)=1$.

B) Separable dispersal kernels
---------------------------------
It is not immediately obvious that the operator $\psi$ will have any eigenfunctions. However, Jentzsch’s theorem guarantees that there is an eigenfunction $u$, provided that the kernel $k$ satisfies some properties (Zhou and Kot 2011). Finding the eigenfunctions and eigenvalues is in general a hard problem to solve. It becomes easier if the kernel $k$ is separable, i.e., there are functions $a_n,b_n$ such that $k(x-y)=\sum_{n=1}^\infty a_n(x)b_n(y)$. In that case, Eq. 5 becomes

$$\lambda u(x)=R_0g\prime (0)f\prime (0)\sum_{n=1}^\infty\left( a_n(x)\int_{-L/2}^{L/2}b_n(y-c)u(y)dy\right)$$

$$\Rightarrow \lambda\int_{-L/2}^{L/2}b_k(x-c)u(x)dx=R_0g\prime (0)f\prime (0)\sum_{n=1}^{\infty}\left(\int_{-L/2}^{L/2}b_n(x-c)u(x)dx\right)\left(\int_{-L/2}^{L/2}a_n(y)b_k(y-c)dy\right)$$

for any $k$

$$\Rightarrow \lambda d_k=R_0g\prime (0)f\prime (0)\sum_{n=1}^\infty A_{nk}d_n$$

where

$$A_{nk}=\int_{-L/2}^{L/2}a_n(x)b_k(x-c)dx \text{ and } d_k=\int_{-L/2}^{L/2}b_k(x-c)u(x)dx$$

Finding the eigenvalues of Eq. 5 then reduces to finding the eigenvalues of the matrix comprised of entires $(A_{nk})_{n,k=1}^\infty$.

To find the equilibrium biomass, we rewrite Eq. 3 using the separable kernel as in (Latore et al.1998):

$$n^*(x)=\int_{-L/2}^{L/2}k(x+c-y)R_0g(f(n^*(y)))dy$$

$$n^*(x)= \int_{-L/2}^{L/2}\left(\sum_{n=1}^\infty a_n(x)b_n(y-c)\right)R_0g(f(n^*(y)))dy$$

$$n^*(x)=\sum_{n=1}^\infty a_n(x) \int_{-L/2}^{L/2}b_n(y-c)R_0g(f(n^*(y)))dy$$

If we define $m_n=\int_{-L/2}^{L/2}b_n(y-c)R_0g(f(n^*(y)))dy$ then we find that

$$n^*(x)=\sum_{n=1}^\infty m_na_n(x) \text { and }$$

$$m_n=\int_{-L/2}^{L/2}b_n(y-c)R_0g\left(f\left(\sum_{n=1}^\infty m_na_n(y)\right)\right)dy$$

Eq 7. allows us to find the $m_n$ numerically and we then find the total equilibrium biomass by integrating $n^*(x)$ over space.

C) Gaussian dispersal kernel
-----------------------------------
The Gaussian dispersal kernel is given by

$$k(x-y)=\frac{1}{2\sqrt{D\pi}}e^{\frac{-(x-y)^2}{4D}},$$

where $D$ is one half the variance of the kernel. This is a separable kernel with $a_n(x)=b_n(x)=\frac{1}{\sqrt{2n!\sqrt{D\pi}}}e^{-x^2/4D}\left(\frac{x}{\sqrt{2D}}\right)^n$ (Latore et al. 1998).

As a first approximation to $k$ we ignore all but the $0^{th}$ terms for $a_n$ and $b_n$ so that Eq. 6 becomes

$$\lambda d_0(c)&=R_0(1-h)A_{00}(c)d_0(c)$$

$$\Rightarrow \lambda=R_0(1-h)A_{00}(c)$$

$$\text{ where } A_{00}(c)=2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]$$

where $\text{erf}$ is the error function. The critical rate of environmental shift $c^*$ and the critical harvesting rate $h^*$ are those values of $c$ and $h$, respectively, that make $\lambda=1$.

D) Sinusoidal dispersal kernel
------------------------------------
The sinusoidal dispersal kernel is given by

$$k(x-y)=\left\{\begin{array}{ccccc}
\frac{w}{2}\cos(w(x-y)) & , & |x-y|\leq\frac{\pi}{2w}
\\ 0 & , & |x-y|>\frac{\pi}{2w}
\end{array}\right.$$

where $L$ is the length of the patch and we assume $\frac{\pi}{2w}>L,c<\frac{\pi}{2w}-L$.

In this case, $k(x-y)=\frac{w}{2}\cos(wx)\cos(w(y-c))+\frac{w}{2}\sin(wx)\sin(w(y-c))$ so that $A_{ij}$ and $d_i$ can be found for $i,j=1,2$ and Eq. 6 reduces to

$$\lambda^2-\left(\frac{R_0(1-h)wL}{2}\cos(wc)\right)\lambda+\frac{R_0^2(1-h)^2}{16}\left(w^2L^2-\sin^2(wL)\right)=0.$$

If we solve for $\lambda$,we find

$$\lambda=(1-h)R_0\left[\frac{wL\cos(wc)}{4}+\frac{1}{4}\sqrt{\sin^2(wL)-w^2L^2\sin^2(wc)}\right].$$

Zhou and Kot (2011) solve for the critical speed, $c^*$, at which the
population will be driven extinct:

$$c^*=c^*(R_0)=\frac{1}{w}\cos^{-1}\left[\frac{16+R_0^2(1-h)^2(w^2L^2-\sin^2(wL))}{8R_0(1-h)wL}\right].$$

In our model, we can additionally solve for the critical harvesting
rate, $h^*$, at which the population will be driven extinct:

$$h^*=1-\frac{1}{R_0}\cdot\frac{4wL}{w^2L^2-\sin^2(wL)}\left[\cos(wc)-\sqrt{\cos^2(wc)-1+\frac{\sin^2(wL)}{w^2L^2}}\right]$$

E) Approximate critical harvesting proportions
--------------------------------------------------------
We will use the following Taylor series to make approximations of the critical harvesting proportions under the two dispersal kernels:

$$\cos(x)=1-\frac{x^2}{2}$$

$$\cos^2(x)=1-x^2$$

$$\sin^2(x)=x^2-\frac{x^4}{3}$$

$$erf(x)=\frac{2}{\sqrt{\pi}}(x-\frac{x^3}{3})$$

$$\exp(x)=1+x+\frac{x^2}{2}$$

For the Gaussian kernel we found

$$h^*=1-\frac{2\sqrt{2}\exp\left(\frac{c^{2}}{8D}\right)}{R_0\left[erf\left(\frac{L-c}{2\sqrt{2D}}\right)-erf\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]}$$

Using the Taylor series and the fact that $D=\frac{\sigma^2}{2}$ where $\sigma^2$ is the variance of the exponential kernel,

$$h^*\sim 1-\frac{\sqrt{2\pi}(1+\frac{c^2}{8D}+\frac{c^4}{128D^2})}{R_0\sqrt{\pi}\left[\frac{L-c}{2\sqrt{2D}}-\frac{(L-c)^3}{3(2\sqrt{2D})^3}-\frac{-L-c}{2\sqrt{2D}}+\frac{(-L-c)^3}{3(2\sqrt{2D})^3)}\right]}$$

$$h^*= 1-\frac{1}{R_0}\cdot\frac{3\sqrt{2\pi}}{8L}\frac{(32\sigma^4+8c^2\sigma^2+c^4)}{\sigma\left(12\sigma^2-(L^2+3c^2)\right)}$$

For the sinusoidal kernel we found

$$h^*=1-\frac{1}{R_0}\cdot\frac{4wL}{w^2L^2-\sin^2(wL)}\left[\cos(wc)-\sqrt{\cos^2(wc)-1+\frac{\sin^2(wL)}{w^2L^2}}\right]$$

Using the Taylor series and the fact that $w=\frac{\sqrt{\frac{\pi^2}{4}-2}}{\sigma}$ where $\sigma^2$ is the variance of the sinusoidal kernel,

$$h^*\sim 1-\frac{1}{R_0}\cdot\frac{12wL}{w^4L^4}\left[1-\frac{w^2c^2}{2}-\sqrt{1-w^2c^2-\frac{w^2L^2}{3}}\right]$$

$$h^*=1-\frac{1}{R_0}\cdot\frac{4\sqrt{3}}{L^3(\pi^2-8)^{3/2}}\cdot\sigma\left[8\sqrt{3}\sigma^2-(\pi^2-8)\sqrt{3}c^2-4\sigma\sqrt{12\sigma^2-(\pi^2-8)(3c^2+L^2)}\right]$$

In the case of both kernels, the critical harvesting proportion can be approximated by a function that looks like

$$h^*\sim1- \frac{1}{R_0}\cdot p(L)q(\sigma^2,c^2,L^2+3c^2)$$

where $p(L)$ is a decreasing function of the length of the viable patch $L$.

F) California Current Black Rockfish Parameterization
---------------------------------------------------------
We parameterize our model for black rockfish (*Sebastes melanops*) in the California Current, with MPAs of spacing and width qualitatively similar to those in the Marine Life Projection Act (MLPA), and with a maximum climate velocity equal to that observed empirically. The parameters and references are provided in Table A1.

Our results with this parameterization are qualitatively similar to the results presented in the main text. In particular, we find the same negative relationship between critical harvesting rate $h^*$ and the climate velocity $c$ (Figure A1) and an essentially additive interaction between the effects of the two stressors on biomass (Figure A2). Additionally, our black rockfish parameterization has the same counterintuitive result that MPAs from which effort is displaced (rather than eliminated) can be worse than no MPA at all (compare Figure A4A and A4D).

G) Protected Area fluctuations
---------------------------------
After the simulations come to equilibrium, the fluctuations in total biomass per generation fluctuate more in reserves that are larger and spaced farther apart than simulations in which the reserves that are smaller and more closely spaced (Figure A3). The large reserves have a slightly larger average population, however large reserves here can induce fluctuations of biomass even in deterministic simulations. Thus we expect if reproduction was stochastic, large reserves spaced far apart would be more likely to result in extinction of the population than more closely spaced, smaller reserves. We find the same effect regardless of whether or not effort remains constant or is removed from the system.
